### 2021-06-03 code/aa-flowers-pilot_analysis.Rmd
add pilot data and start analysis

---
title: "flowers-pilot-analysis"
output:
  html_document: default
  pdf_document: default
---

```{r set-up, include=F}
#knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
#knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
#options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
#library(rlang)
#library(lme4)
#library(brms)
#library(rstan)
#rstan_options(auto_write = TRUE)
#options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}
data_target = "pilot0"
##Data import constants
data_location=paste0("data/",data_target)
date_start=lubridate::ymd('2021-05-27')
image_location=paste0("figs/",data_target)
#model_location="code/models"
```

```{r}
# This was for determining participant bonuses using the version of data with PID
# d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% rename(treatmentId=`_id`)
# 
# d.games <- read_csv(here(data_location, 'games.csv')) %>%
#   select(gameId=`_id`, treatmentId, playerIds) %>%
#   left_join(d.treatments) %>%
#   mutate(playerIds=str_split(playerIds,",")) %>%
#   unnest(playerIds) %>%
#   select(playerId=playerIds, name)
# 
# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   left_join(d.games) %>%
#   select(data.bonus, playerId,id,data.bonus,name) %>%
#     filter(!is.na(name)) %>%
#   mutate(pc_bonus=case_when(
#     name=="fourRotate" ~ 3,
#     name=="threeRotate" ~ 1.5,
#     T ~ 0
#   )) %>%
#   mutate(bonus=round(data.bonus+pc_bonus,2),
#          cost=round(bonus*4/3,2)) %>% write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))
```

# Pre-process

```{r}
extract_context_info <- function(context_key){
  return(str_extract(context_key, "[^.]*$"))
}

d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context)

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
    filter(createdAt >= date_start)

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>%
  select(-ends_with('response'), -ends_with('time')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah, -context) %>% 
  left_join(d.contexts %>% select(gameId, 
                                  repNum = data.repNum, 
                                  blockNum = data.blockNum, 
                                  trialNum = data.trialNum, 
                                  response = label, utility, blinded)) %>%
  mutate(time = as.numeric(time)/1000) %>%
  #get individual utilities 
  group_by(gameId, repNum, blockNum, trialNum, response) %>% 
  mutate(n_resp = n(), ind_winnings = as.numeric(utility)*n_resp)

#only include rounds that finished
rounds_exclude <- d.round_results.raw %>% group_by(gameId,numPlayers,repNum) %>% tally() %>% filter(n!=12) %>% select(gameId,repNum)
d.round_results <-  d.round_results.raw %>% anti_join(rounds_exclude)
```

```{r}
summary <- d.round_results.raw %>% group_by(trialNum, repNum, gameId, numPlayers) %>% 
          # mutate(time= time %|% 180) %>% 
  summarize(max_time=max(time)) %>% 
  group_by(gameId, numPlayers) %>% 
  summarize(total_time=sum(max_time)/60,
            num_rounds=max(repNum+1)) %>% 
  arrange(numPlayers)
#knitr::kable(summary)
message("Full games")
summary %>% filter(num_rounds==12) %>% group_by(numPlayers) %>% tally()
message("Partial games")
summary %>% filter(num_rounds!=12) %>% group_by(numPlayers) %>% tally()
```

# Demographics
```{r demographics}
d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```
```{r feedback}
d.exit.survey %>% select( language, fair, chatUseful, feedback, time) %>% knitr::kable()
```

# Pretty pictures

```{r chat}
#group raw chat by participant
d.chat <- read_csv(here(data_location, "raw_chat.csv")) %>% 
 # filter(!is.chitchat) %>% 
  mutate(text = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', text),
         text = str_squish(text),
         #tangram = gsub('/experiment/tangram_', '', target, fixed=TRUE),
         #tangram = gsub('.png', '', tangram, fixed=TRUE),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  filter(type != "alert")%>%
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars)) %>%
  anti_join(rounds_exclude) %>% 
  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum"))
```

Everything here has bootstrapped 95% CIs.

Should find better curves to fit, but using quadratic to allow for some curvature.

```{r}
# ggplot(d.chat, aes(x=repNum, y=total_num_words, color=role))+
#   facet_wrap(~tangram, nrow=2)+
#   scale_color_brewer(palette="Dark2")+
#      stat_summary(fun.data = "mean_cl_boot")+
#   labs(title="Number of words", y="Number of words", x="Round number")+
#   theme(legend.position="bottom")
ggplot(d.chat, aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
  facet_wrap(~role, nrow=1)+
  scale_color_brewer(palette="Dark2")+
    geom_jitter(alpha=.05)+
    geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  #geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
     stat_summary(fun.data = "mean_cl_boot")+
    scale_y_continuous(limits = c(0,50))+
  labs(title="Number of words", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="bottom")
ggsave(here(image_location, 'words.pdf'), width=6, height=4)
```

```{r}
d.chat %>% filter(role=="speaker") %>% 
    mutate(groupxtangram=str_c(gameId,tangram)) %>% 
  group_by(repNum, numPlayers, gameId,tangram, groupxtangram) %>% 
  summarize(words=sum(total_num_words)) %>% 
ggplot(aes(x=repNum, y=words, color=as.factor(numPlayers)))+
  facet_wrap(~numPlayers, nrow=1)+
  scale_color_brewer(palette="Dark2")+
    geom_line(aes(group=groupxtangram), alpha=.1,method=glm, se=F)+
    geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
    #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  labs(title="Words from speaker per tangram", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="null")
ggsave(here(image_location, 'words_lines.pdf'), width=6, height=4)
```

```{r}
d.chat %>% filter(role=="speaker") %>% 
ggplot(aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
  facet_wrap(~tangram)+
  scale_color_brewer(palette="Dark2")+
    #geom_smooth(method=glm, formula=y~poly(x,2), se=T, alpha=.1)+
      geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
       stat_summary(fun.data = "mean_cl_boot", size=.2)+
  labs(title="Tangram variability", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="bottom")
ggsave(here(image_location, 'words_tangrams.pdf'), width=8, height=6)
```

```{r accuracy}
d.round_results %>% group_by(playerId,repNum, gameId, numPlayers) %>% 
  mutate(correct.num=ifelse(correct,1,0)) %>% 
  ggplot(aes(x=repNum, y=correct.num, color=as.factor(numPlayers)))+
geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
  #geom_point()+
  scale_color_brewer(palette="Dark2")+
  #scale_y_continuous(limits = c(0,1))+
  labs(x="Round Number", y="Fraction correctly selected", title= "Overall accuracy increases over repetitions", color="Player count")+
    theme(legend.position="bottom")
ggsave(here(image_location, 'accuracy.pdf'), width=6, height=4)
```

```{r time}
d.round_results %>% group_by(playerId, repNum, gameId, numPlayers) %>% 
  filter(correct==T) %>% 
  #summarize(time=mean(time)) %>% 
  ggplot(aes(x=repNum, y=time, color=as.factor(numPlayers)))+
  geom_jitter(width=.4, height=0, alpha=.03)+
geom_smooth(method = "glm", formula = y~x,
                      method.args = list(family = gaussian(link = 'log')))+
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
  scale_y_continuous(limits = c(0,180))+
    scale_color_brewer(palette="Dark2")+
  labs(x="Round Number", y="Time to selection in seconds",
       title="People choose faster in later rounds", color="Player count")+
  theme(legend.position = "bottom")
ggsave(here(image_location, 'time.pdf'), width=6, height=4)
```
# Models

```{r, include=F}
model_input <- d.chat %>% filter(role=="speaker") %>% 
  mutate(block=repNum,
         words=total_num_words,
         tangram_group=str_c(tangram, gameId))

priors <- c(
  set_prior("normal(20, 20)", class="Intercept"),
  set_prior("normal(0, 10)", class="b"),
  set_prior("normal(0, 10)", class="sd"),
  set_prior("lkj(1)",       class="cor"))
model <- brm(words ~ block * numPlayers + (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "model_1"),                            prior=priors, control=list(adapt_delta=.95))

```

```{r, include=F}
d.prev.speaker <- d.chat %>% ungroup() %>%  filter(role=="speaker") %>% select(gameId,repNum, tangram, total_num_words_prev=total_num_words)
d.prev.round <- d.chat %>% ungroup() %>% select(playerId, correct, tangram, gameId, repNum) %>% 
  left_join(d.prev.speaker) %>% unique() %>% mutate(repNum=repNum+1)
d.chat.lagged <- d.chat %>%
  ungroup() %>% 
  select(gameId, playerId, trialNum, repNum, playerId, role, tangram, total_num_words, numPlayers) %>%
  left_join(d.prev.round) %>%
  mutate(reduction_word=log(total_num_words)-log(total_num_words_prev)) %>%
  filter(repNum>0) %>%
  filter(role=="speaker") %>%
  mutate(prev_correct_round=correct)
model_input <- d.chat.lagged %>% filter(role=="speaker") %>% 
  mutate(block=repNum,
         words=total_num_words,
         tangram_group=str_c(tangram, gameId),
         was_INcorrect=ifelse(!prev_correct_round,1,0))

priors <- c(
  set_prior("normal(20, 20)", class="Intercept"),
  set_prior("normal(0, 10)", class="b"),
  set_prior("normal(0, 10)", class="sd"),
  set_prior("lkj(1)",       class="cor"))
model_speaker_acc <- brm(words ~ block * numPlayers +block*was_INcorrect+ (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "speaker_acc"),                         prior=priors, control=list(adapt_delta=.95))

```

```{r}
summary(model)
```
```{r}
summary(model_speaker_acc)
```

### 2021-06-04 code/linguistic_analysis_C-D.Rmd
1st draft, points C and D from doc

---
title: "linguistic_analysis_C.Rmd : Linguistic quantity and complexity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

## STEP THREE - Linguistic quantity and complexity

```{r main data, include=TRUE}
data_target = "pilot0"
data_location=paste0("data/",data_target)

d_flowers<- read_csv(here(data_location, "raw_chat.csv"))%>%
 filter(!is.na(text))
```

### 1. Utt Length
```{r utt_length}
#1.

d_flowers <- d_flowers %>%
  mutate(utt_length_words = str_count(text, "\\W+") + 1) %>% 
    group_by(gameId, blockNum, trialNum, repNum, playerId)

d_flowers_utt_length <- d_flowers %>%
      summarise(total_num_words = sum(utt_length_words))

ggplot(d_flowers_utt_length, aes(x=repNum, y=total_num_words, color=as.factor(playerId)))+
  scale_color_brewer(palette="Dark2")+
  geom_jitter(alpha=.05)+
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  stat_summary(fun.data = "mean_cl_boot")+
  scale_y_continuous(limits = c(0,50))+
  labs(title="Number of words", y="Number of words", x="Round number", color="playerId")+
  theme(legend.position="bottom")
```

### 2a. Number of turns
```{r number_of_turns}

d_flowers_n_turns <- d_flowers %>%
  summarise(n=n())

ggplot(d_flowers_n_turns, aes(x=repNum, y=n, color=as.factor(playerId)))+
  scale_color_brewer(palette="Dark2")+
  geom_jitter(alpha=.05)+
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  stat_summary(fun.data = "mean_cl_boot")+
  labs(title="Number of turns", y="Number of turns", x="Round number", color="playerId")+
  theme(legend.position="bottom")

```

```{r number_of_turns2}
## 2b. Number of turns
d_flowers_n_utt_speaker <- d_flowers %>%
  group_by(playerId, repNum, gameId, blockNum) %>%
  summarise(n_utt_speaker=n())

ggplot(d_flowers_n_utt_speaker, 
       aes(x=repNum, y=n_utt_speaker, label=playerId)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_wrap(~playerId, nrow=1)+
  xlab("Round number") + 
  ylab("Number of turns") + 
  theme_bw()
```

### 3. Parts of speech
```{r pos}

library("spacyr")

text <- d_flowers$text
parsed <- spacy_parse(text,pos=TRUE) %>%
  select(doc_id, pos) %>%
  group_by(doc_id) %>%
  mutate(pos = paste(pos, collapse=",")) %>%
  distinct()

d_flowers_pos  <- cbind(d_flowers, parsed) %>%
  separate_rows(pos, convert = TRUE) %>%
  group_by(gameId, blockNum, trialNum, repNum, playerId, pos) %>%
  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
  summarise(pos_count=n()) %>%
  left_join(d_flowers_utt_length) %>%
  mutate(prop_pos=pos_count/total_num_words)

ggplot(d_flowers_pos, 
       aes(x=repNum, y=prop_pos, label=playerId)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_wrap(~pos, nrow=1)+
  xlab("Round number") + 
  ylab("Proportion of pos") + 
  theme_bw()

```

### 4. Lexical Diversity (TTR)

```{r ttr}

```

##STEP FOUR - Feedback items and markers

### 1. Distribution of speech across speakers
```{r word_length}

d_flowers_n_utt_speaker <- d_flowers %>%
  group_by(playerId, repNum, gameId, blockNum) %>%
  summarise(n_utt_speaker=n())

d_flowers_n_utt_game <- d_flowers %>%
  group_by(repNum, gameId, blockNum) %>%
  summarise(n_utt_game=n())

d_flowers_n_utt_dist <- d_flowers_n_utt_speaker %>%
  left_join(d_flowers_n_utt_game) %>%
  mutate(overall_turns=n_utt_speaker/n_utt_game)

ggplot(d_flowers_n_utt_dist, 
       aes(x=repNum, y=overall_turns, label=playerId)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  facet_wrap(~playerId, nrow=1)+
  xlab("Round number") + 
  ylab("Proportion of turns") + 
  theme_bw()

```

### 2. Turn-eliciting questions

```{r cars}
library("stringr")

d_flowers_q <- d_flowers %>% 
 mutate(sentence_type = ifelse(str_detect(text, "\\?"), 
                             "question",
                             "other")) %>%
 group_by(playerId, repNum, gameId, blockNum, sentence_type) %>%
 summarise(n_sentence_Type=n()) %>%
 filter(sentence_type=="question") %>%
 left_join(d_flowers_n_utt_speaker) %>%
 mutate(prop_questions = n_sentence_Type/n_utt_speaker )

ggplot(d_flowers_q, 
       aes(x=repNum, y=prop_questions, label=playerId)) + 
  geom_smooth(method = "lm", formula = y~x) + 
  geom_point(alpha=.1) +
  xlab("Round number") + 
  ylab("Proportion of questions") + 
  theme_bw()

```

### 2021-06-07 code/linguistic_analysis_C-D.Rmd
trialNum

@@
 ---
-title: "linguistic_analysis_C.Rmd : Linguistic quantity and complexity"
+title: "linguistic_analysis_C-D.Rmd : Linguistic quantity and complexity"
 output: html_document
@@
 
-ggplot(d_flowers_utt_length, aes(x=repNum, y=total_num_words, color=as.factor(playerId)))+
+ggplot(d_flowers_utt_length, aes(x=trialNum, y=total_num_words, color=as.factor(playerId)))+
   scale_color_brewer(palette="Dark2")+
+  facet_wrap(~gameId, nrow=1) +
   geom_jitter(alpha=.05)+
@@
 
-ggplot(d_flowers_n_turns, aes(x=repNum, y=n, color=as.factor(playerId)))+
+ggplot(d_flowers_n_turns, aes(x=trialNum, y=n, color=as.factor(playerId)))+
+  facet_wrap(~gameId, nrow=1) +
   scale_color_brewer(palette="Dark2")+
@@
 ```
+### 2b. Number of turns
+```{r number_of_turns2}
 
-```{r number_of_turns2}
-## 2b. Number of turns
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, repNum, gameId, blockNum) %>%
+  group_by(playerId, trialNum, gameId, blockNum) %>%
   summarise(n_utt_speaker=n())
@@
 ggplot(d_flowers_n_utt_speaker, 
-       aes(x=repNum, y=n_utt_speaker, label=playerId)) + 
+       aes(x=trialNum, y=n_utt_speaker, label=playerId)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~playerId, nrow=1)+
+  facet_wrap(~ playerId) +
   xlab("Round number") + 
@@
 ggplot(d_flowers_pos, 
-       aes(x=repNum, y=prop_pos, label=playerId)) + 
+       aes(x=trialNum, y=prop_pos, label=playerId)) + 
   geom_smooth(method = "lm", formula = y~x) + 
@@
 
-##STEP FOUR - Feedback items and markers
+## STEP FOUR - Feedback items and markers
 
@@
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, repNum, gameId, blockNum) %>%
+  group_by(playerId, trialNum, gameId, blockNum) %>%
   summarise(n_utt_speaker=n())
@@
 d_flowers_n_utt_game <- d_flowers %>%
-  group_by(repNum, gameId, blockNum) %>%
+  group_by(trialNum, gameId, blockNum) %>%
   summarise(n_utt_game=n())
@@
 
-ggplot(d_flowers_n_utt_dist, 
-       aes(x=repNum, y=overall_turns, label=playerId)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
-  facet_wrap(~playerId, nrow=1)+
-  xlab("Round number") + 
-  ylab("Proportion of turns") + 
-  theme_bw()
+#ggplot(d_flowers_n_utt_dist, 
+#       aes(x=trialNum, y=overall_turns, label=playerId)) + 
+#  geom_smooth(method = "lm", formula = y~x) + 
+#  geom_point(alpha=.1) +
+#  facet_wrap(~playerId)+
+#  xlab("Round number") + 
+#  ylab("Proportion of turns") + 
+#  theme_bw()
 
@@
 
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(playerId, trialNum, gameId, blockNum) %>%
+  summarise(n_utt_speaker=n())
+
 d_flowers_q <- d_flowers %>% 
@@
                              "other")) %>%
- group_by(playerId, repNum, gameId, blockNum, sentence_type) %>%
+ group_by(playerId, trialNum, gameId, blockNum, sentence_type) %>%
  summarise(n_sentence_Type=n()) %>%
@@
 ggplot(d_flowers_q, 
-       aes(x=repNum, y=prop_questions, label=playerId)) + 
+       aes(x=trialNum, y=prop_questions, label=playerId)) + 
   geom_smooth(method = "lm", formula = y~x) + 
+  facet_wrap(~ playerId) +
   geom_point(alpha=.1) +

### 2021-06-07 code/linguistic_analysis_C-D.Rmd
ttr

@@
 library(here)
+library(magrittr)
+library(qdap)
+library("stringr")
 ```
@@
 
+ttr <- function(text_){
+  text_ <- gsub('[[:punct:] ]+',' ',text_) ####plus remove caps
+  ttr_<- type_token_ratio(text__, n.words=10)
+  return(ttr_)
+}
+
+d_flowers_text <- d_flowers %>%
+    group_by(gameId, blockNum, trialNum, repNum, playerId) %>%
+    mutate(text_block = paste0(text, collapse = " ")) %>%
+    select(-c(text)) %>%
+    distinct()
+
+text_block <- d_flowers_text$text_block
+
+#d_flowers_ttr <- text_block %>%
+#  mutate(ttr_v=ttr(text_block))
+
 ```
@@
 ```{r cars}
-library("stringr")
 

### 2021-06-08 code/linguistic_analysis_C-D.Rmd
blockNum and ttr

--- 2021-06-07 code/linguistic_analysis_C-D.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
 library(tidyverse)
 library(here)
+library("stringr")
 ```
@@
  filter(!is.na(text))
+
 ```
@@
 ```{r utt_length}
-#1.
-
-d_flowers <- d_flowers %>%
-  mutate(utt_length_words = str_count(text, "\\W+") + 1) %>% 
-    group_by(gameId, blockNum, trialNum, repNum, playerId)
 
 d_flowers_utt_length <- d_flowers %>%
-      summarise(total_num_words = sum(utt_length_words))
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, blockNum, trialNum, repNum, playerId)  %>%
+  summarise(total_num_words = sum(utt_length_words),
+            mlu = mean(utt_length_words))
 
-ggplot(d_flowers_utt_length, aes(x=trialNum, y=total_num_words, color=as.factor(playerId)))+
+ggplot(d_flowers_utt_length, aes(x=repNum, y=total_num_words, color=as.factor(playerId)))+
   scale_color_brewer(palette="Dark2")+
-  facet_wrap(~gameId, nrow=1) +
+  facet_grid(cols = vars(blockNum), rows =vars(gameId)) +
   geom_jitter(alpha=.05)+
@@
   scale_y_continuous(limits = c(0,50))+
-  labs(title="Number of words", y="Number of words", x="Round number", color="playerId")+
-  theme(legend.position="bottom")
+  labs(title="Number of words", y="Total number of words", x="Round number", color="playerId")+
+  theme(legend.position="bottom") + 
+  theme_bw()
+
+ggplot(d_flowers_utt_length, aes(x=repNum, y=mlu, color=as.factor(playerId)))+
+  scale_color_brewer(palette="Dark2")+
+  facet_grid(cols = vars(blockNum), rows =vars(gameId)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  scale_y_continuous(limits = c(0,15))+
+  labs(title="MLU", y="MLU", x="Round number", color="playerId")+
+  theme(legend.position="bottom") + 
+  theme_bw()
+
 ```
 
-### 2a. Number of turns
+### 2. Number of turns
 ```{r number_of_turns}
@@
 d_flowers_n_turns <- d_flowers %>%
+  group_by(gameId, blockNum, trialNum, repNum, playerId)  %>%
   summarise(n=n())
@@
 ggplot(d_flowers_n_turns, aes(x=trialNum, y=n, color=as.factor(playerId)))+
-  facet_wrap(~gameId, nrow=1) +
+  facet_grid(rows =vars(gameId)) +
   scale_color_brewer(palette="Dark2")+
@@
   labs(title="Number of turns", y="Number of turns", x="Round number", color="playerId")+
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") + 
+  theme_bw()
 
 ```
+
+#second plot
 ### 2b. Number of turns
+
 ```{r number_of_turns2}
@@
   geom_point(alpha=.1) +
-  facet_wrap(~ playerId) +
+  facet_grid(~playerId) +
   xlab("Round number") + 
@@
 
+ttr <- function(text_){
+  text_ <- tolower(text_)
+  total_words = sapply(strsplit(text_, " "), length)
+  no_duplicates = vapply(lapply(strsplit(text_, " "), unique), paste, character(1L), collapse = " ")
+  total_unique = str_count(no_duplicates, "\\W+") + 1
+  ttr_ = total_unique/total_words
+  return(ttr_)
+}
+
+d_flowers_text <- d_flowers %>%
+    group_by(gameId, blockNum, trialNum, repNum, playerId) %>%
+    mutate(text_block = paste0(text, collapse = " ")) %>%
+    select(-c(text)) %>%
+    distinct()
+
+d_flowers_ttr <- d_flowers_text %>%
+  mutate(ttr_ = ttr(text_block))
+
+#t<-str_split(text_, " ", 1:6)
+#y<- paste(unlist(lapply(t,head,n=5)), collapse=" ")
+
+ggplot(d_flowers_ttr, 
+       aes(x=trialNum, y=ttr_, label=playerId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(gameId)) +
+  xlab("Round number") + 
+  ylab("Type_token ratio") + 
+  theme_bw()
+
 ```
@@
 
-#ggplot(d_flowers_n_utt_dist, 
-#       aes(x=trialNum, y=overall_turns, label=playerId)) + 
-#  geom_smooth(method = "lm", formula = y~x) + 
-#  geom_point(alpha=.1) +
-#  facet_wrap(~playerId)+
-#  xlab("Round number") + 
-#  ylab("Proportion of turns") + 
-#  theme_bw()
+ggplot(d_flowers_n_utt_dist, 
+       aes(x=trialNum, y=overall_turns, label=playerId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_wrap(~playerId)+
+  xlab("Round number") + 
+  ylab("Proportion of turns") + 
+  theme_bw()
 

### 2021-06-08 code/fit_alignment.R
fit alignment code

@@ -1,0 +1,67 @@
+library(ggplot2)
+library(dplyr)
+library(langcog)
+library(readr)
+library(tidyr)
+library(stringr)
+library(magrittr)
+library(directlabels)
+library(lubridate)
+library(lme4)
+library(rstan)
+library(purrr)
+
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+result_files <-  list.files(path = "/Users/lscpuser/Downloads",recursive = TRUE,
+                            all.files = FALSE, full.names = TRUE,
+                            pattern = "output.csv")
+
+setup_alignment_data <- function(corpus_df) {
+  corpus_df %>%
+    group_by(category, subpop) %>%
+    summarise_each(funs(sum), ba, nba, bna, nbna) %>%
+    filter(ba + bna > 1) %>%
+    ungroup() %>%
+    mutate(category_num = as.numeric(as.factor(category)),
+           subpop_num = as.numeric(as.factor(subpop))) 
+}
+
+
+childes_data <- lapply(result_files, read_csv) %>%
+  bind_rows() %>%
+  rename(Speaker = speakerId, Replier = replierId) %>%
+  mutate(subpop = paste(Speaker,  "-", Replier)) %>%
+  filter(Speaker != Replier) 
+
+childes_data <- childes_data %>%
+  setup_alignment_data()
+
+fit_alignment <- function(alignment_df, sd = .5) {
+  
+alignment_data <- list(NumMarkers = length(unique(alignment_df$category)),
+                         NumSubPops = length(unique(alignment_df$pop_)),
+                         NumObservations = nrow(alignment_df),
+                         SpeakerSubPop = alignment_df$subpop_num,
+                         MarkerType = alignment_df$category_num,
+                         NumUtterancesAB = alignment_df$ba + alignment_df$nba,
+                         NumUtterancesNotAB = alignment_df$bna+alignment_df$nbna,
+                         CountsAB = alignment_df$ba,
+                         CountsNotAB = alignment_df$bna,
+                         StdDev=sd )
+
+
+fit_model <- sampling(alignment_model, data = alignment_data, 
+                      iter = 500, warmup = 100, chains = 3 ,
+                      control = list(adapt_delta = .95))
+
+return(fit_model) 
+}
+
+
+alignment_model = stan_model(file = "~/Documents/stanmodel.stan")
+
+alignment_data <- fit_alignment(childes_data) 
+
+

### 2021-06-09 code/fit_alignment.R
path file

@@ -14,9 +14,10 @@ library(purrr)
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-result_files <-  list.files(path = "/Users/lscpuser/Downloads",recursive = TRUE,
-                            all.files = FALSE, full.names = TRUE,
-                            pattern = "output.csv")
+result_files <- "data/pilot0/output.csv"
+#result_files <-  list.files(path = "../data/pilot0/",recursive = TRUE,
+ #                           all.files = FALSE, full.names = TRUE,
+  #                          pattern = "output.csv")
 
 setup_alignment_data <- function(corpus_df) {
   corpus_df %>%

### 2021-06-09 code/fit_alignment.R
figures

@@ -42,7 +42,7 @@ childes_data <- childes_data %>%
 fit_alignment <- function(alignment_df, sd = .5) {
   
 alignment_data <- list(NumMarkers = length(unique(alignment_df$category)),
-                         NumSubPops = length(unique(alignment_df$pop_)),
+                         NumSubPops = length(unique(alignment_df$subpop)),
                          NumObservations = nrow(alignment_df),
                          SpeakerSubPop = alignment_df$subpop_num,
                          MarkerType = alignment_df$category_num,
@@ -52,17 +52,72 @@ alignment_data <- list(NumMarkers = length(unique(alignment_df$category)),
                          CountsNotAB = alignment_df$bna,
                          StdDev=sd )
 
-
-fit_model <- sampling(alignment_model, data = alignment_data, 
-                      iter = 500, warmup = 100, chains = 3 ,
-                      control = list(adapt_delta = .95))
-
-return(fit_model) 
+return(alignment_data) 
 }
+
+alignment_data <- fit_alignment(childes_data) 
 
 
 alignment_model = stan_model(file = "~/Documents/stanmodel.stan")
 
-alignment_data <- fit_alignment(childes_data) 
+fit<- sampling(alignment_model, data = alignment_data, 
+                iter = 500, warmup = 100, chains = 3 ,
+                control = list(adapt_delta = .95))
+
+#fit <- stan(file = alignment_model, data = alignment_data, 
+ #           iter = 500, chains =1 )
+
+mu_notab <- colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)
+mu_ab <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)
+
+d2<- childes_data
+#If the model works, its mu_ values should be "smoothed" estimates of p(B|A) & p(B|notA)
+par(mar = rep(2, 4))
+plot(mu_ab,d2$ba/(d2$ba+d2$nba))
+plot(mu_notab,d2$bna/(d2$bna+d2$nbna))
+
+#checking in on the learned subpopulation values
+eta_ab_subpop <- colMeans(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
[…]

### 2021-07-07 code/aa-flowers-pilot_analysis.Rmd
add technical pilots

@@
 library(here)
+
+theme_set(theme_bw())
+
+data_target = "technical_pilots"
+##Data import constants
+data_location=paste0("data/",data_target)
+date_start=lubridate::ymd('2021-06-30')
+image_location=paste0("figs/",data_target)
+
 #library(rlang)
@@
 #options(mc.cores = parallel::detectCores())
-theme_set(theme_bw())
-ParseJSONColumn <- function(x) {
-  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
-    fromJSON(flatten = T)
-}
-data_target = "pilot0"
-##Data import constants
-data_location=paste0("data/",data_target)
-date_start=lubridate::ymd('2021-05-27')
-image_location=paste0("figs/",data_target)
 #model_location="code/models"
+
 ```
@@
 
-```{r}
+We need:
+1) game info with treatment parameters
+2) context for each round (what were participant's options? max utility based on game conditions?)
+3) round results (what did participants select each round?)
+4) chat (what did individual participants say each round)
+5) feedback/demographics
+
+```{r helper-functions}
+ParseJSONColumn <- function(x) {
+  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
+    fromJSON(flatten = T)
+}
+
 extract_context_info <- function(context_key){
@@
 }
+```
 
+Get info about the games and treatments
+- We need each game, the 
+```{r game-preprocessing}
+d.games.raw <- read_csv(here(data_location, 'games.csv')) %>% 
+  rename(gameId = `_id`) %>% 
+  filter(createdAt >= date_start) %>%
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+
+d.treatments.raw <- read_csv(here(data_location, 'treatments.csv')) %>% 
+  rename(treatmentId = `_id`) %>%
+  #filter(createdAt >= date_start) %>%
+  mutate(factorIds = str_split(factorIds, pattern=",")) %>%
+  unnest(factorIds) %>% rename(factorId = factorIds)
+
+d.factors.raw <- read_csv(here(data_location, 'factors.csv'))%>% 
+  rename(factorId = `_id`)
+
+d.factor_types.raw <- read_csv(here(data_location, 'factor-types.csv')) %>%
+  rename(factorTypeId = "_id",
+         factorName = "name")
+
+d.factors <- d.factors.raw %>% 
+  left_join(d.factor_types.raw %>% 
+              select(factorTypeId, factorName))
+
+d.treatments <- d.treatments.raw %>% 
+  left_join(d.factors %>% 
+              select(factorId, factorName, value)) %>%
+  pivot_wider(id_cols = c(treatmentId, name, createdAt), 
+              names_from = factorName, values_from = value)
+
+d.players <- read_csv(here(data_location, 'players.csv')) %>%
+  rename(playerId = `_id`) %>% 
+  filter(createdAt >= date_start) %>%
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+
+d.games <- d.games.raw %>% left_join(d.treatments %>% 
+                                       select(-createdAt)) %>%
+  janitor::remove_empty() %>%
+  #select out the nonuseful columns
+  select(-c(gameLobbyId, treatmentId, batchId, targetSet, justStarted, submitCode, roundIds)) %>%
+  separate(col = playerIds, sep = ",", into = c("player_1", "player_2", "player_3")) %>% 
+  pivot_longer(cols = starts_with("player_"), names_to = "playerNum", values_to = "playerId") %>%
+  left_join(d.players %>% select(playerId, readyAt)) %>%
+  mutate(game_length = finishedAt - readyAt) %>%
+  group_by(across(gameId:chatEnabled)) %>%
+  summarize(gameLength = max(game_length)) %>% ungroup()
+```
+
+Round info
+For each round, we want to know the context...
+- which flowers were available
+- the utility of each flower
+
+We also want to have a report for each round including...
+- which participants selected which flowers
+- the time the participants took to select
+- the utility of the flower the participants selected
+
+```{r}
+d.round_results.raw <- read_csv(here(data_location,'rounds.csv'), guess_max=10000) %>% 
+  filter(createdAt >= date_start) %>% 
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
+  rename_with ( ~ gsub("room", "player", .x, fixed=T))
+
+#try pivot_longer instead of gather??
+d.round_results <- d.round_results.raw  %>% 
+  select(-chat, -context) %>%
+  mutate_at(vars(starts_with('player')), as.character) %>%
+  pivot_longer(cols = starts_with('player'), 
+               names_to = c('blah', 'playerId', 'info'), 
+               values_to = "value",
+               names_sep = "_",
+               values_drop_na = TRUE) %>%
+  pivot_wider(names_from = "info", values_from = "value") %>%
+  select(-blah) %>%
+  mutate_at(c("time", "utility"), as.numeric) %>%
+  #TODO: some times are negative?? interesting
+  mutate(time = abs(as.numeric(time)/1000)) %>%
+  rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
+  arrange(createdAt, gameId, blockNum, repNum)
+
+#exclude blocks with <12 completed rounds
+rounds_exclude <- d.round_results %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
+d.round_results.final <-  d.round_results %>% anti_join(rounds_exclude)
+```
+```{r}
+summary <- d.round_results %>% 
+  group_by(gameId) %>% 
+  summarize(num_rounds=max(repNum+1)) %>% 
+  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength))
+#knitr::kable(summary)
+message("Full games")
+summary %>% filter(num_rounds==12)
+message("Partial games")
+summary %>% filter(num_rounds!=12) 
+```
+
+```{r}
+d.chat.raw <- d.round_results.raw %>%
+  filter(createdAt >= date_start) %>%
+  anti_join(rounds_exclude) %>%
+  #there will always be an alert for each active round, so drop NA
+  filter(!is.na(chat))%>%
+  select(-c(ends_with('response'), ends_with('time'), ends_with('utility'), context, stageIds)) %>%
+  #mutate(chat = ifelse(is.na(chat), '{}', chat)) %>%
+  rename(roundID = `_id`) %>%
+  mutate(chat = map(chat, .f = ParseJSONColumn)) %>%
+  unnest(chat) %>%
+  arrange(createdAt, gameId, blockNum, repNum) %>% 
+  left_join(d.players %>% select(playerId, name)) %>%
+  left_join(d.round_results %>% select(gameId, blockNum, repNum, trialNum, playerId, playerResponse, playerUtility)) %>%
+  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
+  mutate("participantAction" = ifelse(type == "alert", paste(name, "selected", playerResponse, "for", playerUtility, "point(s)"), NA)) %>%
+  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, participantAction, everything()) %>%
+  write_csv(here(data_location, 'raw_chat.csv'))
+
+#extract the alternative flowers
 d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
@@
                               spread(key, value))) %>%
-  unnest(data.context)
-
-d.games <- read_csv(here(data_location, 'games.csv')) %>% 
-  rename(gameId = `_id`) %>% 
-    filter(createdAt >= date_start)
-
-d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
-  filter(createdAt >= date_start) %>%
-  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
-  rename(row_id = `_id`) %>%
-  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
-  unnest(data.chat) %>%
-  select(-ends_with('response'), -ends_with('time')) %>%
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
-  write_csv(here(data_location, 'raw_chat.csv'))
-
-d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
-  filter(createdAt >= date_start) %>% 
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
-  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
-  select(-chat) %>% 
-  gather(key, value, starts_with('player')) %>% 
-  separate(key, into = c('blah', 'playerId', 'info')) %>% 
-  spread(info, value) %>% 
-  select(-blah, -context) %>% 
-  left_join(d.contexts %>% select(gameId, 
-                                  repNum = data.repNum, 
-                                  blockNum = data.blockNum, 
-                                  trialNum = data.trialNum, 
-                                  response = label, utility, blinded)) %>%
-  mutate(time = as.numeric(time)/1000) %>%
-  #get individual utilities 
-  group_by(gameId, repNum, blockNum, trialNum, response) %>% 
-  mutate(n_resp = n(), ind_winnings = as.numeric(utility)*n_resp)
-
-#only include rounds that finished
-rounds_exclude <- d.round_results.raw %>% group_by(gameId,numPlayers,repNum) %>% tally() %>% filter(n!=12) %>% select(gameId,repNum)
-d.round_results <-  d.round_results.raw %>% anti_join(rounds_exclude)
+  unnest(data.context) %>%
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
 ```
-
-```{r}
-summary <- d.round_results.raw %>% group_by(trialNum, repNum, gameId, numPlayers) %>% 
-          # mutate(time= time %|% 180) %>% 
-  summarize(max_time=max(time)) %>% 
-  group_by(gameId, numPlayers) %>% 
-  summarize(total_time=sum(max_time)/60,
-            num_rounds=max(repNum+1)) %>% 
-  arrange(numPlayers)
-#knitr::kable(summary)
-message("Full games")
-summary %>% filter(num_rounds==12) %>% group_by(numPlayers) %>% tally()
-message("Partial games")
-summary %>% filter(num_rounds!=12) %>% group_by(numPlayers) %>% tally()
-```
+How often do participants select a flower that is hidden from them?
 
@@
 
+Correlation between response time and utility?
+
+```{r}
+d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
+  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+         full_condition = paste0(condition, "-", chat)) %>%
+  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
+  facet_grid(rows = vars(condition), cols = vars(chat)) +
+  geom_point(alpha = .25) + 
+  labs(x = "log time (in seconds)",
+       y = "individual utility",
+       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
+```
+
+Easiest way to do this, is to derive the game start date with the latest participant ready time and the game finishedAt
+
+```{r}
+d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
+  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+         full_condition = paste0(condition, "-", chat)) %>%
+  group_by(gameId, condition, chat, full_condition,gameLength) %>%
+  summarize(totalUtility = sum(playerUtility))%>%
+  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
+  facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
+  geom_point(alpha = .75) + 
+  labs(x = "Game length (in minutes)",
+       y = "total points by group",
+       title = "Length of game in mins and utility", subtitle =  "per game")
+
+d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
+  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+         full_condition = paste0(condition, "-", chat)) %>%
+  group_by(gameId, condition, chat, full_condition,gameLength) %>%
+  summarize(totalUtility = sum(playerUtility))%>%
+  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
+  ggplot(aes(x=log(as.numeric(gameLength, units="mins")), y = as.numeric(totalUtility), color = full_condition)) + 
+  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
+  geom_point(alpha = .75) + 
+  geom_smooth(method=lm, alpha=.3)+
+  labs(x = "Game length (in minutes)",
+       y = "total points by group",
+       title = "Length of game in mins and utility", subtitle =  "per game")
+```
+
 ```{r chat}
 #group raw chat by participant
-d.chat <- read_csv(here(data_location, "raw_chat.csv")) %>% 
+d.chat <- read_csv(here(data_location, "raw_chat.csv"), guess_max=1000) %>% 
+  filter(type == "message") %>%
+  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum")) %>%
  # filter(!is.chitchat) %>% 
@@
          text = str_squish(text),
-         #tangram = gsub('/experiment/tangram_', '', target, fixed=TRUE),
-         #tangram = gsub('.png', '', tangram, fixed=TRUE),
          utt_length_chars = str_length(text), 
          utt_length_words = str_count(text, "\\W+") + 1) %>% 
-  filter(type != "alert")%>%
   group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
@@
             total_num_chars = sum(utt_length_chars)) %>%
-  anti_join(rounds_exclude) %>% 
-  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum"))
+  anti_join(rounds_exclude)
+
 ```
+
+Pick just one game to look at, make sure the plot is working?
+```{r}
+condition = "coopCartel"
+
+d.chat %>% 
+         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
+         filter(condition == "coopCartel",
+                chatEnabled == "true") %>%
+ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
+  geom_jitter(alpha = .5) + 
+  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  theme(legend.position = "none")+
+  labs(title = "Cooperative + Lang", subtitle="Number of words", y="Number of words", x="Round number", color="playerId")
+
+d.chat %>% 
+         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
+         filter(condition == "competCartel",
+                chatEnabled == "true") %>%
+ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
+  geom_jitter(alpha = .5) + 
+  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  theme(legend.position = "none")+
+  labs(title="Competitive + Lang", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
+```
+
+Accuracy
+sum utility for entire group/max possible utility per group (or, the max flower *3)
+Each condition is going to have a different maximum utility
+
+coopcartel will have a max utility = max flower * 3 (players) * 3 (incentive)
+
+competcartel will have a max utility = sum(top_3_flowers)
+```{r}
+d.max_utility <- d.contexts %>% 
+  left_join(d.games %>% select(gameId, condition)) %>%
+  mutate(utility= as.numeric(utility)) %>%
+  group_by(gameId, blockNum, repNum) %>% 
+  slice_max(order_by =utility,n = 3)%>%
+  summarize(competCartel = sum(utility),
+            coopCartel = max(utility * 9)) %>% ungroup() %>%
+  pivot_longer(cols = c(competCartel, coopCartel), names_to = "condition", values_to = "maxUtility")
+
+d.utility <- d.round_results %>%
+         left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% 
+  group_by(gameId, blockNum, repNum, trialNum,condition, chatEnabled) %>% 
+  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
+  left_join(d.max_utility) %>%
+  mutate(prop_utility = group_utility/maxUtility,
+         chat = if_else(chatEnabled=="true", "Lang", "Nonlang"))
+
+d.utility %>% 
+  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
+  geom_jitter(alpha = .4) + 
+  facet_grid(cols = vars(condition), rows = vars(chat), scales = "free_y") + 
+  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  theme(legend.position = "none")+
+  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
+```
+
+-----
 

### 2021-07-09 experiments/flowers-pilot/server/callbacks.js
add coop choose 3 version

@@ -59,23 +59,46 @@ Empirica.onStageEnd((game, round, stage) => {
     const tangrams= round.get('context')
     const scale = game.treatment.scale || 1
     const type = game.treatment.condition
-    _.forEach(_.keys(tangrams), tangram => {
-      let click=[]
+    // 
+    if (type=="coopMulti"){ // this depends on total clicks
+      let rawScore=0
+      let clicked=[]
       _.forEach(players, p => {
-        if (p.get("clicked")==tangram){click.push(p)}
+        clicked.push(p.get("clicked"))
       })
-      let rawScore=tangrams[tangram]["utility"]
-      if (type=="coopCartel"){rawScore=rawScore*click.length}
-      if (type=="competCartel"){rawScore=rawScore/click.length}
+      clicked=_.uniq(clicked) // no duplicates
+      _.forEach(clicked, tangram => {
+        rawScore=rawScore+tangrams[tangram]["utility"]
+      })
+      rawScore=rawScore/players.length
       let scoreIncrement=rawScore*scale*.01
-      _.forEach(click, player=> {
+      _.forEach(players, player => {
         const currScore = player.get("bonus") || 0;      
         player.set("bonus", scoreIncrement + currScore);
         player.set("scoreIncrement", scoreIncrement)
-        player.set("clickLength", click.length)
+        player.set("clickLength", clicked.length) //how many unique flowers were clicked
         player.set("rawScore", rawScore)
       })
-    })   
+    }
+    else{ // these conditions score depends only on others who clicked the same thing
+      _.forEach(_.keys(tangrams), tangram => {
+        let click=[]
+        _.forEach(players, p => {
+          if (p.get("clicked")==tangram){click.push(p)}
+        })
+        let rawScore=tangrams[tangram]["utility"]
+        if (type=="coopCartel"){rawScore=rawScore*click.length}
+        if (type=="competCartel"){rawScore=rawScore/click.length}
+        let scoreIncrement=rawScore*scale*.01
+        _.forEach(click, player=> {
+          const currScore = player.get("bonus") || 0;      
+          player.set("bonus", scoreIncrement + currScore);
+          player.set("scoreIncrement", scoreIncrement)
+          player.set("clickLength", click.length)
+          player.set("rawScore", rawScore)
+        })
+        })
+    }  
     players.forEach(player => {
       round.set('player_' + player._id + '_response', player.get('clicked'));
       round.set('player_' + player._id + '_time', player.get('timeClick'));

### 2021-07-19 experiments/flowers-pilot/client/main.js
finish adding new multi instructions

@@ -15,16 +15,21 @@ import Overview from "./intro/Overview.jsx";
 import SocialInteractionDetails from "./intro/SocialInteractionDetails.jsx";
 
 import BasicInterfaceInstructions from "./intro/instructions/BasicInterfaceInstructions.jsx";
-import BasicInterfaceQuiz from "./intro/quizes/BasicInterfaceQuiz.jsx";
+import BasicInterfaceQuiz from "./intro/quizzes/BasicInterfaceQuiz.jsx";
 
 import AdvancedInterfaceInstructions from "./intro/instructions/AdvancedInterfaceInstructions.jsx";
-import AdvancedInterfaceQuiz from "./intro/quizes/AdvancedInterfaceQuiz.jsx";
+import AdvancedInterfaceQuiz from "./intro/quizzes/AdvancedInterfaceQuiz.jsx";
 
+import GameplayInstructions from "./intro/instructions/GameplayInstructions.jsx";
+import GameplayQuiz from "./intro/quizzes/GameplayQuiz.jsx";
+
+import FinalQuizLang from "./intro/quizzes/FinalQuizLang.jsx";
+import FinalQuizNonLang from "./intro/quizzes/FinalQuizNonLang.jsx";
 
 import MoreAboutBonus from "./intro/MoreAboutBonus.jsx";
 import UIOverview from "./intro/UIOverview.jsx";
-import GroupQuizCompet from "./intro/GroupQuizCompet.jsx";
-import GroupQuizCoop from "./intro/GroupQuizCoop.jsx";
+//import GroupQuizCompet from "./intro/GroupQuizCompet.jsx";
+//import GroupQuizCoop from "./intro/GroupQuizCoop.jsx";
 
 import Round from "./game/Round.jsx";
 import Thanks from "./exit/Thanks.jsx";
@@ -33,7 +38,7 @@ import ExitSurvey from "./exit/ExitSurvey";
 
 import customBreadcrumb from "./game/Breadcrumb.jsx";
 
-Empirica.header(() => null);
+//Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
 Empirica.consent(Consent);
@@ -42,13 +47,23 @@ Empirica.consent(Consent);
 // At this point they have been assigned a treatment. You can return
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
+
[…]

### 2021-07-19 experiments/flowers-pilot/client/main.js
final intro tweaks

@@ -38,7 +38,7 @@ import ExitSurvey from "./exit/ExitSurvey";
 
 import customBreadcrumb from "./game/Breadcrumb.jsx";
 
-//Empirica.header(() => null);
+Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
 Empirica.consent(Consent);

### 2021-07-21 code/linguistic_analysis_C-D.Rmd
add feedback code

@@
 library("stringr")
+library(Hmisc)
+library("spacyr")
+library(ggrepel)
+
 ```
@@
 ```{r main data, include=TRUE}
+d<- read_csv("~/Downloads/raw_chat.csv")
+d <- d %>% filter(type=="message")
+write_csv(d, "~/Downloads/raw_chat1.csv")
+```
+
+```{r main data1, include=TRUE}
 data_target = "pilot0"
@@
 
-d_flowers<- read_csv(here(data_location, "raw_chat.csv"))%>%
+d_flowers<- read_csv(here(data_location, "raw_chat1.csv"))%>%
  filter(!is.na(text))
@@
   theme(legend.position="bottom") + 
-  theme_bw()
+  theme_bw() +
+  ggtitle("Number of words per Round")
 
@@
   theme(legend.position="bottom") + 
-  theme_bw()
+  theme_bw()+
+  ggtitle("MLU per Round")
 
@@
   theme(legend.position="bottom") + 
-  theme_bw()
+  theme_bw() +
+  ggtitle("Number of turns per Round")
 
@@
   xlab("Round number") + 
-  ylab("Number of turns") + 
+  ylab("Number of utterances") + 
   theme_bw()
+
 ```
@@
 ```{r pos}
-
-library("spacyr")
 
@@
   ylab("Type_token ratio") + 
-  theme_bw()
+  theme_bw()+
+  ggtitle("TTR per Round")
 
@@
   ylab("Proportion of turns") + 
-  theme_bw()
+  theme_bw()+
+  ggtitle("Proportion of turns per Round")
 
@@
 
-```{r cars}
+```{r questions}
 
@@
   ylab("Proportion of questions") + 
+  theme_bw()+
+  ggtitle("Proportion of questions per Round")
+
+```
+
+### 1. Distribution of speech across speakers
+```{r feedback}
+
+###politeness markers
+
+###backchanneling: (https://www.reading.ac.uk/AcaDepts/ll/app_ling/internal/Cutrone_vol_2.pdf)
+
+backchanneling<- c("mm", "okay", "uh-huh", "ok", "mm-hm", "uh", "um", "agree", "uhuh", "mmm", "wow", "great", "mm", "hm",  "ummm", "hmmmm", "huh", "un", "um", "ohh", "ooo", "see", "oooo", "ununun", "oh", "ah", "true", "agree", "right", "yeah", "good", "really")
+
+###hedges: (https://www.researchgate.net/publication/280125979_Linguistic_Markers_and_Stylistic_Attributes_of_Hedging_in_English_Academic_Papers_Written_by_Native_and_Non-Native_Speakers_of_English/figures)
+
+hedging <- c("may", "perhaps", "might", "possible", "likely", "possibly", "maybe", "probable", "appear", "seem", "suggest", "sometimes", "seemingly", "apparently", "often", "could", "usually", "likely", "tend", "sometimes", "probably", "primarily", "tendency", "largely")
+
+d_flowers_feedback <- d_flowers %>%
+  group_by(playerId, trialNum, gameId, blockNum) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(text = strsplit(text, " ")) %>%
+  rowwise() %>%
+  mutate(backchannel=list(intersect(backchanneling, text))) %>%
+  mutate(backchannel_length=length(backchannel))%>%
+  mutate(hedge=list(intersect(hedging, text))) %>%
+  mutate(hedge_length=length(hedge))
+
+ggplot(d_flowers_feedback, aes(y=trialNum , x=backchannel_length, label=playerId)) +
+  facet_wrap(~ playerId) +
+  geom_bar(stat="identity")
+
+ggplot(d_flowers_feedback, aes(y=trialNum , x=hedge_length, label=playerId)) +
+  facet_wrap(~ playerId) +
+  geom_bar(stat="identity")
+
+ggplot(d_flowers_feedback, aes(y=trialNum , x=backchannel_length, label=playerId)) +
+  geom_point(alpha = .5)+
+  geom_smooth() +
+  geom_text_repel(aes(label=ifelse(backchannel=="character(0)",  "", backchannel), max.overlaps = 20))+
+  xlim(c(1,5))+
+  theme_bw()
+
+ggplot(d_flowers_feedback, aes(y=trialNum , x=hedge_length, label=playerId)) +
+  geom_point(alpha = .5)+
+  geom_smooth() +
+  geom_text_repel(aes(label=ifelse(hedge=="character(0)", "", hedge), max.overlaps = 20))+
+  xlim(c(1,5))+
   theme_bw()
@@
 ```
+
+```{r minimum_edit_distance, include=TRUE}
+
+```

### 2021-07-21 code/fit_alignment.R
update alignment figures

@@ -1,5 +1,11 @@
 library(ggplot2)
 library(dplyr)
+
+
+#install.packages("devtools")
+#devtools::install_github("langcog/langcog")
+
+
 library(langcog)
 library(readr)
 library(tidyr)
@@ -14,7 +20,7 @@ library(purrr)
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-result_files <- "data/pilot0/output.csv"
+result_files <- "data/pilot0/output_alignment.csv"
 #result_files <-  list.files(path = "../data/pilot0/",recursive = TRUE,
  #                           all.files = FALSE, full.names = TRUE,
   #                          pattern = "output.csv")
@@ -57,8 +63,9 @@ return(alignment_data)
 
 alignment_data <- fit_alignment(childes_data) 
 
+alignment_model = stan_model(file = "code/stanmodel.stan")
 
-alignment_model = stan_model(file = "~/Documents/stanmodel.stan")
+#alignment_model = stan_model(file = "~/Documents/stanmodel.stan")
 
 fit<- sampling(alignment_model, data = alignment_data, 
                 iter = 500, warmup = 100, chains = 3 ,
@@ -78,7 +85,7 @@ plot(mu_notab,d2$bna/(d2$bna+d2$nbna))
 
 #checking in on the learned subpopulation values
 eta_ab_subpop <- colMeans(rstan:::extract(fit,"eta_ab_subpop")$eta_ab_subpop)
-eta_subpop <- colMeans(rstan:::extract(fit,"eta_subpop")$eta_subpop) ######this doesn't exist
+#eta_subpop <- colMeans(rstan:::extract(fit,"eta_subpop")$eta_subpop) ######this doesn't exist
 
 d3 <- d2 %>%
   mutate(model_eta = colMeans(rstan:::extract(fit,"eta_ab_observation")$eta_ab_observation))
@@ -101,15 +108,14 @@ ggplot(aes(x = mean, y = subpop), data = d4) +
   theme_bw(base_size = 14) +
   theme(panel.grid = element_blank())
 
-d4 <- d3 %>%
+d5 <- d3 %>%
   group_by(subpop, category) 
 
 #This is the plot for results/children_eta_bywords.pdf
 ggplot(aes(y = subpop, x = model_eta, colour = category,
-), data=d4) + 
+), data=d5) + 
   geom_point()+
   geom_smooth(method = "loess") +
[…]

### 2021-07-21 experiments/flowers-pilot/client/main.js
fix sorry logic to go to survey

@@ -85,7 +85,7 @@ Empirica.round(Round);
 // exit screen will be shown.
 Empirica.exitSteps((game, player, treatment) => {
   if (player.exitStatus !== "finished") {
-    return [Sorry];
+    return [Sorry,ExitSurvey, Thanks];
   } else {
     if (game.treatment.chatEnabled == "true") {
     //chat exit survey

### 2021-07-22 code/aa-flowers-pilot_analysis.Rmd
add data from technical pilots, full compet run, collab pilot

@@
 
-data_target = "technical_pilots"
+data_target = "collabmulti_pilot0"
 ##Data import constants
 data_location=paste0("data/",data_target)
-date_start=lubridate::ymd('2021-06-30')
-image_location=paste0("figs/",data_target)
+date_start=lubridate::ymd('2021-07-19')
+image_location=paste0(here("figs",data_target))
 
@@
 d.round_results <- d.round_results.raw  %>% 
+  #filter out rounds where no one made a selection
+  filter(!is.na(chat)) %>%
   select(-chat, -context) %>%
@@
   mutate_at(c("time", "utility"), as.numeric) %>%
+  mutate(utility = ifelse(response == "false", 0, utility)) %>%
   #TODO: some times are negative?? interesting
@@
 ```{r}
-summary <- d.round_results %>% 
+game_completion <- d.round_results %>% 
   group_by(gameId) %>% 
-  summarize(num_rounds=max(repNum+1)) %>% 
-  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength))
+  summarize(num_rounds=max(trialNum+1)) %>% 
+  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
+  mutate(gameComplete = ifelse(num_rounds == 48, TRUE, FALSE))
 #knitr::kable(summary)
 message("Full games")
-summary %>% filter(num_rounds==12)
+game_completion %>% filter(num_rounds==48)
 message("Partial games")
-summary %>% filter(num_rounds!=12) 
+game_completion %>% filter(num_rounds!=48) 
 ```
@@
 ```
+
+Recast games with a language USE condition
+If there were fewer than 3 lines of communication, cast langUse = False
+```{r}
+d.games <- d.games %>% left_join(d.chat.raw %>% group_by(gameId) %>% summarize(numChat = sum(!is.na(text))) %>% 
+                                   mutate(langUse = (numChat >5)) %>% select(-numChat)) %>%
+  mutate(langUse = ifelse(is.na(langUse), FALSE, langUse)) %>%
+#add a toddle for completed games
+  left_join(game_completion %>% select(gameId, gameComplete))
+```
+
 How often do participants select a flower that is hidden from them?
@@
 ```{r}
-d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
+d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
+  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+         full_condition = paste0(condition, "-", chat)) %>%
+  ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
+  facet_grid(rows = vars(condition), cols = vars(langUse)) +
+  geom_point(alpha = .25) + 
+  labs(x = "log time (in seconds)",
+       y = "individual utility",
+       title = "Response time in sec and utility", subtitle =  "each participant, each round")
+
+d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
@@
   ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
-  facet_grid(rows = vars(condition), cols = vars(chat)) +
+  facet_grid(rows = vars(condition), cols = vars(langUse)) +
   geom_point(alpha = .25) + 
@@
 ```{r}
-d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
+d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameLength, gameComplete)) %>%
+  filter(gameComplete) %>%
   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
          full_condition = paste0(condition, "-", chat)) %>%
-  group_by(gameId, condition, chat, full_condition,gameLength) %>%
+  group_by(gameId, condition, chat, langUse, full_condition, gameLength) %>%
   summarize(totalUtility = sum(playerUtility))%>%
   ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
-  facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
+  facet_grid(rows = vars(condition), cols = vars(langUse), scales = "free_y") +
   geom_point(alpha = .75) + 
@@
 
-d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
+d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse,gameLength, gameComplete)) %>%
+  filter(gameComplete) %>%
   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
-         full_condition = paste0(condition, "-", chat)) %>%
+         chatUse = ifelse(langUse, "chatUse", "noChat"),
+         full_condition = paste0(condition, "-", chatUse)) %>%
   group_by(gameId, condition, chat, full_condition,gameLength) %>%
@@
   mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
-  ggplot(aes(x=log(as.numeric(gameLength, units="mins")), y = as.numeric(totalUtility), color = full_condition)) + 
+  ggplot(aes(x=as.numeric(gameLength, units="mins"), y = as.numeric(totalUtility), color = full_condition)) + 
   #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
@@
        title = "Length of game in mins and utility", subtitle =  "per game")
+
+ggsave(paste0(image_location, "/game_length.png"))
 ```
@@
 ```{r}
-condition = "coopCartel"
+condition = "competCartel"
 
 d.chat %>% 
-         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
-         filter(condition == "coopCartel",
-                chatEnabled == "true") %>%
+         left_join(d.games %>% select(gameId, createdAt,condition, langUse, chatEnabled, gameComplete)) %>% 
+         filter(langUse, gameComplete) %>%
 ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
@@
   theme(legend.position = "none")+
-  labs(title = "Cooperative + Lang", subtitle="Number of words", y="Number of words", x="Round number", color="playerId")
+  labs(title="Cooperative + Lang Used", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
 
-d.chat %>% 
-         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
-         filter(condition == "competCartel",
-                chatEnabled == "true") %>%
-ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
-  geom_jitter(alpha = .5) + 
-  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
-  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  theme(legend.position = "none")+
-  labs(title="Competitive + Lang", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
+ggsave(paste0(image_location, "/words_over_rounds.png"))
 ```
@@
   summarize(competCartel = sum(utility),
+            coopMulti= sum(utility),
             coopCartel = max(utility * 9)) %>% ungroup() %>%
-  pivot_longer(cols = c(competCartel, coopCartel), names_to = "condition", values_to = "maxUtility")
+  pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")
 
-d.utility <- d.round_results %>%
-         left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% 
-  group_by(gameId, blockNum, repNum, trialNum,condition, chatEnabled) %>% 
+d.utility <- d.round_results.final %>%
+         left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
+  group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
   summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
@@
   mutate(prop_utility = group_utility/maxUtility,
-         chat = if_else(chatEnabled=="true", "Lang", "Nonlang"))
+         langUsed = if_else(langUse, "Lang", "Nonlang"))
 
@@
   ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
-  geom_jitter(alpha = .4) + 
-  facet_grid(cols = vars(condition), rows = vars(chat), scales = "free_y") + 
+  geom_point(alpha = .4) + 
+  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
@@
   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
+
+ggsave(paste0(image_location, "/accuracy.png"))
+
+d.utility %>% 
+  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
+  geom_point(alpha = .4) + 
+  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
+  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  theme(legend.position = "none")+
+  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
+
+ggsave(paste0(image_location, "/accuracy.png"))
+
+d.utility %>% 
+  filter(langUsed == "Lang") %>%
+  ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
+  geom_point(alpha = .4) + 
+  facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
+  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  theme(legend.position = "none")+
+  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)
+
+ggsave(paste0(image_location, "/accuracy.png"))
 ```

### 2021-07-22 code/aa-flowers-pilot_analysis.Rmd
moved preprocessing to script and joined data files

@@
 ```{r set-up, include=F}
-#knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
-#knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
-#options(knitr.table.format = "html")
 library(tidyverse)
@@
 library(here)
-
+source(here("code/preprocessing/00_preprocessing_functions.R"))
 theme_set(theme_bw())
 
-data_target = "collabmulti_pilot0"
 ##Data import constants
-data_location=paste0("data/",data_target)
 date_start=lubridate::ymd('2021-07-19')
-image_location=paste0(here("figs",data_target))
-
-#library(rlang)
-#library(lme4)
-#library(brms)
-#library(rstan)
-#rstan_options(auto_write = TRUE)
-#options(mc.cores = parallel::detectCores())
-#model_location="code/models"
+image_location=here("figs")
 
@@
 
-```{r}
-# This was for determining participant bonuses using the version of data with PID
-# d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% rename(treatmentId=`_id`)
-# 
-# d.games <- read_csv(here(data_location, 'games.csv')) %>%
-#   select(gameId=`_id`, treatmentId, playerIds) %>%
-#   left_join(d.treatments) %>%
-#   mutate(playerIds=str_split(playerIds,",")) %>%
-#   unnest(playerIds) %>%
-#   select(playerId=playerIds, name)
-# 
-# d.players <- read_csv(here(data_location, 'players.csv')) %>%
-#   rename(playerId=`_id`) %>%
-#   left_join(d.games) %>%
-#   select(data.bonus, playerId,id,data.bonus,name) %>%
-#     filter(!is.na(name)) %>%
-#   mutate(pc_bonus=case_when(
-#     name=="fourRotate" ~ 3,
-#     name=="threeRotate" ~ 1.5,
-#     T ~ 0
-#   )) %>%
-#   mutate(bonus=round(data.bonus+pc_bonus,2),
-#          cost=round(bonus*4/3,2)) %>% write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))
+# Pre-process
+```{r process-data, message=FALSE, warning=FALSE}
+data_list <- c("technical_pilots", "collabmulti_pilot0", 
+               "collabmulti_20", "compet_20")
+process_all_data(data_list)
+join_all_data()
 ```
 
-# Pre-process
-
-We need:
-1) game info with treatment parameters
-2) context for each round (what were participant's options? max utility based on game conditions?)
-3) round results (what did participants select each round?)
-4) chat (what did individual participants say each round)
-5) feedback/demographics
-
-```{r helper-functions}
-ParseJSONColumn <- function(x) {
-  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
-    fromJSON(flatten = T)
-}
-
-extract_context_info <- function(context_key){
-  return(str_extract(context_key, "[^.]*$"))
-}
+```{r read-processed-date, message=FALSE}
+d.players <- read_csv(here("data/processed_data/joined_data/players.csv")) %>% distinct()
+d.games <- read_csv(here("data/processed_data/joined_data/games.csv")) %>% distinct()
+d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv")) %>% distinct()
+d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv")) %>% distinct()
 ```
 
-Get info about the games and treatments
-- We need each game, the 
-```{r game-preprocessing}
-d.games.raw <- read_csv(here(data_location, 'games.csv')) %>% 
-  rename(gameId = `_id`) %>% 
-  filter(createdAt >= date_start) %>%
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+```{r}
+#exclude BLOCKS with <12 completed (all 3 participants responded) rounds
+rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
 
-d.treatments.raw <- read_csv(here(data_location, 'treatments.csv')) %>% 
-  rename(treatmentId = `_id`) %>%
-  #filter(createdAt >= date_start) %>%
-  mutate(factorIds = str_split(factorIds, pattern=",")) %>%
-  unnest(factorIds) %>% rename(factorId = factorIds)
-
-d.factors.raw <- read_csv(here(data_location, 'factors.csv'))%>% 
-  rename(factorId = `_id`)
-
-d.factor_types.raw <- read_csv(here(data_location, 'factor-types.csv')) %>%
-  rename(factorTypeId = "_id",
-         factorName = "name")
-
-d.factors <- d.factors.raw %>% 
-  left_join(d.factor_types.raw %>% 
-              select(factorTypeId, factorName))
-
-d.treatments <- d.treatments.raw %>% 
-  left_join(d.factors %>% 
-              select(factorId, factorName, value)) %>%
-  pivot_wider(id_cols = c(treatmentId, name, createdAt), 
-              names_from = factorName, values_from = value)
-
-d.players <- read_csv(here(data_location, 'players.csv')) %>%
-  rename(playerId = `_id`) %>% 
-  filter(createdAt >= date_start) %>%
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
-
-d.games <- d.games.raw %>% left_join(d.treatments %>% 
-                                       select(-createdAt)) %>%
-  janitor::remove_empty() %>%
-  #select out the nonuseful columns
-  select(-c(gameLobbyId, treatmentId, batchId, targetSet, justStarted, submitCode, roundIds)) %>%
-  separate(col = playerIds, sep = ",", into = c("player_1", "player_2", "player_3")) %>% 
-  pivot_longer(cols = starts_with("player_"), names_to = "playerNum", values_to = "playerId") %>%
-  left_join(d.players %>% select(playerId, readyAt)) %>%
-  mutate(game_length = finishedAt - readyAt) %>%
-  group_by(across(gameId:chatEnabled)) %>%
-  summarize(gameLength = max(game_length)) %>% ungroup()
+d.rounds.final <-  d.rounds %>% anti_join(rounds_exclude)
+d.raw_chat.final <- d.raw_chat %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
 ```
 
-Round info
-For each round, we want to know the context...
-- which flowers were available
-- the utility of each flower
-
-We also want to have a report for each round including...
-- which participants selected which flowers
-- the time the participants took to select
-- the utility of the flower the participants selected
-
-```{r}
-d.round_results.raw <- read_csv(here(data_location,'rounds.csv'), guess_max=10000) %>% 
-  filter(createdAt >= date_start) %>% 
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
-  rename_with ( ~ gsub("room", "player", .x, fixed=T))
-
-#try pivot_longer instead of gather??
-d.round_results <- d.round_results.raw  %>% 
-  #filter out rounds where no one made a selection
-  filter(!is.na(chat)) %>%
-  select(-chat, -context) %>%
-  mutate_at(vars(starts_with('player')), as.character) %>%
-  pivot_longer(cols = starts_with('player'), 
-               names_to = c('blah', 'playerId', 'info'), 
-               values_to = "value",
-               names_sep = "_",
-               values_drop_na = TRUE) %>%
-  pivot_wider(names_from = "info", values_from = "value") %>%
-  select(-blah) %>%
-  mutate_at(c("time", "utility"), as.numeric) %>%
-  mutate(utility = ifelse(response == "false", 0, utility)) %>%
-  #TODO: some times are negative?? interesting
-  mutate(time = abs(as.numeric(time)/1000)) %>%
-  rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
-  arrange(createdAt, gameId, blockNum, repNum)
-
-#exclude blocks with <12 completed rounds
-rounds_exclude <- d.round_results %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
-d.round_results.final <-  d.round_results %>% anti_join(rounds_exclude)
-```
 ```{r}

### 2021-07-22 code/preprocessing/00_preprocessing_functions.R

@@ -1,0 +1,175 @@
+# Preprocessing functions for analyzing the adaptive agents pilots
+
+#These functions cover reading in the raw data, converting to tidy formats, filtering out unfinished rounds, and writing tidy processed csvs.
+
+# Libraries
+library(tidyverse)
+library(jsonlite)
+library(here)
+
+# helper functions
+ParseJSONColumn <- function(x) {
+  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
+    fromJSON(flatten = T)
+}
+
+extract_context_info <- function(context_key){
+  return(str_extract(context_key, "[^.]*$"))
+}
+
+get_player_data <- function(data_read_location, date_start) {
+  d.players <- read_csv(here(data_read_location, 'players.csv'), 
+                        col_types = cols(data.timeClick = col_number())) %>%
+    rename(playerId = `_id`) %>% 
+    filter(createdAt >= date_start) %>%
+    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+  return(d.players)
+}
+
+get_game_data <- function(data_read_location, date_start, players_data) {
+  # gets the game information, factors, and treatment variables for each game
+  d.games.raw <- read_csv(here(data_read_location, 'games.csv')) %>% 
+    rename(gameId = `_id`) %>% 
+    filter(createdAt >= date_start) %>%
+    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+  
+  d.treatments.raw <- read_csv(here(data_read_location, 'treatments.csv')) %>% 
+    rename(treatmentId = `_id`) %>%
+    mutate(factorIds = str_split(factorIds, pattern=",")) %>%
+    unnest(factorIds) %>% rename(factorId = factorIds)
+  
+  d.factors.raw <- read_csv(here(data_read_location, 'factors.csv'))%>% 
+    rename(factorId = `_id`)
+  
+  d.factor_types.raw <- read_csv(here(data_read_location, 'factor-types.csv')) %>%
+    rename(factorTypeId = "_id",
+           factorName = "name")
+  
+  d.factors <- d.factors.raw %>% 
+    left_join(d.factor_types.raw %>% 
+                select(factorTypeId, factorName))
+  
+  d.treatments <- d.treatments.raw %>% 
+    left_join(d.factors %>% 
+                select(factorId, factorName, value)) %>%
+    pivot_wider(id_cols = c(treatmentId, name, createdAt), 
+                names_from = factorName, values_from = value)
+  
[…]

### 2021-07-23 experiments/flowers-pilot/client/main.js
add no chat exit survey

@@ -35,7 +35,7 @@ import Round from "./game/Round.jsx";
 import Thanks from "./exit/Thanks.jsx";
 import Sorry from "./exit/Sorry";
 import ExitSurvey from "./exit/ExitSurvey";
-
+import ExitSurveyNochat from "./exit/ExitSurveyNochat.jsx";
 import customBreadcrumb from "./game/Breadcrumb.jsx";
 
 Empirica.header(() => null);
@@ -85,13 +85,18 @@ Empirica.round(Round);
 // exit screen will be shown.
 Empirica.exitSteps((game, player, treatment) => {
   if (player.exitStatus !== "finished") {
-    return [Sorry,ExitSurvey, Thanks];
+    if (game.treatment.chatEnabled == "true") {
+    //chat exit survey
+    return [Sorry, ExitSurvey, Thanks];}
+    else {
+      return [Sorry, ExitSurveyNochat, Thanks];
+    }
   } else {
     if (game.treatment.chatEnabled == "true") {
     //chat exit survey
     return [ExitSurvey, Thanks];}
     else {
-      return [ExitSurvey, Thanks];
+      return [ExitSurveyNochat, Thanks];
     }
   }
 });

### 2021-07-23 code/aa-flowers-pilot_analysis.Rmd
add code for feedback, contexts, and demographics

@@
 source(here("code/preprocessing/00_preprocessing_functions.R"))
-theme_set(theme_bw())
+theme_set(theme_classic())
 
@@
 date_start=lubridate::ymd('2021-07-19')
-image_location=here("figs")
+fig_path=here("figs")
 
@@
 # Pre-process
+Remove joined_data and uncomment the following block to rerun the preprocessing scripts
 ```{r process-data, message=FALSE, warning=FALSE}
@@
 ```{r read-processed-date, message=FALSE}
-d.players <- read_csv(here("data/processed_data/joined_data/players.csv")) %>% distinct()
-d.games <- read_csv(here("data/processed_data/joined_data/games.csv")) %>% distinct()
-d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv")) %>% distinct()
-d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv")) %>% distinct()
+d.players <- read_csv(here("data/processed_data/joined_data/players.csv"),
+                      col_types = cols()) %>% distinct()
+d.games <- read_csv(here("data/processed_data/joined_data/games.csv"),
+                      col_types = cols()) %>% distinct()
+d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv"),
+                      col_types = cols()) %>% distinct()
+d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv"),
+                      col_types = cols()) %>% distinct()
+d.contexts <- read_csv(here("data/processed_data/joined_data/contexts.csv"),
+                      col_types = cols()) %>% distinct()
+d.feedback <- read_csv(here("data/processed_data/joined_data/feedback.csv"),
+                      col_types = cols()) %>% distinct()
+d.demographics <- read_csv(here("data/processed_data/joined_data/demographics.csv"),
+                      col_types = cols()) %>% distinct()
 ```
@@
 #exclude BLOCKS with <12 completed (all 3 participants responded) rounds
+#exclude trial runs
+
 rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
@@
 ```{r}
-game_completion <- d.round_results %>% 
+game_completion <- d.rounds %>% 
   group_by(gameId) %>% 
@@
 
+Game counts
 ```{r}
-d.chat.raw <- d.round_results.raw %>%
-  filter(createdAt >= date_start) %>%
-  anti_join(rounds_exclude) %>%
-  #there will always be an alert for each active round, so drop NA
-  filter(!is.na(chat))%>%
-  select(-c(ends_with('response'), ends_with('time'), ends_with('utility'), context, stageIds)) %>%
-  #mutate(chat = ifelse(is.na(chat), '{}', chat)) %>%
-  rename(roundID = `_id`) %>%
-  mutate(chat = map(chat, .f = ParseJSONColumn)) %>%
-  unnest(chat) %>%
-  arrange(createdAt, gameId, blockNum, repNum) %>% 
-  left_join(d.players %>% select(playerId, name)) %>%
-  left_join(d.round_results %>% select(gameId, blockNum, repNum, trialNum, playerId, playerResponse, playerUtility)) %>%
-  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
-  mutate("participantAction" = ifelse(type == "alert", paste(name, "selected", playerResponse, "for", playerUtility, "point(s)"), NA)) %>%
-  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, participantAction, everything()) %>%
-  write_csv(here(data_location, 'raw_chat.csv'))
-
-#extract the alternative flowers
-d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
-  mutate(data.context = map(data.context, 
-                            .f = ParseJSONColumn)) %>% 
-  select("_id":"data.numPlayers") %>%
-  mutate(data.context = map(data.context,
-                            function(df) df %>% gather(key, value) %>%
-                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
-                              separate(key, into = c("label", "key"), sep = "-") %>%
-                              spread(key, value))) %>%
-  unnest(data.context) %>%
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+#pull out pilots
+d.games %>% group_by(condition, chatEnabled,pilot)%>% tally()
 ```
-
-Recast games with a language USE condition
-If there were fewer than 3 lines of communication, cast langUse = False
-```{r}
-d.games <- d.games %>% left_join(d.chat.raw %>% group_by(gameId) %>% summarize(numChat = sum(!is.na(text))) %>% 
-                                   mutate(langUse = (numChat >5)) %>% select(-numChat)) %>%
-  mutate(langUse = ifelse(is.na(langUse), FALSE, langUse)) %>%
-#add a toddle for completed games
-  left_join(game_completion %>% select(gameId, gameComplete))
-```
-
-How often do participants select a flower that is hidden from them?
 
@@
 ```{r demographics}
-d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
-  filter(createdAt >= date_start) %>%
-  left_join(d.games, by = c('gameId')) %>%
-    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
-```
-```{r feedback}
-d.exit.survey %>% select( language, fair, chatUseful, feedback, time) %>% knitr::kable()
+d.demographics
 ```
 
-# Pretty pictures
+```{r feedback}
+d.feedback
+```
 
-Correlation between response time and utility?
+# Basic Analyses
+## Simple Distributions
 
 ```{r}
-d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
-  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
-         full_condition = paste0(condition, "-", chat)) %>%
-  ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
-  facet_grid(rows = vars(condition), cols = vars(langUse)) +
-  geom_point(alpha = .25) + 
-  labs(x = "log time (in seconds)",
-       y = "individual utility",
-       title = "Response time in sec and utility", subtitle =  "each participant, each round")
-
-d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
-  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
-         full_condition = paste0(condition, "-", chat)) %>%
-  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
-  facet_grid(rows = vars(condition), cols = vars(langUse)) +
-  geom_point(alpha = .25) + 
-  labs(x = "log time (in seconds)",
-       y = "individual utility",
-       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
-```
-
-Easiest way to do this, is to derive the game start date with the latest participant ready time and the game finishedAt
-
-```{r}
-d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameLength, gameComplete)) %>%
-  filter(gameComplete) %>%
-  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
-         full_condition = paste0(condition, "-", chat)) %>%
-  group_by(gameId, condition, chat, langUse, full_condition, gameLength) %>%
-  summarize(totalUtility = sum(playerUtility))%>%
-  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
-  facet_grid(rows = vars(condition), cols = vars(langUse), scales = "free_y") +
-  geom_point(alpha = .75) + 
-  labs(x = "Game length (in minutes)",
-       y = "total points by group",
-       title = "Length of game in mins and utility", subtitle =  "per game")
-
-d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse,gameLength, gameComplete)) %>%
-  filter(gameComplete) %>%
-  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
-         chatUse = ifelse(langUse, "chatUse", "noChat"),
-         full_condition = paste0(condition, "-", chatUse)) %>%
-  group_by(gameId, condition, chat, full_condition,gameLength) %>%
-  summarize(totalUtility = sum(playerUtility))%>%
-  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
-  ggplot(aes(x=as.numeric(gameLength, units="mins"), y = as.numeric(totalUtility), color = full_condition)) + 
-  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
-  geom_point(alpha = .75) + 
-  geom_smooth(method=lm, alpha=.3)+
-  labs(x = "Game length (in minutes)",
-       y = "total points by group",
-       title = "Length of game in mins and utility", subtitle =  "per game")
-
-ggsave(paste0(image_location, "/game_length.png"))
-```
-
-```{r chat}
-#group raw chat by participant
-d.chat <- read_csv(here(data_location, "raw_chat.csv"), guess_max=1000) %>% 
+#get the # of raw points, max adjusted points, game legnth, # of words exchanged
+d.round_word_counts <- d.raw_chat.final %>% 
   filter(type == "message") %>%
-  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum")) %>%
+  full_join(d.rounds.final, c("gameId", "trialNum", "repNum", 
+                              "playerId", "numPlayers", "blockNum")) %>%
  # filter(!is.chitchat) %>% 
@@
             total_num_words = sum(utt_length_words),
-            total_num_chars = sum(utt_length_chars)) %>%
-  anti_join(rounds_exclude)
+            total_num_chars = sum(utt_length_chars))
 
+d.by_game_metrics <- d.rounds.final%>% 
+  left_join(d.games %>% select(-createdAt)) %>%  
+  left_join(d.round_word_counts) %>%
+  left_join(game_completion %>% select(gameId, num_rounds)) %>%
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-")) %>% 
+  #group by game
+  group_by(gameId, name,condition, languageCondition, fullCondition, gameLength,num_rounds) %>%
+  summarize(groupPoints = sum(playerUtility, na.rm = T),
+            groupNumWords = sum(total_num_words, na.rm = T),
+            groupNumChars = sum(total_num_chars, na.rm = T))%>%
+  mutate(groupPoints = ifelse(condition=="coopCartel", 
+                               groupPoints/3, groupPoints)) %>% 
+  #adjust for max
+  mutate(adjustedPoints = case_when(condition == "coopCartel" ~ groupPoints/ ((12*3)*48),
+                                    condition == "coopMulti" ~ groupPoints/ ((12+11+10)*48),
+                                    condition == "competCartel" ~ groupPoints/ ((12+11+10)*48)))
 ```
 
-Pick just one game to look at, make sure the plot is working?
+### Points
+
+Unadjusted
+
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics%>% 
+  ggplot(aes(x = languageCondition, y = groupPoints, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + theme_classic() + theme(legend.position = "none")
+
+ggsave(here(fig_path, "points_distirbution_raw.png"))
+```
+
+Max Adjusted
+
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = languageCondition, y = adjustedPoints, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group (Max Adjusted)", title = "Adjusted Points by Group") + theme_classic() + theme(legend.position = "none")
+
+ggsave(here(fig_path, "points_distirbution_adjusted.png"))
+```
+
+### Time
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = languageCondition, y = gameLength/60, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition))+ labs(y = "Time (In Minutes)", title = "Game Length")+theme_classic() + theme(legend.position = "none")
+```
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = languageCondition, y = log(gameLength/60), color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition))+ labs(y = "Time (In Log Minutes)", title = "Game Length (Log Mins)")+theme_classic() + theme(legend.position = "none")
+```
+
+### Number of Words
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = languageCondition, y = groupNumWords, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged")+theme_classic() + theme(legend.position = "none")
+```
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = languageCondition, y = log(groupNumWords), color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .2, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  facet_grid(cols = vars(condition))+ labs(y = "Log N. Words Exchanged")+theme_classic() + theme(legend.position = "none")
+```
+
+#### Condition Effects on Language Use
 ```{r}
-condition = "competCartel"
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition, languageCondition, condition) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>% 
+  ggplot() + geom_density(aes(x = groupNumWords, fill = condition), alpha = .5, adjust = .7)
+```
 
-d.chat %>% 
-         left_join(d.games %>% select(gameId, createdAt,condition, langUse, chatEnabled, gameComplete)) %>% 
-         filter(langUse, gameComplete) %>%
-ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
-  geom_jitter(alpha = .5) + 
-  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
-  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  theme(legend.position = "none")+
-  labs(title="Cooperative + Lang Used", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
+## Scatters
+### Time x Points
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = gameLength/60, y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic() + theme(legend.position = "none")
+```
 
-ggsave(paste0(image_location, "/words_over_rounds.png"))
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = gameLength/60, y = groupPoints, color = log(groupNumWords))) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic() + scale_color_steps()
+```
+
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>%
+  ggplot(aes(x = gameLength/60, y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic()
+```
+
+Log
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% 
+  ggplot(aes(x = log(gameLength/60), y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+theme_classic() + theme(legend.position = "none")
+```
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% 
+  ggplot(aes(x = log(gameLength/60), y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+theme_classic()
+```
+### Language x Points
+#### Faceted
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(condition))+ labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+theme_classic() + theme(legend.position = "none")
+```
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = groupNumWords, y = adjustedPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(condition))+ labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+theme_classic() + theme(legend.position = "none")
+```
+#### Non-Faceted
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+theme_classic()
+```
+```{r, fig.height=3, fig.width=5}
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = groupNumWords, y = adjustedPoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+theme_classic()
+```
+
+### Progress Over Time
+
+Including language optional games where they decided not to speak
+
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
+  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ theme_classic()
+```
+
+log (n words+1)
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)+1) %>%
+  ggplot(aes(x = repNum, y = log(groupNumWords), color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log (N. Words + 1)")+ theme_classic()
+```
+
+Only games where language was used
+
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
+  filter(groupNumWords > 0) %>%
+  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ theme_classic()
+```
+
+### Accuracy
+#### N. Overlap
+
+```{r}
+# d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
+#   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+#          full_condition = paste0(condition, "-", chat)) %>%
+#   ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
+#   facet_grid(rows = vars(condition), cols = vars(langUse)) +
+#   geom_point(alpha = .25) + 
+#   labs(x = "log time (in seconds)",
+#        y = "individual utility",
+#        title = "Response time in sec and utility", subtitle =  "each participant, each round")
+# 
+# d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
+#   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
+#          full_condition = paste0(condition, "-", chat)) %>%
+#   ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
+#   facet_grid(rows = vars(condition), cols = vars(langUse)) +
+#   geom_point(alpha = .25) + 
+#   labs(x = "log time (in seconds)",
+#        y = "individual utility",
+#        title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
 ```
@@
 ```{r}
-d.max_utility <- d.contexts %>% 
-  left_join(d.games %>% select(gameId, condition)) %>%
-  mutate(utility= as.numeric(utility)) %>%
-  group_by(gameId, blockNum, repNum) %>% 
-  slice_max(order_by =utility,n = 3)%>%
-  summarize(competCartel = sum(utility),
-            coopMulti= sum(utility),
-            coopCartel = max(utility * 9)) %>% ungroup() %>%
-  pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")
-
-d.utility <- d.round_results.final %>%
-         left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
-  group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
-  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
-  left_join(d.max_utility) %>%
-  mutate(prop_utility = group_utility/maxUtility,
-         langUsed = if_else(langUse, "Lang", "Nonlang"))
-
-d.utility %>% 
-  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
-  geom_point(alpha = .4) + 
-  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
-  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  theme(legend.position = "none")+
-  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
-
-ggsave(paste0(image_location, "/accuracy.png"))
-
-d.utility %>% 
-  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
-  geom_point(alpha = .4) + 
-  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
-  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  theme(legend.position = "none")+
-  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
-
-ggsave(paste0(image_location, "/accuracy.png"))
-
-d.utility %>% 
-  filter(langUsed == "Lang") %>%
-  ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
-  geom_point(alpha = .4) + 
-  facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
-  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  theme(legend.position = "none")+
-  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)
-
-ggsave(paste0(image_location, "/accuracy.png"))
+# d.max_utility <- d.contexts %>% 
+#   left_join(d.games %>% select(gameId, condition)) %>%
+#   mutate(utility= as.numeric(utility)) %>%
+#   group_by(gameId, blockNum, repNum) %>% 
+#   slice_max(order_by =utility,n = 3)%>%
+#   summarize(competCartel = sum(utility),
+#             coopMulti= sum(utility),
+#             coopCartel = max(utility * 9)) %>% ungroup() %>%
+#   pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")
+# 
+# 
+# d.utility <- d.round_results.final %>%
+#          left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
+#   group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
+#   summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
+#   left_join(d.max_utility) %>%
+#   mutate(prop_utility = group_utility/maxUtility,
+#          langUsed = if_else(langUse, "Lang", "Nonlang"))
+# 
+# d.utility %>% 
+#   ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
+#   geom_point(alpha = .4) + 
+#   facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
+#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+#   theme(legend.position = "none")+
+#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
+# 
+# ggsave(paste0(image_location, "/accuracy.png"))
+# 
+# d.utility %>% 
+#   ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
+#   geom_point(alpha = .4) + 
+#   facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
+#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+#   theme(legend.position = "none")+
+#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
+# 
+# ggsave(paste0(image_location, "/accuracy.png"))
+# 
+# d.utility %>% 
+#   filter(langUsed == "Lang") %>%
+#   ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
+#   geom_point(alpha = .4) + 
+#   facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
+#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+#   theme(legend.position = "none")+
+#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)
+# 
+# ggsave(paste0(image_location, "/accuracy.png"))
 ```
@@
 #   theme(legend.position="bottom")
-ggplot(d.chat, aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
-  facet_wrap(~role, nrow=1)+
-  scale_color_brewer(palette="Dark2")+
-    geom_jitter(alpha=.05)+
-    geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  #geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
-     stat_summary(fun.data = "mean_cl_boot")+
-    scale_y_continuous(limits = c(0,50))+
-  labs(title="Number of words", y="Number of words", x="Round number", color="Player count")+
-  theme(legend.position="bottom")
-ggsave(here(image_location, 'words.pdf'), width=6, height=4)
+# ggplot(d.chat, aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
+#   facet_wrap(~role, nrow=1)+
+#   scale_color_brewer(palette="Dark2")+
+#     geom_jitter(alpha=.05)+
+#     geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+#   #geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
+#      stat_summary(fun.data = "mean_cl_boot")+
+#     scale_y_continuous(limits = c(0,50))+
+#   labs(title="Number of words", y="Number of words", x="Round number", color="Player count")+
+#   theme(legend.position="bottom")
+# ggsave(here(image_location, 'words.pdf'), width=6, height=4)
 ```
@@
 ```{r}
-d.chat %>% filter(role=="speaker") %>% 
-    mutate(groupxtangram=str_c(gameId,tangram)) %>% 
-  group_by(repNum, numPlayers, gameId,tangram, groupxtangram) %>% 
-  summarize(words=sum(total_num_words)) %>% 
-ggplot(aes(x=repNum, y=words, color=as.factor(numPlayers)))+
-  facet_wrap(~numPlayers, nrow=1)+
-  scale_color_brewer(palette="Dark2")+
-    geom_line(aes(group=groupxtangram), alpha=.1,method=glm, se=F)+
-    geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
-    #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  labs(title="Words from speaker per tangram", y="Number of words", x="Round number", color="Player count")+
-  theme(legend.position="null")
-ggsave(here(image_location, 'words_lines.pdf'), width=6, height=4)
+# d.chat %>% filter(role=="speaker") %>% 
+#     mutate(groupxtangram=str_c(gameId,tangram)) %>% 
+#   group_by(repNum, numPlayers, gameId,tangram, groupxtangram) %>% 
+#   summarize(words=sum(total_num_words)) %>% 
+# ggplot(aes(x=repNum, y=words, color=as.factor(numPlayers)))+
+#   facet_wrap(~numPlayers, nrow=1)+
+#   scale_color_brewer(palette="Dark2")+
+#     geom_line(aes(group=groupxtangram), alpha=.1,method=glm, se=F)+
+#     geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
+#     #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+#   labs(title="Words from speaker per tangram", y="Number of words", x="Round number", color="Player count")+
+#   theme(legend.position="null")
+# ggsave(here(image_location, 'words_lines.pdf'), width=6, height=4)
 ```
@@
 ```{r}
-d.chat %>% filter(role=="speaker") %>% 
-ggplot(aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
-  facet_wrap(~tangram)+
-  scale_color_brewer(palette="Dark2")+
-    #geom_smooth(method=glm, formula=y~poly(x,2), se=T, alpha=.1)+
-      geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
-       stat_summary(fun.data = "mean_cl_boot", size=.2)+
-  labs(title="Tangram variability", y="Number of words", x="Round number", color="Player count")+
-  theme(legend.position="bottom")
-ggsave(here(image_location, 'words_tangrams.pdf'), width=8, height=6)
+# d.chat %>% filter(role=="speaker") %>% 
+# ggplot(aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
+#   facet_wrap(~tangram)+
+#   scale_color_brewer(palette="Dark2")+
+#     #geom_smooth(method=glm, formula=y~poly(x,2), se=T, alpha=.1)+
+#       geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
+#        stat_summary(fun.data = "mean_cl_boot", size=.2)+
+#   labs(title="Tangram variability", y="Number of words", x="Round number", color="Player count")+
+#   theme(legend.position="bottom")
+# ggsave(here(image_location, 'words_tangrams.pdf'), width=8, height=6)
 ```
@@
 ```{r accuracy}
-d.round_results %>% group_by(playerId,repNum, gameId, numPlayers) %>% 
-  mutate(correct.num=ifelse(correct,1,0)) %>% 
-  ggplot(aes(x=repNum, y=correct.num, color=as.factor(numPlayers)))+
-geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
-  #geom_point()+
-  scale_color_brewer(palette="Dark2")+
-  #scale_y_continuous(limits = c(0,1))+
-  labs(x="Round Number", y="Fraction correctly selected", title= "Overall accuracy increases over repetitions", color="Player count")+
-    theme(legend.position="bottom")
-ggsave(here(image_location, 'accuracy.pdf'), width=6, height=4)
+# d.round_results %>% group_by(playerId,repNum, gameId, numPlayers) %>% 
+#   mutate(correct.num=ifelse(correct,1,0)) %>% 
+#   ggplot(aes(x=repNum, y=correct.num, color=as.factor(numPlayers)))+
+# geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
+#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
+#   #geom_point()+
+#   scale_color_brewer(palette="Dark2")+
+#   #scale_y_continuous(limits = c(0,1))+
+#   labs(x="Round Number", y="Fraction correctly selected", title= "Overall accuracy increases over repetitions", color="Player count")+
+#     theme(legend.position="bottom")
+# ggsave(here(image_location, 'accuracy.pdf'), width=6, height=4)
 ```
@@
 ```{r time}
-d.round_results %>% group_by(playerId, repNum, gameId, numPlayers) %>% 
-  filter(correct==T) %>% 
-  #summarize(time=mean(time)) %>% 
-  ggplot(aes(x=repNum, y=time, color=as.factor(numPlayers)))+
-  geom_jitter(width=.4, height=0, alpha=.03)+
-geom_smooth(method = "glm", formula = y~x,
-                      method.args = list(family = gaussian(link = 'log')))+
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
-  scale_y_continuous(limits = c(0,180))+
-    scale_color_brewer(palette="Dark2")+
-  labs(x="Round Number", y="Time to selection in seconds",
-       title="People choose faster in later rounds", color="Player count")+
-  theme(legend.position = "bottom")
-ggsave(here(image_location, 'time.pdf'), width=6, height=4)
+# d.round_results %>% group_by(playerId, repNum, gameId, numPlayers) %>% 
+#   filter(correct==T) %>% 
+#   #summarize(time=mean(time)) %>% 
+#   ggplot(aes(x=repNum, y=time, color=as.factor(numPlayers)))+
+#   geom_jitter(width=.4, height=0, alpha=.03)+
+# geom_smooth(method = "glm", formula = y~x,
+#                       method.args = list(family = gaussian(link = 'log')))+
+#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
+#   scale_y_continuous(limits = c(0,180))+
+#     scale_color_brewer(palette="Dark2")+
+#   labs(x="Round Number", y="Time to selection in seconds",
+#        title="People choose faster in later rounds", color="Player count")+
+#   theme(legend.position = "bottom")
+# ggsave(here(image_location, 'time.pdf'), width=6, height=4)
 ```
@@
 ```{r, include=F}
-model_input <- d.chat %>% filter(role=="speaker") %>% 
-  mutate(block=repNum,
-         words=total_num_words,
-         tangram_group=str_c(tangram, gameId))
-
-priors <- c(
-  set_prior("normal(20, 20)", class="Intercept"),
-  set_prior("normal(0, 10)", class="b"),
-  set_prior("normal(0, 10)", class="sd"),
-  set_prior("lkj(1)",       class="cor"))
-model <- brm(words ~ block * numPlayers + (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "model_1"),                            prior=priors, control=list(adapt_delta=.95))
+# model_input <- d.chat %>% filter(role=="speaker") %>% 
+#   mutate(block=repNum,
+#          words=total_num_words,
+#          tangram_group=str_c(tangram, gameId))
+#          
+# priors <- c(
+#   set_prior("normal(20, 20)", class="Intercept"),
+#   set_prior("normal(0, 10)", class="b"),
+#   set_prior("normal(0, 10)", class="sd"),
+#   set_prior("lkj(1)",       class="cor"))
+# model <- brm(words ~ block * numPlayers + (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "model_1"),                            prior=priors, control=list(adapt_delta=.95))
 
@@
 ```{r, include=F}
-d.prev.speaker <- d.chat %>% ungroup() %>%  filter(role=="speaker") %>% select(gameId,repNum, tangram, total_num_words_prev=total_num_words)
-d.prev.round <- d.chat %>% ungroup() %>% select(playerId, correct, tangram, gameId, repNum) %>% 
-  left_join(d.prev.speaker) %>% unique() %>% mutate(repNum=repNum+1)
-d.chat.lagged <- d.chat %>%
-  ungroup() %>% 
-  select(gameId, playerId, trialNum, repNum, playerId, role, tangram, total_num_words, numPlayers) %>%
-  left_join(d.prev.round) %>%
-  mutate(reduction_word=log(total_num_words)-log(total_num_words_prev)) %>%
-  filter(repNum>0) %>%
-  filter(role=="speaker") %>%
-  mutate(prev_correct_round=correct)
-model_input <- d.chat.lagged %>% filter(role=="speaker") %>% 
-  mutate(block=repNum,
-         words=total_num_words,
-         tangram_group=str_c(tangram, gameId),
-         was_INcorrect=ifelse(!prev_correct_round,1,0))
-
-priors <- c(
-  set_prior("normal(20, 20)", class="Intercept"),
-  set_prior("normal(0, 10)", class="b"),
-  set_prior("normal(0, 10)", class="sd"),
-  set_prior("lkj(1)",       class="cor"))
-model_speaker_acc <- brm(words ~ block * numPlayers +block*was_INcorrect+ (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "speaker_acc"),                         prior=priors, control=list(adapt_delta=.95))
+# d.prev.speaker <- d.chat %>% ungroup() %>%  filter(role=="speaker") %>% select(gameId,repNum, tangram, total_num_words_prev=total_num_words)
+# d.prev.round <- d.chat %>% ungroup() %>% select(playerId, correct, tangram, gameId, repNum) %>% 
+#   left_join(d.prev.speaker) %>% unique() %>% mutate(repNum=repNum+1)
+# d.chat.lagged <- d.chat %>%
+#   ungroup() %>% 
+#   select(gameId, playerId, trialNum, repNum, playerId, role, tangram, total_num_words, numPlayers) %>%
+#   left_join(d.prev.round) %>%
+#   mutate(reduction_word=log(total_num_words)-log(total_num_words_prev)) %>%
+#   filter(repNum>0) %>%
+#   filter(role=="speaker") %>%
+#   mutate(prev_correct_round=correct)
+# model_input <- d.chat.lagged %>% filter(role=="speaker") %>% 
+#   mutate(block=repNum,
+#          words=total_num_words,
+#          tangram_group=str_c(tangram, gameId),
+#          was_INcorrect=ifelse(!prev_correct_round,1,0))
+#          
+# priors <- c(
+#   set_prior("normal(20, 20)", class="Intercept"),
+#   set_prior("normal(0, 10)", class="b"),
+#   set_prior("normal(0, 10)", class="sd"),
+#   set_prior("lkj(1)",       class="cor"))
+# model_speaker_acc <- brm(words ~ block * numPlayers +block*was_INcorrect+ (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "speaker_acc"),                         prior=priors, control=list(adapt_delta=.95))
 
@@
 ```{r}
-summary(model)
+#summary(model)
 ```
 ```{r}
-summary(model_speaker_acc)
+#summary(model_speaker_acc)
 ```

### 2021-07-23 code/preprocessing/00_preprocessing_functions.R

@@ -28,20 +28,20 @@ get_player_data <- function(data_read_location, date_start) {
 
 get_game_data <- function(data_read_location, date_start, players_data) {
   # gets the game information, factors, and treatment variables for each game
-  d.games.raw <- read_csv(here(data_read_location, 'games.csv')) %>% 
+  d.games.raw <- read_csv(here(data_read_location, 'games.csv'), col_types = cols()) %>% 
     rename(gameId = `_id`) %>% 
     filter(createdAt >= date_start) %>%
     rename_with(~ gsub("data.", "", .x, fixed = TRUE))
   
-  d.treatments.raw <- read_csv(here(data_read_location, 'treatments.csv')) %>% 
+  d.treatments.raw <- read_csv(here(data_read_location, 'treatments.csv'),col_types = cols()) %>% 
     rename(treatmentId = `_id`) %>%
     mutate(factorIds = str_split(factorIds, pattern=",")) %>%
     unnest(factorIds) %>% rename(factorId = factorIds)
   
-  d.factors.raw <- read_csv(here(data_read_location, 'factors.csv'))%>% 
+  d.factors.raw <- read_csv(here(data_read_location, 'factors.csv'),col_types = cols())%>% 
     rename(factorId = `_id`)
   
-  d.factor_types.raw <- read_csv(here(data_read_location, 'factor-types.csv')) %>%
+  d.factor_types.raw <- read_csv(here(data_read_location, 'factor-types.csv'),col_types = cols()) %>%
     rename(factorTypeId = "_id",
            factorName = "name")
   
@@ -70,14 +70,17 @@ get_game_data <- function(data_read_location, date_start, players_data) {
                  values_to = "playerId") %>%
     left_join(players_data %>% select(playerId, readyAt)) %>%
     #calculate game time
-    mutate(game_length = finishedAt - readyAt) %>%
+    mutate(game_length = difftime(finishedAt, readyAt, units = "s")) %>%
     group_by(across(gameId:chatEnabled)) %>%
-    summarize(gameLength = max(game_length)) %>% ungroup()
+    summarize(gameLength = min(game_length)) %>% ungroup() %>%
+    mutate(pilot = ifelse(grepl("Pilot", name),TRUE,FALSE))
+  
   return(d.games)
 }
 
 get_round_data <- function(data_read_location, date_start) {
-  d.round_results.raw <- read_csv(here(data_read_location,'rounds.csv'), guess_max=10000) %>% 
[…]

### 2021-07-23 experiments/flowers-pilot/client/main.js
fix no chat exit

@@ -84,7 +84,7 @@ Empirica.round(Round);
 // If you don't return anything, or do not define this function, a default
 // exit screen will be shown.
 Empirica.exitSteps((game, player, treatment) => {
-  if (player.exitStatus !== "finished") {
+   if (player.exitStatus !== "finished") {
     if (game.treatment.chatEnabled == "true") {
     //chat exit survey
     return [Sorry, ExitSurvey, Thanks];}

### 2021-07-23 code/aa-flowers-pilot_analysis.Rmd
add multi non-lang condition

@@
 data_list <- c("technical_pilots", "collabmulti_pilot0",
-               "collabmulti_20", "compet_20")
+               "collabmulti_20", "compet_20", "multi_nonlang_20")
 process_all_data(data_list)
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = gameLength/60, y = groupPoints, color = log(groupNumWords))) + 
+  ggplot(aes(x = gameLength/60, y = groupPoints, color = log(groupNumWords+1))) + 
   geom_point(alpha = .5) +
@@
 ```
+log num words
 
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
+  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ theme_classic()
+```
 ### Accuracy

### 2021-07-26 code/linguistic_analysis_C-D.Rmd
plots for new pilot

--- 2021-06-07 code/linguistic_analysis_C-D.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
 library(tidyverse)
 library(here)
-library(magrittr)
-library(qdap)
 library("stringr")
+library(Hmisc)
+library("spacyr")
+library(ggrepel)
+
 ```
@@
 ```{r main data, include=TRUE}
-data_target = "pilot0"
-data_location=paste0("data/",data_target)
+d<- read_csv("/Users/lscpuser/Documents/AA-flowers2/AA-flowers/data/processed_data/joined_data/filtered_raw_chat.csv")
+d_flowers <- d %>% 
+  filter(type=="message") %>%
+  filter(!is.na(text))
+#write_csv(d, "~/Downloads/raw_chat2.csv")
 
-d_flowers<- read_csv(here(data_location, "raw_chat.csv"))%>%
- filter(!is.na(text))
+d_flowers <- d_flowers %>% 
+  filter(!condition=="coopCartel")
+
+```
+
+```{r main data1, include=TRUE}
+#data_target = "pilot0"
+#data_location=paste0("data/",data_target)
+
+#d_flowers<- read_csv(here(data_location, "raw_chat2.csv"))%>%
+ #filter(!is.na(text))
+
 ```
@@
 ```{r utt_length}
-#1.
 
-d_flowers <- d_flowers %>%
-  mutate(utt_length_words = str_count(text, "\\W+") + 1) %>% 
-    group_by(gameId, blockNum, trialNum, repNum, playerId)
+d_flowers_utt_length_block <- d_flowers %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, blockNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words),
+            mlu = mean(utt_length_words))
 
-d_flowers_utt_length <- d_flowers %>%
-      summarise(total_num_words = sum(utt_length_words))
+#, color=as.factor(gameId)
 
-ggplot(d_flowers_utt_length, aes(x=trialNum, y=total_num_words, color=as.factor(playerId)))+
-  scale_color_brewer(palette="Dark2")+
-  facet_wrap(~gameId, nrow=1) +
+#jpeg("~/Desktop/plots/nwords_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=total_num_words))+
+  facet_grid(rows =vars(condition)) +
+  geom_point()+
   geom_jitter(alpha=.05)+
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  scale_y_continuous(limits = c(0,50))+
-  labs(title="Number of words", y="Number of words", x="Round number", color="playerId")+
-  theme(legend.position="bottom")
-```
+  labs(title="Number of words", y="Total number of words", x="Block number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of words per block")#+
+ # theme(text = element_text(size = 30))
+#color="gameId" cols = vars(blockNum),
+#dev.off()
 
-### 2a. Number of turns
-```{r number_of_turns}
-
-d_flowers_n_turns <- d_flowers %>%
-  summarise(n=n())
-
-ggplot(d_flowers_n_turns, aes(x=trialNum, y=n, color=as.factor(playerId)))+
-  facet_wrap(~gameId, nrow=1) +
-  scale_color_brewer(palette="Dark2")+
+#jpeg("~/Desktop/plots/mlu_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=mlu))+
+  facet_grid( rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  labs(title="Number of turns", y="Number of turns", x="Round number", color="playerId")+
-  theme(legend.position="bottom")
+ # scale_y_continuous(limits = c(0,15))+
+  labs(title="MLU", y="MLU", x="Block number")+
+  theme(legend.position="bottom") + 
+  theme_bw()+
+  ggtitle("MLU per block")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_utt_length_trial <- d_flowers %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, trialNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words),
+            mlu = mean(utt_length_words))
+
+#jpeg("~/Desktop/plots/nwords_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=total_num_words))+
+  facet_grid(rows =vars(condition)) +
+  geom_point()+
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(title="Number of words", y="Total number of words", x="Trial number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of words per trial")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+#jpeg("~/Desktop/plots/mlu_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=mlu))+
+  facet_grid( rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+ # scale_y_continuous(limits = c(0,15))+
+  labs(title="MLU", y="MLU", x="Trial number")+
+  theme(legend.position="bottom") + 
+  theme_bw()+
+  ggtitle("MLU per trial")#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_utt_length_rep <- d_flowers %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, repNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words),
+            mlu = mean(utt_length_words))
+
+#jpeg("~/Desktop/plots/nword_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=total_num_words))+
+  facet_grid(rows =vars(condition)) +
+  geom_point()+
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(title="Number of words", y="Total number of words", x="Rep number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of words per rep")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+#jpeg("~/Desktop/plots/mlu_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=mlu))+
+  facet_grid( rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+ # scale_y_continuous(limits = c(0,15))+
+  labs(title="MLU", y="MLU", x="Rep number")+
+  theme(legend.position="bottom") + 
+  theme_bw()+
+  ggtitle("MLU per rep")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
 
 ```
+
+### 2. Number of turns
+```{r number_of_turns}
+
+d_flowers_n_turns_rep <- d_flowers %>%
+  group_by(gameId, condition, repNum)  %>%
+  summarise(n=n())
+
+#jpeg("~/Desktop/plots/nturn_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_turns_rep, aes(x=repNum, y=n))+
+  facet_grid(rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(title="Number of turns", y="Number of turns", x="Rep number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of turns per Rep")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_turns_trial <- d_flowers %>%
+  group_by(gameId, condition, trialNum)  %>%
+  summarise(n=n())
+
+#jpeg("~/Desktop/plots/nturn_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_turns_trial, aes(x=trialNum, y=n))+
+  facet_grid(rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(title="Number of turns", y="Number of turns", x="Trial number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of turns per trial")#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_turns_block <- d_flowers %>%
+  group_by(gameId, condition, blockNum)  %>%
+  summarise(n=n())
+
+#jpeg("~/Desktop/plots/nturn_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_turns_block, aes(x=blockNum, y=n))+
+  facet_grid(rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(title="Number of turns", y="Number of turns", x="Block number")+
+  theme(legend.position="bottom") + 
+  theme_bw() +
+  ggtitle("Number of turns per block")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+```
+
+#second plot
 ### 2b. Number of turns
+
 ```{r number_of_turns2}
@@
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, trialNum, gameId, blockNum) %>%
+  group_by( trialNum, gameId, condition) %>%
   summarise(n_utt_speaker=n())
 
+#jpeg("~/Desktop/plots/nutts_trial.jpg", width = 950, height = 950)
 ggplot(d_flowers_n_utt_speaker, 
-       aes(x=trialNum, y=n_utt_speaker, label=playerId)) + 
+       aes(x=trialNum, y=n_utt_speaker)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~ playerId) +
-  xlab("Round number") + 
-  ylab("Number of turns") + 
-  theme_bw()
+  facet_grid(~condition) +
+  xlab("Trial number") + 
+  ylab("Number of utterances") + 
+  theme_bw()#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by( blockNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+#jpeg("~/Desktop/plots/nutts_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_utt_speaker, 
+       aes(x=blockNum, y=n_utt_speaker)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(~condition) +
+  xlab("Block number") + 
+  ylab("Number of utterances") + 
+  theme_bw()#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by( repNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+#jpeg("~/Desktop/plots/nutts_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_utt_speaker, 
+       aes(x=repNum, y=n_utt_speaker)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(~condition) +
+  xlab("Rep number") + 
+  ylab("Number of utterances") + 
+  theme_bw()#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
 ```
@@
   separate_rows(pos, convert = TRUE) %>%
-  group_by(gameId, blockNum, trialNum, repNum, playerId, pos) %>%
+  group_by(condition, blockNum, gameId, pos) %>%
   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
   summarise(pos_count=n()) %>%
-  left_join(d_flowers_utt_length) %>%
+  left_join(d_flowers_utt_length_block) %>%
   mutate(prop_pos=pos_count/total_num_words)
 
+#jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
 ggplot(d_flowers_pos, 
-       aes(x=trialNum, y=prop_pos, label=playerId)) + 
+       aes(x=blockNum, y=prop_pos, label=gameId)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~pos, nrow=1)+
-  xlab("Round number") + 
+  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  xlab("Block number") + 
   ylab("Proportion of pos") + 
-  theme_bw()
+  theme_bw()#+
+ # theme(text = element_text(size = 30))
+#dev.off()
 
+d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+  separate_rows(pos, convert = TRUE) %>%
+  group_by(condition, trialNum, gameId, pos) %>%
+  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+  summarise(pos_count=n()) %>%
+  left_join(d_flowers_utt_length_trial) %>%
+  mutate(prop_pos=pos_count/total_num_words)
+
+#jpeg("~/Desktop/plots/pos_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_pos, 
+       aes(x=trialNum, y=prop_pos, label=gameId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  xlab("Trial number") + 
+  ylab("Proportion of pos") + 
+  ylim(0,5)+
+  theme_bw()#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+  separate_rows(pos, convert = TRUE) %>%
+  group_by(condition, repNum, gameId, pos) %>%
+  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+  summarise(pos_count=n()) %>%
+  left_join(d_flowers_utt_length_rep) %>%
+  mutate(prop_pos=pos_count/total_num_words)
+
+#jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_pos, 
+       aes(x=repNum, y=prop_pos, label=gameId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  xlab("Rep number") + 
+  ylab("Proportion of pos") +
+  ylim(0,5)+
+  theme_bw()#+
+ # theme(text = element_text(size = 30))
+#dev.off() 
 ```
@@
 ttr <- function(text_){
-  text_ <- gsub('[[:punct:] ]+',' ',text_) ####plus remove caps
-  ttr_<- type_token_ratio(text__, n.words=10)
+  text_ <- tolower(text_)
+  total_words = sapply(strsplit(text_, " "), length)
+  no_duplicates = vapply(lapply(strsplit(text_, " "), unique), paste, character(1L), collapse = " ")
+  total_unique = str_count(no_duplicates, "\\W+") + 1
+  ttr_ = total_unique/total_words
   return(ttr_)
@@
 
-d_flowers_text <- d_flowers %>%
-    group_by(gameId, blockNum, trialNum, repNum, playerId) %>%
+d_flowers_text_trial <- d_flowers %>%
+    group_by(gameId, trialNum, condition) %>%
     mutate(text_block = paste0(text, collapse = " ")) %>%
@@
 
-text_block <- d_flowers_text$text_block
+d_flowers_ttr_trial <- d_flowers_text_trial %>%
+  mutate(ttr_ = ttr(text_block))
 
-#d_flowers_ttr <- text_block %>%
-#  mutate(ttr_v=ttr(text_block))
+#t<-str_split(text_, " ", 1:6)
+#y<- paste(unlist(lapply(t,head,n=5)), collapse=" ")
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_ttr_trial, 
+       aes(x=trialNum, y=ttr_)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(condition)) +
+  xlab("Trial number") + 
+  ylab("Type_token ratio") + 
+  ylim(0,1)+
+  theme_bw()+
+  ggtitle("TTR per trial")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_text_rep <- d_flowers %>%
+    group_by(gameId, repNum, condition) %>%
+    mutate(text_block = paste0(text, collapse = " ")) %>%
+    select(-c(text)) %>%
+    distinct()
+
+d_flowers_ttr_rep <- d_flowers_text_rep %>%
+  mutate(ttr_ = ttr(text_block))
+
+#jpeg("~/Desktop/plots/ttr_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_ttr_rep, 
+       aes(x=repNum, y=ttr_)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(condition)) +
+  xlab("Rep number") + 
+  ylab("Type_token ratio") + 
+  ylim(0,1)+
+  theme_bw()+
+  ggtitle("TTR per rep")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_text_block <- d_flowers %>%
+    group_by(gameId, blockNum, condition) %>%
+    mutate(text_block = paste0(text, collapse = " ")) %>%
+    select(-c(text)) %>%
+    distinct()
+
+d_flowers_ttr_block <- d_flowers_text_block %>%
+  mutate(ttr_ = ttr(text_block))
+
+#jpeg("~/Desktop/plots/ttr_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_ttr_block, 
+       aes(x=blockNum, y=ttr_)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_grid(rows =vars(condition)) +
+  xlab("Block number") + 
+  ylab("Type_token ratio") + 
+  ylim(0,1)+
+  theme_bw()+
+  ggtitle("TTR per block")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
 
@@
 ## STEP FOUR - Feedback items and markers
-
 ### 1. Distribution of speech across speakers
@@
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, trialNum, gameId, blockNum) %>%
+  group_by(playerId, trialNum, gameId, condition) %>%
   summarise(n_utt_speaker=n())
@@
 d_flowers_n_utt_game <- d_flowers %>%
-  group_by(trialNum, gameId, blockNum) %>%
+  group_by(trialNum, gameId, condition) %>%
   summarise(n_utt_game=n())
@@
 
-#ggplot(d_flowers_n_utt_dist, 
-#       aes(x=trialNum, y=overall_turns, label=playerId)) + 
-#  geom_smooth(method = "lm", formula = y~x) + 
-#  geom_point(alpha=.1) +
-#  facet_wrap(~playerId)+
-#  xlab("Round number") + 
-#  ylab("Proportion of turns") + 
-#  theme_bw()
+#jpeg("~/Desktop/plots/proputts_trial_speaker.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_utt_dist, 
+       aes(x=trialNum, y=overall_turns, label=playerId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_wrap(~condition)+
+  xlab("Trial number") + 
+  ylab("Proportion of turns") + 
+  theme_bw()+
+  ggtitle("Proportion of turns per trial and speaker")#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(playerId, blockNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+d_flowers_n_utt_game <- d_flowers %>%
+  group_by(blockNum, gameId, condition) %>%
+  summarise(n_utt_game=n())
+
+d_flowers_n_utt_dist <- d_flowers_n_utt_speaker %>%
+  left_join(d_flowers_n_utt_game) %>%
+  mutate(overall_turns=n_utt_speaker/n_utt_game)
+
+#jpeg("~/Desktop/plots/proputts_block_speaker.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_utt_dist, 
+       aes(x=blockNum, y=overall_turns, label=playerId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_wrap(~condition)+
+  xlab("Block number") + 
+  ylab("Proportion of turns") + 
+  theme_bw()+
+  ggtitle("Proportion of turns per block and speaker")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(playerId, repNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+d_flowers_n_utt_game <- d_flowers %>%
+  group_by(repNum, gameId, condition) %>%
+  summarise(n_utt_game=n())
+
+d_flowers_n_utt_dist <- d_flowers_n_utt_speaker %>%
+  left_join(d_flowers_n_utt_game) %>%
+  mutate(overall_turns=n_utt_speaker/n_utt_game)
+
+#jpeg("~/Desktop/plots/proputts_rep_speaker.jpg", width = 950, height = 950)
+ggplot(d_flowers_n_utt_dist, 
+       aes(x=repNum, y=overall_turns, label=playerId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  geom_point(alpha=.1) +
+  facet_wrap(~condition)+
+  xlab("Rep number") + 
+  ylab("Proportion of turns") + 
+  theme_bw()+
+  ggtitle("Proportion of turns per rep and speaker")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
 
@@
 
-```{r cars}
+```{r questions}
 
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, trialNum, gameId, blockNum) %>%
+  group_by(trialNum, gameId, condition) %>%
   summarise(n_utt_speaker=n())
@@
                              "other")) %>%
- group_by(playerId, trialNum, gameId, blockNum, sentence_type) %>%
+ group_by(trialNum, gameId, sentence_type) %>%
  summarise(n_sentence_Type=n()) %>%
@@
 
+#jpeg("~/Desktop/plots/propquestions_trial.jpg", width = 950, height = 950)
 ggplot(d_flowers_q, 
-       aes(x=trialNum, y=prop_questions, label=playerId)) + 
+       aes(x=trialNum, y=prop_questions, label=gameId)) + 
   geom_smooth(method = "lm", formula = y~x) + 
-  facet_wrap(~ playerId) +
-  geom_point(alpha=.1) +
-  xlab("Round number") + 
+  facet_wrap(~ condition) +
+  geom_point(alpha=.5) +
+  xlab("Trial number") + 
   ylab("Proportion of questions") + 
-  theme_bw()
+  theme_bw()+
+  ggtitle("Proportion of questions per trial")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+#####
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(repNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+d_flowers_q <- d_flowers %>% 
+ mutate(sentence_type = ifelse(str_detect(text, "\\?"), 
+                             "question",
+                             "other")) %>%
+ group_by(repNum, gameId, sentence_type) %>%
+ summarise(n_sentence_Type=n()) %>%
+ filter(sentence_type=="question") %>%
+ left_join(d_flowers_n_utt_speaker) %>%
+ mutate(prop_questions = n_sentence_Type/n_utt_speaker )
+
+#jpeg("~/Desktop/plots/propquestions_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_q, 
+       aes(x=repNum, y=prop_questions, label=gameId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  facet_wrap(~ condition) +
+  geom_point(alpha=.5) +
+  xlab("Rep number") + 
+  ylab("Proportion of questions") + 
+  theme_bw()+
+  ggtitle("Proportion of questions per rep")#+
+#theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(blockNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+d_flowers_q <- d_flowers %>% 
+ mutate(sentence_type = ifelse(str_detect(text, "\\?"), 
+                             "question",
+                             "other")) %>%
+ group_by(blockNum, gameId, sentence_type) %>%
+ summarise(n_sentence_Type=n()) %>%
+ filter(sentence_type=="question") %>%
+ left_join(d_flowers_n_utt_speaker) %>%
+ mutate(prop_questions = n_sentence_Type/n_utt_speaker )
+
+#jpeg("~/Desktop/plots/propquestions_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_q, 
+       aes(x=blockNum, y=prop_questions, label=gameId)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+  facet_wrap(~ condition) +
+  geom_point(alpha=.5) +
+  xlab("Block number") + 
+  ylab("Proportion of questions") + 
+  theme_bw()+
+  ggtitle("Proportion of questions per block")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
 
 ```
+
+### 1. Distribution of speech across speakers
+```{r feedback}
+
+###politeness markers
+
+###backchanneling: (https://www.reading.ac.uk/AcaDepts/ll/app_ling/internal/Cutrone_vol_2.pdf)
+
+backchanneling<- c("mm", "okay", "uh-huh", "ok", "mm-hm", "uh", "um", "agree", "uhuh", "mmm", "wow", "great", "mm", "hm",  "ummm", "hmmmm", "huh", "un", "um", "ohh", "ooo", "see", "oooo", "ununun", "oh", "ah", "true", "agree", "right", "yeah", "good", "really")
+
+###hedges: (https://www.researchgate.net/publication/280125979_Linguistic_Markers_and_Stylistic_Attributes_of_Hedging_in_English_Academic_Papers_Written_by_Native_and_Non-Native_Speakers_of_English/figures)
+
+hedging <- c("may", "perhaps", "might", "possible", "likely", "possibly", "maybe", "probable", "appear", "seem", "suggest", "sometimes", "seemingly", "apparently", "often", "could", "usually", "likely", "tend", "sometimes", "probably", "primarily", "tendency", "largely")
+
+d_flowers_nwords <- d_flowers %>%
+  group_by(trialNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, trialNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words))
+
+d_flowers_feedback<- d_flowers %>%
+  group_by(trialNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%          
+  mutate(text = strsplit(text, " ")) %>%
+  rowwise() %>%
+  mutate(backchannel=list(intersect(backchanneling, text))) %>%
+  mutate(backchannel_length=length(backchannel))%>%
+  mutate(hedge=list(intersect(hedging, text))) %>%
+  mutate(hedge_length=length(hedge)) %>%
+  group_by(gameId, trialNum, condition)  %>%
+  summarise(total_num_backchannel = sum(backchannel_length),
+            total_num_hedge = sum(hedge_length)) %>%
+  left_join(d_flowers_nwords) %>%
+  mutate(prop_backchannel_words= total_num_backchannel/total_num_words) %>%
+  mutate(prop_hedge_words= total_num_hedge/total_num_words)
+
+#jpeg("~/Desktop/plots/propbackchannel_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=trialNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Trial number") + 
+  ylab("Proportion of backchannel words") + 
+  theme_bw()+
+  ggtitle("Proportion of backchannel words per trial")#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+#jpeg("~/Desktop/plots/prophedge_trial.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=trialNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Trial number") + 
+  ylab("Proportion of hedge words") + 
+  theme_bw()+
+  ggtitle("Proportion of hedge words per trial")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+#ggplot(d_flowers_feedback, aes(y=trialNum , x=backchannel_length, label=gameId)) +
+#  geom_point(alpha = .5)+
+#  geom_smooth() +
+#  geom_text_repel(aes(label=ifelse(backchannel=="character(0)",  "", backchannel), #max.overlaps = 20))+
+#  xlim(c(1,5))+
+#  theme_bw()+
+#  facet_wrap(~ condition)
+
+#ggplot(d_flowers_feedback, aes(y=trialNum , x=hedge_length, label=gameId)) +
+#  geom_point(alpha = .5)+
+#  geom_smooth() +
+#  geom_text_repel(aes(label=ifelse(hedge=="character(0)", "", hedge), max.overlaps = 20))+
+#  xlim(c(1,5))+
+#  theme_bw()+
+#  facet_wrap(~ condition)
+
+d_flowers_nwords <- d_flowers %>%
+  group_by(blockNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, blockNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words))
+
+d_flowers_feedback<- d_flowers %>%
+  group_by(blockNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%          
+  mutate(text = strsplit(text, " ")) %>%
+  rowwise() %>%
+  mutate(backchannel=list(intersect(backchanneling, text))) %>%
+  mutate(backchannel_length=length(backchannel))%>%
+  mutate(hedge=list(intersect(hedging, text))) %>%
+  mutate(hedge_length=length(hedge)) %>%
+  group_by(gameId, blockNum, condition)  %>%
+  summarise(total_num_backchannel = sum(backchannel_length),
+            total_num_hedge = sum(hedge_length)) %>%
+  left_join(d_flowers_nwords) %>%
+  mutate(prop_backchannel_words= total_num_backchannel/total_num_words) %>%
+  mutate(prop_hedge_words= total_num_hedge/total_num_words)
+
+#jpeg("~/Desktop/plots/propbackchannel_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=blockNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Block number") + 
+  ylab("Proportion of backchannel words") + 
+  theme_bw()+
+  ggtitle("Proportion of backchannel words per block")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+#jpeg("~/Desktop/plots/prophedge_block.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=blockNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Block number") + 
+  ylab("Proportion of hedge words") + 
+  theme_bw()+
+  ggtitle("Proportion of hedge words per block")#+
+#  theme(text = element_text(size = 30))
+#dev.off()
+
+d_flowers_nwords <- d_flowers %>%
+  group_by(repNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, repNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words))
+
+d_flowers_feedback<- d_flowers %>%
+  group_by(repNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%          
+  mutate(text = strsplit(text, " ")) %>%
+  rowwise() %>%
+  mutate(backchannel=list(intersect(backchanneling, text))) %>%
+  mutate(backchannel_length=length(backchannel))%>%
+  mutate(hedge=list(intersect(hedging, text))) %>%
+  mutate(hedge_length=length(hedge)) %>%
+  group_by(gameId, repNum, condition)  %>%
+  summarise(total_num_backchannel = sum(backchannel_length),
+            total_num_hedge = sum(hedge_length)) %>%
+  left_join(d_flowers_nwords) %>%
+  mutate(prop_backchannel_words= total_num_backchannel/total_num_words) %>%
+  mutate(prop_hedge_words= total_num_hedge/total_num_words)
+
+#jpeg("~/Desktop/plots/propbackchannel_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=repNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Rep number") + 
+  ylab("Proportion of backchannel words") + 
+  theme_bw()+
+  ggtitle("Proportion of backchannel words per rep")#+
+ # theme(text = element_text(size = 30))
+#dev.off()
+
+#jpeg("~/Desktop/plots/prophedge_rep.jpg", width = 950, height = 950)
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=repNum, label=gameId)) +
+  facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Rep number") + 
+  ylab("Proportion of hedge words") + 
+  theme_bw()+
+  ggtitle("Proportion of hedge words per rep")#+
+  #theme(text = element_text(size = 30))
+#dev.off()
+
+```
+
+```{r minimum_edit_distance, include=TRUE}
+
+```

### 2021-07-27 code/linguistic_analysis_C-D.Rmd
separate conditions by color

@@
   filter(type=="message") %>%
-  filter(!is.na(text))
-#write_csv(d, "~/Downloads/raw_chat2.csv")
+  filter(!is.na(text)) %>% 
+  filter(!condition=="coopCartel")
+write_csv(d, "~/Downloads/raw_chat3.csv")
 
@@
 #jpeg("~/Desktop/plots/nwords_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=total_num_words))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=total_num_words, color=condition))+
+  #facet_grid(rows =vars(condition)) +
   geom_point()+
@@
 #jpeg("~/Desktop/plots/mlu_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=mlu))+
-  facet_grid( rows =vars(condition)) +
+ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=mlu,  color=condition))+
+#  facet_grid( rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 #jpeg("~/Desktop/plots/nwords_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=total_num_words))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=total_num_words,  color=condition))+
+#  facet_grid(rows =vars(condition)) +
   geom_point()+
@@
 #jpeg("~/Desktop/plots/mlu_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=mlu))+
-  facet_grid( rows =vars(condition)) +
+ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=mlu,  color=condition))+
+#  facet_grid( rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 #jpeg("~/Desktop/plots/nword_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=total_num_words))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=total_num_words,  color=condition))+
+#  facet_grid(rows =vars(condition)) +
   geom_point()+
@@
 #jpeg("~/Desktop/plots/mlu_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=mlu))+
-  facet_grid( rows =vars(condition)) +
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=mlu,  color=condition))+
+#  facet_grid( rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 #jpeg("~/Desktop/plots/nturn_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_turns_rep, aes(x=repNum, y=n))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_n_turns_rep, aes(x=repNum, y=n,  color=condition))+
+ # facet_grid(rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 #jpeg("~/Desktop/plots/nturn_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_turns_trial, aes(x=trialNum, y=n))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_n_turns_trial, aes(x=trialNum, y=n,  color=condition))+
+  #facet_grid(rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 #jpeg("~/Desktop/plots/nturn_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_turns_block, aes(x=blockNum, y=n))+
-  facet_grid(rows =vars(condition)) +
+ggplot(d_flowers_n_turns_block, aes(x=blockNum, y=n,  color=condition))+
+ # facet_grid(rows =vars(condition)) +
   geom_jitter(alpha=.05)+
@@
 ggplot(d_flowers_n_utt_speaker, 
-       aes(x=trialNum, y=n_utt_speaker)) + 
+       aes(x=trialNum, y=n_utt_speaker,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(~condition) +
+ # facet_grid(~condition) +
   xlab("Trial number") + 
@@
 ggplot(d_flowers_n_utt_speaker, 
-       aes(x=blockNum, y=n_utt_speaker)) + 
+       aes(x=blockNum, y=n_utt_speaker,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(~condition) +
+ # facet_grid(~condition) +
   xlab("Block number") + 
@@
 ggplot(d_flowers_n_utt_speaker, 
-       aes(x=repNum, y=n_utt_speaker)) + 
+       aes(x=repNum, y=n_utt_speaker,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(~condition) +
+#  facet_grid(~condition) +
   xlab("Rep number") + 
@@
 ggplot(d_flowers_pos, 
-       aes(x=blockNum, y=prop_pos, label=gameId)) + 
+       aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  facet_grid( cols=vars(pos)) +
   xlab("Block number") + 
@@
 ggplot(d_flowers_pos, 
-       aes(x=trialNum, y=prop_pos, label=gameId)) + 
+       aes(x=trialNum, y=prop_pos, label=gameId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  facet_grid( cols=vars(pos)) +
   xlab("Trial number") + 
@@
 ggplot(d_flowers_pos, 
-       aes(x=repNum, y=prop_pos, label=gameId)) + 
+       aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition), cols=vars(pos)) +
+  facet_grid(cols=vars(pos)) +
   xlab("Rep number") + 
@@
 ggplot(d_flowers_ttr_trial, 
-       aes(x=trialNum, y=ttr_)) + 
+       aes(x=trialNum, y=ttr_,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition)) +
+ # facet_grid(rows =vars(condition)) +
   xlab("Trial number") + 
@@
 ggplot(d_flowers_ttr_rep, 
-       aes(x=repNum, y=ttr_)) + 
+       aes(x=repNum, y=ttr_,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition)) +
+ # facet_grid(rows =vars(condition)) +
   xlab("Rep number") + 
@@
 ggplot(d_flowers_ttr_block, 
-       aes(x=blockNum, y=ttr_)) + 
+       aes(x=blockNum, y=ttr_,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_grid(rows =vars(condition)) +
+ # facet_grid(rows =vars(condition)) +
   xlab("Block number") + 
@@
 ggplot(d_flowers_n_utt_dist, 
-       aes(x=trialNum, y=overall_turns, label=playerId)) + 
+       aes(x=trialNum, y=overall_turns, label=playerId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~condition)+
+#  facet_wrap(~condition)+
   xlab("Trial number") + 
@@
 ggplot(d_flowers_n_utt_dist, 
-       aes(x=blockNum, y=overall_turns, label=playerId)) + 
+       aes(x=blockNum, y=overall_turns, label=playerId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~condition)+
+ # facet_wrap(~condition)+
   xlab("Block number") + 
@@
 ggplot(d_flowers_n_utt_dist, 
-       aes(x=repNum, y=overall_turns, label=playerId)) + 
+       aes(x=repNum, y=overall_turns, label=playerId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
   geom_point(alpha=.1) +
-  facet_wrap(~condition)+
+ # facet_wrap(~condition)+
   xlab("Rep number") + 
@@
 ggplot(d_flowers_q, 
-       aes(x=trialNum, y=prop_questions, label=gameId)) + 
+       aes(x=trialNum, y=prop_questions, label=gameId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
-  facet_wrap(~ condition) +
+ # facet_wrap(~ condition) +
   geom_point(alpha=.5) +
@@
 ggplot(d_flowers_q, 
-       aes(x=repNum, y=prop_questions, label=gameId)) + 
+       aes(x=repNum, y=prop_questions, label=gameId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
-  facet_wrap(~ condition) +
+#  facet_wrap(~ condition) +
   geom_point(alpha=.5) +
@@
 ggplot(d_flowers_q, 
-       aes(x=blockNum, y=prop_questions, label=gameId)) + 
+       aes(x=blockNum, y=prop_questions, label=gameId,  color=condition)) + 
   geom_smooth(method = "lm", formula = y~x) + 
-  facet_wrap(~ condition) +
+  #facet_wrap(~ condition) +
   geom_point(alpha=.5) +
@@
 #jpeg("~/Desktop/plots/propbackchannel_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=trialNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=trialNum, label=gameId,  color=condition)) +
+ # facet_wrap(~ condition) +
   geom_point(alpha = .5)+
@@
 #jpeg("~/Desktop/plots/prophedge_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=trialNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=trialNum, label=gameId,  color=condition)) +
+ # facet_wrap(~ condition) +
   geom_point(alpha = .5)+
@@
 #jpeg("~/Desktop/plots/propbackchannel_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=blockNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=blockNum, label=gameId,  color=condition)) +
+#  facet_wrap(~ condition) +
   geom_point(alpha = .5)+
@@
 #jpeg("~/Desktop/plots/prophedge_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=blockNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=blockNum, label=gameId,  color=condition)) +
+#  facet_wrap(~ condition) +
   geom_point(alpha = .5)+
@@
 #jpeg("~/Desktop/plots/propbackchannel_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=repNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=repNum, label=gameId,  color=condition)) +
+  #facet_wrap(~ condition) +
   geom_point(alpha = .5)+
@@
 #jpeg("~/Desktop/plots/prophedge_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=repNum, label=gameId)) +
-  facet_wrap(~ condition) +
+ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=repNum, label=gameId,  color=condition)) +
+#  facet_wrap(~ condition) +
   geom_point(alpha = .5)+

### 2021-07-27 code/aa-flowers-pilot_analysis.Rmd
add prelim figs

--- 2021-07-23 code/aa-flowers-pilot_analysis.Rmd
@@
 source(here("code/preprocessing/00_preprocessing_functions.R"))
-theme_set(theme_classic())
+
+#add pretty fonts
+library(showtext)
+font_add_google(name = "Assistant", family = "Assistant")
+font_add_google(name = "Roboto", family = "Roboto")
+showtext_auto()
 
@@
 fig_path=here("figs")
+```
 
+```{r}
+#colors/theme
+bay_palette <- function(num_values) {
+  bay_colors <- c(indigo = "#00496F",
+                  teal = "#0F85A0",
+                        green = "#6B9555",
+                        citron = "#ACB64E",
+                        yellow = "#edd746",
+                        orange = "#E58C35",
+                        flame = "#E1672D",
+                        red = "#BD3922",
+                        maroon = "#903B3D",
+                        orchid = "#623D57")
+  color_order <- c("teal", "red","green", "flame",  "orchid","indigo", "yellow", "maroon", "citron", "orange")
+  num_colors <- length(color_order)
+  if (num_values < num_colors) {
+    unname(bay_colors[Filter(
+      function(color) color %in% color_order[1:num_values],
+      names(bay_colors)
+    )])
+  } else {
+    color_indeces <- 0:(num_values - 1) %% num_colors
+    unname(bay_colors[color_indeces + 1])
+  }
+}
+
+scale_colour_bay <- function(...) {
+  ggplot2::discrete_scale("colour", "bay", bay_palette, ...)
+}
+
+scale_color_bay <- function(...) {
+  scale_colour_bay(...)
+}
+
+scale_fill_bay <- function(...) {
+  ggplot2::discrete_scale("fill", "bay", bay_palette, ...)
+}
+
+theme_jmank <- function (base_size = 11, base_family = "Assistant", base_line_size = base_size/22, 
+    base_rect_size = base_size/22) 
+{
+    theme_classic(base_size = base_size, base_family = base_family, 
+        base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace% 
+        theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
+            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", 
+                size = rel(1)), legend.key = element_blank(), 
+            strip.background = element_rect(fill = "white", colour = "black", 
+                size = rel(2)), complete = TRUE) + 
+    theme(plot.title = element_text(face = "bold"))
+}
 ```
@@
 ```{r process-data, message=FALSE, warning=FALSE}
-data_list <- c("technical_pilots", "collabmulti_pilot0",
-               "collabmulti_20", "compet_20")
-process_all_data(data_list)
-join_all_data()
+# data_list <- c("technical_pilots", "collabmulti_pilot0",
+#                "collabmulti_20", "compet_20", "multi_nonlang_20")
+# process_all_data(data_list)
+# join_all_data()
 ```
@@
 #exclude trial runs
-
+# exclude 
 rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
 
-d.rounds.final <-  d.rounds %>% anti_join(rounds_exclude)
-d.raw_chat.final <- d.raw_chat %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
+d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, condition)) %>% 
+  filter(condition != "coopCartel")%>% anti_join(rounds_exclude)
+d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, condition)) %>% 
+  filter(condition != "coopCartel") %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
 ```
@@
 
-d.by_game_metrics <- d.rounds.final%>% 
-  left_join(d.games %>% select(-createdAt)) %>%  
+d.max_round <- d.contexts %>%
+  group_by(gameId, repNum, blockNum, trialNum) %>%
+  slice_max(utility, n = 3) %>%
+  summarise(roundMax = sum(utility))
+
+d.by_round_metrics <- d.rounds.final  %>% 
   left_join(d.round_word_counts) %>%
-  left_join(game_completion %>% select(gameId, num_rounds)) %>%
-  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
-         fullCondition = paste(condition, languageCondition, sep = "-")) %>% 
-  #group by game
-  group_by(gameId, name,condition, languageCondition, fullCondition, gameLength,num_rounds) %>%
+  group_by(gameId, trialNum, repNum, blockNum) %>%
   summarize(groupPoints = sum(playerUtility, na.rm = T),
             groupNumWords = sum(total_num_words, na.rm = T),
-            groupNumChars = sum(total_num_chars, na.rm = T))%>%
-  mutate(groupPoints = ifelse(condition=="coopCartel", 
-                               groupPoints/3, groupPoints)) %>% 
-  #adjust for max
-  mutate(adjustedPoints = case_when(condition == "coopCartel" ~ groupPoints/ ((12*3)*48),
-                                    condition == "coopMulti" ~ groupPoints/ ((12+11+10)*48),
-                                    condition == "competCartel" ~ groupPoints/ ((12+11+10)*48)))
+            groupNumChars = sum(total_num_chars, na.rm = T),
+            numFlowers = n_distinct(playerResponse)) %>%
+  left_join(d.max_round) %>%
+  left_join(game_completion %>% select(gameId, gameComplete)) %>%
+  left_join(d.games %>% select(-createdAt)) %>%  
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         roundPointsAdjusted = groupPoints/roundMax)
+
+d.by_block_metrics <- d.by_round_metrics %>% 
+  group_by(gameId, name, gameComplete, condition, 
+           languageCondition, fullCondition, gameLength, blockNum) %>%
+  summarize(blockPoints = sum(groupPoints, na.rm = T),
+            blockNumWords = sum(groupNumWords, na.rm = T),
+            blockNumChars = sum(groupNumChars, na.rm = T),
+            blockMaxPoints = sum(roundMax),
+            blockNumFlowers = sum(numFlowers)) %>%
+  mutate(blockPointsAdjusted = blockPoints/blockMaxPoints,
+         meanNumFlowers = blockNumFlowers*3/(12*3))
+
+d.by_game_metrics <- d.by_block_metrics %>%
+  #from this point, only look at completed games
+    filter(gameComplete) %>%
+  #group by game
+  group_by(gameId, name, condition, languageCondition, fullCondition, gameLength) %>%
+  summarize(gamePoints = sum(blockPoints, na.rm = T),
+            gameNumWords = sum(blockNumWords, na.rm = T),
+            gameNumChars = sum(blockNumChars, na.rm = T),
+            gameMaxPoints = sum(blockMaxPoints),
+            gameNumFlowers = sum(blockNumFlowers)) %>%
+  mutate(pointsAdjusted = gamePoints/gameMaxPoints,
+         meanNumFlowers = gameNumFlowers*3/ (12*4*3))
 ```
@@
 d.by_game_metrics%>% 
-  ggplot(aes(x = languageCondition, y = groupPoints, color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  ggplot(aes(x = languageCondition, y = gamePoints, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + theme_classic() + theme(legend.position = "none")
+  ylim(0, 1400)+
+  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")
 
-ggsave(here(fig_path, "points_distirbution_raw.png"))
+ggsave(here(fig_path, "points_distirbution_raw.png"), width = 3, height = 5)
 ```
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = adjustedPoints, color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  ggplot(aes(x = languageCondition, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group (Max Adjusted)", title = "Adjusted Points by Group") + theme_classic() + theme(legend.position = "none")
+  facet_grid(cols = vars(condition)) + 
+  labs(x = "Language Condition", 
+       y = "Points Earned by Group (Max Adjusted)", 
+       title = "Points by Group") + 
+  ylim(0,1) +
+  scale_color_bay() + theme_jmank() + theme(legend.position = "none")
 
-ggsave(here(fig_path, "points_distirbution_adjusted.png"))
+ggsave(here(fig_path, "points_distirbution_adjusted.png"),  width = 3, height = 5)
 ```
@@
 ### Time
-```{r, fig.height=3, fig.width=5}
+```{r, fig.height=5, fig.width=3}
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = gameLength/60, color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  ggplot(aes(x = languageCondition, y = gameLength/60, 
+             color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "Time (In Minutes)", title = "Game Length")+theme_classic() + theme(legend.position = "none")
+  facet_grid(cols = vars(condition))+ 
+  labs(y = "Time (In Minutes)", title = "Game Length", x = "Language Condition")+
+  ylim(0, 90)+
+  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+ggsave(here(fig_path, "time_distirbution.png"),  width = 3, height = 5)
+
 ```
+
 ```{r, fig.height=3, fig.width=5}
@@
   ggplot(aes(x = languageCondition, y = log(gameLength/60), color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "Time (In Log Minutes)", title = "Game Length (Log Mins)")+theme_classic() + theme(legend.position = "none")
+  facet_grid(cols = vars(condition)) + 
+  ylim(0, 4.5)+
+  labs(y = "Time (In Log Minutes)", title = "Game Length (Log Mins)")+
+  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+
+ggsave(here(fig_path, "time_log_distirbution.png"))
 ```
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = groupNumWords, color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  ggplot(aes(x = languageCondition, y = gameNumWords, color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged")+theme_classic() + theme(legend.position = "none")
+  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged", x = "Language Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+
+ggsave(here(fig_path, "numwords_distirbution.png"), height = 5, width = 3)
+
 ```
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = log(groupNumWords), color = condition, shape = languageCondition)) + 
-  geom_jitter(width = .2, alpha = .5) +
+  ggplot(aes(x = languageCondition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
+  geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "Log N. Words Exchanged")+theme_classic() + theme(legend.position = "none")
+  facet_grid(cols = vars(condition))+ labs(y = "Log N. Words Exchanged")+
+  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+ggsave(here(fig_path, "numwords_log_distirbution.png"))
+
 ```
 
-#### Condition Effects on Language Use
-```{r}
-d.round_word_counts %>%left_join(d.games %>% 
-                                   select(gameId, condition,chatEnabled)) %>% 
-  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
-         fullCondition = paste(condition, languageCondition, sep = "-"),
-         block_name = paste0("Block ", blockNum)) %>%
-    filter(languageCondition == "Lang") %>%
-  group_by(gameId, fullCondition, languageCondition, condition) %>% 
-  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>% 
-  ggplot() + geom_density(aes(x = groupNumWords, fill = condition), alpha = .5, adjust = .7)
-```
-
-## Scatters
+## Covariate
 ### Time x Points
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = gameLength/60, y = groupPoints, color = condition)) + 
+  ggplot(aes(x = gameNumWords, y = gameLength/60, color = condition, shape = languageCondition)) + 
   geom_point(alpha = .5) +
   geom_smooth(method=lm)+ 
-  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic() + theme(legend.position = "none")
+  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
+  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+ggsave(here(fig_path, "time_words_scatter.png"))
 ```
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = gameLength/60, y = groupPoints, color = log(groupNumWords))) + 
+  ggplot(aes(x = gameLength/60, y = gamePoints, color = log(gameNumWords+1))) + 
   geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic() + scale_color_steps()
+  geom_smooth(method=lm, color = "black")+
+  facet_grid(cols = vars(condition)) + 
+  labs(x = "Game length (in Minutes)", 
+       y = "Points Earned by Group",
+       color = "Log Num. Words") + 
+  theme_jmank() +  scale_color_steps(low = "#0F85A0", high = "#BD3922")
+
+ggsave(here(fig_path, "all_correlated.png"), height = 3, width = 5)
+```
+
+```{r, fig.height=3, fig.width=5}
+ms <- d.by_game_metrics %>%
+  group_by(condition, languageCondition) %>%
+  summarise(mean_len = mean(gameLength/60),
+            ci_len = 1.96 * sd(gameLength/60) / sqrt(length(gameLength)),
+            mean_points = mean(gamePoints),
+            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))
+
+d.by_game_metrics %>%
+  ggplot(aes(x = gameLength/60, y = gamePoints, color = condition, shape = languageCondition)) +#, size = log(groupNumWords+1))) + 
+  geom_smooth(aes(group = condition), method = "lm") +
+  geom_point(alpha = .2) +
+  geom_pointrange(data = ms, aes(x = mean_len, y = mean_points, 
+                                 ymin = mean_points - ci_points,
+                                 ymax = mean_points + ci_points)) +
+  geom_errorbarh(data = ms, aes(x = mean_len, y = mean_points, 
+                                 xmin = mean_len - ci_len,
+                                 xmax = mean_len + ci_len), width = 0) + 
+  # ylim(0,1500) + 
+  xlim(0,90) + 
+  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
+  ggthemes::scale_color_solarized() + 
+  scale_color_bay() + theme_jmank() +  
+  theme(legend.position = "bottom")
+
+ggsave(here(fig_path, "time_points_scatter_nofacet.png"))
+
 ```
@@
 d.by_game_metrics %>% 
-  ggplot(aes(x = gameLength/60, y = groupPoints, color = condition)) + 
+  ggplot(aes(x = log(gameLength/60), y = gamePoints, color = condition)) + 
   geom_point(alpha = .5) +
   geom_smooth(method=lm)+
-  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+theme_classic()
+  labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+scale_color_bay() + theme_jmank()
+
+ggsave(here(fig_path, "log_time_points_scatter_nofacet.png"))
+```
+### Language x Points
+
+```{r}
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>%
+  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition)) + 
+  geom_point(alpha = .5) +
+  geom_smooth(method=lm)+
+  labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+
+ms <- d.by_game_metrics %>%
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
+  group_by(condition, fullCondition, languageCondition) %>%
+  summarise(mean_nw = mean(gameNumWords),
+            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
+            mean_points = mean(gamePoints),
+            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))
+
+d.by_game_metrics %>%
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
+  #filter(langUse) %>%
+  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = languageCondition)) +
+  #geom_smooth(aes(group = fullCondition), method = "lm") +
+  geom_smooth(aes(group = fullCondition), method=glm, formula=y~poly(x,2), alpha=.3)+
+  geom_point( alpha = .5) +
+  geom_pointrange(data = ms, aes(x = mean_nw, y = mean_points,
+                                ymin = mean_points - ci_points,
+                                ymax = mean_points + ci_points)) +
+  geom_errorbarh(data = ms, aes(x = mean_nw, y = mean_points,
+                                xmin = mean_nw - ci_nw,
+                                xmax = mean_nw + ci_nw), width = 0) +
+  # ylim(0,1500) + 
+  labs(x = "Words Exchanged", y = "Points Earned", title = ) + 
+  scale_color_bay() + theme_jmank() +  
+  theme(legend.position = "bottom")
 ```
 
-Log
+```{r}
+ms <- d.by_game_metrics %>%
+  mutate(langUse = ifelse(groupNumWords > 0, TRUE, FALSE)) %>% 
+  group_by(condition, langUse) %>%
+  summarise(mean_nw = mean(groupNumWords),
+            ci_nw = 1.96 * sd(groupNumWords) / sqrt(length(gameLength)),
+            mean_points = mean(groupPoints),
+            ci_points = 1.96 * sd(groupPoints) / sqrt(length(gameLength)))
+
+d.by_game_metrics %>%
+  mutate(langUse = ifelse(groupNumWords > 0, TRUE, FALSE)) %>% 
+  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition, shape = langUse)) +
+  #geom_smooth(aes(group = condition), method = "lm") +
+  geom_point(alpha = .5) +
+  geom_pointrange(data = ms, aes(x = mean_nw, y = mean_points, 
+                                 ymin = mean_points - ci_points,
+                                 ymax = mean_points + ci_points)) +
+  geom_errorbarh(data = ms, aes(x = mean_nw, y = mean_points, 
+                                 xmin = mean_nw - ci_nw,
+                                 xmax = mean_nw + ci_nw), width = 0) + 
+  # ylim(0,1500) + 
+  labs(x = "Words Echanged", y = "Points Earned") + 
+  scale_color_bay() + theme_jmank() +  
+  theme(legend.position = "bottom")
+```
+
 ```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% 
-  ggplot(aes(x = log(gameLength/60), y = groupPoints, color = condition)) + 
+d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>%
+  ggplot(aes(x= gameNumWords, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
   geom_point(alpha = .5) +
   geom_smooth(method=lm)+ 
-  facet_grid(cols = vars(condition))+ labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+theme_classic() + theme(legend.position = "none")
-```
-```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% 
-  ggplot(aes(x = log(gameLength/60), y = groupPoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+theme_classic()
-```
-### Language x Points
-#### Faceted
-```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  facet_grid(cols = vars(condition))+ labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+theme_classic() + theme(legend.position = "none")
-```
-```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = groupNumWords, y = adjustedPoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  facet_grid(cols = vars(condition))+ labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+theme_classic() + theme(legend.position = "none")
-```
-#### Non-Faceted
-```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+theme_classic()
-```
-```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = groupNumWords, y = adjustedPoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+theme_classic()
+  labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+scale_color_bay() + 
+  theme_jmank() +  theme(legend.position = "none")
+
+ggsave(here(fig_path, "nwords_points_scatter.png"), height = 3, width = 3)
 ```
@@
 ### Progress Over Time
+
+tried to make a gif? :/
+```{r}
+trial_num = 1
+randomGame = sample(game_completion %>% filter(chat == T, gameComplete == T) %>% pull(gameId), 1)
+
+save_plot <- function(rep_Num, block_Num, randomGame) {
+  full_data <- d.by_round_metrics %>% 
+    filter(repNum == rep_Num, blockNum == block_Num) %>%
+  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% filter(chatEnabled)
+
+  to_plot <- full_data %>% ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
+                                      shape = condition, color = condition)) + 
+    geom_point(alpha = .2) + 
+    geom_point(data = full_data %>% filter(gameId == randomGame), aes(x = groupNumWords, y = roundPointsAdjusted, 
+                                      shape = condition, color = condition), alpha = 1) + 
+    ggtitle(paste0("Block ", blockNum, ", Trial ", repNum))+
+  theme_jmank() + scale_color_bay()
+
+  ggsave(to_plot, file = here(fig_path, "animate", paste0(block_Num, "_", rep_Num, ".png")), height = 3, width = 5)
+  }
+#mapply(save_plot, seq(0,11), rep(0, 12), rep("sTX6ksbxqfnfsmzxe", 12))
+```
 
@@
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ theme_classic()
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
+ggsave(here(fig_path, "nwords_over_time.png"))
 ```
@@
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log (N. Words + 1)")+ theme_classic()
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank()
+ggsave(here(fig_path, "nwords_log_over_time.png"))
+
 ```
@@
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ theme_classic()
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
 ```
 
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
+  filter(groupNumWords > 0) %>%
+  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
+```
+
+log num words
+
+```{r, fig.height=3, fig.width=5}
+d.round_word_counts %>%left_join(d.games %>% 
+                                   select(gameId, condition, chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(languageCondition == "Lang") %>%
+  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>%
+
+  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
+  #filter(groupNumWords >1) %>%
+  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  ylim(0, 6) +
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
+```
 ### Accuracy
-#### N. Overlap
+#### N. Flowers Selected each round
+```{r}
+d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
+  summarize(num_unique_flowers = n_distinct(playerResponse),
+            mean_response_time = mean(time_sec, na.rm = T))  %>%
+  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(condition != "coopCartel") %>%
+  ggplot(aes(x = repNum, y = num_unique_flowers, color = condition, linetype = languageCondition)) + 
+  geom_point(alpha = .1) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name), rows = vars(condition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()
+```
+
+```{r}
+d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
+  summarize(num_unique_flowers = n_distinct(playerResponse),
+            mean_response_time = mean(time_sec, na.rm = T))  %>%
+  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(condition != "coopCartel") %>%
+  ggplot(aes(x = repNum, y = log(mean_response_time), color = condition, linetype = languageCondition)) + 
+  geom_point(alpha = .1) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time")+ scale_color_bay() + theme_jmank()
+```
+
+#More Accurate proportion possible/accuracy
+
+Are they getting better over games?
+```{r}
+d.by_round_metrics %>% left_join(d.games %>% select(gameId, condition, chatEnabled))
+```
 

### 2021-07-28 code/aa-flowers-pilot_analysis.Rmd
final tweaks to figures before talk

@@
 
-```{r set-up, include=F}
-library(tidyverse)
+```{r set-up, message=FALSE, include=F}
+
 library(jsonlite)
 library(here)
+library(tidyverse)
+
 source(here("code/preprocessing/00_preprocessing_functions.R"))
-
+source(here("code/theme_specs.R"))
 #add pretty fonts
@@
 fig_path=here("figs")
-```
-
-```{r}
-#colors/theme
-bay_palette <- function(num_values) {
-  bay_colors <- c(indigo = "#00496F",
-                  teal = "#0F85A0",
-                        green = "#6B9555",
-                        citron = "#ACB64E",
-                        yellow = "#edd746",
-                        orange = "#E58C35",
-                        flame = "#E1672D",
-                        red = "#BD3922",
-                        maroon = "#903B3D",
-                        orchid = "#623D57")
-  color_order <- c("teal", "red","green", "flame",  "orchid","indigo", "yellow", "maroon", "citron", "orange")
-  num_colors <- length(color_order)
-  if (num_values < num_colors) {
-    unname(bay_colors[Filter(
-      function(color) color %in% color_order[1:num_values],
-      names(bay_colors)
-    )])
-  } else {
-    color_indeces <- 0:(num_values - 1) %% num_colors
-    unname(bay_colors[color_indeces + 1])
-  }
-}
-
-scale_colour_bay <- function(...) {
-  ggplot2::discrete_scale("colour", "bay", bay_palette, ...)
-}
-
-scale_color_bay <- function(...) {
-  scale_colour_bay(...)
-}
-
-scale_fill_bay <- function(...) {
-  ggplot2::discrete_scale("fill", "bay", bay_palette, ...)
-}
-
-theme_jmank <- function (base_size = 11, base_family = "Assistant", base_line_size = base_size/22, 
-    base_rect_size = base_size/22) 
-{
-    theme_classic(base_size = base_size, base_family = base_family, 
-        base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace% 
-        theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
-            panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", 
-                size = rel(1)), legend.key = element_blank(), 
-            strip.background = element_rect(fill = "white", colour = "black", 
-                size = rel(2)), complete = TRUE) + 
-    theme(plot.title = element_text(face = "bold"))
-}
 ```
@@
 
+Rename conditions for readability
+```{r}
+d.games <- d.games %>% mutate(conditionName = condition, 
+                              condition = case_when(condition == "coopMulti" ~ "Cooperative",
+                                         condition == "competCartel" ~ "Competitive",
+                                         condition == "coopCartel" ~ "Old Cooperative"))
+d.games
+```
+
 ```{r}
@@
 
-d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, condition)) %>% 
-  filter(condition != "coopCartel")%>% anti_join(rounds_exclude)
-d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, condition)) %>% 
-  filter(condition != "coopCartel") %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
+d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
+  filter(conditionName != "coopCartel")%>% anti_join(rounds_exclude)
+d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
+  filter(conditionName != "coopCartel") %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
 ```
@@
   mutate(gameComplete = ifelse(num_rounds == 48, TRUE, FALSE))
+
 #knitr::kable(summary)
@@
 #pull out pilots
-d.games %>% group_by(condition, chatEnabled,pilot)%>% tally()
+d.games %>% group_by(conditionName, chatEnabled,pilot)%>% tally()
 ```
@@
 d.feedback
+d.feedback %>% ggplot(aes(x = human)) + geom_bar(stat = "count") + theme_jmank()
 ```
@@
 # Basic Analyses
-## Simple Distributions
-
 ```{r}
@@
  # filter(!is.chitchat) %>% 
-  mutate(text = gsub("\\n", '', fixed = T, text),
-         text = gsub("[/?/.]", ' ', text),
+  mutate(text = gsub('[[:punct:] ]+',' ',text),
          text = str_squish(text),
@@
   group_by(gameId, repNum, blockNum, trialNum) %>%
+  #top 3 flowers
   slice_max(utility, n = 3) %>%
@@
 
+## Simple Distributions
 ### Points
@@
 ```
-
+# All Correlated
 ```{r, fig.height=3, fig.width=5}
 d.by_game_metrics %>%
-  ggplot(aes(x = gameLength/60, y = gamePoints, color = log(gameNumWords+1))) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm, color = "black")+
+  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
+  geom_point(alpha = .8) +
+  geom_smooth(aes(group = condition), method=lm, color = "black")+
   facet_grid(cols = vars(condition)) + 
@@
        y = "Points Earned by Group",
-       color = "Log Num. Words") + 
-  theme_jmank() +  scale_color_steps(low = "#0F85A0", high = "#BD3922")
+       color = "Log Num. Words",
+       shape = "Language") + 
+  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")
 
@@
   ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = languageCondition)) +
-  #geom_smooth(aes(group = fullCondition), method = "lm") +
-  geom_smooth(aes(group = fullCondition), method=glm, formula=y~poly(x,2), alpha=.3)+
+  geom_smooth(aes(group = fullCondition), method = "lm") +
+  #geom_smooth(aes(group = fullCondition), method=glm, formula=y~poly(x,2), alpha=.3)+
   geom_point( alpha = .5) +
@@
 ms <- d.by_game_metrics %>%
-  mutate(langUse = ifelse(groupNumWords > 0, TRUE, FALSE)) %>% 
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
   group_by(condition, langUse) %>%
-  summarise(mean_nw = mean(groupNumWords),
-            ci_nw = 1.96 * sd(groupNumWords) / sqrt(length(gameLength)),
-            mean_points = mean(groupPoints),
-            ci_points = 1.96 * sd(groupPoints) / sqrt(length(gameLength)))
+  summarise(mean_nw = mean(gameNumWords),
+            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
+            mean_points = mean(gamePoints),
+            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))
 
 d.by_game_metrics %>%
-  mutate(langUse = ifelse(groupNumWords > 0, TRUE, FALSE)) %>% 
-  ggplot(aes(x = groupNumWords, y = groupPoints, color = condition, shape = langUse)) +
+  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
+  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = langUse)) +
   #geom_smooth(aes(group = condition), method = "lm") +
@@
 tried to make a gif? :/
-```{r}
+```{r eval=FALSE, include=FALSE}
 trial_num = 1
@@
     filter(languageCondition == "Lang") %>%
-  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  group_by(gameId, condition,block_name, trialNum, repNum) %>% 
   summarize(groupNumWords = sum(total_num_words, na.rm = T)+1) %>%
-  ggplot(aes(x = repNum, y = log(groupNumWords), color = fullCondition)) + 
+  ggplot(aes(x = repNum, y = log(groupNumWords), color = condition)) + 
   geom_point(alpha = .2) +
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank()
-ggsave(here(fig_path, "nwords_log_over_time.png"))
+  facet_grid(cols = vars(block_name)) + 
+  labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")
 
+ggsave(here(fig_path, "nwords_log_over_time.png"), height = 3, width = 5)
 ```
@@
   facet_grid(cols = vars(block_name), rows = vars(condition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()
+
+ggsave(ggsave(here(fig_path, "collisions.png")))
 ```
@@
 ```{r}
-d.by_round_metrics %>% left_join(d.games %>% select(gameId, condition, chatEnabled))
+d.by_round_metrics %>% left_join(d.games %>% 
+                                   select(gameId, condition, chatEnabled))%>%
+  mutate(blockName = paste0("Block ", blockNum))%>% 
+  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
+             color = condition)) + 
+  geom_point(alpha = .25) + theme_jmank() + scale_color_bay() + facet_wrap(~blockName) +
+  labs(title = "Number of points per trial")
 ```
+
+# Language Results
 

### 2021-07-28 code/linguistic_analysis_C-D.Rmd
final tweaks to figures before talk

@@
 
+source("./theme_specs.R")
 ```
@@
 ```{r main data, include=TRUE}
-d<- read_csv("/Users/lscpuser/Documents/AA-flowers2/AA-flowers/data/processed_data/joined_data/filtered_raw_chat.csv")
+d<- read_csv("../data/processed_data/joined_data/filtered_raw_chat.csv")
 d_flowers <- d %>% 
@@
   filter(!is.na(text)) %>% 
-  filter(!condition=="coopCartel")
-write_csv(d, "~/Downloads/raw_chat3.csv")
+  filter(!conditionName=="coopCartel")
+
+write_csv(d, "../data/processed_data/raw_chat3.csv")
 
 d_flowers <- d_flowers %>% 
-  filter(!condition=="coopCartel")
+  filter(!conditionName=="coopCartel")
 
@@
   theme(legend.position="bottom") + 
-  theme_bw() +
+  theme_jmank() + scale_color_bay() +
   ggtitle("Number of words per block")#+
@@
  # scale_y_continuous(limits = c(0,15))+
-  labs(title="MLU", y="MLU", x="Block number")+
+  labs(title="Mean Length of Utterance", y="MLU", x="Block number")+
   theme(legend.position="bottom") + 
-  theme_bw()+
-  ggtitle("MLU per block")#+
+  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
 #  theme(text = element_text(size = 30))
 #dev.off() 
+ggsave("../figs/mean_utt_length_block.png", height = 3, width = 3)
 
@@
   theme(legend.position="bottom") + 
-  theme_bw() +
+  theme_jmank() + scale_color_bay() +
   ggtitle("Number of words per trial")#+
@@
  # scale_y_continuous(limits = c(0,15))+
-  labs(title="MLU", y="MLU", x="Trial number")+
-  theme(legend.position="bottom") + 
-  theme_bw()+
-  ggtitle("MLU per trial")#+
+  labs(title="Mean Length of Utterance", y="MLU", x="Trial number")+
+  theme_jmank() + scale_color_bay() + theme(legend.position="bottom") #+ 
   #theme(text = element_text(size = 30))
 #dev.off()
+
+ggsave("../figs/mean_utt_length.png", height = 5, width = 4)
 
@@
   theme(legend.position="bottom") + 
-  theme_bw() +
+  theme_jmank() + scale_color_bay() +
   ggtitle("Number of words per rep")#+
@@
   theme(legend.position="bottom") + 
-  theme_bw()+
+  theme_jmank() + scale_color_bay() +
   ggtitle("MLU per rep")#+
@@
 
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=log(mlu+1),  color=condition))+
+#  facet_grid( rows =vars(condition)) +
+  geom_jitter(alpha=.05)+
+  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
+  stat_summary(fun.data = "mean_cl_boot")+
+ # scale_y_continuous(limits = c(0,15))+
+  labs(title="MLU", y="MLU", x="Rep number")+
+  theme(legend.position="bottom") + 
+  theme_jmank() + scale_color_bay() +
+  ggtitle("MLU per rep")#+
+ # theme(text = element_text(size = 30))
+#dev.off() 
 ```
@@
   theme(legend.position="bottom") + 
-  theme_bw() +
+  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
   ggtitle("Number of turns per block")#+
@@
 
+ggsave("../figs/num_turns_block.png", height = 3, width = 3)
 ```

### 2021-07-28 code/theme_specs.R

@@ -1,0 +1,50 @@
+#Theming
+#colors/theme
+bay_palette <- function(num_values) {
+  bay_colors <- c(indigo = "#00496F",
+                  teal = "#0F85A0",
+                  green = "#6B9555",
+                  citron = "#ACB64E",
+                  yellow = "#edd746",
+                  orange = "#E58C35",
+                  flame = "#E1672D",
+                  red = "#BD3922",
+                  maroon = "#903B3D",
+                  orchid = "#623D57")
+  color_order <- c("teal", "red","green", "flame",  "orchid","indigo", "yellow", "maroon", "citron", "orange")
+  num_colors <- length(color_order)
+  if (num_values < num_colors) {
+    unname(bay_colors[Filter(
+      function(color) color %in% color_order[1:num_values],
+      names(bay_colors)
+    )])
+  } else {
+    color_indeces <- 0:(num_values - 1) %% num_colors
+    unname(bay_colors[color_indeces + 1])
+  }
+}
+
+scale_colour_bay <- function(...) {
+  ggplot2::discrete_scale("colour", "bay", bay_palette, ...)
+}
+
+scale_color_bay <- function(...) {
+  scale_colour_bay(...)
+}
+
+scale_fill_bay <- function(...) {
+  ggplot2::discrete_scale("fill", "bay", bay_palette, ...)
+}
+
+theme_jmank <- function (base_size = 11, base_family = "Roboto", base_line_size = base_size/22, 
+                         base_rect_size = base_size/22) 
+{
+  theme_classic(base_size = base_size, base_family = base_family, 
+                base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace% 
+    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
+          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", 
+                                                                       size = rel(1)), legend.key = element_blank(), 
+          strip.background = element_rect(fill = "white", colour = "black", 
+                                          size = rel(2)), complete = TRUE) + 
+    theme(plot.title = element_text(face = "bold"))
+}

### 2021-07-29 code/aa-flowers-pilot_analysis.Rmd
created a human-readable transcript to share with group

@@
 
+#exclude cooperative condition
 d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
   filter(conditionName != "coopCartel")%>% anti_join(rounds_exclude)
+
 d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
   filter(conditionName != "coopCartel") %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
+```
+
+Clean up a sharable data csv
+```{r}
+joined_data <- d.raw_chat.final %>% select(-c(index, createdAt, 
+                               submitted, conditionName)) %>% 
+  left_join(d.rounds.final %>% 
+              select(gameId, playerId,repNum, blockNum, trialNum, 
+                     playerResponse, playerEarnings = playerUtility)) %>% 
+  left_join(d.players %>% select(playerId, playerName = name)) %>%
+  left_join(d.contexts %>% select(gameId, repNum, blockNum, trialNum, playerResponse = label, responseRawUtility = utility, blinded))
 ```
@@
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time")+ scale_color_bay() + theme_jmank()
+  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time") + scale_color_bay() + theme_jmank()
 ```

### 2021-07-29 code/aa-flowers-pilot_analysis.Rmd
and the code for the transcript

@@
                                submitted, conditionName)) %>% 
+  # merge chat condition
+  left_join(d.games %>% select(gameId, chatEnabled)) %>%
+  #merge player selections aligned with chat
   left_join(d.rounds.final %>% 
@@
                      playerResponse, playerEarnings = playerUtility)) %>% 
+  #merge player names
   left_join(d.players %>% select(playerId, playerName = name)) %>%
-  left_join(d.contexts %>% select(gameId, repNum, blockNum, trialNum, playerResponse = label, responseRawUtility = utility, blinded))
+  #merge the raw utility for the selections
+  left_join(d.contexts %>% select(gameId, repNum, blockNum, trialNum, 
+                                  playerResponse = label, responseRawUtility = utility, blinded)) %>%
+  #to avoid confusion, drop any round info thats actually a message
+   mutate(playerResponse = ifelse(type == "message", NA, playerResponse),
+         playerEarnings = ifelse(type == "message", NA, playerEarnings),
+         responseRawUtility = ifelse(type == "message", NA, responseRawUtility),
+         blinded = ifelse(type == "message", NA, blinded)) %>%
+  select(gameId, condition, chatEnabled,roundID, repNum:trialNum, playerId, 
+         playerName, text,  type, playerResponse:responseRawUtility, blinded,everything()) %>% write_csv(here("data/interpretable_transcript.csv"))
 ```

### 2021-08-16 experiments/flowers-pilot/server/main.js
add shuffle

@@ -16,7 +16,7 @@ function chooseImages(flowers, low, high, number, players, blinded){
   const f=_.slice(_.shuffle(flowers),0,number)
   const l=_.slice(_.shuffle(low),0,number-1) // choose n-1 from the normal range
   const h=_.slice(_.shuffle(high),0,1) //choose 1 high
-  const u=_.concat(l,h)
+  const u=_.shuffle(_.concat(l,h))
   const values=_.zipWith(f,u,blinds, (a,b,c)=>_.assign({},a,b,c))
   const names=_.map(f, (a)=>a.label)
   const all=_.zipObject(names,values)

### 2021-08-16 experiments/flowers-pilot/server/callbacks.js
move player defs to start for if games terminate early

@@ -19,6 +19,23 @@ Empirica.onGameStart((game) => {
     player.set("nameColor", nameColors[i]);
     player.set("bonus", 0);
   });
+
+  const flowers= game.get("flowers")
+  const flower_names = _.zipObject(flowers.map(a => a.label),flowers.map( ()=> ""))
+  const otherflowers=game.get("otherflowers")
+  //const otherflower_names = _.zipObject(otherflowers.map(a => a.label),otherflowers.map( ()=> ""))
+
+  players.forEach((player, i) => {
+    //player.set("tangrams", _.shuffle(targets));
+    player.set("flower_labels", flower_names)
+    player.set("flowers", _.shuffle(flowers))
+    let othersample = _.slice(_.shuffle(otherflowers),0,4) //randomly choose 4 to show
+    player.set("otherflowers", othersample) 
+    player.set("otherflower_labels",_.zipObject(othersample.map(a => a.label),othersample.map( ()=> "")))
+    let otherplayers=_.filter(_.map(game.players,(a) => a.get("name")), a => a!==player.get("name"))
+    player.set("p1",otherplayers[0])
+    player.set("p2",otherplayers[1])
+  });
 });
 
 // onRoundStart is triggered before each round starts, and before onStageStart.
@@ -123,22 +140,7 @@ Empirica.onGameEnd((game) => {
 
   console.debug("The game", game._id, "has ended");
    //const targets = Map(game.get('context'));
-   const flowers= game.get("flowers")
-   const flower_names = _.zipObject(flowers.map(a => a.label),flowers.map( ()=> ""))
-   const otherflowers=game.get("otherflowers")
-   //const otherflower_names = _.zipObject(otherflowers.map(a => a.label),otherflowers.map( ()=> ""))
-
-   players.forEach((player, i) => {
-     //player.set("tangrams", _.shuffle(targets));
-     player.set("flower_labels", flower_names)
-     player.set("flowers", _.shuffle(flowers))
-     let othersample = _.slice(_.shuffle(otherflowers),0,4) //randomly choose 4 to show
-     player.set("otherflowers", othersample) 
-     player.set("otherflower_labels",_.zipObject(othersample.map(a => a.label),othersample.map( ()=> "")))
-     let otherplayers=_.filter(_.map(game.players,(a) => a.get("name")), a => a!==player.get("name"))
-     player.set("p1",otherplayers[0])
-     player.set("p2",otherplayers[1])
-   });
+  
  });
  
 

### 2021-08-17 experiments/flowers-pilot/server/callbacks.js
comment unneeded code

@@ -166,7 +166,7 @@ Empirica.onGameEnd((game) => {
 
 // // onSet is called when the experiment code call the .set() method
 // // on games, rounds, stages, players, playerRounds or playerStages.
-Empirica.onSet(
+/*Empirica.onSet(
   (
     game,
     round,
@@ -182,7 +182,7 @@ Empirica.onSet(
     if (key === "clicked") {
     }
   }
-);
+);*/
 
 // // onSet is called when the experiment code call the `.append()` method
 // // on games, rounds, stages, players, playerRounds or playerStages.

### 2021-08-17 code/fit_alignment.R
alignment update to compare across conditions

@@ -1,7 +1,3 @@
-library(ggplot2)
-library(dplyr)
-
-
 #install.packages("devtools")
 #devtools::install_github("langcog/langcog")
 
@@ -16,32 +12,51 @@ library(lubridate)
 library(lme4)
 library(rstan)
 library(purrr)
+library(dplyr)
+library(ggplot2)
 
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-result_files <- "data/pilot0/output_alignment.csv"
+##############prepare data 
+
+result_files <- "data/processed_data/joined_data/output_alignment.csv"
 #result_files <-  list.files(path = "../data/pilot0/",recursive = TRUE,
- #                           all.files = FALSE, full.names = TRUE,
-  #                          pattern = "output.csv")
+#all.files = FALSE, full.names = TRUE, pattern = "output.csv")
+
+#add condition to alignment data
+o<- read_csv("data/processed_data/joined_data/filtered_raw_chat.csv")
+coop_players <- o %>% 
+  filter(condition == "coopMulti") 
+list_coop_players <- unique(coop_players$playerId)
+comp_players <- o %>% 
+  filter(condition == "competCartel") 
+list_comp_players <- unique(comp_players$playerId)
+
 
 setup_alignment_data <- function(corpus_df) {
   corpus_df %>%
     group_by(category, subpop) %>%
     summarise_each(funs(sum), ba, nba, bna, nbna) %>%
-    filter(ba + bna > 1) %>%
+    filter(as.numeric(ba) + as.numeric(bna) > 1) %>%
     ungroup() %>%
     mutate(category_num = as.numeric(as.factor(category)),
            subpop_num = as.numeric(as.factor(subpop))) 
 }
 
-
 childes_data <- lapply(result_files, read_csv) %>%
   bind_rows() %>%
   rename(Speaker = speakerId, Replier = replierId) %>%
-  mutate(subpop = paste(Speaker,  "-", Replier)) %>%
-  filter(Speaker != Replier) 
+ # mutate(subpop = paste(Speaker,  "-", Replier)) %>%
+  mutate(subpop= ifelse(Replier %in% list_coop_players, "coop", ifelse(Replier %in% list_comp_players, "comp", ifelse(Speaker %in% list_comp_players, "comp",ifelse(Speaker %in% list_coop_players, "coop","no"))))) %>% 
+ # filter(Speaker != Replier) %>%
[…]

### 2021-08-18 experiments/flowers-pilot/client/main.js
update instruction text and images

@@ -51,7 +51,9 @@ Empirica.header(() => null);
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
 
-  const steps = [Overview, BasicInterfaceInstructions, BasicInterfaceQuiz, AdvancedInterfaceInstructions, AdvancedInterfaceQuiz, GameplayInstructions, GameplayQuiz];
+  const steps = [Overview, BasicInterfaceInstructions, BasicInterfaceQuiz,
+    AdvancedInterfaceInstructions, AdvancedInterfaceQuiz,
+    GameplayInstructions, GameplayQuiz];
   if (treatment.chatEnabled) {
     steps.push(FinalQuizLang);
   }
@@ -69,8 +71,8 @@ Empirica.introSteps((game, treatment) => {
   //}
   //steps.push(MoreAboutBonus);
 
-  //return steps;
-  return[]
+  return steps;
+  //return[]
 });
 
 // The Round component containing the game UI logic.

### 2021-08-18 experiments/flowers-pilot/client/main.js
add other color test to nonlang ending

@@ -99,9 +99,9 @@ Empirica.exitSteps((game, player, treatment) => {
   } else {
     if (game.treatment.chatEnabled) {
     //chat exit survey
-    return [PostTestInstructions,OwnColorTest,OtherColorTest,ExitSurvey, Thanks];}
+    return [PostTestInstructions, OwnColorTest, OtherColorTest,ExitSurvey, Thanks];}
     else {
-      return [ExitSurveyNochat, Thanks];
+      return [PostTestInstructions, OtherColorTest, ExitSurveyNochat, Thanks];
     }
   }
 });

### 2021-08-18 code/aa-flowers-pilot_analysis.Rmd
move previous game data into data/version_0 and version preprocessing code

@@
   mutate(blockName = paste0("Block ", blockNum))%>% 
-  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
+  ggplot(aes(x = log(groupNumWords+1), y = roundPointsAdjusted, 
              color = condition)) + 
@@
   labs(title = "Number of points per trial")
+```
+
+### Selecting Hidden Flowers?
+
+```{r}
+d.rounds.final %>% left_join(d.games %>% 
+                                   select(gameId, condition, chatEnabled)) %>%
+  left_join(d.contexts %>% select(gameId, blockNum, trialNum, repNum, playerResponse = label, blinded)) %>%
+  mutate(select_blinded = playerId == blinded) %>% 
+  group_by(gameId, blockNum, trialNum, repNum, chatEnabled, condition) %>%
+  summarize(numSelectHidden = sum(select_blinded)) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"),
+         block_name = paste0("Block ", blockNum)) %>%
+    filter(condition != "coopCartel") %>%
+  ggplot(aes(x = repNum, y = numSelectHidden, color = condition, linetype = languageCondition)) + 
+  geom_point(alpha = .1) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "N. Select Hidden")+ scale_color_bay() + theme_jmank()
 ```

### 2021-08-18 code/simulate_game.Rmd
move previous game data into data/version_0 and version preprocessing code

---
title: "simulate_game"
author: "Jessica Mankewitz"
date: "7/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup}
library(here)
source(here("code/theme_specs.R"))
library(png)
library(ggimage)
library(tidyverse)
```

```{r pressure, echo=FALSE}
d.players <- read_csv(here("data/processed_data/joined_data/players.csv"),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here("data/processed_data/joined_data/games.csv"),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv"),
                      col_types = cols()) %>% distinct()
d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv"),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here("data/processed_data/joined_data/contexts.csv"),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here("data/processed_data/joined_data/feedback.csv"),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here("data/processed_data/joined_data/demographics.csv"),
                      col_types = cols()) %>% distinct()

d.events <- read.csv(here("data/interpretable_transcript.csv"))
```

1) Figure out plotting t0 state of board (flowers available + maybe hidden markers?)
2) Figure out plotting t final state of board (selections)
3) Figure out plotting selections through events

For each round, need list of flowers available, assign coordinates?

```{r}
#available flowers are stored in d.contexts, which also has the image paths (woo!)
library(cowplot)

d.contexts %>%
  filter(gameId == "ExXQbCn35Xznf9tM5", trialNum == 0) %>%
  mutate(fullLocation = paste0(here(), "experiment/yellow_similar/image_06676.jpg")) %>%
  ggdraw() + draw_image(fullLocation)
```

#simulate utility structures?
# how often is each flower in the top 3 if 6 flowers are sampled from the 12 without replacement and the worth is assigned with a normal distribution?

```{r}
num_games <- 14400
#6 flowers for 24 trials for 100 games

utility_samples_low <- replicate(num_games,sample(1:17, size = 5, replace = FALSE, prob = 17:1))
dimnames(utility_samples_low) <- list(flower_num = 1:5, sample_num = 1:num_games)

flower_utility <- replicate(num_games,sample(18:20, size = 1, replace = FALSE))

d.utility_samples <- as.data.frame(as.table(utility_samples_low)) %>% 
  rename(flower_utility = Freq) %>% 
  mutate(flower_num = as.numeric(flower_num), 
         sample_num = as.numeric(sample_num)) %>%
  rbind(as.data.frame(flower_utility) %>% 
              mutate(flower_num = as.numeric(6), 
                     sample_num = as.numeric(1:num_games))) %>% arrange(sample_num, flower_num)

d.utility_samples$game <- rep(1:600, each=6*24)
```

```{r}
#flower samples

flower_samples <- replicate(num_games, sample(1:12, size = 6, replace = FALSE))
dimnames(flower_samples) <- list(flower_num = 1:6, sample_num = 1:num_games)

d.flower_samples <- as.data.frame(as.table(flower_samples)) %>% 
  rename(flower_name = Freq) %>% 
  mutate(flower_name = paste0("flower_", flower_name),
         flower_num = as.numeric(flower_num),
         sample_num = as.numeric(sample_num))

d.sampled_context = d.flower_samples %>% left_join(d.utility_samples) %>% rename(trial = sample_num)
```

#how often is each flower the highest?
```{r}

#top flower
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 1) %>% 
  group_by(game, flower_name) %>%
  summarize(count = n()) %>%
  group_by(flower_name) %>%
  summarize(mean = mean(count),
            median = median(count))

#in the top 3
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 3) %>% 
  group_by(game, flower_name) %>%
  summarize(count = n()) %>%
  group_by(flower_name) %>%
  summarize(mean = mean(count),
            sd = sd(count))

#estimated points (communication)
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 3) %>% 
  summarize(trial_utility = sum(flower_utility)) %>%
  group_by(game) %>%
  summarize(mean_game_points = mean(trial_utility),
            sd_game_points = sd(trial_utility))

#penalty amount, if collision get n points
collision_penalty = 10
#estimated points with collisions
d.sampled_context %>% group_by(game, trial) %>%
  #take top 2
  slice_max(order_by = flower_utility, n = 2) %>% 
  #take smaller of top 2 (participant with the highest hidden)
  slice_min(order_by = flower_utility, n = 1) %>% 
  mutate(trial_utility = sum(flower_utility+(collision_penalty*2))) %>%
  group_by(game) %>%
  summarize(mean_game_points = mean(trial_utility),
            sd_game_points = sd(trial_utility))
```

### 2021-08-19 code/aa-flowers-pilot_analysis.Rmd
add pilot data

@@
 ```{r process-data, message=FALSE, warning=FALSE}
-# data_list <- c("technical_pilots", "collabmulti_pilot0",
-#                "collabmulti_20", "compet_20", "multi_nonlang_20")
-# process_all_data(data_list)
-# join_all_data()
+data_list <- c("pilots")
+process_all_data(data_list)
+join_all_data()
 ```
@@
 # exclude 
-rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
+rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=18) %>% select(gameId,blockNum)
 
@@
 
-d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  filter(conditionName != "coopCartel") %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
+d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
 ```
@@
 ```{r}
-joined_data <- d.raw_chat.final %>% select(-c(index, createdAt, 
-                               submitted, conditionName)) %>% 
-  # merge chat condition
-  left_join(d.games %>% select(gameId, chatEnabled)) %>%
-  #merge player selections aligned with chat
-  left_join(d.rounds.final %>% 
-              select(gameId, playerId,repNum, blockNum, trialNum, 
-                     playerResponse, playerEarnings = playerUtility)) %>% 
-  #merge player names
-  left_join(d.players %>% select(playerId, playerName = name)) %>%
-  #merge the raw utility for the selections
+interpretable_data <- d.raw_chat.final %>% 
+  add_chat_info(d.players, d.rounds.final, d.games) %>% 
     left_join(d.contexts %>% select(gameId, repNum, blockNum, trialNum, 
-                                  playerResponse = label, responseRawUtility = utility, blinded)) %>%
+                                  playerResponse = label, 
+                                  responseRawUtility = utility, blinded)) %>%
    #to avoid confusion, drop any round info thats actually a message
    mutate(playerResponse = ifelse(type == "message", NA, playerResponse),
-         playerEarnings = ifelse(type == "message", NA, playerEarnings),
+         playerUtility = ifelse(type == "message", NA, playerUtility),
          responseRawUtility = ifelse(type == "message", NA, responseRawUtility),
          blinded = ifelse(type == "message", NA, blinded)) %>%
-  select(gameId, condition, chatEnabled,roundID, repNum:trialNum, playerId, 
-         playerName, text,  type, playerResponse:responseRawUtility, blinded,everything()) %>% write_csv(here("data/interpretable_transcript.csv"))
+  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, 
+         participantAction, type, repNum, 
+         blockNum, time, everything(), - submitted, -index, -createdAt) %>% write_csv(here("data/interpretable_transcript.csv"))
 ```
@@
   left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
-  mutate(gameComplete = ifelse(num_rounds == 48, TRUE, FALSE))
+  mutate(gameComplete = ifelse(num_rounds == 24, TRUE, FALSE))
 
@@
 message("Full games")
-game_completion %>% filter(num_rounds==48)
+game_completion %>% filter(num_rounds==24)
 message("Partial games")
-game_completion %>% filter(num_rounds!=48) 
+game_completion %>% filter(num_rounds!=24) 
 ```
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  ylim(0, 1400)+
   facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

### 2021-08-19 code/preprocessing/00_preprocessing_functions.R

@@ -2,6 +2,10 @@
 
 #These functions cover reading in the raw data, converting to tidy formats, filtering out unfinished rounds, and writing tidy processed csvs.
 
+#note: The most recent version of this is tweaked to support the most recent version of the data collection. 
+
+#The previous data (pre-timestamp recording, post-test quiz, and clovers) is at commit 5bbeb59490dffe6f52483b97956afa4b779aca02
+
 # Libraries
 library(tidyverse)
 library(jsonlite)
@@ -53,7 +57,8 @@ get_game_data <- function(data_read_location, date_start, players_data) {
     left_join(d.factors %>% 
                 select(factorId, factorName, value)) %>%
     pivot_wider(id_cols = c(treatmentId, name, createdAt), 
-                names_from = factorName, values_from = value)
+                names_from = factorName, values_from = value) %>%
+    mutate(numBlocks = as.numeric(numBlocks))
   
   d.games <- d.games.raw %>% 
     #add treatment variables
@@ -63,7 +68,7 @@ get_game_data <- function(data_read_location, date_start, players_data) {
     #select out the nonuseful columns
     select(-c(gameLobbyId, treatmentId, batchId, targetSet, 
               justStarted, submitCode, roundIds)) %>%
-    #separate out player info
+    #separate out player info, to reliably calculate game time
     separate(col = playerIds, sep = ",", into = c("player_1", "player_2", "player_3")) %>% 
     pivot_longer(cols = starts_with("player_"), 
                  names_to = "playerNum", 
@@ -120,23 +125,37 @@ get_raw_chat <- function(data_read_location, date_start) {
               ends_with('time'), 
               ends_with('utility'), 
               context, stageIds)) %>%
-    #mutate(chat = ifelse(is.na(chat), '{}', chat)) %>%
     rename(roundID = `_id`) %>%
     mutate(chat = map(chat, .f = ParseJSONColumn)) %>%
-    unnest(chat)
+    unnest(chat) %>%
+    mutate(time = as.numeric(time)/1000)
+  
   return(d.chat.raw)
 }
 
 add_chat_info <- function(raw_chat_data, player_data, round_data, game_data) {
-  chat.info <- raw_chat_data %>%
+  # reallign the message and selection timestamps
+  # get the chat timestamp first alert
+  min_chat_alert <- raw_chat_data %>% group_by(gameId, trialNum) %>%
+    filter(type == "selectionAlert") %>% 
+    summarize(first_chat_time = min(time))
+  
+  # get get the selection timestamp first alert
[…]

### 2021-08-19 code/aa-flowers-pilot_analysis.Rmd
add collaborative and competitive first round game data

@@
 ```{r process-data, message=FALSE, warning=FALSE}
-data_list <- c("pilots")
+data_list <- c("pilots", "collab_lang_20", "compet_lang_20")
 process_all_data(data_list)
-join_all_data()
+#join_all_data()
 ```
@@
 #pull out pilots
-d.games %>% group_by(conditionName, chatEnabled,pilot)%>% tally()
+game_completion %>% group_by(condition, chat, gameComplete)%>% tally()
 ```
@@
 ### Time
-```{r, fig.height=5, fig.width=3}
+```{r, fig.height=3, fig.width=5}
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = gameLength/60, 
+  ggplot(aes(x = condition, y = gameLength/60, 
              color = condition, shape = languageCondition)) + 
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ 
-  labs(y = "Time (In Minutes)", title = "Game Length", x = "Language Condition")+
-  ylim(0, 90)+
+  #facet_grid(cols = vars(condition))+ 
+  labs(y = "Time (In Minutes)", title = "Game Length", x = "Condition")+
   scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
-ggsave(here(fig_path, "time_distirbution.png"),  width = 3, height = 5)
+ggsave(here(fig_path, "time_distirbution.png"),  width = 5, height = 3)
 
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = gameNumWords, color = condition, shape = languageCondition)) + 
+  ggplot(aes(x = condition, y = gameNumWords, color = condition, shape = languageCondition)) + 
   geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged", x = "Language Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+  #facet_grid(cols = vars(condition))+ 
+  labs(y = "N. Words", title = "Number of Words Exchanged", x = "Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
 
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = languageCondition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
+  ggplot(aes(x = condition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
   geom_jitter(width = .1, alpha = .5) +
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "Log N. Words Exchanged")+
+  #facet_grid(cols = vars(condition))+ 
+  labs(y = "Log N. Words Exchanged")+
   scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
@@
 
-```{r}
+```{r, fig.width=5, fig.height=3}
 ms <- d.by_game_metrics %>%
   mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
-  group_by(condition, langUse) %>%
+  group_by(condition) %>%
   summarise(mean_nw = mean(gameNumWords),
@@
   mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
-  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = langUse)) +
+  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition)) +
   #geom_smooth(aes(group = condition), method = "lm") +
@@
   # ylim(0,1500) + 
-  labs(x = "Words Echanged", y = "Points Earned") + 
+  labs(x = "Words Exchanged (Game)", y = "Points Earned (Game)") + 
   scale_color_bay() + theme_jmank() +  
@@
   mutate(blockName = paste0("Block ", blockNum))%>% 
-  ggplot(aes(x = log(groupNumWords+1), y = roundPointsAdjusted, 
+  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
              color = condition)) + 

### 2021-08-19 code/preprocessing/00_preprocessing_functions.R

@@ -23,10 +23,12 @@ extract_context_info <- function(context_key){
 
 get_player_data <- function(data_read_location, date_start) {
   d.players <- read_csv(here(data_read_location, 'players.csv'), 
-                        col_types = cols(data.timeClick = col_number())) %>%
+                        col_types = cols(data.timeClick = col_number(),
+                                         data.rawScore = col_number())) %>%
     rename(playerId = `_id`) %>% 
     filter(createdAt >= date_start) %>%
-    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
+    select(-c(clicked:clickLength)) 
   return(d.players)
 }
 
@@ -182,7 +184,8 @@ get_alternative_context <- function(data_read_location, date_start) {
 }
 
 get_demographics <- function(data_read_location, date_start, games_data) {
-  read_csv(here(data_read_location, 'player-inputs.csv')) %>%
+  read_csv(here(data_read_location, 'player-inputs.csv'), 
+           col_types = cols(data.age = col_number())) %>%
     filter(createdAt >= date_start) %>%
   rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% select(gameId:education)
 }

### 2021-08-20 code/aa-flowers-pilot_analysis.Rmd
add nonlang conditions for comparison

--- 2021-08-19 code/aa-flowers-pilot_analysis.Rmd
@@
 ```{r process-data, message=FALSE, warning=FALSE}
-data_list <- c("pilots")
+data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
 process_all_data(data_list)
-join_all_data()
+#join_all_data()
 ```
@@
 #exclude BLOCKS with <12 completed (all 3 participants responded) rounds
-#exclude trial runs
-# exclude 
-rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=18) %>% select(gameId,blockNum)
+# exclude incomplete games (not just incomplete rounds)
+
+games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=3*24) %>% select(gameId)
 
@@
 d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  filter(conditionName != "coopCartel")%>% anti_join(rounds_exclude)
+  filter(conditionName != "coopCartel") %>% anti_join(games_exclude)
 
-d.raw_chat.final <- d.raw_chat %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
+d.raw_chat.final <- d.raw_chat %>% 
+  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
+  anti_join(games_exclude) %>% 
+  write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
 ```
@@
 ```{r}
+#how many games are we left with
 game_completion <- d.rounds %>% 
@@
 #pull out pilots
-d.games %>% group_by(conditionName, chatEnabled,pilot)%>% tally()
+d.rounds.final %>% left_join(d.games %>% select(gameId, chatEnabled)) %>% group_by(condition, chatEnabled) %>% 
+  summarize(n_games = n_distinct(gameId)) %>% arrange(condition, chatEnabled)
 ```
@@
 # Basic Analyses
-```{r}
-#get the # of raw points, max adjusted points, game legnth, # of words exchanged
-d.round_word_counts <- d.raw_chat.final %>% 
+
+Break out the by-player statistics for each round
+```{r player-stats}
+
+#player language per round
+d.player_word_counts <- d.raw_chat.final %>% 
   filter(type == "message") %>%
@@
 
+# player points per round
+d.by_player_metrics <- d.rounds.final  %>% 
+  left_join(d.player_word_counts) %>%
+  left_join(game_completion %>% 
+              select(gameId, gameComplete)) %>%
+  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"))
+
+```
+
+Round info
+```{r round-stats}
+#get the # of raw points, max adjusted points, game legnth, # of words exchanged
 d.max_round <- d.contexts %>%
@@
 
-d.by_round_metrics <- d.rounds.final  %>% 
-  left_join(d.round_word_counts) %>%
+d.by_round_metrics <- d.by_player_metrics  %>% 
   group_by(gameId, trialNum, repNum, blockNum) %>%
@@
   left_join(d.max_round) %>%
-  left_join(game_completion %>% select(gameId, gameComplete)) %>%
-  left_join(d.games %>% select(-createdAt)) %>%  
+  left_join(game_completion %>% 
+              select(gameId, gameComplete)) %>%
+  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
   mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
@@
          roundPointsAdjusted = groupPoints/roundMax)
+```
 
+by block
+```{r}
 d.by_block_metrics <- d.by_round_metrics %>% 
-  group_by(gameId, name, gameComplete, condition, 
+  group_by(gameId, gameComplete, condition, 
            languageCondition, fullCondition, gameLength, blockNum) %>%
@@
   mutate(blockPointsAdjusted = blockPoints/blockMaxPoints,
-         meanNumFlowers = blockNumFlowers*3/(12*3))
+         meanNumFlowers = blockNumFlowers*3/(6*3))
+```
 
+by game
+```{r by-game}
 d.by_game_metrics <- d.by_block_metrics %>%
@@
   #group by game
-  group_by(gameId, name, condition, languageCondition, fullCondition, gameLength) %>%
+  group_by(gameId, condition, languageCondition, fullCondition, gameLength) %>%
   summarize(gamePoints = sum(blockPoints, na.rm = T),
@@
   mutate(pointsAdjusted = gamePoints/gameMaxPoints,
-         meanNumFlowers = gameNumFlowers*3/ (12*4*3))
+         meanNumFlowers = gameNumFlowers*3/ (6*4*3))
 ```
 
+```{r}
+d.by_player_metrics %>% group_by(gameId, condition, languageCondition, playerId) %>% 
+  summarize(mean_player_utility = mean(playerUtility)) %>%
+  group_by(condition, languageCondition) %>% 
+  summarize_at("mean_player_utility", funs(mean, min, max, sd))
+```
 ## Simple Distributions
@@
 ```{r, fig.height=3, fig.width=5}
-d.by_game_metrics%>% 
+d.by_game_metrics %>% 
   ggplot(aes(x = languageCondition, y = gamePoints, color = condition, shape = languageCondition)) + 
@@
 ### Time
-```{r, fig.height=5, fig.width=3}
+```{r, fig.height=3, fig.width=5}
 d.by_game_metrics %>%
@@
   facet_grid(cols = vars(condition))+ 
-  labs(y = "Time (In Minutes)", title = "Game Length", x = "Language Condition")+
-  ylim(0, 90)+
+  labs(y = "Time (In Minutes)", title = "Game Length", x = "Condition")+
   scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
-ggsave(here(fig_path, "time_distirbution.png"),  width = 3, height = 5)
+ggsave(here(fig_path, "time_distirbution.png"),  width = 5, height = 3)
 
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "N. Words", title = "Number of Words Exchanged", x = "Language Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
+  facet_grid(cols = vars(condition))+ 
+  labs(y = "N. Words", title = "Number of Words Exchanged", x = "Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
 
@@
   stat_summary(fun.data = "mean_cl_boot")+
-  facet_grid(cols = vars(condition))+ labs(y = "Log N. Words Exchanged")+
+  facet_grid(cols = vars(condition))+ 
+  labs(y = "Log N. Words Exchanged")+
   scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
@@
 d.by_game_metrics %>%
-  ggplot(aes(x = gameNumWords, y = gameLength/60, color = condition, shape = languageCondition)) + 
+  ggplot(aes(y = gamePoints, x = gameLength/60, color = condition, shape = languageCondition)) + 
   geom_point(alpha = .5) +
@@
 ```{r}
-d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
-  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>%
-  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition)) + 
-  geom_point(alpha = .5) +
-  geom_smooth(method=lm)+
-  labs(x = "N. Words Exchanged", y = "Group Points (Raw)")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
-
 ms <- d.by_game_metrics %>%
@@
   labs(x = "Words Exchanged", y = "Points Earned", title = ) + 
-  scale_color_bay() + theme_jmank() +  
-  theme(legend.position = "bottom")
-```
-
-```{r}
-ms <- d.by_game_metrics %>%
-  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
-  group_by(condition, langUse) %>%
-  summarise(mean_nw = mean(gameNumWords),
-            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
-            mean_points = mean(gamePoints),
-            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))
-
-d.by_game_metrics %>%
-  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
-  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = langUse)) +
-  #geom_smooth(aes(group = condition), method = "lm") +
-  geom_point(alpha = .5) +
-  geom_pointrange(data = ms, aes(x = mean_nw, y = mean_points, 
-                                 ymin = mean_points - ci_points,
-                                 ymax = mean_points + ci_points)) +
-  geom_errorbarh(data = ms, aes(x = mean_nw, y = mean_points, 
-                                 xmin = mean_nw - ci_nw,
-                                 xmax = mean_nw + ci_nw), width = 0) + 
-  # ylim(0,1500) + 
-  labs(x = "Words Echanged", y = "Points Earned") + 
   scale_color_bay() + theme_jmank() +  
@@
 
-Including language optional games where they decided not to speak
+Including language _optional_ games where they decided not to speak
 
 ```{r, fig.height=3, fig.width=5}
-d.round_word_counts %>%left_join(d.games %>% 
+d.player_word_counts %>%
+  left_join(d.games %>% 
               select(gameId, condition,chatEnabled)) %>% 
@@
     filter(languageCondition == "Lang") %>%
-  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
+  group_by(gameId, fullCondition, block_name, trialNum, repNum) %>% 
   summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
@@
 ```{r, fig.height=3, fig.width=5}
-d.round_word_counts %>%left_join(d.games %>% 
+d.player_word_counts %>%left_join(d.games %>% 
                                    select(gameId, condition,chatEnabled)) %>% 
@@
 
-Only games where language was used
+Only games where language was _used_
 
 ```{r, fig.height=3, fig.width=5}
-d.round_word_counts %>%left_join(d.games %>% 
+d.player_word_counts %>%left_join(d.games %>% 
                                    select(gameId, condition,chatEnabled)) %>% 
@@
 ```
-
+log num words
 ```{r, fig.height=3, fig.width=5}
-d.round_word_counts %>%left_join(d.games %>% 
+d.player_word_counts %>%left_join(d.games %>% 
                                    select(gameId, condition,chatEnabled)) %>% 
@@
   geom_smooth(method=lm)+
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
+  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log N. Words")+ scale_color_bay() + theme_jmank()
 ```
 
-log num words
-
-```{r, fig.height=3, fig.width=5}
-d.round_word_counts %>%left_join(d.games %>% 
-                                   select(gameId, condition, chatEnabled)) %>% 
-  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
-         fullCondition = paste(condition, languageCondition, sep = "-"),
-         block_name = paste0("Block ", blockNum)) %>%
-    filter(languageCondition == "Lang") %>%
-  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>%
-
-  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
-  #filter(groupNumWords >1) %>%
-  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
-  geom_point(alpha = .2) +
-  geom_smooth(method=lm)+
-  ylim(0, 6) +
-  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
-```
 ### Accuracy
@@
 ```{r}
-d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
-  summarize(num_unique_flowers = n_distinct(playerResponse),
-            mean_response_time = mean(time_sec, na.rm = T))  %>%
-  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
-  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
-         fullCondition = paste(condition, languageCondition, sep = "-"),
-         block_name = paste0("Block ", blockNum)) %>%
-    filter(condition != "coopCartel") %>%
-  ggplot(aes(x = repNum, y = num_unique_flowers, color = condition, linetype = languageCondition)) + 
+d.by_round_metrics %>% 
+  mutate(block_name = paste0("Block ", blockNum)) %>%
+  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
   geom_point(alpha = .1) +
@@
 
-ggsave(ggsave(here(fig_path, "collisions.png")))
+ggsave(ggsave(here(fig_path, "collisions_conditionfacet.png")))
+```
+```{r}
+d.by_round_metrics %>% 
+  mutate(block_name = paste0("Block ", blockNum)) %>%
+  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
+  geom_point(alpha = .1) +
+  geom_smooth(method=lm)+
+  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()
+
+ggsave(ggsave(here(fig_path, "collisions_language_facet.png")))
 ```
 
+Response Time
 ```{r}
 d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
-  summarize(num_unique_flowers = n_distinct(playerResponse),
-            mean_response_time = mean(time_sec, na.rm = T))  %>%
+  summarize(mean_response_time = mean(time_sec, na.rm = T))  %>%
   left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
@@
 ```{r}
-d.by_round_metrics %>% left_join(d.games %>% 
-                                   select(gameId, condition, chatEnabled))%>%
+d.by_round_metrics %>%
   mutate(blockName = paste0("Block ", blockNum))%>% 
-  ggplot(aes(x = log(groupNumWords+1), y = roundPointsAdjusted, 
+  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
              color = condition)) + 
@@
   labs(title = "Number of points per trial")
+```
+### Over/underpowered participants
+
+```{r}
+d.point_propotions <- d.by_player_metrics %>% 
+  #points for entire game
+  group_by(condition, languageCondition, gameId, playerId) %>% 
+  summarize(player_total = sum(playerUtility)) %>% 
+  group_by(condition, gameId) %>% mutate(game_total = sum(player_total)) %>% ungroup() %>%
+  mutate(player_proportion = player_total/game_total)
+
+#min
+min_prop <- d.point_propotions %>% 
+  group_by(condition, languageCondition, gameId) %>% 
+  #lowest earned
+  slice_min(order_by = player_proportion, n = 1) %>% 
+  filter(condition == "Competitive") %>% select(-condition) %>% 
+  mutate(player_placement = "Lowest Earning")
+
+#max
+max_prop <- d.point_propotions %>% 
+  group_by(condition, languageCondition, gameId) %>% 
+  slice_max(order_by = player_proportion, n = 1) %>% 
+  filter(condition == "Competitive") %>% 
+  select(-condition) %>% 
+  mutate(player_placement = "Highest Earning")
+
+rbind(min_prop, max_prop) %>%
+  ggplot(aes(x = languageCondition, y = player_proportion, color = condition)) + 
+  geom_hline(yintercept=.33, linetype = "dashed", alpha = .5)+
+  geom_jitter(width = .1, alpha = .5) +
+  facet_grid(cols = vars(player_placement)) + 
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(x = "Participant Placement", y = "Points earned by participant in place") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")
+
+```
+### Entropy
+
+```{r}
+entropy_func <- function(p1, p2, p3){
+  return((p1 * log(p1)) + (p2 * log(p2)) + (p3 * log(p3)))
+  #return((entropy_list[1] * log(entropy_list[1])) + (entropy_list[2] * log(entropy_list[2])) + (entropy_list[3] * log(entropy_list[3])))
+}
+d.game_entropies <- d.point_propotions %>% select(condition, languageCondition, gameId, player_proportion) %>% 
+  mutate(p_num = rep(c("p1","p2","p3"), nrow(d.point_propotions)/3)) %>% 
+  pivot_wider(names_from = p_num, values_from = player_proportion) %>% 
+  mutate(entropy = 0-entropy_func(p1, p2, p3)) %>% select(-c(p1,p2,p3))
+
+d.game_entropies %>% filter(condition == "Competitive") %>% 
+  ggplot(aes(x = languageCondition, y = entropy)) + 
+  geom_hline(yintercept=(0-entropy_func(1/3, 1/3, 1/3)), linetype = "dashed", alpha = .5)+
+  #minimum is tricky? this is 12, 1, and 2 which is the least fair a game _could possible_ be, but isnt adjusted per game
+  #geom_hline(yintercept=(0-entropy_func(12/15, 1/15, 2/15)), linetype = "dashed", alpha = .5)+
+  geom_jitter(width = .1, alpha = .5) +
+  stat_summary(fun.data = "mean_cl_boot")+
+  labs(x = "Condition", y = "Entropy of Player Scores (by game)") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")
 ```

### 2021-08-24 code/aa-flowers-pilot_analysis.Rmd
update code bits

@@
 
+What is the spread of participant earnings over entire games, is there a diff here?
+
 ```{r}
@@
 ```
+What is the spread of participant earnings in ind. trials (average earning per participant per trial)
+```{r}
+d.by_player_metrics %>% group_by(condition, languageCondition) %>% 
+  summarize_at("playerUtility", funs(mean, min, max, sd))
+```
+
 ## Simple Distributions

### 2021-08-31 code/aa-flowers-pilot_analysis.Rmd
add player points fig

@@
                                          condition == "competCartel" ~ "Competitive",
-                                         condition == "coopCartel" ~ "Old Cooperative"))
+                                         condition == "coopCartel" ~ "Old Cooperative",
+                                         TRUE ~ condition))
 d.games
@@
 
+## Deference
+How often does a player select a "lower ranked" target?
+* Get the rankings of the flowers _visible_ to each participant
+** Need a set of rows with each flower visible to each participant and the flower utilities, arrange by utility (grouped by trial and participant ID), assign 1-4 ranking, drop flowers that werent selected
+
+NB: What about cases where a participant selected a flower hidden from them? If we drop both the flowers that are hidden and the flowers that arent selected, we wont have a selection for that trial
+
+[] fix roundId
+
+```{r deference}
+trial_deference <- d.by_player_metrics %>% 
+  rename(roundID = "_id") %>% 
+  full_join(d.contexts %>% 
+              select(roundID, blinded, label, utility)) %>%
+  filter(blinded != playerId) %>%
+  group_by(playerId, roundID) %>%
+  summarize(deference = utility[label == playerResponse] != max(utility))
+
+d.deference <- d.by_player_metrics %>% 
+  rename(roundID = "_id") %>% 
+  left_join(trial_deference)
+```
+
+```{r}
+d.deference %>% filter(!is.na(deference)) %>%
+  ggplot(aes(x = trialNum, y = as.numeric(deference), color = condition)) + 
+  geom_smooth(method = "lm") + 
+  geom_jitter(width = .1, height = .1, alpha = .05) +
+  facet_grid(.~languageCondition)+ scale_color_bay() + theme_jmank()
+```
+
+## Post-Test
+```{r post-test}
+
+```
+
 ## Simple Distributions
-### Points
+### Individual Points
+```{r, fig.height=3, fig.width=5}
+d.by_player_metrics %>%
+  group_by(languageCondition, condition, playerId) %>%
+  summarize(playerPoints = sum(playerUtility)) %>%
+  ggplot(aes(x = languageCondition, y = playerPoints, color = condition)) + 
+  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")
+```
+
+### Game Points
 
@@
 ```
+
+#Points per individual (per trial)
 

### 2021-09-06 code/aa-flowers-pilot_analysis.Rmd
add prelim data tweaks and ms analyses

@@
 ---
+This notebook runs the preprocessing scripts, saves the data, and generates many summary analyses and plots
 
@@
 library(showtext)
-font_add_google(name = "Assistant", family = "Assistant")
-font_add_google(name = "Roboto", family = "Roboto")
+#font_add_google(name = "Assistant", family = "Assistant")
+#font_add_google(name = "Roboto", family = "Roboto")
 showtext_auto()
@@
 data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
-process_all_data(data_list)
+#process_all_data(data_list)
 #join_all_data()
@@
 d.games <- d.games %>% mutate(conditionName = condition, 
-                              condition = case_when(condition == "coopMulti" ~ "Cooperative",
-                                         condition == "competCartel" ~ "Competitive",
+                              condition = case_when(condition == "coopMulti" ~ "Shared Utilities",
+                                         condition == "competCartel" ~ "Individual Utilities",
                                          condition == "coopCartel" ~ "Old Cooperative",
@@
 #pull out pilots
-d.rounds.final %>% left_join(d.games %>% select(gameId, chatEnabled)) %>% group_by(condition, chatEnabled) %>% 
-  summarize(n_games = n_distinct(gameId)) %>% arrange(condition, chatEnabled)
+game_completion %>% group_by(condition, chat, gameComplete) %>% 
+  summarize(n_games = n_distinct(gameId)) %>% arrange(-gameComplete, -chat, condition)
 ```
@@
 ```{r demographics}
-d.demographics
+d.demographics_clean <- d.demographics %>% filter(gameId %in% game_completion$gameId) %>%
+  mutate(gender_code = case_when(gender %in% c("Female", "female", "f", "Woman","Femaile", "FEMALE", "Trans female", "Femal", "woman") ~ "Female",
+                                gender %in% c("Male", "male", "MALE") ~ "Male",
+                                is.na(gender) ~ "Unspecified",
+                                gender %in% c("Non-Binary", "non-binary", "Non-binary")~ "Non-Binary", gender %in% c("agender") ~ "Agender", TRUE ~ "Other"))
+
+d.demographics_clean %>% count(gender_code)
+
+nrow(d.demographics_clean)
 ```
@@
 
+## Word Counts
+```{r}
+d.raw_chat.final %>% filter(!is.na(text))
+
+sum(d.by_game_metrics$gameNumWords)
+```
+
 ## Deference
@@
 ## Simple Distributions
-### Individual Points
+#Points per individual (per trial) 
 ```{r, fig.height=3, fig.width=5}
-d.by_player_metrics %>%
+#average points earned by participants per round
+player_points_earned <- d.by_player_metrics %>%
   group_by(languageCondition, condition, playerId) %>%
-  summarize(playerPoints = sum(playerUtility)) %>%
+  summarize(playerPoints = mean(playerUtility)) %>%
   ggplot(aes(x = languageCondition, y = playerPoints, color = condition)) + 
   geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
-  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
+  scale_color_bay() + theme_jmank() + 
+  labs(y = "Average Player Points", x = "", color = "Utility Condition") +theme(legend.position = "bottom")
+
+player_points_earned
 ```
@@
 
-#Points per individual (per trial)
-
 Max Adjusted
@@
        title = "Points by Group") + 
-  ylim(0,1) +
+  #ylim(0,1) +
   scale_color_bay() + theme_jmank() + theme(legend.position = "none")
@@
 ```
+
 ```{r, fig.height=3, fig.width=5}
@@
 ```{r, fig.height=3, fig.width=5}
-d.by_game_metrics %>%
+all_correlated <- d.by_game_metrics %>%
   ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
   geom_point(alpha = .8) +
-  geom_smooth(aes(group = condition), method=lm, color = "black")+
+  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
   facet_grid(cols = vars(condition)) + 
@@
        y = "Points Earned by Group",
-       color = "Log Num. Words",
+       color = "Log N. Words",
        shape = "Language") + 
@@
 
+all_correlated
+
 ggsave(here(fig_path, "all_correlated.png"), height = 3, width = 5)
+```
+
+# MS Figs
+
+```{r}
+
+library(cowplot)
+all_correlated
+player_points_earned
+
+plot_grid(player_points_earned,all_correlated , labels = c('A', 'B'), label_size = 12)
+
+ggsave(here(fig_path, "combined_fig.png"), height = 3, width = 10, rel_)
 ```

### 2021-09-06 code/aa_mic_submission_analyses.Rmd
add prelim data tweaks and ms analyses

---
title: "aa_mic_submission_analyses"
author: "Jessica Mankewitz"
date: "9/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook reads in the preprocessed and joined data, runs the basic processing/summarizing code, and generates figures and numbers for the 2021 MIC workshop submission.

```{r setup}
library(here)
library(tidyverse)

#for plotting
source(here("code/theme_specs.R"))

#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
```
# Summary Processing
## Read data

```{r read-data}
d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here(paste(read_data_path, 
                                "rounds.csv", 
                                sep="")),
                      col_types = cols()) %>% distinct()
d.raw_chat <- read_csv(here(paste(read_data_path, 
                                  "raw_chat.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here(paste(read_data_path, 
                                  "contexts.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here(paste(read_data_path, 
                                  "feedback.csv", 
                                  sep = "")),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here(paste(read_data_path,
                                      "demographics.csv", 
                                      sep="")),
                      col_types = cols()) %>% distinct()
```

## Rename Conditions
Rename conditions for readability/alignment with ms
```{r}
d.games <- d.games %>% 
  mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName))
```

## Identify complete games

For the purposes of this ms, we're removing all unfinished games from the analyses 
```{r}
n_players = 3
n_rounds = 24
games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=n_players*n_rounds) %>% select(gameId)

d.rounds.final <-  d.rounds %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  filter(conditionName != "coopCartel") %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId))

d.raw_chat.final <- d.raw_chat %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId)) 
```

# Summary Stats
## Game counts
```{r}
d.games %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId)) %>%
  group_by(-game_complete, condition, chatEnabled) %>%
  count()
```

## Game lengths
## Bonus Payments

# Points x Time x Num. Words

# Avg. Points Per Round

# Language Reduction

# Georgia Analyses

### 2021-09-06 code/linguistic_analysis_C-D.Rmd
add nonlang conditions for comparison
add prelim data tweaks and ms analyses

@@
 #dev.off() 
-ggsave("../figs/mean_utt_length_block.png", height = 3, width = 3)
+#ggsave("../figs/mean_utt_length_block.png", height = 3, width = 3)
 
@@
 ### 3. Parts of speech
-```{r pos}
+```{r pos, eval=FALSE, include=FALSE}
 
-text <- d_flowers$text
-parsed <- spacy_parse(text,pos=TRUE) %>%
-  select(doc_id, pos) %>%
-  group_by(doc_id) %>%
-  mutate(pos = paste(pos, collapse=",")) %>%
-  distinct()
-
-d_flowers_pos  <- cbind(d_flowers, parsed) %>%
-  separate_rows(pos, convert = TRUE) %>%
-  group_by(condition, blockNum, gameId, pos) %>%
-  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
-  summarise(pos_count=n()) %>%
-  left_join(d_flowers_utt_length_block) %>%
-  mutate(prop_pos=pos_count/total_num_words)
-
-#jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
-ggplot(d_flowers_pos, 
-       aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
-  facet_grid( cols=vars(pos)) +
-  xlab("Block number") + 
-  ylab("Proportion of pos") + 
-  theme_bw()#+
- # theme(text = element_text(size = 30))
-#dev.off()
-
-d_flowers_pos  <- cbind(d_flowers, parsed) %>%
-  separate_rows(pos, convert = TRUE) %>%
-  group_by(condition, trialNum, gameId, pos) %>%
-  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
-  summarise(pos_count=n()) %>%
-  left_join(d_flowers_utt_length_trial) %>%
-  mutate(prop_pos=pos_count/total_num_words)
-
-#jpeg("~/Desktop/plots/pos_trial.jpg", width = 950, height = 950)
-ggplot(d_flowers_pos, 
-       aes(x=trialNum, y=prop_pos, label=gameId,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
-  facet_grid( cols=vars(pos)) +
-  xlab("Trial number") + 
-  ylab("Proportion of pos") + 
-  ylim(0,5)+
-  theme_bw()#+
-  #theme(text = element_text(size = 30))
-#dev.off()
-
-d_flowers_pos  <- cbind(d_flowers, parsed) %>%
-  separate_rows(pos, convert = TRUE) %>%
-  group_by(condition, repNum, gameId, pos) %>%
-  filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
-  summarise(pos_count=n()) %>%
-  left_join(d_flowers_utt_length_rep) %>%
-  mutate(prop_pos=pos_count/total_num_words)
-
-#jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_pos, 
-       aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
-  facet_grid(cols=vars(pos)) +
-  xlab("Rep number") + 
-  ylab("Proportion of pos") +
-  ylim(0,5)+
-  theme_bw()#+
- # theme(text = element_text(size = 30))
-#dev.off() 
+# text <- d_flowers$text
+# parsed <- spacy_parse(text,pos=TRUE) %>%
+#   select(doc_id, pos) %>%
+#   group_by(doc_id) %>%
+#   mutate(pos = paste(pos, collapse=",")) %>%
+#   distinct()
+#   
+# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+#   separate_rows(pos, convert = TRUE) %>%
+#   group_by(condition, blockNum, gameId, pos) %>%
+#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+#   summarise(pos_count=n()) %>%
+#   left_join(d_flowers_utt_length_block) %>%
+#   mutate(prop_pos=pos_count/total_num_words)
+#         
+# #jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
+# ggplot(d_flowers_pos, 
+#        aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
+#   geom_smooth(method = "lm", formula = y~x) + 
+#   geom_point(alpha=.1) +
+#   facet_grid( cols=vars(pos)) +
+#   xlab("Block number") + 
+#   ylab("Proportion of pos") + 
+#   theme_bw()#+
+#  # theme(text = element_text(size = 30))
+# #dev.off() 
+# 
+# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+#   separate_rows(pos, convert = TRUE) %>%
+#   group_by(condition, trialNum, gameId, pos) %>%
+#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+#   summarise(pos_count=n()) %>%
+#   left_join(d_flowers_utt_length_trial) %>%
+#   mutate(prop_pos=pos_count/total_num_words)
+#         
+# #jpeg("~/Desktop/plots/pos_trial.jpg", width = 950, height = 950)
+# ggplot(d_flowers_pos, 
+#        aes(x=trialNum, y=prop_pos, label=gameId,  color=condition)) + 
+#   geom_smooth(method = "lm", formula = y~x) + 
+#   geom_point(alpha=.1) +
+#   facet_grid( cols=vars(pos)) +
+#   xlab("Trial number") + 
+#   ylab("Proportion of pos") + 
+#   ylim(0,5)+
+#   theme_bw()#+
+#   #theme(text = element_text(size = 30))
+# #dev.off() 
+# 
+# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+#   separate_rows(pos, convert = TRUE) %>%
+#   group_by(condition, repNum, gameId, pos) %>%
+#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+#   summarise(pos_count=n()) %>%
+#   left_join(d_flowers_utt_length_rep) %>%
+#   mutate(prop_pos=pos_count/total_num_words)
+#         
+# #jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
+# ggplot(d_flowers_pos, 
+#        aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
+#   geom_smooth(method = "lm", formula = y~x) + 
+#   geom_point(alpha=.1) +
+#   facet_grid(cols=vars(pos)) +
+#   xlab("Rep number") + 
+#   ylab("Proportion of pos") +
+#   ylim(0,5)+
+#   theme_bw()#+
+#  # theme(text = element_text(size = 30))
+# #dev.off() 
 ```
@@
 ```
-
-```{r minimum_edit_distance, include=TRUE}
-
-```

### 2021-09-06 code/preprocessing/rl_preprocessing.R

@@ -1,0 +1,25 @@
+# One time script for preprocessing the RL agent model results
+
+library(tidyverse)
+library(here)
+library(stringr)
+
+d.rl_data.raw <- read_csv(here("data/raw_data/RL_experiment_results_raw.csv"), skip=1)
+
+d.rl_data <- d.rl_data.raw %>% 
+  select(everything(), -contains("overall")) %>%
+  pivot_longer(cols = -any_of("Name"), names_to = "measure") %>%
+  #replace "overall" with something with an understore so I can string split
+  #string split everything into a list? 
+  mutate(agent = substr(measure, str_length("train_")+1, str_length("train_agent_1")),
+         measure = substr(measure, 
+                              str_length("train_agent_1")+2, 
+                              str_length(measure))) %>%
+  separate(Name, into = c("condition", "languageCondition", "trial"), sep = "_") %>%
+  pivot_wider(names_from = "measure", values_from = "value") %>%
+  mutate(condition_raw = condition, 
+         condition = ifelse(condition_raw == "joint", "Shared Utilities", "Individual Utilities"),
+         languageConditionRaw = languageCondition,
+         languageCondition =ifelse(languageConditionRaw == "contcomm", "Lang", "Nonlang"))
+  write_csv(here("data/processed_data/RL_experiment_results.csv"))
+

### 2021-09-06 code/theme_specs.R

@@ -36,10 +36,10 @@ scale_fill_bay <- function(...) {
   ggplot2::discrete_scale("fill", "bay", bay_palette, ...)
 }
 
-theme_jmank <- function (base_size = 11, base_family = "Roboto", base_line_size = base_size/22, 
+theme_jmank <- function (base_size = 11, base_line_size = base_size/22, 
                          base_rect_size = base_size/22) 
 {
-  theme_classic(base_size = base_size, base_family = base_family, 
+  theme_classic(base_size = base_size, 
                 base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace% 
     theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", 

### 2021-09-08 code/fit_alignment.R
update fit_alignment for new utilities data

@@ -27,10 +27,10 @@ result_files <- "data/processed_data/joined_data/output_alignment.csv"
 #add condition to alignment data
 o<- read_csv("data/processed_data/joined_data/filtered_raw_chat.csv")
 coop_players <- o %>% 
-  filter(condition == "coopMulti") 
+  filter(condition == "Shared Utilities") 
 list_coop_players <- unique(coop_players$playerId)
 comp_players <- o %>% 
-  filter(condition == "competCartel") 
+  filter(condition == "Individual Utilities") 
 list_comp_players <- unique(comp_players$playerId)
 
 
@@ -46,9 +46,10 @@ setup_alignment_data <- function(corpus_df) {
 
 childes_data <- lapply(result_files, read_csv) %>%
   bind_rows() %>%
+  distinct() %>%
   rename(Speaker = speakerId, Replier = replierId) %>%
  # mutate(subpop = paste(Speaker,  "-", Replier)) %>%
-  mutate(subpop= ifelse(Replier %in% list_coop_players, "coop", ifelse(Replier %in% list_comp_players, "comp", ifelse(Speaker %in% list_comp_players, "comp",ifelse(Speaker %in% list_coop_players, "coop","no"))))) %>% 
+  mutate(subpop= ifelse(Replier %in% list_coop_players, "Shared Utilities", ifelse(Replier %in% list_comp_players, "Individual Utilities", ifelse(Speaker %in% list_comp_players, "Individual Utilities",ifelse(Speaker %in% list_coop_players, "Shared Utilities","no"))))) %>% 
  # filter(Speaker != Replier) %>%
   filter(subpop !="no") 
 
@@ -111,11 +112,34 @@ d3$model_mu <- log(colMeans(rstan:::extract(fit,"mu_ab")$mu_ab)) -
 d3$model_dnm <- colMeans(rstan:::extract(fit,"mu_ab")$mu_ab) - 
   colMeans(rstan:::extract(fit,"mu_notab")$mu_notab)
 
+write.csv(d3, "data/processed_data/joined_data/output_stan_model.csv")
 
 d4 <- d3 %>%
   group_by(subpop) %>%
   multi_boot_standard("model_eta", na.rm = F)
 
+
+options(ggrepel.max.overlaps = 20)
+ggplot(aes(y = model_eta, x = subpop, colour = subpop,
+), data=d3) + 
+  #geom_line(aes(group=category))+
+  geom_point(data=d3)+
[…]

### 2021-09-08 code/linguistic_analysis_C-D.Rmd
update rmd with figs and new format utilities data

@@
 library(ggrepel)
+library(cowplot)
+library(grid)
+library(gridExtra)
 
@@
   filter(type=="message") %>%
-  filter(!is.na(text)) %>% 
-  filter(!conditionName=="coopCartel")
-
-write_csv(d, "../data/processed_data/raw_chat3.csv")
-
-d_flowers <- d_flowers %>% 
-  filter(!conditionName=="coopCartel")
+  filter(!is.na(text))
 
@@
 
-### 1. Utt Length
+### Utt Length (MLU)
 ```{r utt_length}
@@
 
-#, color=as.factor(gameId)
-
-#jpeg("~/Desktop/plots/nwords_block.jpg", width = 950, height = 950)
 ggplot(d_flowers_utt_length_block, aes(x=blockNum, y=total_num_words, color=condition))+
@@
 
-#jpeg("~/Desktop/plots/nwords_trial.jpg", width = 950, height = 950)
 ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=total_num_words,  color=condition))+
@@
   theme_jmank() + scale_color_bay() +
-  ggtitle("Number of words per trial")#+
-#  theme(text = element_text(size = 30))
-#dev.off()
+  ggtitle("Number of words per trial")
 
-#jpeg("~/Desktop/plots/mlu_trial.jpg", width = 950, height = 950)
 ggplot(d_flowers_utt_length_trial, aes(x=trialNum, y=mlu,  color=condition))+
@@
   #theme(text = element_text(size = 30))
-#dev.off()
 
@@
 #jpeg("~/Desktop/plots/nword_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=total_num_words,  color=condition))+
-#  facet_grid(rows =vars(condition)) +
-  geom_point()+
-  geom_jitter(alpha=.05)+
-  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  stat_summary(fun.data = "mean_cl_boot")+
-  labs(title="Number of words", y="Total number of words", x="Rep number")+
-  theme(legend.position="bottom") + 
-  theme_jmank() + scale_color_bay() +
-  ggtitle("Number of words per rep")#+
-#  theme(text = element_text(size = 30))
-#dev.off()
-
-#jpeg("~/Desktop/plots/mlu_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=mlu,  color=condition))+
+ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=log(mlu+1),  color=condition))+
 #  facet_grid( rows =vars(condition)) +
@@
 
-ggplot(d_flowers_utt_length_rep, aes(x=repNum, y=log(mlu+1),  color=condition))+
-#  facet_grid( rows =vars(condition)) +
-  geom_jitter(alpha=.05)+
-  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  stat_summary(fun.data = "mean_cl_boot")+
- # scale_y_continuous(limits = c(0,15))+
-  labs(title="MLU", y="MLU", x="Rep number")+
-  theme(legend.position="bottom") + 
-  theme_jmank() + scale_color_bay() +
-  ggtitle("MLU per rep")#+
- # theme(text = element_text(size = 30))
-#dev.off() 
 ```
 
-### 2. Number of turns
+### Number of turns
 ```{r number_of_turns}
-
-d_flowers_n_turns_rep <- d_flowers %>%
-  group_by(gameId, condition, repNum)  %>%
-  summarise(n=n())
-
-#jpeg("~/Desktop/plots/nturn_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_turns_rep, aes(x=repNum, y=n,  color=condition))+
- # facet_grid(rows =vars(condition)) +
-  geom_jitter(alpha=.05)+
-  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
-  stat_summary(fun.data = "mean_cl_boot")+
-  labs(title="Number of turns", y="Number of turns", x="Rep number")+
-  theme(legend.position="bottom") + 
-  theme_bw() +
-  ggtitle("Number of turns per Rep")#+
-#  theme(text = element_text(size = 30))
-#dev.off()
 
@@
 
-ggsave("../figs/num_turns_block.png", height = 3, width = 3)
+#ggsave("../figs/num_turns_block.png", height = 3, width = 3)
 ```
-
-#second plot
-### 2b. Number of turns
 
@@
 
+```
+
+```{r number_of_turns3}
+
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by( repNum, gameId, condition) %>%
+  group_by(trialNum, blockNum, gameId, condition) %>%
   summarise(n_utt_speaker=n())
 
-#jpeg("~/Desktop/plots/nutts_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_utt_speaker, 
-       aes(x=repNum, y=n_utt_speaker,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
-#  facet_grid(~condition) +
-  xlab("Rep number") + 
-  ylab("Number of utterances") + 
-  theme_bw()#+
-#  theme(text = element_text(size = 30))
-#dev.off()
+turnb_function <- function(d_flowers_n_utt_speaker, c, q) {
+  p<-ggplot2::ggplot(d_flowers_n_utt_speaker, aes(y=n_utt_speaker, x=trialNum, label=gameId, color=condition)) + xlim(c,q) +
+  geom_point(alpha = .5)+
+  geom_smooth(method = "lm", formula = y~x)+
+  xlab("Trial number") + 
+  theme_bw()+
+  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+return(p)
+}
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+a<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==0), 0,5) + ylab("Number of utterances")+  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank())+ labs(subtitle = "Block1")
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+b<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==1), 6,11) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block2")
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+c<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==2), 12,17) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block3")
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+d<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==3), 18,23) +theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block4")
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<- plot_grid(a,b,c,d, legend,  nrow = 1, ncol=5, align = "v")
+
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Number of utterances", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
 
@@
 
-### 3. Parts of speech
-```{r pos, eval=FALSE, include=FALSE}
+### Parts of speech
+```{r pos, eval=TRUE, include=TRUE}
 
-# text <- d_flowers$text
-# parsed <- spacy_parse(text,pos=TRUE) %>%
-#   select(doc_id, pos) %>%
-#   group_by(doc_id) %>%
-#   mutate(pos = paste(pos, collapse=",")) %>%
-#   distinct()
+d_flowers_uttlength_nwords <- d_flowers %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%
+  mutate(utt_length_words = sapply(strsplit(text, " "), length)) %>% 
+  group_by(gameId, blockNum, trialNum, condition)  %>%
+  summarise(total_num_words = sum(utt_length_words),
+            mlu = mean(utt_length_words))
+
+ text <- d_flowers$text
+ parsed <- spacy_parse(text,pos=TRUE) %>%
+   select(doc_id, pos) %>%
+   group_by(doc_id) %>%
+   mutate(pos = paste(pos, collapse=",")) %>%
+   distinct()
 #   
@@
 #   group_by(condition, blockNum, gameId, pos) %>%
-#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+ #  filter(pos=="NOUN"| pos=="VERB") %>% #| pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
   # summarise(pos_count=n()) %>%
-#   left_join(d_flowers_utt_length_block) %>%
+  # left_join(d_flowers_uttlength_block) %>%
   # mutate(prop_pos=pos_count/total_num_words)
-#         
+
 # #jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
-# ggplot(d_flowers_pos, 
+ #ggplot(d_flowers_pos, 
 #        aes(x=blockNum, y=prop_pos, label=gameId,  color=condition)) + 
@@
 # #dev.off() 
-# 
+## 
 # d_flowers_pos  <- cbind(d_flowers, parsed) %>%
@@
 #   group_by(condition, trialNum, gameId, pos) %>%
-#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
+#   filter(pos=="NOUN"| pos=="VERB" ) %>% #| pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
 #   summarise(pos_count=n()) %>%
@@
 #   theme_bw()#+
-#   #theme(text = element_text(size = 30))
+   #theme(text = element_text(size = 30))
 # #dev.off() 
 #
-# d_flowers_pos  <- cbind(d_flowers, parsed) %>%
-#   separate_rows(pos, convert = TRUE) %>%
-#   group_by(condition, repNum, gameId, pos) %>%
-#   filter(pos=="NOUN"| pos=="VERB" | pos=="ADJ"| pos=="ADP"| pos=="DET" | pos=="PRON") %>%
-#   summarise(pos_count=n()) %>%
-#   left_join(d_flowers_utt_length_rep) %>%
-#   mutate(prop_pos=pos_count/total_num_words)
-#         
-# #jpeg("~/Desktop/plots/pos_rep.jpg", width = 950, height = 950)
-# ggplot(d_flowers_pos, 
-#        aes(x=repNum, y=prop_pos, label=gameId, color=condition)) + 
-#   geom_smooth(method = "lm", formula = y~x) + 
-#   geom_point(alpha=.1) +
-#   facet_grid(cols=vars(pos)) +
-#   xlab("Rep number") + 
-#   ylab("Proportion of pos") +
-#   ylim(0,5)+
-#   theme_bw()#+
-#  # theme(text = element_text(size = 30))
-# #dev.off() 
+
+text <- d_flowers$text
+parsed <- spacy_parse(text,pos=TRUE) %>%
+  select(doc_id, pos) %>%
+  group_by(doc_id) %>%
+  mutate(pos = paste(pos, collapse=",")) %>%
+  distinct()
+
+d_flowers_pos  <- cbind(d_flowers, parsed) %>%
+  separate_rows(pos, convert = TRUE) %>%
+  group_by(condition, blockNum, trialNum, gameId, pos) %>%
+  filter(pos=="NOUN"| pos=="VERB" ) %>% # pos=="ADJ"| pos=="DET" | pos=="PRON"
+  summarise(pos_count=n()) %>%
+  left_join(d_flowers_uttlength_nwords) %>%
+  mutate(prop_pos=pos_count/total_num_words)
+
+#jpeg("~/Desktop/plots/pos_block.jpg", width = 950, height = 950)
+pos_function <-function(d_flowers_pos, c,q){
+
+ggplot(d_flowers_pos, 
+       aes(x=trialNum, y=prop_pos, label=gameId, color=condition, linetype=pos)) + 
+  geom_smooth(method = "lm", formula = y~x) + 
+ # geom_point(alpha=.5) +
+ # facet_grid(cols=vars(condition)) +
+ # ylim(0,1)+
+  xlim(c,q)+
+  theme_bw() 
+}
+
+a<- pos_function(d_flowers_pos %>% filter(blockNum==0), 0,5) +  theme(legend.position="none", axis.title.y = element_blank(),axis.title.x = element_blank())+
+   labs(subtitle = "Block1") + ylim(0.0, 0.4)
+b<- pos_function(d_flowers_pos %>% filter(blockNum==1), 6,11) +  theme(legend.position="none", axis.title.y = element_blank(),axis.text.y = element_blank(), axis.title.x = element_blank())+
+   labs(subtitle = "Block2")+ ylim(0.0, 0.4)
+c<- pos_function(d_flowers_pos %>% filter(blockNum==2), 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.title.x = element_blank(),axis.text.y = element_blank())+
+   labs(subtitle = "Block3")+ ylim(0.0, 0.4)
+d<- pos_function(d_flowers_pos %>% filter(blockNum==3), 18,23) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+
+   labs(subtitle = "Block4")+ ylim(0.0, 0.4)
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d, legend, nrow = 1, ncol=5,  align = "v")
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Proportion Verbs, Nouns", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
 ```
 
-### 4. Lexical Diversity (TTR)
+### Lexical Diversity (TTR)
 
@@
 
-d_flowers_text_rep <- d_flowers %>%
-    group_by(gameId, repNum, condition) %>%
-    mutate(text_block = paste0(text, collapse = " ")) %>%
-    select(-c(text)) %>%
-    distinct()
-
-d_flowers_ttr_rep <- d_flowers_text_rep %>%
-  mutate(ttr_ = ttr(text_block))
-
-#jpeg("~/Desktop/plots/ttr_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_ttr_rep, 
-       aes(x=repNum, y=ttr_,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
- # facet_grid(rows =vars(condition)) +
-  xlab("Rep number") + 
-  ylab("Type_token ratio") + 
-  ylim(0,1)+
-  theme_bw()+
-  ggtitle("TTR per rep")#+
- # theme(text = element_text(size = 30))
-#dev.off()
-
 d_flowers_text_block <- d_flowers %>%
@@
 
-## STEP FOUR - Feedback items and markers
-### 1. Distribution of speech across speakers
+```{r ttr2}
+d_flowers_text_trial <- d_flowers %>%
+    group_by(gameId, trialNum, blockNum, condition) %>%
+    mutate(text_block = paste0(text, collapse = " ")) %>%
+    select(-c(text)) %>%
+    distinct()
+
+d_flowers_ttr_trial <- d_flowers_text_trial %>%
+  mutate(ttr_ = ttr(text_block))
+
+ttr_function <- function(d_flowers_ttr_trial, c, q) {
+  p<-ggplot2::ggplot(d_flowers_ttr_trial, aes(y=ttr_, x=trialNum, label=gameId, color=condition)) + xlim(c,q) + ylim(0.5, 1.0)+
+  geom_point(alpha = .5)+
+  geom_smooth(method = "lm", formula = y~x)+
+  theme_bw()+
+  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+return(p)
+}
+
+#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
+a<- ttr_function(d_flowers_ttr_trial %>% filter(blockNum==0), 0,5) + ylab("Type token ration")+  theme(legend.position="none",  axis.title.x = element_blank(), axis.title.y = element_blank())+ labs(subtitle = "Block1")
+
+b<- ttr_function(d_flowers_ttr_trial %>% filter(blockNum==1), 6,11) + theme(legend.position="none", axis.title.x = element_blank(), axis.title.y = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block2")+ylim(0.5, 1.0)
+
+c<- ttr_function(d_flowers_ttr_trial %>% filter(blockNum==2), 12,17) + theme(legend.position="none", axis.title.x = element_blank(), axis.title.y = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block3")+ylim(0.5, 1.0)
+
+d<- ttr_function(d_flowers_ttr_trial %>% filter(blockNum==3), 18,23) + labs(subtitle = "Block4")+ theme(legend.position="none", axis.title.x = element_blank(), axis.title.y = element_blank(),axis.text.y = element_blank())+ylim(0.5, 1.0)
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d, legend, nrow = 1, ncol=5, align = "v")
+#dev.off() 
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Type token ratio", rot=90)
+p<- grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
+#ggsave(file="../figs/ttr.png", p)
+
+```
+
+### Distribution of speech across speakers
 ```{r word_length}
@@
 
+```
+
+```{r word_length2}
 d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(playerId, repNum, gameId, condition) %>%
+  group_by(playerId, trialNum, blockNum, gameId, condition) %>%
   summarise(n_utt_speaker=n())
@@
 d_flowers_n_utt_game <- d_flowers %>%
-  group_by(repNum, gameId, condition) %>%
+  group_by(trialNum, blockNum, gameId, condition) %>%
   summarise(n_utt_game=n())
@@
 
-#jpeg("~/Desktop/plots/proputts_rep_speaker.jpg", width = 950, height = 950)
-ggplot(d_flowers_n_utt_dist, 
-       aes(x=repNum, y=overall_turns, label=playerId,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
- # facet_wrap(~condition)+
-  xlab("Rep number") + 
-  ylab("Proportion of turns") + 
+turn_per_plot_function <- function(d_flowers_n_utt_dist, overall_turns, c,q) {
+  p<-ggplot2::ggplot(d_flowers_n_utt_dist, aes(y={{overall_turns}}, x=trialNum, label=playerId)) +
+  geom_point(alpha = .5)+
+  xlim(c,q) +
+  facet_wrap(~condition)+
+  geom_smooth(method = "lm", formula = y~x)+
   theme_bw()+
-  ggtitle("Proportion of turns per rep and speaker")#+
- # theme(text = element_text(size = 30))
-#dev.off()
+  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),axis.title.y = element_blank(),axis.title.x = element_blank())
+return(p)
+}
+
+#jpeg("~/Desktop/plots/proputts_trial_speaker.jpg", width = 950, height = 950)
+a<-turn_per_plot_function(d_flowers_n_utt_dist %>% filter(blockNum==0), overall_turns, 0,5) + ylab("Proportion of turns")+ labs(subtitle = "Block1") + theme_bw() + theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank())
+
+b<-turn_per_plot_function(d_flowers_n_utt_dist %>% filter(blockNum==1), overall_turns, 6,11) +  theme(legend.position="none", axis.text.y=element_blank(),axis.title.x = element_blank(),axis.title.y = element_blank())+ labs(subtitle = "Block2")+ theme_bw()
+
+c<-turn_per_plot_function(d_flowers_n_utt_dist%>% filter(blockNum==2), overall_turns, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), axis.title.x = element_blank())+ labs(subtitle = "Block3")+ theme_bw()
+
+d<-turn_per_plot_function(d_flowers_n_utt_dist%>% filter(blockNum==3), overall_turns, 18,23) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(), axis.text.y=element_blank(), plot.tag.position = "topright")+ labs(subtitle = "Block4")+ theme_bw()
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d,  nrow = 2, ncol=3,  legend, align = "v")
+
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Proportion of turns", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
 
@@
 
-### 2. Turn-eliciting questions
+### Questions
 
@@
 
-#####
-
-d_flowers_n_utt_speaker <- d_flowers %>%
-  group_by(repNum, gameId, condition) %>%
-  summarise(n_utt_speaker=n())
-
-d_flowers_q <- d_flowers %>% 
- mutate(sentence_type = ifelse(str_detect(text, "\\?"), 
-                             "question",
-                             "other")) %>%
- group_by(repNum, gameId, sentence_type) %>%
- summarise(n_sentence_Type=n()) %>%
- filter(sentence_type=="question") %>%
- left_join(d_flowers_n_utt_speaker) %>%
- mutate(prop_questions = n_sentence_Type/n_utt_speaker )
-
-#jpeg("~/Desktop/plots/propquestions_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_q, 
-       aes(x=repNum, y=prop_questions, label=gameId,  color=condition)) + 
-  geom_smooth(method = "lm", formula = y~x) + 
-#  facet_wrap(~ condition) +
-  geom_point(alpha=.5) +
-  xlab("Rep number") + 
-  ylab("Proportion of questions") + 
-  theme_bw()+
-  ggtitle("Proportion of questions per rep")#+
-#theme(text = element_text(size = 30))
-#dev.off()
+##
 
@@
 
-### 1. Distribution of speech across speakers
+```{r questions2}
+
+d_flowers_n_utt_speaker <- d_flowers %>%
+  group_by(trialNum, blockNum, gameId, condition) %>%
+  summarise(n_utt_speaker=n())
+
+d_flowers_q <- d_flowers %>% 
+ mutate(sentence_type = ifelse(str_detect(text, "\\?"), 
+                             "question",
+                             "other")) %>%
+ group_by(trialNum, blockNum, gameId, sentence_type) %>%
+ summarise(n_sentence_Type=n()) %>%
+ filter(sentence_type=="question") %>%
+ left_join(d_flowers_n_utt_speaker) %>%
+ mutate(prop_questions = n_sentence_Type/n_utt_speaker )
+
+word_per_plot_function <- function(d_flowers, prop_words, c, q) {
+  p<-ggplot2::ggplot(d_flowers, aes(y={{prop_words}}, x=trialNum, label=gameId, color=condition)) + ylim(0.0,0.20)+ xlim(c,q) +
+  geom_point(alpha = .5)+
+  geom_smooth(method = "lm", formula = y~x)+
+  theme_bw()+
+  xlab("Trial number") + 
+  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+return(p)
+}
+
+#jpeg("~/Desktop/plots/propquestions_trial.jpg", width = 950, height = 950)
+a<- word_per_plot_function(d_flowers_q %>% filter(blockNum==0), prop_questions, 0,5) + ylab("Proportion of questions")+  theme(legend.position="none", axis.title.x = element_blank(), axis.title.y = element_blank())+ labs(subtitle = "Block1")
+
+b<-word_per_plot_function(d_flowers_q %>% filter(blockNum==1), prop_questions, 6,11) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), axis.title.x = element_blank())+ labs(subtitle = "Block2")
+
+c<-word_per_plot_function(d_flowers_q%>% filter(blockNum==2), prop_questions, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), axis.title.x = element_blank())+ labs(subtitle = "Block3")
+
+d<-word_per_plot_function(d_flowers_q%>% filter(blockNum==3), prop_questions, 18,23) +  theme(legend.position="none",axis.title.y = element_blank(), axis.text.y=element_blank(),axis.title.x = element_blank(), plot.tag.position = "topright")+ labs(subtitle = "Block4")
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d,  nrow = 1, ncol=5, legend, align = "v")
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Proportion of questions", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
+```
+
+### Feedback words (backchannel, hedge)
 ```{r feedback}
 
-###politeness markers
+backchanneling<- c("mm", "okay", "uh-huh", "ok", "mm-hm", "uh", "um", "agree", "uhuh", "mmm", "wow", "great", "mm", "hm",  "ummm", "hmmmm", "huh", "un", "um", "ohh", "ooo", "see", "oooo", "ununun", "oh", "ah", "true", "agree", "right", "yeah", "good", "really") #backchanneling: (https://www.reading.ac.uk/AcaDepts/ll/app_ling/internal/Cutrone_vol_2.pdf)
 
-###backchanneling: (https://www.reading.ac.uk/AcaDepts/ll/app_ling/internal/Cutrone_vol_2.pdf)
-
-backchanneling<- c("mm", "okay", "uh-huh", "ok", "mm-hm", "uh", "um", "agree", "uhuh", "mmm", "wow", "great", "mm", "hm",  "ummm", "hmmmm", "huh", "un", "um", "ohh", "ooo", "see", "oooo", "ununun", "oh", "ah", "true", "agree", "right", "yeah", "good", "really")
-
-###hedges: (https://www.researchgate.net/publication/280125979_Linguistic_Markers_and_Stylistic_Attributes_of_Hedging_in_English_Academic_Papers_Written_by_Native_and_Non-Native_Speakers_of_English/figures)
-
-hedging <- c("may", "perhaps", "might", "possible", "likely", "possibly", "maybe", "probable", "appear", "seem", "suggest", "sometimes", "seemingly", "apparently", "often", "could", "usually", "likely", "tend", "sometimes", "probably", "primarily", "tendency", "largely")
+hedging <- c("may", "perhaps", "might", "possible", "likely", "possibly", "maybe", "probable", "appear", "seem", "suggest", "sometimes", "seemingly", "apparently", "often", "could", "usually", "likely", "tend", "sometimes", "probably", "primarily", "tendency", "largely") #based on https://www.researchgate.net/publication/280125979_Linguistic_Markers_and_Stylistic_Attributes_of_Hedging_in_English_Academic_Papers_Written_by_Native_and_Non-Native_Speakers_of_English/figures
 
@@
 
+```
+
+```{r feedback2}
+
 d_flowers_feedback<- d_flowers %>%
-  group_by(repNum, gameId, condition) %>%
+  group_by(trialNum, blockNum, gameId, condition) %>%
   mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%          
@@
   mutate(hedge_length=length(hedge)) %>%
-  group_by(gameId, repNum, condition)  %>%
+  group_by(gameId, trialNum, blockNum, condition)  %>%
   summarise(total_num_backchannel = sum(backchannel_length),
             total_num_hedge = sum(hedge_length)) %>%
-  left_join(d_flowers_nwords) %>%
+  left_join(d_flowers_uttlength_nwords) %>%
   mutate(prop_backchannel_words= total_num_backchannel/total_num_words) %>%
@@
 
-#jpeg("~/Desktop/plots/propbackchannel_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_backchannel_words , x=repNum, label=gameId,  color=condition)) +
-  #facet_wrap(~ condition) +
-  geom_point(alpha = .5)+
-  geom_smooth()+
-  xlab("Rep number") + 
-  ylab("Proportion of backchannel words") + 
-  theme_bw()+
-  ggtitle("Proportion of backchannel words per rep")#+
- # theme(text = element_text(size = 30))
-#dev.off()
+a<- word_per_plot_function(d_flowers_feedback %>% filter(blockNum==0), prop_backchannel_words, 0,5) +  theme(legend.position="none",  axis.title.y = element_blank(), axis.title.x = element_blank())+
+   labs(subtitle = "Block1")
 
-#jpeg("~/Desktop/plots/prophedge_rep.jpg", width = 950, height = 950)
-ggplot(d_flowers_feedback, aes(y=prop_hedge_words , x=repNum, label=gameId,  color=condition)) +
+b<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==1), prop_backchannel_words, 6,11) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), axis.title.x = element_blank())+ labs(subtitle = "Block2")
+
+c<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==2), prop_backchannel_words, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), axis.title.x = element_blank())+ labs(subtitle = "Block3")
+
+d<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==3), prop_backchannel_words, 18,23) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), plot.tag.position = "topright", axis.title.x = element_blank())+ labs(subtitle = "Block4")
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d,  nrow = 1, ncol=5, legend, align = "v")
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Backchannel words", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
+a<- word_per_plot_function(d_flowers_feedback %>% filter(blockNum==0), prop_hedge_words, 0,5) +  theme(legend.position="none", axis.title.y = element_blank())+
+   labs(subtitle = "Block1")+ylim(0.0, 0.05)
+
+b<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==1), prop_hedge_words, 6,11) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank())+ labs(subtitle = "Block2")+ylim(0.0, 0.05)
+
+c<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==2), prop_hedge_words, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank())+ labs(subtitle = "Block3")+ylim(0.0, 0.05)
+
+d<-word_per_plot_function(d_flowers_feedback%>% filter(blockNum==3), prop_hedge_words, 18,23) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(), plot.tag.position = "topright")+ labs(subtitle = "Block4")+ylim(0.0, 0.05)
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+
+p<-plot_grid(a,b,c,d, legend,  nrow = 1, ncol=5,align = "v")
+
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("Hedge words", rot=90)
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
+```
+
+### LIWC (positive, negative) categories
+
+```{r LIWC categories, include=TRUE, warning=FALSE}
+
+liwc<- read_csv("/Users/lscpuser/Downloads/liwc1.csv")
+liwc <- liwc[!is.na(liwc$Negemo), ]
+liwc <- liwc[!is.na(liwc$Posemo), ]
+
+d_flowers_liwc<- d_flowers %>%
+  group_by(blockNum, trialNum, gameId, condition) %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text)) %>%          
+  mutate(text = strsplit(text, " ")) %>%
+  rowwise() %>%
+  mutate(positive=list(intersect(liwc$Posemo, text))) %>%
+  mutate(positive_length=length(positive))%>%
+  mutate(negative=list(intersect(liwc$Negemo, text))) %>%
+  mutate(negative_length=length(negative)) %>%
+  group_by(gameId, trialNum, blockNum, condition)  %>%
+  summarise(total_num_positive = sum(positive_length),
+            total_num_negative = sum(negative_length)) %>%
+  left_join(d_flowers_uttlength_nwords) %>%
+  mutate(prop_negative_words= total_num_negative/total_num_words) %>%
+  mutate(prop_positive_words= total_num_positive/total_num_words)
+
+#######positive LIWC
+ggplot(d_flowers_liwc, aes(y=prop_positive_words , x=trialNum, label=gameId,  color=condition)) +
  # facet_wrap(~ condition) +
@@
   geom_smooth()+
-  xlab("Rep number") + 
-  ylab("Proportion of hedge words") + 
+  xlab("Trial number") + 
+  ylab("Proportion of positive words") + 
   theme_bw()+
-  ggtitle("Proportion of hedge words per rep")#+
+  ggtitle("Proportion of positive words per trial")#+
+  #theme(text = element_text(size = 30))
+
+a<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==0), prop_positive_words, 0,5) +  theme(legend.position="none",  axis.title.y = element_blank(),axis.title.x = element_blank())+
+   labs(subtitle = "Block1") +  ylab("LIWC positive words")
+
+b<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==1), prop_positive_words, 6,11) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(),axis.title.x = element_blank())+ labs(subtitle = "Block2")
+
+c<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==2), prop_positive_words, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(),axis.title.x = element_blank())+ labs(subtitle = "Block3")
+
+d<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==3), prop_positive_words, 18,23) +  theme(legend.position="none", axis.title.y = element_blank(), axis.text.y=element_blank(),axis.title.x = element_blank(), plot.tag.position = "topright")+ labs(subtitle = "Block4")
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  d +  theme(legend.position = "right")
+)
+p<-plot_grid(a,b,c,d, nrow = 1, legend,  ncol=5, align = "v")
+
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("LIWC positive words", rot=90)
+p<-grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+ggsave(file="../figs/liwc_pos.png", p)
+
+######### negative LIWC
+ggplot(d_flowers_liwc, aes(y=prop_negative_words , x=trialNum, label=gameId,  color=condition)) +
+ # facet_wrap(~ condition) +
+  geom_point(alpha = .5)+
+  geom_smooth()+
+  xlab("Trial number") + 
+  ylab("Proportion of negative words") + 
+  theme_bw()+
+  ggtitle("Proportion of negative words per trial")#+
   #theme(text = element_text(size = 30))
@@
 
+e<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==0), prop_negative_words, 0,5) +  theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y=element_blank()) +
+   labs(subtitle = "Block1")  +ylim(0.0, 0.05)
+
+f<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==1), prop_negative_words, 6,11) +  theme(legend.position="none", axis.title.y = element_blank(),axis.title.x = element_blank(), axis.text.y=element_blank())+ labs(subtitle = "Block2")+ylim(0.0, 0.05)
+
+g<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==2), prop_negative_words, 12,17) +  theme(legend.position="none", axis.title.y = element_blank(),axis.title.x = element_blank(), axis.text.y=element_blank())+ labs(subtitle = "Block3")+ylim(0.0, 0.05)
+
+h<-word_per_plot_function(d_flowers_liwc%>% filter(blockNum==3), prop_negative_words, 18,23) +  theme(legend.position="none", axis.title.y = element_blank(),axis.title.x = element_blank(), axis.text.y=element_blank(), plot.tag.position = "topright")+ labs(subtitle = "Block4")+ylim(0.0, 0.05)
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  h +  theme(legend.position = "right")
+)
+
+p<- plot_grid(e,f,g,h, legend, nrow = 1, ncol=5, align = "v")
+x.grob <- textGrob("Trial Number")
+y.grob <- textGrob("LIWC negative words", rot=90)
+
+#add to plot
+grid.arrange(arrangeGrob(p, left = y.grob, bottom = x.grob))
+
 ```
+### Alignment
+
+```{r alignment, include=TRUE, warning=FALSE}
+library(langcog)
+library(ggpubr)
+d3<- read_csv("../data/processed_data/joined_data/output_stan_model.csv")
+
+d4 <- d3 %>%
+  group_by(subpop) %>%
+  multi_boot_standard("model_eta", na.rm = F)
+
+options(ggrepel.max.overlaps = 50)
+ggplot(aes(y = model_eta, x = subpop, colour = subpop,
+), data=d3) + 
+  #geom_line(aes(group=category))+
+  geom_point(data=d3)+
+  #geom_smooth(method = "loess") +
+  theme_bw(base_size = 14) +
+  geom_text_repel(aes(label=category))+
+  theme(panel.grid = element_blank()) +
+  labs(title='Model-estimated Alignment for speaker pairs, with mean and CI',
+       y='Alignment (delta log-odds)',
+       x='Conditions',
+       colour='Alignment')+
+  geom_pointrange(data=d4, mapping=aes(y = mean, x = subpop, ymax = ci_upper, ymin = ci_lower), col="black")+
+ylim(-0.5,0.5)+
+  grids(linetype = "dashed")
+
+ggsave("../figs/alignment.png", height = 3, width = 10)

### 2021-09-16 code/aa-flowers-pilot_analysis.Rmd
add ms analyses

@@
 d.feedback
-d.feedback %>% ggplot(aes(x = human)) + geom_bar(stat = "count") + theme_jmank()
+d.feedback %>% ggplot(aes(x = human)) + geom_bar(stat = "count")
 ```
@@
 
-[] fix roundId
-
 ```{r deference}
 trial_deference <- d.by_player_metrics %>% 
-  rename(roundID = "_id") %>% 
   full_join(d.contexts %>% 
-              select(roundID, blinded, label, utility)) %>%
+              select(roundId, blinded, label, utility)) %>%
   filter(blinded != playerId) %>%
-  group_by(playerId, roundID) %>%
+  group_by(playerId, roundId) %>%
   summarize(deference = utility[label == playerResponse] != max(utility))
@@
 d.deference <- d.by_player_metrics %>% 
-  rename(roundID = "_id") %>% 
   left_join(trial_deference)

### 2021-09-16 code/aa_mic_submission_analyses.Rmd
add ms analyses

@@
 
-```{r setup}
+```{r}
 library(here)
@@
 n_rounds = 24
-games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=n_players*n_rounds) %>% select(gameId)
+
+game_completion <- d.rounds %>% 
+  group_by(gameId) %>% 
+  tally() %>% 
+  mutate(game_complete = n==n_players*n_rounds) %>%
+  rename(n_trials = n)
+
+d.games.final <- d.games %>% 
+  left_join(game_completion) %>% 
+  replace_na(list(game_complete = FALSE, n_trials = 0))
+
+d.games.conditions <- d.games.final %>% 
+  select(gameId, color, condition, chatEnabled, game_complete) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"))
 
 d.rounds.final <-  d.rounds %>% 
-  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  filter(conditionName != "coopCartel") %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId))
+  left_join(d.games.final %>% 
+              select(gameId, conditionName, condition,color, chatEnabled, game_complete)) %>% 
+  filter(conditionName != "coopCartel",
+         game_complete)
 
 d.raw_chat.final <- d.raw_chat %>% 
-  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId)) 
+  left_join(d.games.final %>% 
+              select(gameId, conditionName, condition, game_complete)) %>%
+  filter(game_complete)
+
 ```
@@
 ```{r}
-d.games %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId)) %>%
+d.games.final %>% 
   group_by(-game_complete, condition, chatEnabled) %>%
@@
 ## Game lengths
+```{r}
+d.games.final %>% 
+filter(game_complete) %>%
+  group_by(condition, chatEnabled) %>%
+  mutate(gameLength_minutes = gameLength/60) %>%
+  summarize_at("gameLength_minutes", funs(min, max, mean))
+```
+
 ## Bonus Payments
+Max bonus payment
+```{r}
+max(d.players$bonus, na.rm=T)
+```
+## Hourly compensation 
+```{r}
+#I just pulled these averages from prolific
+(35.81*16 + 25.75 * 47 + 26.38 * 47 + 24.76*18 + 10.17* 18 + 12.64 * 18 + 11.79 * 68 + 11.24*66)/(16+47+47+18+18+18+68+66)
 
-# Points x Time x Num. Words
+```
 
-# Avg. Points Per Round
+# Summarizing Data
 
+We'll pull out the summary statistics with 3 views: 
+* by player (points earned, number of utterances, number of words)
+  ** selection metrics: select hidden, non-top flower
+* by round (group points earned, max possible points earned in round, number of words exchanged by group)
+* by game (total points, max possible points in game)
+
+```{r}
+#player language per round
+d.player_word_counts <- d.raw_chat.final %>% 
+  filter(type == "message") %>%
+  mutate(text = gsub('[[:punct:] ]+',' ',text),
+         text = str_squish(text),
+         utt_length_chars = str_length(text), 
+         utt_length_words = str_count(text, "\\W+") + 1) %>% 
+  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
+  summarize(text = paste0(text, collapse = ' - '),
+            total_num_words = sum(utt_length_words),
+            total_num_chars = sum(utt_length_chars))
+
+# player points per round
+d.by_player_metrics <- d.rounds.final  %>% 
+  left_join(d.player_word_counts) %>%
+  left_join(d.games.conditions) %>% 
+  replace_na(list(total_num_words = 0, total_num_chars = 0))
+
+#add deference measure and flower ranking
+deference_rank <-  d.by_player_metrics %>%  
+  full_join(d.contexts %>% 
+              select(roundId, blinded, label, utility)) %>%
+  filter(blinded != playerId) %>%
+  group_by(playerId, roundId) %>%
+  arrange(-utility) %>%
+  mutate(selectionRank = c(0:3)) %>%
+  filter(playerResponse == label) %>%
+  mutate(deference = selectionRank != 1) %>%
+  select(roundId, playerId,deference, selectionRank)
+
+d.by_player_metrics <- d.by_player_metrics %>% 
+  left_join(d.contexts %>% select(roundId, playerResponse = label, 
+                                  selectedHidden = blinded,
+                                  selectionUtility = utility)) %>%
+  mutate(selectedHidden = selectedHidden == playerId) %>%
+  left_join(deference_rank)
+```
+
+Calculate the best possible per round
+
+```{r max_round}
+d.max_round <- d.contexts %>%
+  group_by(gameId, repNum, blockNum, trialNum) %>%
+  #top 3 flowers
+  slice_max(utility, n = 3) %>%
+  summarise(roundMax = sum(utility))
+```
+
+```{r}
+d.by_round_metrics <- d.by_player_metrics  %>% 
+  group_by(gameId, trialNum, roundId) %>%
+  summarize(roundUtility = sum(playerUtility, na.rm = T),
+            numWords = sum(total_num_words, na.rm = T),
+            numChars = sum(total_num_chars, na.rm = T),
+            numFlowers = n_distinct(playerResponse),
+            roundLength = max(time_sec)) %>%
+  left_join(d.max_round) %>%
+  mutate(pointsAdjusted = roundUtility/roundMax) %>% 
+  left_join(d.games.conditions)
+```
+```{r}
+d.by_game_metrics <- d.by_round_metrics  %>% 
+  group_by(gameId) %>%
+  summarize(gameUtility = sum(roundUtility, na.rm = T),
+            numWords = sum(numWords, na.rm = T),
+            numChars = sum(numChars, na.rm = T),
+            maxPossPoints = sum(roundMax)) %>%
+  mutate(pointsAdjusted = gameUtility/maxPossPoints) %>% 
+  left_join(d.games.conditions) %>%
+  left_join(d.games.final %>% select(gameId, gameLength))
+```
+
+# Merge in RL Experiments
+
+```{r read-rl}
+d.rl_metrics <- read_csv(here("data/processed_data/RL_experiment_results.csv"))
+
+#get in line with player summary statistics?
+
+d.rl.player_metrics <- d.rl_metrics %>% 
+  mutate(gameId = paste(condition_raw, languageConditionRaw, trial, sep = "-"), 
+         playerId = paste0(trial, "-",agent)) %>%
+  select(playerId, languageCondition, condition, avgPlayerPoints = reward) %>% 
+  mutate(rl_agent = "Agent")
+
+```
+
+# Fig 1 & 2
+## RL Point Comp
+
+### All Players
+```{r}
+d.by_player_metrics %>%
+  group_by(languageCondition, condition, playerId) %>%
+  summarize(avgPlayerPoints = mean(playerUtility)) %>% 
+  mutate(rl_agent = "Human") %>%
+  rbind(d.rl.player_metrics) %>%
+  ggplot(aes(x = condition, y = avgPlayerPoints, color = rl_agent)) + 
+  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
+  scale_color_brewer(palette="Dark2")+ theme_jmank() + 
+  facet_wrap(~languageCondition)+ 
+  theme(legend.position = "bottom", axis.title.x=element_blank()) +
+  labs(y = "Average Player Points", color = "Player Type")
+```
+### Avg Points over game
+Do the same but by average points per round across players
+
+```{r}
+d.rl.game_metrics <- d.rl_metrics %>% 
+  mutate(gameId = paste(condition_raw, languageConditionRaw, trial, sep = "-"), 
+         playerId = paste0(trial, "-",agent)) %>%
+  group_by(gameId, languageCondition, condition) %>% 
+  summarize(avgRoundPoints = sum(reward)) %>%
+  mutate(rl_agent = "Agent")
+
+d.rl.game_metrics
+```
+
+{} to match other plot, facet by utility condition, color by agent type?
+```{r}
+rl_comparison <- d.by_round_metrics %>%
+  group_by(languageCondition, condition, gameId) %>%
+  summarize(avgRoundPoints = mean(roundUtility)) %>% 
+  mutate(rl_agent = "Human") %>%
+  rbind(d.rl.game_metrics) %>%
+  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
+  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
+  scale_color_bay()+ theme_jmank() + 
+ # labs(y = "Average Points", color = "Player Type") +
+  labs(y = "Average Points per Round", x = "Language Condition", color = "Utility Condition") +
+  facet_wrap(~rl_agent)+ 
+  theme(legend.position = "bottom")
+
+rl_comparison
+```
+```{r}
+d.by_round_metrics %>%
+  group_by(languageCondition, condition, gameId) %>%
+  summarize(avgRoundPoints = mean(roundUtility)) %>% 
+  mutate(rl_agent = "Human") %>%
+  rbind(d.rl.game_metrics) %>%
+  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
+  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
+  scale_color_brewer(palette="Dark2")+ theme_jmank() + 
+ # labs(y = "Average Points", color = "Player Type") +
+  labs(y = "Average Points per Round", color = "Player Type", x = "Language Condition") +
+  facet_wrap(~rl_agent)+ 
+  theme(legend.position = "bottom")
+```
+
+## Points x Time x Num. Words
+```{r}
+lang_by_points <- d.by_game_metrics %>% 
+  ggplot(aes(x = gameLength/60, y = gameUtility, shape = languageCondition, color = log(numWords))) + 
+  geom_point(alpha = .8) +
+  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
+  facet_grid(cols = vars(condition)) + 
+  labs(x = "Game length (in Minutes)", 
+       y = "Points Earned by Group",
+       color = "Log N. Words",
+       shape = "Language") + 
+  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + theme(legend.position = "bottom")
+lang_by_points
+```
+
+### Rewards Blurb
+
+```{r}
+d.by_game_metrics %>% mutate(lang_used = numWords > 0) %>%
+  group_by(languageCondition, lang_used) %>% count()
+
+max(d.by_game_metrics$numWords)
+```
 # Language Reduction
+Want number of words exchanged over the game
+ More utterances in individual, (same turns)
 
-# Georgia Analyses
+ shared utilities
+
+ more questions in shared?
+```{r}
+lang_reduction <- d.by_round_metrics %>%
+  filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = trialNum, y = numWords, color = condition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  #geom_smooth(method=lm, formula=y~log(x))+
+   labs(x = "Trial Number", y = "N. Words", color = "Utility Condition")+ 
+  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")# + scale_y_continuous(trans='log10')
+
+utterances <- d.raw_chat.final %>% 
+  left_join(d.games.conditions) %>%
+  filter(type == "message") %>%
+  select(gameId, trialNum, playerId, text) %>%
+  #this drops the NAs, we want to keep instances where there was a selection but no message ugh
+  mutate(text = gsub('[[\']+]|[[:punct:] ]+',' ',text),
+         text = str_squish(text),
+         utt_length_chars = str_length(text), 
+         utt_length_words = str_count(text, "\\W+")+1) %>%
+  group_by(gameId, trialNum, playerId) %>%
+  summarize(utt_length_chars = list(utt_length_chars),
+            utt_length_words = list(utt_length_words))
+
+expand_utt <- d.raw_chat.final %>%
+  select(gameId, trialNum, playerId) %>% 
+  distinct() %>%
+  full_join(utterances)
+
+no_lang_index <- sapply(expand_utt$utt_length_chars, is.null)
+expand_utt$utt_length_chars[no_lang_index] <- 0
+expand_utt$utt_length_words[no_lang_index] <- 0
+
+utt_reduction <- expand_utt %>% 
+  unnest(cols = c(utt_length_chars, utt_length_words)) %>%
+  left_join(d.games.conditions) %>%
+  filter(languageCondition == "Lang", game_complete) %>%
+  group_by(gameId, trialNum, languageCondition, condition) %>%
+  summarize(mean_utt_length = mean(utt_length_words))
+
+utt_reduction %>%  ggplot(aes(x = trialNum+1, y = mean_utt_length, color = condition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  labs(x = "Trial Number", y = "Average N. Words per Utterance", color = "Utility Condition")+ 
+  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")
+
+#points increase as language decreases?
+d.by_round_metrics %>%
+  filter(languageCondition == "Lang") %>%
+  ggplot(aes(x = trialNum, y = roundUtility, color = condition)) + 
+  geom_point(alpha = .5) + 
+  geom_smooth(method=lm)+
+  scale_color_bay() + theme_jmank()
+
+```
+
+# Final Figs
+
+```{r}
+library(patchwork)
+lang_by_points
+rl_comparison
+lang_reduction
+
+lang_by_points + lang_reduction +plot_annotation(tag_levels = 'A')#, labels = c('A', 'B'), label_size = 12, aligh = "hv")
+
+ggsave(here(fig_path, "points_earned_reduction_joined.png"), height = 4, width = 11)
+
+rl_comparison
+ggsave(here(fig_path, "rl_comparison.png"), height = 3, width = 5)
+```
+
+# Number of questions
+Adapted from Georgia's report
+
+```{r}
+question_words <- c("who",  "what",  "where",  "why",  "how")
+d_messages <- d.raw_chat.final %>% 
+  filter(type=="message") %>%
+  filter(!is.na(text))
+
+d_messages <- d_messages %>% 
+  mutate(question_mark = ifelse(str_detect(text, "\\?"),TRUE,FALSE),
+         question_word = grepl(paste(question_words, collapse="|"), tolower(text)),
+         any_question = ifelse((question_mark | question_word), TRUE, FALSE))
+
+d_messaged_summarized <- d_messages %>% group_by(gameId, trialNum, condition) %>% 
+  summarize(total_utt = n(), 
+            n_question_marks = sum(question_mark), 
+            n_question_words = sum(question_word),
+            n_questions = sum(any_question),
+            proportion_q = n_question_marks/total_utt)
+
+d_messaged_summarized %>%
+  ggplot(aes(x = trialNum, y = proportion_q, color = condition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  #geom_smooth(method=lm, formula=y~log(x))+
+   labs(x = "Trial Number", y = "Prop of Questions", color = "Utility Condition")+ 
+  scale_color_bay() + theme_jmank() + theme(legend.position = "none") + ggtitle("All Trials")
+
+```

### 2021-09-16 code/linguistic_analysis_C-D.Rmd
add ms analyses

@@
 knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
-library(tidyverse)
+
 library(here)
@@
 library(gridExtra)
-
+library(tidyverse)
 source("./theme_specs.R")
@@
   filter(!is.na(text)) 
-
 ```
@@
 #  facet_grid(rows =vars(condition)) +
-  geom_point()+
-  geom_jitter(alpha=.05)+
+  geom_point(alpha = .25)+
   geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
@@
 #  facet_grid( rows =vars(condition)) +
-  geom_jitter(alpha=.05)+
+  geom_jitter(alpha=.25)+
   geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
@@
 
+#rep num is irrelevant here?
 #jpeg("~/Desktop/plots/nword_rep.jpg", width = 950, height = 950)
@@
   #facet_grid(rows =vars(condition)) +
-  geom_jitter(alpha=.05)+
+  geom_jitter(alpha=.25)+
   geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
@@
   theme(legend.position="bottom") + 
-  theme_bw() +
+  scale_color_bay() +
+  theme_jmank() +
   ggtitle("Number of turns per trial")#+
@@
   labs(title="Number of turns", y="Number of turns", x="Block number")+
-  theme(legend.position="bottom") + 
-  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
+  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")+
   ggtitle("Number of turns per block")#+
@@
   geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
+  geom_point(alpha=.25) +
  # facet_grid(~condition) +
@@
   ylab("Number of utterances") + 
-  theme_bw()#+
+  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
  # theme(text = element_text(size = 30))
@@
   ylab("Number of utterances") + 
-  theme_bw()#+
+  theme_jmank() + scale_color_bay() + theme(legend.position = "bottom")
 #  theme(text = element_text(size = 30))
@@
   xlab("Trial number") + 
-  theme_bw()+
+theme_jmank() + scale_color_bay() +  
     theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
@@
 #jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
-a<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==0), 0,5) + ylab("Number of utterances")+  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank())+ labs(subtitle = "Block1")
+a<- turnb_function(d_flowers_n_utt_speaker %>% 
+                     filter(blockNum==0), 0,5) + 
+  ylab("Number of utterances")+  
+  theme(legend.position="none",axis.title.y = element_blank(),
+        axis.title.x = element_blank())+ labs(subtitle = "Block1")
 
 #jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
-b<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==1), 6,11) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block2")
+b<- turnb_function(d_flowers_n_utt_speaker %>% 
+                     filter(blockNum==1), 6,11) +  
+  theme(legend.position="none",axis.title.y = element_blank(),
+        axis.title.x = element_blank(),
+        axis.text.y = element_blank())+ 
+  labs(subtitle = "Block2")
 
 #jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
-c<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==2), 12,17) +  theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block3")
+c<- turnb_function(d_flowers_n_utt_speaker %>% 
+                     filter(blockNum==2), 12,17) +  
+  theme(legend.position="none",axis.title.y = element_blank(),
+        axis.title.x = element_blank(),
+        axis.text.y = element_blank())+ 
+  labs(subtitle = "Block3")
 
 #jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
-d<- turnb_function(d_flowers_n_utt_speaker %>% filter(blockNum==3), 18,23) +theme(legend.position="none",axis.title.y = element_blank(),axis.title.x = element_blank(),axis.text.y = element_blank())+ labs(subtitle = "Block4")
+d<- turnb_function(d_flowers_n_utt_speaker %>% 
+                     filter(blockNum==3), 18,23) +
+  theme(legend.position="none",axis.title.y = element_blank(),
+        axis.title.x = element_blank(),axis.text.y = element_blank())+ 
+  labs(subtitle = "Block4")
 
@@
 
+#TODO I dont really understand why this figure is done this way?
 ```
@@
 ### Parts of speech
-```{r pos, eval=TRUE, include=TRUE}
+```{r pos, eval=FALSE, include=FALSE}
 
@@
  text <- d_flowers$text
- parsed <- spacy_parse(text,pos=TRUE) %>%
-   select(doc_id, pos) %>%
-   group_by(doc_id) %>%
-   mutate(pos = paste(pos, collapse=",")) %>%
-   distinct()
+ #I dont have spacy in an env for this project, cant reproduce will have to add later -jm TODO
+ # parsed <- spacy_parse(text,pos=TRUE) %>%
+ #   select(doc_id, pos) %>%
+ #   group_by(doc_id) %>%
+ #   mutate(pos = paste(pos, collapse=",")) %>%
+ #   distinct()
 #   
@@
   geom_smooth(method = "lm", formula = y~x) + 
-  geom_point(alpha=.1) +
+  geom_point(alpha=.05) +
  # facet_grid(rows =vars(condition)) +
-  xlab("Trial number") + 
-  ylab("Type_token ratio") + 
+  xlab("Trial Number") + 
+  ylab("Type to Token Ratio (TTR)") + 
   ylim(0,1)+
-  theme_bw()+
-  ggtitle("TTR per trial")#+
+  scale_color_bay() + theme_jmank() +
+  theme(legend.position = "bottom")
  # theme(text = element_text(size = 30))
 #dev.off()
+
+ggsave(here(fig_path, "ttr_trials.png"), height = 3, width = 3)
 

### 2021-09-16 code/preprocessing/00_preprocessing_functions.R

@@ -109,7 +109,7 @@ get_round_data <- function(data_read_location, date_start) {
     mutate(utility = ifelse(response == "false", 0, utility)) %>%
     #TODO: some times are negative?? interesting
     mutate(time = abs(as.numeric(time)/1000)) %>%
-    rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
+    rename(time_sec = time, playerResponse = response, playerUtility = utility, roundId = "_id") %>%
     arrange(createdAt, gameId, blockNum, repNum)
   return(d.round_results)
 }
@@ -136,8 +136,6 @@ get_raw_chat <- function(data_read_location, date_start) {
 }
 
 add_chat_info <- function(raw_chat_data, player_data, round_data, game_data) {
-  # reallign the message and selection timestamps
-  # get the chat timestamp first alert
   min_chat_alert <- raw_chat_data %>% group_by(gameId, trialNum) %>%
     filter(type == "selectionAlert") %>% 
     summarize(first_chat_time = min(time))
@@ -179,7 +177,7 @@ get_alternative_context <- function(data_read_location, date_start) {
                                 spread(key, value))) %>%
     unnest(data.context) %>%
     rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
-    rename(roundID = `_id`)
+    rename(roundId = `_id`)
   return(d.context)
 }
 
@@ -187,7 +185,7 @@ get_demographics <- function(data_read_location, date_start, games_data) {
   read_csv(here(data_read_location, 'player-inputs.csv'), 
            col_types = cols(data.age = col_number())) %>%
     filter(createdAt >= date_start) %>%
-  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% select(gameId:education)
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% select(playerId:education)
 }
 
 get_feedback <- function(data_read_location, date_start) {

### 2021-09-16 code/preprocessing/rl_preprocessing.R

@@ -20,6 +20,7 @@ d.rl_data <- d.rl_data.raw %>%
   mutate(condition_raw = condition, 
          condition = ifelse(condition_raw == "joint", "Shared Utilities", "Individual Utilities"),
          languageConditionRaw = languageCondition,
-         languageCondition =ifelse(languageConditionRaw == "contcomm", "Lang", "Nonlang"))
+         languageCondition =ifelse(languageConditionRaw == "contcomm", "Lang", "Nonlang"),
+         rl_agent = TRUE) %>%
   write_csv(here("data/processed_data/RL_experiment_results.csv"))
 

### 2021-09-16 code/linguistic_analysis_C-D.Rmd
mattr instead of simple type token ratio

@@
 
-#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
 a<- turnb_function(d_flowers_n_utt_speaker %>% 
@@
 
-#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
 b<- turnb_function(d_flowers_n_utt_speaker %>% 
@@
 
-#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
 c<- turnb_function(d_flowers_n_utt_speaker %>% 
@@
 
-#jpeg("~/Desktop/plots/ttr_trial.jpg", width = 950, height = 950)
 d<- turnb_function(d_flowers_n_utt_speaker %>% 
@@
 
+library(koRpus)
+install.koRpus.lang(lang=c("en"))
+library(koRpus.lang.en)
+
+MATTR <- function(txt, window=100, char=FALSE, ...){
+  if(isTRUE(char)){
+    char.value <- "MATTR"
+  } else {
+    char.value <- c()
+  }
+  results <- lex.div(txt=txt, window=window, measure="MATTR", char=char.value)
+  return(results)
+}
+
 ttr <- function(text_){
@@
 
+ttrMATTR <- function(text_){
+  text_ <- tolower(text_)
+  total_words = sapply(strsplit(text_, " "), length)
+  if (total_words >21){
+  text_obj <- tokenize(text_, format="obj", lang="en")
+  ttr_<- MATTR(text_obj, window=21)[1]
+  return(ttr_)
+  }
+  else {
+  ttr_<- 0
+  }
+}
+
 d_flowers_text_trial <- d_flowers %>%
@@
     mutate(text_block = paste0(text, collapse = " ")) %>%
-    select(-c(text)) %>%
+    #select(-c(text, time, playerId)) %>%
+    select(gameId, trialNum, condition, text_block) %>%
     distinct()
@@
 d_flowers_ttr_trial <- d_flowers_text_trial %>%
-  mutate(ttr_ = ttr(text_block))
-
-#t<-str_split(text_, " ", 1:6)
-#y<- paste(unlist(lapply(t,head,n=5)), collapse=" ")
+  mutate(ttr_ = ttrMATTR(text_block))
 
@@
   xlab("Trial Number") + 
-  ylab("Type to Token Ratio (TTR)") + 
+  ylab("Moving Average Type to Token Ratio (MATTR)") + 
   ylim(0,1)+

### 2021-09-16 code/aa_mic_submission_analyses.Rmd
final fig tweaking

@@
  # labs(y = "Average Points", color = "Player Type") +
-  labs(y = "Average Points per Round", x = "Language Condition", color = "Utility Condition") +
+  labs(y = "Average Points per Trial", x = "Language Condition", color = "Utility Condition") +
   facet_wrap(~rl_agent)+ 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        #legend.box="vertical",
+        legend.margin=margin(),
+        text = element_text(size=15),
+        axis.text = element_text(size = 12),
+        legend.title=element_text(size=12),
+        legend.text=element_text(size=12),
+        strip.text = element_text(size=15))
 
@@
   ggplot(aes(x = gameLength/60, y = gameUtility, shape = languageCondition, color = log(numWords))) + 
-  geom_point(alpha = .8) +
+  geom_point(alpha = .8, size = 4) +
   geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
@@
        color = "Log N. Words",
-       shape = "Language") + 
-  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + theme(legend.position = "bottom")
+       shape = "") + 
+  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + 
+  theme(legend.position = "bottom",
+        #legend.box="vertical",
+        legend.margin=margin(),
+        text = element_text(size=15),
+        axis.text = element_text(size = 12),
+        legend.title=element_text(size=12),
+        legend.text=element_text(size=12),
+        strip.text = element_text(size=15))
 lang_by_points
@@
 # Language Reduction
-Want number of words exchanged over the game
- More utterances in individual, (same turns)
 
- shared utilities
-
- more questions in shared?
 ```{r}
+# number of words exchanged decreases over time
 lang_reduction <- d.by_round_metrics %>%
   filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = trialNum, y = numWords, color = condition)) + 
+  mutate(utility_cond_short = ifelse(condition == "Shared Utilities", "Shared", "Individual")) %>%
+  ggplot(aes(x = trialNum, y = numWords, color = utility_cond_short)) + 
   geom_point(alpha = .2) +
@@
    labs(x = "Trial Number", y = "N. Words", color = "Utility Condition")+ 
-  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")# + scale_y_continuous(trans='log10')
+  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom") + 
+   theme(legend.position = "bottom",
+        text = element_text(size=15),
+        axis.text = element_text(size = 12),
+        legend.title=element_text(size=12),
+        legend.text=element_text(size=12),
+        strip.text = element_text(size=15))
+lang_reduction
 
@@
 
+# normal MLU also decreases
+utt_mlu <- utterances %>% 
+  unnest(c(utt_length_chars, utt_length_words)) %>%
+  left_join(d.games.conditions) %>%
+  group_by(trialNum, gameId) %>%
+  mutate(mlu = mean(utt_length_words)) %>% 
+  ggplot(aes(x = trialNum, y = mlu, color = condition)) + 
+  geom_point(alpha = .2) +
+  geom_smooth(method=lm)+
+  labs(x = "Trial Number", y = "Mean Length of Utterance", color = "Utility Condition")+ 
+  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")
+
+# MLU including trials with 0 "length" utterances
 expand_utt <- d.raw_chat.final %>%
@@
 
-#points increase as language decreases?
+#points increase (slightly) over time
 d.by_round_metrics %>%
   filter(languageCondition == "Lang") %>%
-  ggplot(aes(x = trialNum, y = roundUtility, color = condition)) + 
+  ggplot(aes(x = trialNum, y = pointsAdjusted, color = condition)) + 
   geom_point(alpha = .5) + 
@@
 
+```
+Point x Trial correlation
+```{r}
+library(psych)
+corel <- d.by_round_metrics %>%
+  filter(languageCondition == "Lang") %>% ungroup() %>%
+  select(trialNum, pointsAdjusted, numWords)
+corr.test(corel$trialNum, corel$pointsAdjusted)
 ```
@@
 library(patchwork)
+utt_mlu
 lang_by_points

### 2021-09-16 code/preprocessing/00_preprocessing_functions.R

@@ -194,6 +194,17 @@ get_feedback <- function(data_read_location, date_start) {
     rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% select(c(playerId, correctness:time))
 }
 
+get_posttest <- function(data_read_location, date_start) {
+  d.players <- read_csv(here(data_read_location, 'players.csv'), 
+                        col_types = cols(data.timeClick = col_number(),
+                                         data.rawScore = col_number())) %>%
+    rename(playerId = `_id`) %>% 
+    filter(createdAt >= date_start) %>%
+    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
+    select(-c(clicked:clickLength)) 
+  return(d.players)
+}
+
 preprocess_dataset <- function(data_name, date_start = lubridate::ymd('2021-01-01')) {
   data_read_location = here("data/raw_data", data_name)
   data_write_location = here("data/processed_data", data_name)

### 2022-06-10 code/prep_annotate.Rmd
process data for annotating flower referents

--- 2021-09-06 code/aa_mic_submission_analyses.Rmd
@@
 ---
-title: "aa_mic_submission_analyses"
-author: "Jessica Mankewitz"
-date: "9/6/2021"
+title: "prep_annotating"
+author: "vboyce"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+knitr::opts_chunk$set(echo = F)
 
-This notebook reads in the preprocessed and joined data, runs the basic processing/summarizing code, and generates figures and numbers for the 2021 MIC workshop submission.
-
-```{r setup}
 library(here)
@@
 #for plotting
-source(here("code/theme_specs.R"))
 
@@
 n_rounds = 24
-games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=n_players*n_rounds) %>% select(gameId)
+
+game_completion <- d.rounds %>% 
+  group_by(gameId) %>% 
+  tally() %>% 
+  mutate(game_complete = n==n_players*n_rounds) %>%
+  rename(n_trials = n)
+
+d.games.final <- d.games %>% 
+  left_join(game_completion) %>% 
+  replace_na(list(game_complete = FALSE, n_trials = 0))
+
+d.games.conditions <- d.games.final %>% 
+  select(gameId, color, condition, chatEnabled, game_complete) %>% 
+  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
+         fullCondition = paste(condition, languageCondition, sep = "-"))
 
 d.rounds.final <-  d.rounds %>% 
-  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  filter(conditionName != "coopCartel") %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId))
+  left_join(d.games.final %>% 
+              select(gameId, conditionName, condition,color, chatEnabled, game_complete)) %>% 
+  filter(conditionName != "coopCartel",
+         game_complete)
 
 d.raw_chat.final <- d.raw_chat %>% 
-  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId)) 
+  left_join(d.games.final %>% 
+              select(gameId, conditionName, condition, game_complete)) %>%
+  filter(game_complete)
+
 ```
 
-# Summary Stats
-## Game counts
+# Want to find a "best" game to start with
+
 ```{r}
-d.games %>% 
-  mutate(game_complete = !(gameId %in% games_exclude$gameId)) %>%
-  group_by(-game_complete, condition, chatEnabled) %>%
-  count()
+d.rounds.final %>% filter(game_complete) %>% 
+  group_by(gameId) %>% summarise(utils=sum(playerUtility)) %>% arrange(desc(utils))
+
+# let's look at gameId == QHkfso2yMgbbziE7r first
 ```
 
-## Game lengths
-## Bonus Payments
+```{r}
 
-# Points x Time x Num. Words
+good_rounds <- d.rounds.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
+  select(playerId,playerResponse,roundId,gameId) %>% 
+  separate(playerResponse, into=c("color","pResp"))
 
-# Avg. Points Per Round
+good_contexts <- d.contexts %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
+  select(roundId, gameId, label, utility) %>% 
+  separate(label, into=c("color", "num")) %>% 
+  mutate(labelutil=str_c(num,":",utility)) %>% 
+  group_by(roundId, gameId, color) %>% 
+  arrange(desc(utility)) %>% 
+  summarize(flowers=str_c(labelutil,collapse=", "))
 
-# Language Reduction
+good_chat <- d.raw_chat.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
+  left_join(d.players %>% select(playerId, name)) %>% 
+  rename(roundId=roundID) %>% 
+  left_join(good_contexts) %>% 
+  left_join(good_rounds) %>% 
+  mutate(rowid=row_number()) %>% 
+  select(rowid,trialNum,name, pResp, text, flowers, color) %>% 
+  filter(!is.na(text)) %>% 
+  write_tsv("test.tsv")
 
-# Georgia Analyses
+```

### 2022-06-14 code/prep_annotate.Rmd
play analysis from 1 group

@@
 knitr::opts_chunk$set(echo = F)
+knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
 
@@
 #for plotting
-
+theme_set(theme_bw())
 #Data import constants
@@
 ```
+
+# Sample analysis
+
+To test that we're annotating usefully, try to do initial analysis on sample annotation.
+
+Next steps:
+- add in post-test reference expressions
+- sbert and look at convergence / divergence
+- inter-group differences/ correlations
+
+```{r}
+sample <- read_tsv("post_annotate_sample.tsv")
+
+words <- sample %>% filter(!is.na(span)) %>% 
+  mutate(numchar=str_length(span),
+         numword=str_count(span, "\\W+") + 1)
+```
+
+How many words of referring language are there per round?
+
+```{r}
+
+sum_words <- words %>% group_by(trialNum) %>% summarize(sum_char=sum(numchar),
+                                           sum_word=sum(numword))
+
+  ggplot(sum_words,aes(x=trialNum,y=sum_char))+geom_point()+geom_smooth()
+  ggplot(sum_words,aes(x=trialNum,y=sum_word))+geom_point()+geom_smooth()
+
+```
+
+How long are individual referring expressions?
+
+```{r}
+
+  ggplot(words,aes(x=trialNum,y=numchar))+geom_point()+geom_smooth()
+  ggplot(words,aes(x=trialNum,y=numword))+geom_point()+geom_smooth()
+
+```
+
+How many referring expressions are there?
+```{r}
+  ggplot(words,aes(x=trialNum))+geom_bar()
+
+ggplot(words,aes(x=as.factor(referent)))+geom_bar()
+```
+
+```{r}
+
+  ggplot(words,aes(x=trialNum,y=numchar, color=as.factor(referent)))+geom_point()+geom_smooth(method="lm",se=F)
+  ggplot(words,aes(x=trialNum,y=numword, color=as.factor(referent)))+geom_point()+geom_smooth(method="lm", se=F)
+
+```

### 2022-06-21 code/prep_annotate.Rmd
sample graphs from 1 annotated game

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F)
+knitr::opts_chunk$set(echo = F, message=F, warning=F)
 knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
@@
 ```
+See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing (purple flowers) for the flower number - image correspondence.
+
+```{r}
+
+sample %>% filter(!is.na(span)) %>% select(trialNum, name, referent, span) %>% arrange(referent) %>% knitr::kable(align="l")
+```

### 2022-06-23 code/prep_annotate.Rmd
prep all text for annotations

@@
 read_data_path <- "data/processed_data/joined_data/"
+annotation_path <- "data/annotations"
 ```
@@
   select(rowid,trialNum,name, pResp, text, flowers, color) %>% 
+  filter(!is.na(text))
+
+```
+# Write all games to tsv for coding
+```{r}
+
+good_rounds <- d.rounds.final %>% 
+  select(playerId,playerResponse,roundId,gameId) %>% 
+  separate(playerResponse, into=c("color","pResp"))
+
+good_contexts <- d.contexts %>% 
+  select(roundId, gameId, label, utility) %>% 
+  separate(label, into=c("color", "num")) %>% 
+  mutate(labelutil=str_c(num,":",utility)) %>% 
+  group_by(roundId, gameId, color) %>% 
+  arrange(desc(utility)) %>% 
+  summarize(flowers=str_c(labelutil,collapse=", "))
+
+good_chat <- d.raw_chat.final %>% 
+  left_join(d.players %>% select(playerId, name)) %>% 
+  rename(roundId=roundID) %>% 
+  left_join(good_contexts) %>% 
+  left_join(good_rounds) %>% 
+  mutate(rowid=row_number(),
+         span="",
+         referent="",) %>% 
+  select(rowid,trialNum,name, pResp, text,span, referent, flowers, color) %>% 
   filter(!is.na(text)) %>% 
-  write_tsv("test.tsv")
+  write_tsv(here(annotation_path, "pre_annotation.tsv"))
 

### 2022-06-23 code/prep_annotate.Rmd
sample for inter-annotator agreement

@@
 ```{r}
-d.rounds.final %>% filter(game_complete) %>% 
-  group_by(gameId) %>% summarise(utils=sum(playerUtility)) %>% arrange(desc(utils))
+d.rounds.final %>% filter(game_complete) %>% filter(chatEnabled) %>% 
+  group_by(gameId, conditionName,condition, ) %>% summarise(utils=sum(playerUtility)) %>% arrange(desc(utils))
 
@@
          referent="",) %>% 
-  select(rowid,trialNum,name, pResp, text,span, referent, flowers, color) %>% 
+  select(gameId,rowid,trialNum,name, pResp, text,span, referent, flowers, color) %>% 
   filter(!is.na(text)) %>% 
@@
 
+```
+# Choose games for agreement coding
+
+```{r}
+agreement_sample <- good_chat %>% filter(gameId %in% c("Rq47E8uzWFKm4CBPJ","Etbyk29DvPTqmSJM7")) %>% write_tsv(here(annotation_path, "pre_agreement.tsv"))
 ```

### 2022-09-22 code/prep_annotate.Rmd
add full annotations, qc on them

@@
 ```
+# Interannotator agreement
+
+```{r}
+vb <- read_csv(here(annotation_path, "vb_agreement.csv")) %>% rename(vb_span=span,vb_referent=referent) %>% 
+  mutate(vb_span=ifelse(str_starts(vb_span, "the "), str_sub(vb_span, 5, -1), vb_span)) %>% 
+  mutate(row=row_number())
+bl <- read_csv(here(annotation_path, "bl_agreement.csv")) %>% rename(bl_span=span,bl_referent=referent) %>% 
+  mutate(bl_span=str_sub(bl_span, 2,-2)) %>% #remove quote marks
+  mutate(row=row_number())
+
+both <- vb %>% full_join(bl) %>% 
+  mutate(same_span=ifelse(vb_span==bl_span, 1,0),
+         same_ref=ifelse(vb_referent==bl_referent, 1,0)) %>% 
+  select(text,vb_span, bl_span, vb_referent,bl_referent, same_span, same_ref)
+
+differ <- both %>% 
+  filter(same_span==0)
+
+count <- both %>% group_by(same_ref) %>% tally()
+
+```
+
+# Quality control
+We can run some sanity checks on annotations --
+
+First - the identified flower should be one of the available options.
+Second - the span should (usually) be a substring of the text
+
+```{r}
+annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
+  filter(!is.na(span)) |> 
+  mutate(flower_list=str_split(flowers,", "),
+         flower_list=map(flower_list,~str_replace(.,":.+","")),
+         referent_char=as.character(referent),
+         referent_good=map2(referent_char,flower_list, ~.x %in% .y),
+         span_good=map2(text,span, ~str_detect(.x,coll(.y))))
+
+filter(annotated,referent_good==F) |> View() # Note: originally 35, caught some errors, one that remains I believe is actually talking about an image from prior round!
+
+filter(annotated,span_good==F) |> View() #Note: dowm from more that were incorrectly copied. The ones left are either as a result of duplicate spacing not copying
+# or intentional when someone wrote description "is my best" continued description
+
+```
 

### 2022-09-22 code/prep_annotate.Rmd
add sbert of annotations

@@
 
-```{r}
+```{r, include=F}
 agreement_sample <- good_chat %>% filter(gameId %in% c("Rq47E8uzWFKm4CBPJ","Etbyk29DvPTqmSJM7")) %>% write_tsv(here(annotation_path, "pre_agreement.tsv"))
@@
 
-```{r}
+```{r, include=F}
 vb <- read_csv(here(annotation_path, "vb_agreement.csv")) %>% rename(vb_span=span,vb_referent=referent) %>% 
@@
 
-filter(annotated,referent_good==F) |> View() # Note: originally 35, caught some errors, one that remains I believe is actually talking about an image from prior round!
+#filter(annotated,referent_good==F)  |> View() # Note: originally 35, caught some errors, one that remains I believe is actually talking about an image from prior round!
 
-filter(annotated,span_good==F) |> View() #Note: dowm from more that were incorrectly copied. The ones left are either as a result of duplicate spacing not copying
+#filter(annotated,span_good==F)  |> View() #Note: dowm from more that were incorrectly copied. The ones left are either as a result of duplicate spacing not copying
 # or intentional when someone wrote description "is my best" continued description
@@
 
-# Sample analysis
+** Note ** : there is probably inconsistency in how dubiously referential things like "same as last time" or "that one" are coded. Actually descriptions are probably well coded.
+All flowers were identified, some with lower justification/certainty than others.
 
-To test that we're annotating usefully, try to do initial analysis on sample annotation.
+# Sbert prep
 
-Next steps:
-- add in post-test reference expressions
-- sbert and look at convergence / divergence
-- inter-group differences/ correlations
+It's unclear how we want to do sbert -- one way would be to group by speaker+target and concatenate, the other would be to respect newlines.
 
 ```{r}
-sample <- read_tsv("post_annotate_sample.tsv")
+pre_sbert <- annotated |> filter(!is.na(normalized_span)) |> 
+  select(gameId, trialNum,name,normalized_span, referent,color) |> 
+  write_csv(here(annotation_path, "pre_sbert.csv"))
 
-words <- sample %>% filter(!is.na(span)) %>% 
-  mutate(numchar=str_length(span),
-         numword=str_count(span, "\\W+") + 1)
-```
-
-How many words of referring language are there per round?
-
-```{r}
-
-sum_words <- words %>% group_by(trialNum) %>% summarize(sum_char=sum(numchar),
-                                           sum_word=sum(numword))
-
-  ggplot(sum_words,aes(x=trialNum,y=sum_char))+geom_point()+geom_smooth()
-  ggplot(sum_words,aes(x=trialNum,y=sum_word))+geom_point()+geom_smooth()
+pre_sbert_concat <- pre_sbert |> group_by(gameId, trialNum, name, referent, color) |> 
+    summarize(normalized_span=str_c(normalized_span, collapse=" ")) |> 
+  write_csv(here(annotation_path, "pre_sbert_concat.csv"))
 
@@
 
-How long are individual referring expressions?
-
-```{r}
-
-  ggplot(words,aes(x=trialNum,y=numchar))+geom_point()+geom_smooth()
-  ggplot(words,aes(x=trialNum,y=numword))+geom_point()+geom_smooth()
-
+```{r, include=F, eval=F}
+library(reticulate)
+np <- import("numpy")
+mat = np$load(here(annotation_path,'post_sbert.npy'))
+saveRDS(mat,here(annotation_path,'post_sbert.RData'))
 ```
 
-How many referring expressions are there?
-```{r}
-  ggplot(words,aes(x=trialNum))+geom_bar()
-
-ggplot(words,aes(x=as.factor(referent)))+geom_bar()
+```{r, include=F, eval=F}
+library(reticulate)
+np <- import("numpy")
+mat = np$load(here(annotation_path,'post_sbert_concat.npy'))
+saveRDS(mat,here(annotation_path,'post_sbert_concat.RData'))
 ```
-
-```{r}
-
-  ggplot(words,aes(x=trialNum,y=numchar, color=as.factor(referent)))+geom_point()+geom_smooth(method="lm",se=F)
-  ggplot(words,aes(x=trialNum,y=numword, color=as.factor(referent)))+geom_point()+geom_smooth(method="lm", se=F)
-
-```
-See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing (purple flowers) for the flower number - image correspondence.
-
-```{r}
-
-sample %>% filter(!is.na(span)) %>% select(trialNum, name, referent, span) %>% arrange(referent) %>% knitr::kable(align="l")
-```

### 2022-09-22 code/annotation_analysis.Rmd
first pass at analysis

---
title: "Flowers text analysis"
author: "vboyce"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))

library(here)
library(tidyverse)

#for plotting
theme_set(theme_bw())
#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
annotation_path <- "data/annotations"
```

# VB to do notes
- Sbert
- compare with post-games stuff (and sbert that as well)
- consider being more consistent about what "the same one" or "that one" or etc got annotated
- inter-group differences/ correlations
- correlates of success

# Intro

I have annotated the transcripts of the full games marking what parts are referential (descriptive of a flower) and which flower (as best I can tell) it refers to.

See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing for the flower number - image correspondences.

# Analysis

```{r}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName)

annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
  filter(!is.na(span)) |> 
  mutate(numchar=str_length(span),
         numword=str_count(span, "\\W+") + 1) |> 
  left_join(d.games)
```

## How many words of referring language are there per round?

```{r}

sum_words <- annotated %>% group_by(trialNum, gameId,conditionName) %>% summarize(sum_char=sum(numchar),
                                           sum_word=sum(numword))

  ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
    #geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)+
    geom_smooth(aes(group=gameId), method="lm", se=F)+
    geom_jitter(alpha=.5)

    ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
    geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)

      ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    #geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)+
    geom_smooth(aes(group=gameId), method="lm", se=F)+
    geom_jitter(alpha=.5)

    ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)

```
There are reduction trends (not sure if linear fit is correct or not) for most groups, some serious group differences, also noise.

## How long are individual referring expressions?

```{r}

  ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
  ggplot(annotated,aes(x=trialNum,y=numword))+geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)

```
Individual referring expressions tend to get shorter.

## How many referring expressions are there?
```{r}

annotated |> group_by(gameId, trialNum, conditionName) |> tally() |> 
  ggplot(aes(x=trialNum, y=n))+geom_jitter()+geom_smooth()+facet_grid(.~conditionName)

annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
  group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
```
Mostly the number of referring expressions is relatively constant, but the expressions themselves shorten.

# Sbert stuff
(currently copied from tangrams stuff)
```{r, include=F,}
sbert_concat <- read_csv(here(annotation_path,"pre_sbert_concat.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_concat.RData')))  %>% as_tibble()

```

```{r, include=F,}
sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()

```

```{r, include=F}

F_mat <- sbert_concat %>% select(starts_with("V")) %>% as.matrix() #Features
M_mat <- sbert_concat %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
```

```{r helpers}
# note: cor expects features to be in columns so we transpose
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_within_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          .$repNum)) %>%
    mutate(rep1 = as.numeric(dim1), 
           rep2 = as.numeric(dim2)) 
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}
```

### 2022-10-13 code/annotation_analysis.Rmd
start modeling

@@
 author: "vboyce"
-output: html_document
+output: html_document:
+    toc: true
 ---
@@
 library(tidyverse)
-
+library(rstanarm)
+library(rstan)
+library(tidybayes)
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
 #for plotting
@@
 annotation_path <- "data/annotations"
+model_path <- "code/models"
+
+show_summary_rstan <- function(model){
+  badlist <- c("accept_stat__","divergent__","energy__","n_leapfrog__","sigma", "stepsize__","treedepth__")
+  intervals <- gather_draws(model, `[a-zA-z_()]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
+    mutate(lower=round(.lower, digits=2),
+           upper=round(.upper, digits=2),
+           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
+           Term=.variable,
+           Estimate=round(.value, digits=2)) %>% 
+    select(Term, Estimate, `Credible Interval`)
+
+  intervals
+}
+
+show_summary <- function(model){
+  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
+
+  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
+    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
+    group_by(.variable) %>% 
+    summarize(pct_above_0=mean(above_0)) %>% 
+   mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
+    left_join(intervals, by=".variable") %>% 
+    mutate(lower=round(.lower, digits=2),
+           upper=round(.upper, digits=2),
+           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
+           Term=str_sub(.variable, 3, -1),
+           Estimate=round(.value, digits=2)) %>% 
+    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)
+
+  stats
+}
+
+stats <- function(model, row){
+  str_c(model[row,1],": Est=", model[row,2], ", CrI=", model[row,3])
+}
+
+stats_text  <- function(model, row){
+  str_c( model[row,2], " (CrI=", model[row,3],")")
+}
 ```
 
-# VB to do notes
-- Sbert
-- compare with post-games stuff (and sbert that as well)
-- consider being more consistent about what "the same one" or "that one" or etc got annotated
-- inter-group differences/ correlations
-- correlates of success
-
-# Intro
+# Note
 
@@
 See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing for the flower number - image correspondences.
+
+# Cogsci notes
+
+Key points:
+
+- Many of the key reduction findings from tangrams generalize to this situation.
+Specifically, we see utterance reduction over time and w/i group convergence for each image and divergence between images.
+This situation is different in that we have different stimuli (more natural) and the set up is collaborative and more free-form in what is talked about.
+These patterns hold for both the individual and group payoff structures.
+
+- One difference we see is that groups don't diverge (from each other). This may be dependent on stimulus properities (are there universal features of some of the images?) and group dynamics
+
+- Conclusion: The key reference game findings have some generalizability.
+Settings like this one may be useful for encouraging discussion of a set of images and setting up partial knowledge situations.
 
@@
          numword=str_count(span, "\\W+") + 1) |> 
-  left_join(d.games)
+  left_join(d.games) |> 
+  mutate(flower=str_c(referent,color),
+         player=str_c(gameId, name)) |> 
+  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, conditionName)
 ```
@@
 
-  ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
-    #geom_line(aes(group=gameId))+
-    facet_grid(.~conditionName)+
-    geom_smooth(aes(group=gameId), method="lm", se=F)+
-    geom_jitter(alpha=.5)
+  # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
+  #   #geom_line(aes(group=gameId))+
+  #   facet_grid(.~conditionName)+
+  #   geom_smooth(aes(group=gameId), method="lm", se=F)+
+  #   geom_jitter(alpha=.5)
 
-    ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
-    geom_line(aes(group=gameId))+
-    facet_grid(.~conditionName)
-
+    # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
+    # geom_line(aes(group=gameId))+
+    # facet_grid(.~conditionName)
+    # 
+    # 
       ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
@@
 
-    ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
-    geom_line(aes(group=gameId))+
-    facet_grid(.~conditionName)
+    # ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
+    # geom_line(aes(group=gameId))+
+    # facet_grid(.~conditionName)
+    #
 
@@
 
-  ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
+ # ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
   ggplot(annotated,aes(x=trialNum,y=numword))+geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
@@
   ggplot(aes(x=trialNum, y=n))+geom_jitter()+geom_smooth()+facet_grid(.~conditionName)
-
-annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
-  group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
+# 
+# annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
+#   group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
 ```
@@
 
+## Model
+
+```{r,eval=F, include=F}
+
+m1 <- stan_lmer(numword ~ trialNum + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
+
+```
+
+```{r}
+m1 <- read_rds(here(model_path,"m1.rds"))
+m1
+#prior_summary(m1)
+show_summary_rstan(m1)
+```
+
 # Sbert stuff
-(currently copied from tangrams stuff)
+
+Some pre-commentary on analytic choices:
+- it's unclear whether we should be looking at referential statements individually (i.e.
+treating anything that occured as a separate utterance, after a line break, as separate) or concatenating everything Laju said about flower yellow3 into one statement.
+I think the first is generally better since sometimes a flower is described twice (perhaps in the same way), as in "the best values are 3 in a row and big fluffy" ...
+"okay, I'll take big fluffy".
+But there are also times when someone has a line break before continuing/clarifying a description.
+For now we go with single utterances.
+
+- Unlike tangrams where there was a very clear cut every item gets described once every round by every group, here we have no such guarantees!
+So there's a big question of what scale to use for "time block".
+In some within game cases it might make sense to compare utterances within a single round (for distinctiveness of different descriptions or if multiple describe the same flower differently).
+It also might make sense to treat round as the semi-continuous variable that it is for some modelling where we can eventually see that farther away pairs do something.
+But for now, I'm going to divide the game into quarters (sets of 6 rounds) and pretend that we have 4 time points.
+This is arbitrary and preliminary but it's easy to do viz with.
+
 ```{r, include=F,}
@@
 ```{r, include=F,}
-sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()
+d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
+                      col_types = cols()) %>% distinct() |> 
+    mutate(conditionName = condition, 
+         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
+                               conditionName == "competCartel" ~ "Individual Utilities",
+                               conditionName == "coopCartel" ~ "Old Cooperative",
+                               TRUE ~ conditionName)) |> 
+  select(gameId,conditionName)
+
+sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble() |> 
+  left_join(d.games)
 
@@
 
-F_mat <- sbert_concat %>% select(starts_with("V")) %>% as.matrix() #Features
-M_mat <- sbert_concat %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
+F_mat <- sbert_single %>% select(starts_with("...")) %>% as.matrix() #Features
+M_mat <- sbert_single %>% select(-starts_with("...")) %>% mutate(feature_ind=row_number()) |> 
+  mutate(quarter=case_when(
+    trialNum < 6 ~ 1,
+    trialNum < 12 ~ 2,
+    trialNum < 18 ~ 3,
+    trialNum < 24 ~ 4
+  ))
 ```
@@
 ```
+
+## Do flower labels diversify within games?
+
+Hypothesis: descriptions for different flowers get more different over time (within game)
+
+```{r }
+
+flower_distinctive <- M_mat %>% 
+  group_by(gameId,quarter, conditionName) %>% 
+  mutate(combinedId=referent) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(referent1!=referent2) %>% 
+  ungroup()
+```
+
+```{r}
+ggplot(flower_distinctive, aes(x=quarter,y=sim,color=conditionName))+
+      stat_summary(aes(group=str_c(gameId), color=conditionName),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
+```
+
+## Do groups diverge on labels?
+
+Hypothesis: descriptions for the same flower between games get more different over time (maybe, for some flowers)
+
+```{r }
+
+game_distinctive <- M_mat %>% 
+  group_by(quarter, color, referent) %>% 
+  mutate(combinedId=gameId) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  rename(gameId1=dim1,gameId2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(gameId1!=gameId2) %>% 
+  ungroup()
+```
+
+```{r}
+ggplot(game_distinctive, aes(x=quarter,y=sim,color=color))+
+      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
+```
+
+Wide variability -- some flowers might get more consistent names than others?
+
+## Do group members converge with each other?
+
+Hypothesis: players within the same game will label the same flower more similarly later in the game than earlier.
+
+```{r }
+
+player_similar <- M_mat %>% 
+  group_by(quarter, color, referent, gameId) %>% 
+  mutate(combinedId=name) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  rename(name1=dim1,name2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(name1!=name2) %>% 
+  ungroup()
+```
+
+```{r}
+ggplot(player_similar, aes(x=quarter,y=sim,color=color))+
+      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
+```

### 2022-10-13 code/prep_annotate.Rmd
do sbert on post-quiz names

@@
 ```
+
+# Sbert prep on end stuff
+
+```{r}
+library(jsonlite)
+
+ParseJSONColumn <- function(x) {
+  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
+    fromJSON(flatten = T)
+}
+d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
+                      col_types = cols()) %>% distinct() |> 
+  select(playerId,own=flower_labels,other=otherflower_labels) |> 
+  inner_join(d.rounds.final |> select(gameId, playerId, conditionName, color) |> unique()) |> 
+  mutate(own= ifelse(is.na(own), '{}', own)) %>%
+  mutate(other = ifelse(is.na(other), '{}', other)) %>%
+  mutate(own = map(own, .f = ParseJSONColumn)) %>% 
+  mutate(other =  map(other, .f = ParseJSONColumn)) %>% 
+  unnest(cols=c(own,other), names_sep=":") %>% 
+  pivot_longer(cols=starts_with("own")|starts_with("other")) %>% 
+  filter(!is.na(value)) %>% 
+  filter(value!="") %>% 
+  separate(name, into=c("seen","color","num")) %>% 
+  rename(text=value) %>% 
+  mutate(text = gsub("[/?/.]", ' ', text),
+         text = str_squish(text),
+         utt_length_chars = str_length(text), 
+         utt_length_words = str_count(text, "\\W+") + 1) |> write_csv(here(annotation_path,"post_test.csv"))
+
+```
+
+```{r, include=F, eval=F}
+library(reticulate)
+np <- import("numpy")
+mat = np$load(here(annotation_path,'post_sbert_post_test.npy'))
+saveRDS(mat,here(annotation_path,'post_sbert_post_test.RData'))
+```

### 2022-10-13 code/annotation_analysis.Rmd
poke at analysis of end game labels

@@
 ```
+
+# Within end-stuff
+
+```{r}
+sbert_post_test <- read_csv(here(annotation_path,"post_test.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData')))  %>% as_tibble()
+
+F_mat <- sbert_post_test %>% select(starts_with("...")) %>% as.matrix() #Features
+M_mat <- sbert_post_test %>% select(-starts_with("...")) %>% mutate(feature_ind=row_number(), flower=str_c(color,num))
+```
+
+## Within flower
+```{r}
+
+end_distinctive <- M_mat %>% 
+  group_by(flower, color) %>% 
+  mutate(combinedId=str_c(gameId,"_",seen)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("gameId1", "seen1")) |> 
+  separate(dim2, c("gameId2", "seen2")) |> 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  ungroup()
+```
+
+```{r}
+
+end_distinctive |> mutate(same_group=ifelse(gameId1==gameId2, 
+                                            "same_group","different_group"),
+                          same_color=case_when(
+  seen1=="own" & seen2=="own" ~ "own",
+  seen2=="other" & seen1=="other" ~ "other",
+  T ~"mixed"
+))|> ggplot(aes(group=flower, x=color, y=sim, color=str_c(same_group,same_color)))+     
+  stat_summary(aes(group=str_c(same_group, same_color)),fun.data = "mean_cl_boot", position = position_dodge(width=.3), geom="pointrange")
+
+```
+This is really poorly labelled. Needs stats.
+
+## Between flowers
+```{r}
+
+end_between <- M_mat %>% 
+  group_by(color) %>% 
+  mutate(combinedId=str_c(playerId, "_",gameId,"_",seen, "_", flower)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("player1","gameId1", "seen1" ,"flower1")) |> 
+  separate(dim2, c("player2", "gameId2", "seen2", "flower2")) |> 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(flower1!=flower2) |> 
+  ungroup()
+```
+
+```{r}
+
+end_between |> mutate(same_group=case_when(
+  player1==player2 ~ "same_person",
+  gameId1==gameId2 ~ "same_group",
+  T ~ "diferent_group"
+), same_color=case_when(
+  seen1=="own" & seen2=="own" ~ "own",
+  seen2=="other" & seen1=="other" ~ "other",
+  T ~"mixed"
+))|> ggplot(aes(x=color, y=sim, color=str_c(same_group,same_color)))+     
+  stat_summary(aes(group=str_c(same_group, same_color)),fun.data = "mean_cl_boot", position = position_dodge(width=.3), geom="pointrange")+facet_grid(.~same_group)
+
+```
+
+# TODO: between game & end-game

### 2022-10-14 code/annotation_analysis.Rmd
work on helper funcs for game x end sim (not entirely working!)

@@
 author: "vboyce"
-output: html_document:
+output:
+  html_document:
     toc: true
+    toc_depth: 2
 ---
@@
 
-make_within_df <- function(M_mat, F_mat, method) {
-  M_mat %>%
-    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
-                          .$repNum)) %>%
-    mutate(rep1 = as.numeric(dim1), 
-           rep2 = as.numeric(dim2)) 
-}
-
 make_across_df <- function(M_mat, F_mat, method) {
@@
 
+### TODO model
 ## Do groups diverge on labels?
@@
 Wide variability -- some flowers might get more consistent names than others?
+### TODO model
 
@@
 ```
+### TODO model
 
@@
 This is really poorly labelled. Needs stats. 
+### TODO models
 
@@
 ```
+### TODO models
 
 # TODO: between game & end-game
+
+```{r}
+
+d.raw_chat <- read_csv(here(paste(read_data_path, 
+                                  "raw_chat.csv", 
+                                  sep="")),
+                      col_types = cols()) %>% distinct() |> 
+  select(gameId, playerId) |> unique()
+
+d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
+                      col_types = cols()) %>% distinct() |> select(playerId,name) |> inner_join(d.raw_chat)
+
+sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble() |> 
+   left_join(d.players)
+```
+
+```{r to-end-helpers}
+
+#here we only want sims between an in-game utt and an end description
+# note: cor expects features to be in columns so we transpose
+
+pairwiseCosine <- function(matrix1,matrix2) {
+    apply(matrix1,2,function(x){
+        apply(matrix2,2,lsa::cosine,x)
+    })
+}
+get_sim_matrix = function(df, F_mat, method = 'cosine') {
+  df_game <- df |> filter(type=="game")
+  df_end <- df |> filter(type=="end")
+  feats_game = F_mat[df_game$feature_ind,]
+  print(t(feats_game))
+  feats_end = F_mat[df_end$feature_ind,]
+  print(t(feats_end))
+  if(method == 'cor') {
+    return(cor(t(feats_game), t(feats_end), method = 'pearson'))
+  }  else if (method == 'cosine') {
+    return(as.matrix(pairwiseCosine(t(feats_game), t(feats_end))))
+  } else {
+    stop(paste0('unknown method', method))
+  }
+}
+
+# note this does de-duplicated version
+flatten_sim_matrix <- function(cormat, ids_game, ids_end) {
+  data.frame(
+    dim1 = ids_end[row(cormat)],
+    dim2 = ids_game[col(cormat)],
+    sim  = as.numeric(cormat)
+  ) %>%
+    mutate(dim1 = as.character(dim1),
+           dim2 = as.character(dim2))
+}
+
+make_across_df <- function(M_mat, F_mat, method) {
+  M_mat %>%
+    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
+                          as.character(filter(., type=="game")$combinedId), as.character(filter(., type=="end")$combinedId)))
+}
+
+```
+
+** Does not work with grouped data frames, works with non-grouped dfs ** 
+```{r}
+#test
+test_df1 <- tibble(color=c("blue","green"),id=c("game1","game2"), val1=c(.5,.7), val2=c(.9,.4)) |> mutate(type="game")
+test_df2 <- tibble(color=c("blue", "green"),id=c("end1", "end2"), val1=c(1,-1), val2=c(-1,-1)) |> mutate(type="end")
+test <- test_df1 |> union(test_df2) 
+F_mat <- test %>% select(starts_with("val")) %>% as.matrix() #Features
+M_mat <- test%>% select(-starts_with("val")) %>% mutate(feature_ind=row_number(), combinedId=id) |> group_by(color)
+
+a <- make_across_df(M_mat, F_mat, method="cosine")
+
+# want there to be one thing that gets grouped (M_mat) and then we separate it by type
+
+sol <- tibble(dim1=c("end1","end2","end1","end2"), 
+              dim2=c("game1", "game1", "game2","game2"),
+              sim=c(lsa::cosine(c(.5,.9),c(1,-1)),
+                    lsa::cosine(c(.5, .9), c(-1,-1)),
+                    lsa::cosine(c(.7,.4),c(1,-1)),
+                    lsa::cosine(c(.7,.4),c(-1,-1))))
+```
+
+# TODO: Correlates of success
+
+# TODO: scale of variation between groups

### 2022-11-16 code/annotation_analysis.Rmd
work on draft cogsci figs, much sbert

--- 2022-10-13 code/annotation_analysis.Rmd
@@
 library(tidybayes)
+library(viridis)
+library(cowplot)
 rstan_options(auto_write = TRUE)
@@
 
-# Analysis
+# Prep
 
@@
                                TRUE ~ conditionName)) |> 
-  select(gameId,conditionName)
+  select(gameId,conditionName, condition)
 
@@
          player=str_c(gameId, name)) |> 
-  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, conditionName)
+  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)
 ```
 
-## How many words of referring language are there per round?
+# How much data
+
+```{r, message=T}
+
+message("total utterances")
+annotated |> tally()
+
+message("Utterances / game")
+
+annotated |> group_by(gameId,condition) |> tally() |> pull(n) |> summary()
+
+message("Utterances / flower")
+annotated |> group_by(flower) |> tally() |> pull(n) |> summary()
+
+```
+
+# Word count basics
+
+Take away -- across the condition manipulation, the number of utterances doesn't decrease, but the length of each one does.
+
+Graph could use finessing if going in paper.
+
+```{r, fig.height=4, fig.width=8}
+
+sum_words <- annotated %>% group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
+                                           sum_word=sum(numword))
+
+     total <-  ggplot(sum_words,aes(x=trialNum+1,y=sum_word, color=condition))+
+   # geom_smooth(method="lm", se=T)+
+            geom_smooth(method="lm", formula=y~poly(x,2), se=T)+
+
+        geom_jitter(alpha=.1, width=.2, height=.5)+
+        labs(y="Referential Words per trial per game", x="Trial Number", color="")+
+        theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+    legend <- get_legend(total)
+
+    all_words <- total+theme(legend.position = "none")
+
+        words <- ggplot(annotated,aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.1)+ geom_smooth()+
+          labs(y="Words per reference expression", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+        utts <- annotated |> group_by(gameId, trialNum, condition) |> tally() |> 
+  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.02)+geom_smooth()+
+          labs(y="Reference expressions  per trial per game", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+        top <- plot_grid(all_words, words, utts,nrow=1)
+        plot_grid(top, legend, nrow=2, rel_heights=c(1,.1))
+
+```
+
+# How many different flowers described / trial?
+
+We can probably give this as summary stats and not as graph
+
+Per game
 
@@
 
-sum_words <- annotated %>% group_by(trialNum, gameId,conditionName) %>% summarize(sum_char=sum(numchar),
-                                           sum_word=sum(numword))
+annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> 
+  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.2)+geom_smooth()+
+          labs(y="Distinct flowers referenced per trial", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+```
 
-  # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
-  #   #geom_line(aes(group=gameId))+
-  #   facet_grid(.~conditionName)+
-  #   geom_smooth(aes(group=gameId), method="lm", se=F)+
-  #   geom_jitter(alpha=.5)
-
-    # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
-    # geom_line(aes(group=gameId))+
-    # facet_grid(.~conditionName)
-    # 
-    # 
-      ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
-    #geom_line(aes(group=gameId))+
-    facet_grid(.~conditionName)+
-    geom_smooth(aes(group=gameId), method="lm", se=F)+
-    geom_jitter(alpha=.5)
-
-    # ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
-    # geom_line(aes(group=gameId))+
-    # facet_grid(.~conditionName)
-    #
-
-```
-There are reduction trends (not sure if linear fit is correct or not) for most groups, some serious group differences, also noise.
-
-## How long are individual referring expressions?
+Per person
 
 ```{r}
-
- # ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
-  ggplot(annotated,aes(x=trialNum,y=numword))+geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
-
+annotated |> group_by(gameId, trialNum, condition, player) |> select(flower) |> unique() |> tally() |> 
+  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.1)+geom_smooth()+
+          labs(y="Distinct flowers referenced per player per trial", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
 ```
-Individual referring expressions tend to get shorter.
-
-## How many referring expressions are there?
-```{r}
-
-annotated |> group_by(gameId, trialNum, conditionName) |> tally() |> 
-  ggplot(aes(x=trialNum, y=n))+geom_jitter()+geom_smooth()+facet_grid(.~conditionName)
-# 
-# annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
-#   group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
-```
-Mostly the number of referring expressions is relatively constant, but the expressions themselves shorten.
 
@@
 
-m1 <- stan_lmer(numword ~ trialNum + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
+m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
 
@@
 
-# Sbert stuff
+# Sbert notes
 
@@
 ```{r, include=F,}
-sbert_concat <- read_csv(here(annotation_path,"pre_sbert_concat.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_concat.RData')))  %>% as_tibble()
-
-```
-
-```{r, include=F,}
 d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
@@
                                TRUE ~ conditionName)) |> 
-  select(gameId,conditionName)
+  select(gameId,conditionName, condition)
 
-sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble() |> 
-  left_join(d.games)
+ d.raw_chat <- read_csv(here(paste(read_data_path, 
+                                   "raw_chat.csv", 
+                                   sep="")),
+                       col_types = cols()) %>% distinct() |> 
+   select(gameId, playerId) |> unique()
 
-```
+ d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
+                       col_types = cols()) %>% distinct() |> select(playerId,name) |> inner_join(d.raw_chat)
 
-```{r, include=F}
+ sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv"))|> 
+    left_join(d.players) |> left_join(d.games) |> mutate(seen="game") |> 
+   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()
 
-F_mat <- sbert_single %>% select(starts_with("...")) %>% as.matrix() #Features
-M_mat <- sbert_single %>% select(-starts_with("...")) %>% mutate(feature_ind=row_number()) |> 
-  mutate(quarter=case_when(
-    trialNum < 6 ~ 1,
-    trialNum < 12 ~ 2,
-    trialNum < 18 ~ 3,
-    trialNum < 24 ~ 4
+sbert_post_test <- read_csv(here(annotation_path,"post_test.csv")) |> left_join(d.games) |> 
+  mutate(trialNum=25) |> 
+  select(playerId, gameId, condition, text,referent=num, color, trialNum, seen)%>% bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData')))  %>% as_tibble()
+
+sbert_all <- sbert_single |> union(sbert_post_test)
+F_mat <- sbert_all %>% select(starts_with("...")) %>% as.matrix() #Features
+M_mat <- sbert_all %>% 
+  select(-starts_with("...")) %>% 
+  mutate(feature_ind=row_number(),
+         sixth=case_when(
+           trialNum < 4 ~ 1,
+          trialNum < 8 ~ 2,
+          trialNum < 12 ~ 3,
+          trialNum < 16 ~ 4,
+          trialNum < 20 ~ 5,
+          trialNum < 24 ~ 6,
+          trialNum > 24 ~ 7
   ))
@@
 
-## Do flower labels diversify within games?
+# Sbert pictures
 
-Hypothesis: descriptions for different flowers get more different over time (within game)
+These have quadratic smooths fit onto them, to allow for curvature, but no reason to choose this curve and not another.
+
+## During game
 
@@
 
-flower_distinctive <- M_mat %>% 
-  group_by(gameId,quarter, conditionName) %>% 
-  mutate(combinedId=referent) %>% 
+# by trial
+flower_all <- M_mat %>% 
+  filter(seen %in% c("game")) |> 
+  group_by(color,condition) %>% 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId)) %>% 
   make_across_df(F_mat, 'cosine') %>% 
-  rename(referent1=dim1,referent2=dim2) %>% 
+  separate(dim1, c("ref1", "trialNum1","game1","player1"), convert=T) |> 
+  separate(dim2, c("ref2", "trialNum2", "game2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
   mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
-  filter(referent1!=referent2) %>% 
+  #filter(ref1!=ref2) %>% 
+  mutate(earlier=ifelse(trialNum1>trialNum2,trialNum2, trialNum1) |> as.numeric(),
+         later=ifelse(trialNum1>trialNum2, trialNum1,trialNum2) |> as.numeric()) |> 
   ungroup()
@@
 ```{r}
-ggplot(flower_distinctive, aes(x=quarter,y=sim,color=conditionName))+
-      stat_summary(aes(group=str_c(gameId), color=conditionName),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
+flower_all|> 
+  filter(later-earlier<3) |> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same person",
+    game1==game2 ~ "same game",
+    T ~ "different game"),
+    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower")
+  ) |> 
+ggplot( aes(x=later,y=sim,color=condition))+     
+  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
+  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
 ```
 
-## Do groups diverge on labels?
+Comparing utterances within rounds that were up to 2 apart, round coded as the later (so 10-12,11-12 and 12-12 will all contribute to "12").
+Up to 2 apart is somewhat arbitrary, not sure what the right rolling window to do is.
 
-Hypothesis: descriptions for the same flower between games get more different over time (maybe, for some flowers)
+Take aways:
+ - descriptions to the same flower within game (either same person or not) become more similar
+ - descriptions to the same flower across games become (slightly) less similar
+ - descriptions to different flowers become less similar (regardless of between/within game)
+
+## Between game and end
+
+We can treat what someone wrote at the end as the convention and then compare how similar this is to earlier utterances by them and their group mates.
+
+Big question is whether the end utts need cleaning (which probably yes?)
+
+We see that for the same flower, similarity to the end utt increases over blocks both within individual and indiv - other group mate (this is all within a game)
 
@@
 
-game_distinctive <- M_mat %>% 
-  group_by(quarter, color, referent) %>% 
-  mutate(combinedId=gameId) %>% 
+# by trial
+to_end <- M_mat %>% 
+  filter(seen %in% c("game", "own")) |> 
+  group_by(gameId,color,condition) %>% 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",seen,"_",playerId)) %>% 
   make_across_df(F_mat, 'cosine') %>% 
-  rename(gameId1=dim1,gameId2=dim2) %>% 
+  separate(dim1, c("ref1", "trialNum1","seen1","player1"), convert=T) |> 
+  separate(dim2, c("ref2", "trialNum2", "seen2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
   mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
-  filter(gameId1!=gameId2) %>% 
+  filter(seen1!=seen2) %>% 
+  mutate(earlier=ifelse(trialNum1>trialNum2,trialNum2, trialNum1) |> as.numeric(),
+         later=ifelse(trialNum1>trialNum2, trialNum1,trialNum2) |> as.numeric()) |> 
+  ungroup()
+```
+
+Comparing utterances earlier to post-game descriptions.
+
+General question of how to display these things -- what should be color v faceting and what to do for the dots.
+
+```{r}
+to_end|> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same person",
+    T ~ "same game"),
+        same_flower=ifelse(ref1==ref2, "Same flower", "diff flower")
+) |> 
+ggplot( aes(x=earlier,y=sim,color=condition))+     
+  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
+  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+```
+
+##  Within end-stuff
+
+We can look at how converged people are (or are not) by comparing what they say at the end.
+
+Each person was asked for names ( ~ "how would you label this to your group") for the 12 images they were seeing and then for 4 from a different color palette.
+
+Own = color you saw all game
+
+Other = the color you didn't see
+
+Mixed = own for one person, other for other person
+
+```{r }
+
+# by trial
+in_end <- M_mat %>% 
+  filter(seen %in% c("other", "own")) |> 
+  group_by(color,condition) %>% 
+  mutate(combinedId=str_c(referent, "_",seen,"_", gameId, "_",playerId)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("ref1", "seen1","game1","player1"), convert=T) |> 
+  separate(dim2, c("ref2",  "seen2","game2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
   ungroup()
@@
 ```{r}
-ggplot(game_distinctive, aes(x=quarter,y=sim,color=color))+
-      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
+in_end|> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same person",
+    game1==game2 ~ "same game",
+    T ~ "different game"),
+    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower"),
+    same_color=case_when(
+  seen1=="own" & seen2=="own" ~ "own",
+  seen2=="other" & seen1=="other" ~ "other",
+  T ~"mixed"
+  ))|> 
+ggplot( aes(x=same_color,y=sim,color=condition))+     
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3))+
+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_color_brewer(palette="Dark2")
 ```
 
-Wide variability -- some flowers might get more consistent names than others?
+Not sure how to show spread here. Or how to do a viz that allows for comparisons.
 
-## Do group members converge with each other?
+Take aways:
 
-Hypothesis: players within the same game will label the same flower more similarly later in the game than earlier.
+- People use more similar descriptions for two flowers that weren't in their set than those that were (checks out -- in set ones have been described and diverged)
+- This same thing holds (with overall lower sims) for group-mates
+- Group mates are more similar in how they describe flowers they've seen (v not) (at least in shared utils)
+- not sure how to interpret the diff game -- this is the baseline? Maybe it's just that you're more similar for the same flower versus different regardless? (also this is baseline for other panels)
 
-```{r }
+# Future things
 
-player_similar <- M_mat %>% 
-  group_by(quarter, color, referent, gameId) %>% 
-  mutate(combinedId=name) %>% 
-  make_across_df(F_mat, 'cosine') %>% 
-  rename(name1=dim1,name2=dim2) %>% 
-  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
-  filter(name1!=name2) %>% 
-  ungroup()
-```
+- have we tried log space? 
+- Could do more end game -- earlier comparisons
+- tSNE??
+- Models of (some of ?) what's shown in forgoing graphs
+- CLIP ????
+- clean up end utts? 
+- are there interesting lang -- performance connections ?
 
-```{r}
-ggplot(player_similar, aes(x=quarter,y=sim,color=color))+
-      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
-```
+## Boring diligence
+- Double check that non-talkers aren't in this data at all (end game joins). Consider coding and grouping end stuff by what color they usually saw during the game
+- do sbert a different way and confirm (split v aggregate?)

### 2022-12-02 write-ups/cogsci23_paper/manuscript.Rmd
start manuscript

---
title: "TODO"
bibliography: [library.bib, aaflowers.bib]
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA
    \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\ Department of Educational Psychology, 1025 W. Johnson Street \\ Madison, WI 53706 USA}

abstract: >
    TODO

keywords: >
    Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
library(here)
library(tidyverse)
library(rstanarm)
library(rstan)
library(tidybayes)
library(viridis)
library(cowplot)
#rstan_options(auto_write = TRUE)
#options(mc.cores = parallel::detectCores())
#for plotting
theme_set(theme_bw())

#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
annotation_path <- "data/annotations"
model_path <- "code/models"
```

TODO introduction

# Methods
This paper presents a reanalysis of data from @mankewitz2021. We describe both the original data collection and the additional data processing done for reanalysis.

## Data Acquisition
@mankewitz2021 had participants play a real-time coordination game in groups of three, implemented using Empirica [@almaatouqEmpiricaVirtualLab2020]. On each trial, each group saw a set of 6 flower images. Each participant could see the values for four of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant. Players could coordinate and discuss using a chat box, and each selected a flower. If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they got a low reward instead. Thus, participants were incentivized to coordinate on selecting different flowers. The rewards earned over the course of game translated in a monetary bonus for the participants at the end of the game.

In *individual utility* games, each player earned points for the flowers they selected; in the *shared utility* games, the points earned were averaged together, and all players in a game got the same reward. This made for slightly different incentives; in an individual game, players want to maximize the rewards of flowers they select, and only care about avoiding collisions with other players' selections; in a shared game, players want their teammates to select different high reward flowers, and are indifferent on who selects the highest one.

Each game was assigned a color of flower (white, red, yellow, purple) and the flower images were drawn from a set of 12 for that color, so players saw the same flowers repeatedly across the game, in different combinations. Each game consisted of 24 trials. The use of different flowers of the same color was to create a situation where players needed to refer to the images but didn't have preset names for them in the context, and thus would need to collaboratively develop shared referring expressions.

After the game, players were asked how they would describe each of the images they had played with to their teammates. They were also asked how they would describe each of 4 images from a different color set than the one they had used.

TODO number of games going into this data -- Mankewitz says 18 indiv and 21 shared

## Textual annotations

We annotated the chat transcripts to extract all referring expressions and identify which flower image each expression referred to. TODO interannotator agreements.

We then embedded each of these referring expressions using SBERT cite! in order to compare between referring expressions in semantic space. We use cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.

# Results

## Reduction of referring expressions

## Convergence of referring expressions

## End expressions

# Discussion

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
#img <- png::readPNG("figs/walrus.png")
#grid::grid.raster(img)
```

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
#img <- png::readPNG("figs/lab_logo_stanford.png")
#grid::grid.raster(img)
```

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
                aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                       caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2022-12-07 write-ups/cogsci23_paper/manuscript.Rmd
draft intro

@@
 title: "TODO"
-bibliography: [library.bib, aaflowers.bib]
+bibliography: [ aaflowers.bib]
 csl: apa6.csl
@@
 library(tidyverse)
-library(rstanarm)
-library(rstan)
+#library(rstanarm)
+#library(rstan)
 library(tidybayes)
@@
 TODO introduction
+
+So people often care about reference or whatever.
+Referring expressions are crucial to interactive use of language so we can know what the interlocuter is talking about.
+Many goals of conversation rely on joint reference as a part of them -- for instance in order to negotiate, even from an adversarial perspective, you still need to have terms that refer to the same things in order to do the negotiation.
+Joint reference is a cooperative act, needed for both cooperative and non-cooperative acts.
+
+In some cases joint reference is easy b/c there are conventional meanings of words that refer to the objects.
+However, this isn't always the case, either because objects are novel and don't have names or because objects are similar to one another and so while they might be nameable out of context, in context there are not conventional names.
+
+How people develop ad-hoc shared conventional referring expressions for objects without conventional names has primarily been studied in the realm of dyadic reference games.
+Two participants see a set of images, and the speaker sees a particular one highlighted as the one to pick out next (or they have them in the correct order and the listener needs to reorder theirs to match).
+This is a task explicitly about referring to the objects that is highly structured.
+There are assigned roles (one person has all the knowledge) and a clear indication of which one to talk about.
+Selecting or ordering objects is pretty simple.
+
+This has been a very useful microcosm for studying referring expressions, expectations about conventions, and what happens with new listeners or speakers.
+The key phenomena observed across these experiments are that pairs develop partner specific referring expressions -- the utterances produced by the speaker to refer to objects get shorter over repetition, the accuracy-speed trade off improves (?
+maybe don't mention this b/c it's not what we look at), and the terms that develop are partner specific -- people converge on terms within partnerships, but differ between partnerships on the terms used.
+
+Implicitly, these features are thought to be about how referring expressions develop across some wider range of situations where people interact repeatedly in ways that need reference to some objects that do not already have conventional names that are sufficient for reference.
+Basically it needs to be a situation where people don't at the outset have a high degree of agreement on how to refer to the thing.
+
+Needs more transition here.
+
+@mankewitz2021 tried to bridge the gap between datasets of negotiation (like let's make a deal?) where there is negotiation over easily nameable objects, and the reference game literature where there is reference (but no negotiation) to hard to name objects.
+They created a 3 person game where players had to select what flowers to grow from a set of 6 shown on the screen.
+The values of the flowers were partially hidden, and flowers were all the same color (but different shapes/species).
+Each player could select a flower, but it was only worth the shown value if one player selected it; if multiple did, they got a much lower reward.
+Flower images repeated across the game.
+@mankewitz intended this set-up so that players would need to refer the flowers in order to negotiate who would take what, but in a context where there wasn't a flower to talk about and there wasn't a designated speaker.
+They also had two conditions, one where players within a group won points together (and thus had fully aligned incentives) and another where they won points as individuals (and didn't).
+The paper found ....
+a slight decrease in language over the course of hte game, possibly consistent with reduced referring expressions, but also consistent with developing a negotiation pattern.
+
+This builds on the ref game literature in a few dimensions.
+The key difference is the more free-form format where reference is implicitly motivated by the task, rather than being an explicit part of the task, and the equal positions of the players (each has the same amount of knowledge) rather than an asymmetric teacher/student.
+Other differences include that the images are natural photos, which has been used in ref games (CITE that management paper), but is not the typical and the presence of more than two participants (which is done, but not modal, citations!).
+
+Here we extract the referring expressions from the chat logs of @mankewitz in order to look at reduction as well as semantic trends.
+We see whether the phenomena found in classic dyadic iterated reference games extend to this more naturalistic domain, and whether they differ across the individual and shared incentive conditions.
+
+TODO list of key points.
 
@@
 
-We annotated the chat transcripts to extract all referring expressions and identify which flower image each expression referred to. TODO interannotator agreements.
+We annotated the chat transcripts to extract all referring expressions and identify which flower image each expression referred to.
+We also spellchecked and corrected the referring expressions.
+Annotations were done primarily by the first author, with some done by a research assistant.
+A fraction of the games were coded by both to assess interannotator agreement -- on ZZ expressions, there was agreement on YY on what the referring expression was, and on XX out of QQ on which flower was being described.
+TODO interannotator agreements.
 

### 2022-12-07 write-ups/cogsci23_paper/manuscript.Rmd
start results text

@@
 
-TODO introduction
+# introduction
 
@@
 
+We extracted a total of ZZ referring expressions,
+
 We then embedded each of these referring expressions using SBERT cite!
@@
 
+We assess the list of key points given at the beginning.
+
 ## Reduction of referring expressions
+First thing to check is the key phenomena that over the course of iterated references, the referring expressions get shorter.
+We see that in Fig ??.
+Across both conditions, the number of referring words per trial/game decreases, which is driven by a decrease in the length of each referring expression.
+The number of referring expressions remains constant (although shows a group difference).
+
+On average, in each trial, ZZ distinct flowers were referred to (this stayed relatively constant) with most trials having 2 or 3 referred to ()
+TODO! double check that this is handling 0s appropriately!!
+Each player usualy referred to one, with a minority referring to 2 or 3.
+This is consistent with a strategy where players say which flower they are selecting; higher numbers probably involve more coordination and info sharing.
+
+TODO include model!
 
@@
 # Discussion
+
+Many of the key reduction findings from tangrams generalize to this situation.
+Specifically, we see utterance reduction over time and w/i group convergence for each image and divergence between images.
+This situation is different in that we have different stimuli (more natural) and the set up is collaborative and more free-form in what is talked about.
+These patterns hold for both the individual and group payoff structures.
+
+One difference we see is that groups don’t diverge (from each other). This may be dependent on stimulus properities (are there universal features of some of the images?) and group dynamics
+
+Conclusion: The key reference game findings have some generalizability. Settings like this one may be useful for encouraging discussion of a set of images and setting up partial knowledge situations.
 

### 2022-12-07 code/prep_annotate.Rmd
work on draft cogsci figs, much sbert
rerun sbert post test

@@
 
-#filter(annotated,span_good==F)  |> View() #Note: dowm from more that were incorrectly copied. The ones left are either as a result of duplicate spacing not copying
+#filter(annotated,span_good==F) #Note: dowm from more that were incorrectly copied. The ones left are either as a result of duplicate spacing not copying
 # or intentional when someone wrote description "is my best" continued description
@@
 
+```{r}
+end <- read_csv(here(annotation_path,"post_test_annotated.csv")) |> 
+  filter(!is.na(spellchecked)) |> write_csv(here(annotation_path,"post_test_pre_sbert.csv"))
+
+```
+
 ```{r, include=F, eval=F}

### 2022-12-07 code/annotation_analysis.Rmd
wrangling adding zeros and code cleanup

@@
 ```{r}
-d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
-                      col_types = cols()) %>% distinct() |> 
+d.games <- read_csv(here(read_data_path, "games.csv")) |> 
                         mutate(conditionName = condition, 
@@
                                TRUE ~ conditionName)) |> 
+  filter(chatEnabled) |> 
   select(gameId,conditionName, condition)
 
+n_players = 3
+n_rounds = 24
+
+d.rounds <- read_csv(here(read_data_path,"rounds.csv"))
+d.players <- read_csv(here(read_data_path, "players.csv")) |> select(playerId,name)
+game_completion <- d.rounds %>% 
+  group_by(gameId) %>% 
+  tally() %>% 
+  mutate(game_complete = n==n_players*n_rounds) %>%
+  select(gameId, game_complete)
+
+d.games.final <- d.games %>% 
+  left_join(game_completion) %>% 
+  replace_na(list(game_complete = FALSE, n_trials = 0)) |> 
+  filter(game_complete)
+
+d.rounds.final <-  d.rounds %>% 
+  inner_join(d.games.final) %>% 
+  filter(conditionName != "coopCartel",
+         game_complete) |> 
+  left_join(d.players) |> 
+  #left_join(d.games) |> 
+  select(gameId,trialNum, name, condition)
+
 annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
-  filter(!is.na(span)) |> 
+  filter(!is.na(normalized_span)) |> 
+  inner_join(d.games.final) |> 
   mutate(numchar=str_length(span),
-         numword=str_count(span, "\\W+") + 1) |> 
-  left_join(d.games) |> 
+         numword=ifelse(numchar==0,0,str_count(span, "\\W+") + 1)) |> 
   mutate(flower=str_c(referent,color),
          player=str_c(gameId, name)) |> 
-  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)
+  select(gameId, player,name, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)
+
 ```
@@
 message("total utterances")
-annotated |> tally()
+annotated |> filter(numword>0) |>  tally()
 
@@
 
-annotated |> group_by(gameId,condition) |> tally() |> pull(n) |> summary()
+annotated |> filter(numword>0) |> group_by(gameId,condition) |> tally() |> pull(n) |> summary()
 
 message("Utterances / flower")
-annotated |> group_by(flower) |> tally() |> pull(n) |> summary()
+annotated |> filter(numword>0) |> group_by(flower) |> tally() |> pull(n) |> summary()
 
@@
 
-sum_words <- annotated %>% group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
+sum_words <- annotated %>% filter(numword>0) |>  group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
                                            sum_word=sum(numword))
@@
 
-        words <- ggplot(annotated,aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.1)+ geom_smooth()+
+        words <- ggplot(annotated |> filter(numword>0),aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.1)+ geom_smooth()+
           labs(y="Words per reference expression", x="Trial Number")+
@@
 
-        utts <- annotated |> group_by(gameId, trialNum, condition) |> tally() |> 
+        utts <- annotated |> filter(numword>0) |> group_by(gameId, trialNum, condition) |> tally() |> 
   ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.02)+geom_smooth()+
@@
 ```{r}
+possible <- d.rounds.final |> select(-name) |> unique()
 
-annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> 
+annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
+  mutate(n=ifelse(is.na(n),0,n)) |> 
   ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.2)+geom_smooth()+
@@
 ```{r}
-annotated |> group_by(gameId, trialNum, condition, player) |> select(flower) |> unique() |> tally() |> 
+
+annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
+  mutate(n=ifelse(is.na(n),0,n)) |>
+
   ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.1)+geom_smooth()+
@@
 ```{r, include=F,}
-d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
-                      col_types = cols()) %>% distinct() |> 
-    mutate(conditionName = condition, 
-         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
-                               conditionName == "competCartel" ~ "Individual Utilities",
-                               conditionName == "coopCartel" ~ "Old Cooperative",
-                               TRUE ~ conditionName)) |> 
-  select(gameId,conditionName, condition)
 
@@
  sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv"))|> 
-    left_join(d.players) |> left_join(d.games) |> mutate(seen="game") |> 
-   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()
+    left_join(d.players) |> left_join(d.games.final) |> mutate(seen="game") |> 
+   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) |> bind_cols(readRDS(here(annotation_path,'post_sbert.RData'))  %>% as_tibble())
 
-sbert_post_test <- read_csv(here(annotation_path,"post_test.csv")) |> left_join(d.games) |> 
+sbert_post_test <- read_csv(here(annotation_path,"post_test_pre_sbert.csv")) |> 
   mutate(trialNum=25) |> 
-  select(playerId, gameId, condition, text,referent=num, color, trialNum, seen)%>% bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData')))  %>% as_tibble()
+ bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData'))  %>% as_tibble()) |> inner_join(d.games) |> 
+   select(playerId, gameId, condition, text,referent=num, color, trialNum, seen, starts_with("V"))
 
@@
 M_mat <- sbert_all %>% 
-  select(-starts_with("...")) %>% 
+  select(-starts_with("V")) %>% 
   mutate(feature_ind=row_number(),

### 2022-12-08 code/annotation_analysis.Rmd
start adding figs

@@
 library(tidyverse)
-library(rstanarm)
+#library(rstanarm)
 library(rstan)
@@
 
+possible <- d.rounds.final |> select(-name) |> unique()
+
 sum_words <- annotated %>% filter(numword>0) |>  group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
-                                           sum_word=sum(numword))
+                                           sum_word=sum(numword)) |> full_join(possible) |> 
+  mutate(sum_char=ifelse(is.na(sum_char),0,sum_char),
+         sum_word=ifelse(is.na(sum_word),0,sum_word))
 
@@
 
-        utts <- annotated |> filter(numword>0) |> group_by(gameId, trialNum, condition) |> tally() |> 
-  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.02)+geom_smooth()+
+        utts <- annotated |> filter(numword>0) |>
+          group_by(gameId, trialNum, condition) |> tally() |> full_join(possible) |> 
+          mutate(n=ifelse(is.na(n),0,n)) |> 
+  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.05)+geom_smooth()+
           labs(y="Reference expressions  per trial per game", x="Trial Number")+
@@
 
+```{r}
+possible <- d.rounds.final |> select(-name) |> unique()
+
+per_game <-  annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+
+mean_flowers <- per_game |> group_by(condition) |> summarize(mean=mean(n))
+
+per_game |> group_by(condition, n) |> tally() |> rename("Distinct flowers"=n, Trials=nn) |> 
+  pivot_wider(names_from=condition, values_from=Trials)
+
+per_person <- annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+
+mean_person <- per_person |> group_by(condition) |> summarize(mean=mean(n))
+
+per_person |> group_by(condition, n) |> tally() |> rename("Distinct flowers"=n, Trials=nn) |> 
+  pivot_wider(names_from=condition, values_from=Trials)
+```
 Per person
@@
 
-```{r}
+```{r, eval=F, include=F}
 m1 <- read_rds(here(model_path,"m1.rds"))
@@
 sbert_all <- sbert_single |> union(sbert_post_test)
-F_mat <- sbert_all %>% select(starts_with("...")) %>% as.matrix() #Features
+F_mat <- sbert_all %>% select(starts_with("V")) %>% as.matrix() #Features
 M_mat <- sbert_all %>% 
   select(-starts_with("V")) %>% 
-  mutate(feature_ind=row_number(),
-         sixth=case_when(
-           trialNum < 4 ~ 1,
-          trialNum < 8 ~ 2,
-          trialNum < 12 ~ 3,
-          trialNum < 16 ~ 4,
-          trialNum < 20 ~ 5,
-          trialNum < 24 ~ 6,
-          trialNum > 24 ~ 7
-  ))
+  mutate(feature_ind=row_number())
 ```
@@
 
-```{r }
+```{r}
 
@@
   group_by(color,condition) %>% 
-  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId)) %>% 
-  make_across_df(F_mat, 'cosine') %>% 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId))
+  make_across_df(F_mat, 'cosine') 
   separate(dim1, c("ref1", "trialNum1","game1","player1"), convert=T) |> 

### 2022-12-08 write-ups/cogsci23_paper/manuscript.Rmd
start adding figs

@@
 library(cowplot)
+library(lsa)
 #rstan_options(auto_write = TRUE)
@@
 model_path <- "code/models"
+
+stats <- function(model, row){
+  str_c(model[row,1],": Est=", model[row,2], ", CrI=", model[row,3])
+}
+
+stats_text  <- function(model, row){
+  str_c( model[row,2], " (CrI=", model[row,3],")")
+}
+```
+
+```{r}
+d.games <- read_csv(here(read_data_path, "games.csv")) |> 
+                        mutate(conditionName = condition, 
+         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
+                               conditionName == "competCartel" ~ "Individual Utilities",
+                               conditionName == "coopCartel" ~ "Old Cooperative",
+                               TRUE ~ conditionName)) |> 
+  filter(chatEnabled) |> 
+  select(gameId,conditionName, condition)
+
+n_players = 3
+n_rounds = 24
+
+d.rounds <- read_csv(here(read_data_path,"rounds.csv"))
+d.players <- read_csv(here(read_data_path, "players.csv")) |> select(playerId,name)
+game_completion <- d.rounds %>% 
+  group_by(gameId) %>% 
+  tally() %>% 
+  mutate(game_complete = n==n_players*n_rounds) %>%
+  select(gameId, game_complete)
+
+d.games.final <- d.games %>% 
+  left_join(game_completion) %>% 
+  replace_na(list(game_complete = FALSE, n_trials = 0)) |> 
+  filter(game_complete)
+
+d.rounds.final <-  d.rounds %>% 
+  inner_join(d.games.final) %>% 
+  filter(conditionName != "coopCartel",
+         game_complete) |> 
+  left_join(d.players) |> 
+  #left_join(d.games) |> 
+  select(gameId,trialNum, name, condition)
+
+annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
+  filter(!is.na(normalized_span)) |> 
+  inner_join(d.games.final) |> 
+  mutate(numchar=str_length(span),
+         numword=ifelse(numchar==0,0,str_count(span, "\\W+") + 1)) |> 
+  mutate(flower=str_c(referent,color),
+         player=str_c(gameId, name)) |> 
+  select(gameId, player,name, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)
+
 ```
@@
 # Methods
+
+TODO do methods get pictures?
+
 This paper presents a reanalysis of data from @mankewitz2021. We describe both the original data collection and the additional data processing done for reanalysis.
@@
 
-TODO number of games going into this data -- Mankewitz says 18 indiv and 21 shared
+Following @Mankewitz2021 we only included games where players finished all 24 trials in the game.
+We only included games where participants had access ot the chat box (@Mankewitz2021 also had a no-chat baseline condition).
+This left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the Individual Utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the Shared Utilities Condition.
 
 ## Textual annotations
+
+```{r, include=F}
+vb <- read_csv(here(annotation_path, "vb_agreement.csv")) %>% rename(vb_span=span,vb_referent=referent) %>% 
+  mutate(vb_span=ifelse(str_starts(vb_span, "the "), str_sub(vb_span, 5, -1), vb_span)) %>% 
+  mutate(row=row_number())
+bl <- read_csv(here(annotation_path, "bl_agreement.csv")) %>% rename(bl_span=span,bl_referent=referent) %>% 
+  mutate(bl_span=str_sub(bl_span, 2,-2)) %>% #remove quote marks
+  mutate(row=row_number())
+
+both <- vb %>% full_join(bl) %>% 
+  mutate(same_span=ifelse(vb_span==bl_span, 1,0),
+         same_ref=ifelse(vb_referent==bl_referent, 1,0)) %>% 
+  select(text,vb_span, bl_span, vb_referent,bl_referent, same_span, same_ref) |> 
+  filter(!(is.na(vb_span)&is.na(bl_span)))
+
+raw_rows <- vb |> select(text) |> nrow()
+vb |> select(gameId) |> unique() |> nrow()
+
+agree_span <- both %>% group_by(same_span) |> tally() |> filter(same_span==1) |> pluck(2)
+
+count <- both %>% filter(!is.na(bl_referent))
+
+agree_target <- count |> group_by(same_ref) %>% tally() |> filter(same_ref==1) |> pluck(2)
+
+```
 
@@
 Annotations were done primarily by the first author, with some done by a research assistant.
-A fraction of the games were coded by both to assess interannotator agreement -- on ZZ expressions, there was agreement on YY on what the referring expression was, and on XX out of QQ on which flower was being described.
-TODO interannotator agreements.
 
-We extracted a total of ZZ referring expressions,
+Two games consisting of `r raw_rows` utterances were annotated by both annotators.
+`r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
+The second annotator coded the target of the referring expression for `r nrow(count)` of the utterances.
+Of these, the two annotators agreed on the target in `r agree_target` (`r round(agree_target/nrow(count)*100)`%) of cases.
+We take this high level of interannotator agreement as an indication that the reference spans and targets were identified in a reliable way.
+
+We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions.
 
@@
 ## Reduction of referring expressions
-First thing to check is the key phenomena that over the course of iterated references, the referring expressions get shorter.
-We see that in Fig ??.
-Across both conditions, the number of referring words per trial/game decreases, which is driven by a decrease in the length of each referring expression.
-The number of referring expressions remains constant (although shows a group difference).
+```{r wordcount, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
+A: Total words of referring language per game per trial.
+B: Words in each reference expression by trial.
+C: Total number of reference expressions per game per trial.", fig.height=2, fig.width=7}
 
-On average, in each trial, ZZ distinct flowers were referred to (this stayed relatively constant) with most trials having 2 or 3 referred to ()
-TODO! double check that this is handling 0s appropriately!!
-Each player usualy referred to one, with a minority referring to 2 or 3.
-This is consistent with a strategy where players say which flower they are selecting; higher numbers probably involve more coordination and info sharing.
+possible <- d.rounds.final |> select(-name) |> unique()
+
+sum_words <- annotated %>% filter(numword>0) |>  group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
+                                           sum_word=sum(numword)) |> full_join(possible) |> 
+  mutate(sum_char=ifelse(is.na(sum_char),0,sum_char),
+         sum_word=ifelse(is.na(sum_word),0,sum_word))
+
+     total <-  ggplot(sum_words,aes(x=trialNum+1,y=sum_word, color=condition))+
+   # geom_smooth(method="lm", se=T)+
+            geom_smooth(method="lm", formula=y~poly(x,2), se=T)+
+
+        geom_jitter(alpha=.05, width=.2, height=.5)+
+        labs(y="Words in reference \n expressions per trial", x="Trial Number", color="")+
+        theme(legend.position="bottom", legend.title=element_blank(), axis.title = element_text(size=8))+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")+
+        guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
+
+    legend <- get_legend(total)
+
+    all_words <- total+theme(legend.position = "none")
+
+        words <- ggplot(annotated |> filter(numword>0),aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.01)+ geom_smooth()+
+          labs(y="Words per reference", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank(), axis.title = element_text(size=8))+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+        utts <- annotated |> filter(numword>0) |>
+          group_by(gameId, trialNum, condition) |> tally() |> full_join(possible) |> 
+          mutate(n=ifelse(is.na(n),0,n)) |> 
+  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.05)+geom_smooth()+
+          labs(y="References per trial", x="Trial Number")+
+        theme(legend.position="none", legend.title=element_blank(), axis.title = element_text(size=8))+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+        top <- plot_grid(all_words, words, utts,nrow=1, labels="AUTO")
+        plot_grid(top, legend, nrow=2, rel_heights=c(1,.15))
+
+```
+
+```{r}
+possible <- d.rounds.final |> select(-name) |> unique()
+
+per_game <-  annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+
+mean_flowers <- per_game |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
+  select(condition, summ)
+
+per_person <- annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+mean_person <- per_person |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
+  select(condition, summ)
+```
+Consistent with the results from dyadic iterated reference games, the amount of referring language decreases over the course of the game.
+As shown in Fig @ref(fig:wordcount)A, the number of words of referring language decreases across the game in both conditions.
+This is due to the individual referring expressions getting shorter (@ref(fig:wordcount)B), while the total number of referring epxressions remains that same (@ref(fig:wordcount)C).
 
@@
 
+Another coarse metric for how referring expressions are being produced is to look at how many different flowers are talked about each round.
+There are 6 visible on the screen, and participants are incentivized to each pick a different one.
+In general, each trial contained references to 2 or 3 distinct flowers (Individual Utilities: mean of `r  mean_flowers |> pluck(2,1)`, Shared Utilities: `r mean_flowers |> pluck(2,2)`).
+Most players referred to one flower each round (Individual Utilities: `r  mean_person |> pluck(2,1)`, Shared Utilities: `r mean_person |> pluck(2,2)`).
+
+This pattern is consistent with each person saying what flower they plan on choosing; some games talked less than this, and others talked more as players queried the worths of various flowers or confirmed the plan of who would pick what.
+
 ## Convergence of referring expressions
+```{r, include=F,}
+
+ d.raw_chat <- read_csv(here(paste(read_data_path, 
+                                   "raw_chat.csv", 
+                                   sep="")),
+                       col_types = cols()) %>% distinct() |> 
+   select(gameId, playerId) |> unique()
+
+ d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
+                       col_types = cols()) %>% distinct() |> select(playerId,name) |> inner_join(d.raw_chat)
+
+ sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv"))|> 
+    left_join(d.players) |> left_join(d.games.final) |> mutate(seen="game") |> 
+   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) |> bind_cols(readRDS(here(annotation_path,'post_sbert.RData'))  %>% as_tibble())
+
+sbert_post_test <- read_csv(here(annotation_path,"post_test_pre_sbert.csv")) |> 
+  mutate(trialNum=25) |> 
+ bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData'))  %>% as_tibble()) |> inner_join(d.games) |> 
+   select(playerId, gameId, condition, text,referent=num, color, trialNum, seen, starts_with("V"))
+
+sbert_all <- sbert_single |> union(sbert_post_test)
+F_mat <- sbert_all %>% select(starts_with("V")) %>% as.matrix() #Features
+M_mat <- sbert_all %>% 
+  select(-starts_with("V")) %>% 
+  mutate(feature_ind=row_number())
+```
+
+```{r helpers}
+# note: cor expects features to be in columns so we transpose
+get_sim_matrix = function(df, F_mat, method = 'cosine') {
+  feats = F_mat[df$feature_ind,]
+  if(method == 'cor') {
+    return(cor(t(feats), method = 'pearson'))
+  } else if (method == 'euclidean') {
+    return(as.matrix(dist(feats, method = 'euclidean')))
+  } else if (method == 'cosine') {
+    return(as.matrix(lsa::cosine(t(feats))))
+  } else {
+    stop(paste0('unknown method', method))
+  }
+}
+
+# note this does de-duplicated version
+flatten_sim_matrix <- function(cormat, ids) {
+  ut <- upper.tri(cormat)
+  data.frame(
+    dim1 = ids[row(cormat)[ut]],
+    dim2 = ids[col(cormat)[ut]],
+    sim  = as.numeric(cormat[ut])
+  ) %>%
+    mutate(dim1 = as.character(dim1),
+           dim2 = as.character(dim2))
+}
+
+make_across_df <- function(M_mat, F_mat, method) {
+  M_mat %>%
+    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
+                          as.character(.$combinedId)))
+}
+```
+
+```{r, cache=T}
+
+# by trial
+flower_all <- M_mat %>% 
+  filter(seen %in% c("game")) |> 
+  group_by(color,condition) %>% 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("ref1", "trialNum1","game1","player1"), convert=T) |> 
+  separate(dim2, c("ref2", "trialNum2", "game2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  #filter(ref1!=ref2) %>% 
+  mutate(earlier=ifelse(trialNum1>trialNum2,trialNum2, trialNum1) |> as.numeric(),
+         later=ifelse(trialNum1>trialNum2, trialNum1,trialNum2) |> as.numeric()) |> 
+  ungroup()
+```
+
+```{r during, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
+A: Total words of referring language per game per trial.
+B: Words in each reference expression by trial.
+C: Total number of reference expressions per game per trial.", fig.height=6, fig.width=7}
+flower_all|> 
+  filter(later-earlier<3) |> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same person",
+    game1==game2 ~ "same game",
+    T ~ "different game"),
+    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower")
+  ) |> 
+ggplot( aes(x=later,y=sim,color=condition))+     
+  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
+  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")
+
+```
 

### 2022-12-09 code/annotation_analysis.Rmd
some model summaries

@@
 library(tidyverse)
-#library(rstanarm)
+library(rstanarm)
 library(rstan)
@@
   badlist <- c("accept_stat__","divergent__","energy__","n_leapfrog__","sigma", "stepsize__","treedepth__")
-  intervals <- gather_draws(model, `[a-zA-z_()]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
+  intervals <- gather_draws(model, `[a-zA-z_(): ]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
     mutate(lower=round(.lower, digits=2),
@@
 m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
-
+show_summary_rstan(m1) |> write_rds(here(model_path,"word_model.rds"))
 ```
@@
 m1
-#prior_summary(m1)
-show_summary_rstan(m1)
+
 ```
@@
   group_by(color,condition) %>% 
-  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId))
-  make_across_df(F_mat, 'cosine') 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",gameId,"_",playerId)) |> 
+  make_across_df(F_mat, 'cosine') |> 
   separate(dim1, c("ref1", "trialNum1","game1","player1"), convert=T) |> 
@@
         scale_color_brewer(palette="Dark2")
+
+```
+### Model
+
+```{r, eval=F}
+within_game_flower <- flower_all |> 
+  filter(game1==game2) |> 
+  filter(ref1==ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(same_player=ifelse(player1==player2,1,0),
+         flower=str_c(color,ref1)) |> 
+  select(sim, condition, earlier, same_player, game1, flower )
+
+m_samegame_sameflower <-  stan_lmer(sim ~ condition*earlier*same_player + 
+                                      (1|game1)+(1|flower), data=within_game_flower) |> write_rds(here(model_path,"m2.rds"))
+
+show_summary_rstan(m_samegame_sameflower) |> write_rds(here(model_path,"samegame_sameflower_model.rds"))
+
+```
+
+```{r, eval=F}
+within_game_between_flower <- flower_all |> 
+  filter(game1==game2) |> 
+  filter(ref1!=ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(same_player=ifelse(player1==player2,1,0)) |> 
+  select(sim, condition, earlier, same_player, game1)
+
+m_samegame_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=within_game_between_flower) |> write_rds(here(model_path,"m3.rds"))
+
+show_summary_rstan(m_samegame_diffflower) |> write_rds(here(model_path,"samegame_diffflower_model.rds"))
+
+```
+
+```{r}
+between_game_within_flower <- flower_all |> 
+  filter(game1!=game2) |> 
+  filter(ref1==ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(flower=str_c(color,ref1)) |> 
+  select(sim, condition, earlier, flower)
+
+m_diffgame_sameflower <-  stan_glm(sim ~ condition*earlier, data=between_game_within_flower) |> write_rds(here(model_path,"m4.rds"))
+
+show_summary_rstan(m_diffgame_sameflower) |> write_rds(here(model_path,"diffgame_sameflower_model.rds"))
 

### 2022-12-09 write-ups/cogsci23_paper/manuscript.Rmd
some model summaries

@@
 ```
+
+# TODOs
+redraft intro & add citations
+add figs
+learn stan and run models
+redraft methods
+write analysis
+write discussion
+abstract & front matter
 
@@
 
+## SBERT embeddings
+
 We then embedded each of these referring expressions using SBERT cite!
@@
 We use cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
+
+For the pairwise comparisons of utterances during the game, we take the similarities for utterances that are within 2 trials of each other (ex.
+an utterance from trial 10 is compared with utterances from trials 8,9,10,11,12).
+We do this because different flowers were shown to different groups on different trials.
+This width is a compromise between having enough pairs of utterances and still being able to treat them as coming from one time point in the game.
 
@@
 ```
+
+```{r,eval=F, include=F}
+
+m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
+show_summary_rstan(m1) |> write_rds(here(model_path,"word_model.rds"))
+```
+
+```{r}
+
+word_count <- read_rds(here(model_path,"word_model.rds"))
+```
+
 Consistent with the results from dyadic iterated reference games, the amount of referring language decreases over the course of the game.
@@
 
+```{r}
+wi_g_wi_f <- read_rds(here(model_path,"samegame_sameflower_model.rds"))
+
+#wi_g_bet_f <- read_rds(here(model_path,"samegame_diffflower_model.rds"))
+
+#bet_g_wi_f <- read_rds(here(model_path,"diffgame_sameflower_model.rds"))
+```
+
+### Within games
+TODO we compare games within 2 rounds of each other (ex. trial 10 compared to 8, 9, 10, 11, 12)
+
+One of the key claims is that within a partnership, the referring expressions should converge for the same image, but get more different for different images.
+Convergence would be increasing similiarity (higher cosine similarity in later rounds), divergence would be decreasing similarity in later rounds).
+
+For the same flower within the same game, similarity increased over rounds (trial: `r stats_text(wi_g_wi_f,6)`).
+Utterances were more similar if they were produced by the same participant (`r stats_text(wi_g_wi_f,8)`), although this did not interact with trial number (`r stats_text(wi_g_wi_f,7)`).
+Games in the shared utilities condition may have had higher similarities overall (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or same participant versus different participant (`r stats_text(wi_g_wi_f,5)`).
+
+same person 
+condition
+interactions w/ trial
+
+### Between games
+
 ## End expressions

### 2022-12-09 code/annotation_analysis.Rmd
more models

@@
 ```
+### Model
 
+```{r, eval=F}
+to_end_same_flower <- to_end |> 
+  filter(ref1==ref2) |> 
+  mutate(same_player=ifelse(player1==player2,1,0),
+         flower=str_c(color,ref1)) |> 
+  select(sim, condition, earlier, same_player, gameId, flower)
+
+m_to_end_sameflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_same_flower) |> write_rds(here(model_path,"m5.rds"))
+
+show_summary_rstan(m_to_end_sameflower) |> write_rds(here(model_path,"to_end_sameflower_model.rds"))
+
+```
+
+```{r, eval=F}
+to_end_diff_flower <- to_end |> 
+  filter(ref1!=ref2) |> 
+  mutate(same_player=ifelse(player1==player2,1,0)) |> 
+  select(sim, condition, earlier, same_player, gameId)
+
+m_to_end_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_diff_flower) |> write_rds(here(model_path,"m5.rds"))
+
+show_summary_rstan(m_to_end_diffflower) |> write_rds(here(model_path,"to_end_diffflower_model.rds"))
+
+```
 ##  Within end-stuff

### 2022-12-09 write-ups/cogsci23_paper/manuscript.Rmd
more models

@@
 stats_text  <- function(model, row){
-  str_c( model[row,2], " (CrI=", model[row,3],")")
+  str_c( model[row,2], " CrI=", model[row,3],"")
 }
@@
 add figs
-learn stan and run models
+consider rerunning wi_game bet flower with a mixed effect (was being slow, so didn't) and other models
+
 redraft methods
@@
 
-```{r during, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
-A: Total words of referring language per game per trial.
-B: Words in each reference expression by trial.
-C: Total number of reference expressions per game per trial.", fig.height=6, fig.width=7}
+```{r during, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Cosine similarities between SBERT embeddings of utterances produced within 2 trials of each other.
+Utterances were paired with other utterances from different games; from the same game, but a different utterer speaker; or from the same speaker (vertical panels).
+Utterances were either in reference to different flowers or both in reference to the same flower (horizontal panels).", fig.height=4, fig.width=10, out.width="100%"}
 flower_all|> 
@@
   mutate(distance=case_when(
-    player1==player2 ~ "Same person",
-    game1==game2 ~ "same game",
-    T ~ "different game"),
-    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower")
+    player1==player2 ~ "Same Person",
+    game1==game2 ~ "Same Game",
+    T ~ "Different Games"),
+    same_flower=ifelse(ref1==ref2, "Same Flower", "Different Flowers")
   ) |> 
@@
         scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        scale_color_brewer(palette="Dark2")
+        scale_color_brewer(palette="Dark2")+
+  labs(x="Trial Number", y="Cosine Similarity")+
+  guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
 
@@
 
-#wi_g_bet_f <- read_rds(here(model_path,"samegame_diffflower_model.rds"))
+wi_g_bet_f <- read_rds(here(model_path,"samegame_diffflower_model.rds"))
 
-#bet_g_wi_f <- read_rds(here(model_path,"diffgame_sameflower_model.rds"))
+bet_g_wi_f <- read_rds(here(model_path,"diffgame_sameflower_model.rds"))
 ```
@@
 ### Within games
+
 TODO we compare games within 2 rounds of each other (ex. trial 10 compared to 8, 9, 10, 11, 12)
@@
 
-same person 
-condition
-interactions w/ trial
+Different flowers within the same game have diverging descriptions that get less similar over time (trial: `r stats_text(wi_g_bet_f,6)`).
+Terms are overall slightly less similar in Shared Utility groups (`r stats_text(wi_g_bet_f,2)`), but this does not interact with trial (`r stats_text(wi_g_bet_f,3)`).
+There is not an effect of being said by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) or condition (`r stats_text(wi_g_bet_f,5)`).
 
@@
 
+Utterances describing the same flower in different games diverged slightly (trial: `r stats_text(bet_g_wi_f,4)`).
+There was a condition difference with shared utility conditions having more similar descriptions across games (`r stats_text(bet_g_wi_f,2)`); this did not interact with trial (`r stats_text(bet_g_wi_f,3)`).
+
+TODO: what is our possibly explanation for this?!
+
+This may be a fluke, as there are not a huge number of games in either condition, but if it's robust, it could be that the shared utility framing leads to less idiosyncratic descriptions and instead towards more universal priors.
+TODO this should be hedged *a lot* but we can also cite the study about large groupy things and the crab-like images 
+```{r }
+
+# by trial
+to_end <- M_mat %>% 
+  filter(seen %in% c("game", "own")) |> 
+  group_by(gameId,color,condition) %>% 
+  mutate(combinedId=str_c(referent,"_", trialNum, "_",seen,"_",playerId)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("ref1", "trialNum1","seen1","player1"), convert=T) |> 
+  separate(dim2, c("ref2", "trialNum2", "seen2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(seen1!=seen2) %>% 
+  mutate(earlier=ifelse(trialNum1>trialNum2,trialNum2, trialNum1) |> as.numeric(),
+         later=ifelse(trialNum1>trialNum2, trialNum1,trialNum2) |> as.numeric()) |> 
+  ungroup()
+```
+
+```{r toend, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.", fig.height=4, fig.width=6, out.width="50%"}
+to_end|> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same Person",
+    T ~ "Same Game - Different Person"),
+        same_flower=ifelse(ref1==ref2, "Same flower", "Different flower")
+) |> 
+ggplot( aes(x=earlier,y=sim,color=condition))+     
+  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
+  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        scale_color_brewer(palette="Dark2")+
+  guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
+
+```
+
+```{r}
+to_e_wi_f <- read_rds(here(model_path,"to_end_sameflower_model.rds"))
+
+to_e_bet_f <- read_rds(here(model_path,"to_end_diffflower_model.rds"))
+```
 ## End expressions
+
+### game to end
+As a complementary way of looking at how referring expressions change over time as conventions are formed, we can take the post-game flower descriptions (each player was asked to provide what they would call each flower to their teammates) as the conventions.
+
+Descriptions become more similar to the conventions over time  (trial: `r stats_text(to_e_wi_f,6)`).
+Utterances were more similar to the convention given by the same person (versu from a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
+Games in the shared utilities condition had utterances that were more similar to their conventions (`r stats_text(to_e_wi_f,2)`), but this did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
+
+There was not a significant divergence from the conventions for other flowers over time (trial: `r stats_text(to_e_bet_f,6)`).
+Terms are overall slightly less similar in Shared Utility groups (`r stats_text(to_e_bet_f,2)`), but this does not interact with trial (`r stats_text(to_e_bet_f,3)`).
+There is not an effect of being said by the same person (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
+Similarity to a different convention was slightly higher if was said by the same person in a Shared Utility group (`r stats_text(to_e_bet_f,5)`).
+
+### within end
+want to look for more similar with a group than between groups
+
+do we want to look at the "other" (flower not in game) at all? No?
 

### 2023-01-08 write-ups/cogsci23_paper/manuscript.Rmd
add figure of methods

@@
 date_start=lubridate::ymd('2021-07-19')
-fig_path=here("figs")
+fig_path=here("write-ups/cogsci23_paper/figs")
 read_data_path <- "data/processed_data/joined_data/"
@@
 
-# TODOs
-redraft intro & add citations
-add figs
-consider rerunning wi_game bet flower with a mixed effect (was being slow, so didn't) and other models
-
-redraft methods
-write analysis
-write discussion
-abstract & front matter
-
 # introduction
@@
 Needs more transition here. 
+```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Player interface.
+Panel A shows the selection phase, where each participant sees 6 flowers, 4 with value bars.
+Panel B shows the feedback stage, after all players have selected.
+When multiple players select the same flower, they recieve a lower value rather than what is shown.
+\\label{interface}", cache=FALSE}
+
+selection <- png::readPNG(here(fig_path, "selection.png"))
+
+feedback <- png::readPNG(here(fig_path, "feedback.png"))
+
+a <- ggdraw() + draw_image(selection)
+b <- ggdraw()+draw_image(feedback)
+
+plot_grid(a,b, labels="AUTO")
+```
 
@@
 
-TODO do methods get pictures?
-
 This paper presents a reanalysis of data from @mankewitz2021. We describe both the original data collection and the additional data processing done for reanalysis.
@@
 ## Data Acquisition
-@mankewitz2021 had participants play a real-time coordination game in groups of three, implemented using Empirica [@almaatouqEmpiricaVirtualLab2020].
-On each trial, each group saw a set of 6 flower images.
-Each participant could see the values for four of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant.
-Players could coordinate and discuss using a chat box, and each selected a flower.
-If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they got a low reward instead.
-Thus, participants were incentivized to coordinate on selecting different flowers.
+As described in @mankewitz2021, participants played a real-time coordination game in groups of three, implemented using Empirica [@almaatouqEmpiricaVirtualLab2020].
+On each trial, each group saw a set of 6 flower images (Figure \ref{interface}A) .
+Each participant saw the values for four of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant.
+Players could coordinate and discuss using a chat box before each selected a flower.
+If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they each got a lower reward instead (Figure \ref{interface}B).
+This incentivized participants to communicate about the flowers in order to coordinate on selecting differenct flowers.
 The rewards earned over the course of game translated in a monetary bonus for the participants at the end of the game.
@@
 In *individual utility* games, each player earned points for the flowers they selected; in the *shared utility* games, the points earned were averaged together, and all players in a game got the same reward.
-This made for slightly different incentives; in an individual game, players want to maximize the rewards of flowers they select, and only care about avoiding collisions with other players' selections; in a shared game, players want their teammates to select different high reward flowers, and are indifferent on who selects the highest one.
+This made for slightly different incentives; in an individual game, players wanted to maximize the rewards of flowers they selected, and only cared about avoiding collisions with other players' selections; in a shared game, players wanted their teammates to select different high reward flowers, and were indifferent on who selected the highest one.
 
@@
 Each game consisted of 24 trials.
-The use of different flowers of the same color was to create a situation where players needed to refer to the images but didn't have preset names for them in the context, and thus would need to collaboratively develop shared referring expressions.
+The use of different flowers of the same color created situations where players did not have established names for the flowers in context and needed to develop shared referring expressions to clearly communicate with their partners.
 
@@
 
-Following @Mankewitz2021 we only included games where players finished all 24 trials in the game.
-We only included games where participants had access ot the chat box (@Mankewitz2021 also had a no-chat baseline condition).
-This left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the Individual Utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the Shared Utilities Condition.
+## Exclusions 
+Following @mankewitz2021 we only included games where players finished all 24 trials in the game.
+We only included games where participants had access tt the chat box (@mankewitz2021 also had a no-chat baseline condition).
+This left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the individual utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the shared utilities Condition.
 
@@
 
-We then embedded each of these referring expressions using SBERT cite!
-in order to compare between referring expressions in semantic space.
-We use cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
+We embedded each of the extracted referring expressions using SBERT [@reimers2019], and used cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
 
-For the pairwise comparisons of utterances during the game, we take the similarities for utterances that are within 2 trials of each other (ex.
-an utterance from trial 10 is compared with utterances from trials 8,9,10,11,12).
-We do this because different flowers were shown to different groups on different trials.
-This width is a compromise between having enough pairs of utterances and still being able to treat them as coming from one time point in the game.
+For the pairwise comparisons of utterances during the game, we took the similarities for utterances that were within 2 trials of each other (ex.
+an utterance from trial 10 was compared with utterances from trials 8,9,10,11,12).
+Presentation of flowers was randomized, so different groups saw different flowers combinations on the same number trial.
+The width of the comparison window was a compromise between having enough pairs of utterances while still being able to treat them as coming from one time point in the game.
 
@@
 B: Words in each reference expression by trial.
-C: Total number of reference expressions per game per trial.", fig.height=2, fig.width=7}
+C: Total number of reference expressions per game per trial.
+\\label{wordcount}", fig.height=2, fig.width=7}
 
@@
 Consistent with the results from dyadic iterated reference games, the amount of referring language decreases over the course of the game.
-As shown in Fig @ref(fig:wordcount)A, the number of words of referring language decreases across the game in both conditions.
-This is due to the individual referring expressions getting shorter (@ref(fig:wordcount)B), while the total number of referring epxressions remains that same (@ref(fig:wordcount)C).
+As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game in both conditions.
+This is due to the individual referring expressions getting shorter (Figure \ref{wordcount}B), while the total number of referring expressions remains that same (Figure \ref{wordcount}C).
 
@@
 Utterances were paired with other utterances from different games; from the same game, but a different utterer speaker; or from the same speaker (vertical panels).
-Utterances were either in reference to different flowers or both in reference to the same flower (horizontal panels).", fig.height=4, fig.width=10, out.width="100%"}
+Utterances were either in reference to different flowers or both in reference to the same flower (horizontal panels).
+\\label{during}", fig.height=4, fig.width=10, out.width="100%"}
 flower_all|> 
@@
 
-```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
-x <- 0:100
-y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
-
-ggplot2::ggplot(data = data.frame(x, y), 
-                aes(x = x, y = y)) + 
-  geom_point() + 
-  geom_smooth(method = "lm")
-```
-
-```{r xtable, results="asis"}
-n <- 100
-x <- rnorm(n)
-y <- 2*x + rnorm(n)
-out <- lm(y ~ x)
-
-tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
-                       caption = "This table prints across one column.")
-
-print(tab1, type="latex", comment = F, table.placement = "H")
-```
-
 # Acknowledgements

### 2023-01-09 code/annotation_analysis.Rmd
semi-working draft

@@
 
+```{r }
+
+# by trial
+in_end <- M_mat %>% 
+  filter(seen %in% c("own")) |> 
+  group_by(color,condition) %>% 
+  mutate(combinedId=str_c(referent, "_",seen,"_", gameId, "_",playerId)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("ref1", "seen1","game1","player1"), convert=T) |> 
+  separate(dim2, c("ref2",  "seen2","game2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  ungroup()
+```
+
 ```{r}
 in_end|> 
+  filter(player2!=player1) |> 
   mutate(distance=case_when(
-    player1==player2 ~ "Same person",
     game1==game2 ~ "same game",
     T ~ "different game"),
-    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower"),
-    same_color=case_when(
-  seen1=="own" & seen2=="own" ~ "own",
-  seen2=="other" & seen1=="other" ~ "other",
-  T ~"mixed"
-  ))|> 
-ggplot( aes(x=same_color,y=sim,color=condition))+     
+    same_flower=ifelse(ref1==ref2, "Same flower", "diff flower"))|> 
+ggplot( aes(x=condition,y=sim,color=condition))+ geom_point(alpha=.01)+    
   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3))+
@@
 
+### Model
+
+```{r, eval=F}
+within_end <- in_end |> 
+  filter(seen1=="own" & seen2=="own") |> 
+  filter(player1!=player2) |> 
+  mutate(same_flower=ifelse(ref1==ref2, 1, 0),
+         same_game=ifelse(game1==game2, 1,0)) |> 
+  select(sim, condition, same_flower, same_game)
+
+m_within_end <-  stan_glm(sim ~ condition*same_flower*same_game, data=within_end) |> write_rds(here(model_path,"m6.rds"))
+
+show_summary_rstan(m_within_end) |> write_rds(here(model_path,"within_end.rds"))
+
+```
 Not sure how to show spread here. Or how to do a viz that allows for comparisons.

### 2023-01-09 write-ups/cogsci23_paper/manuscript.Rmd
semi-working draft

@@
 
-So people often care about reference or whatever.
-Referring expressions are crucial to interactive use of language so we can know what the interlocuter is talking about.
-Many goals of conversation rely on joint reference as a part of them -- for instance in order to negotiate, even from an adversarial perspective, you still need to have terms that refer to the same things in order to do the negotiation.
-Joint reference is a cooperative act, needed for both cooperative and non-cooperative acts.
+Successful communication is grounded in a shared understanding of what utterances mean in the context they are produced in.
+In many cases conventional word meanings are enough, but there are also contexts where objects to be referred to are not easy to distinguish.
+In these situations, establishing reference is more difficult, but no less important.
 
-In some cases joint reference is easy b/c there are conventional meanings of words that refer to the objects.
-However, this isn't always the case, either because objects are novel and don't have names or because objects are similar to one another and so while they might be nameable out of context, in context there are not conventional names.
+The formation of these ad-hoc referring expressions has been studied extensively in the context of dyadic reference games [@clarkReferringCollaborativeProcess1986, ADD MORE CITES].
+In these, two participants see a set of images (often abstract shapes) and one person is tasked with identifying an image so the other person can pick it out.
+People can be very successful at this task, achieving high levels of matches.
+Over repeated reference to the same images, pairs develop shorthand names for the images, leading to shorter descriptions in later repetitions.
+These nicknames conventionalize with a pair, as they tend to stick to the same description, but are ideosyncratic and differ between pairs.
 
-How people develop ad-hoc shared conventional referring expressions for objects without conventional names has primarily been studied in the realm of dyadic reference games.
-Two participants see a set of images, and the speaker sees a particular one highlighted as the one to pick out next (or they have them in the correct order and the listener needs to reorder theirs to match).
-This is a task explicitly about referring to the objects that is highly structured.
-There are assigned roles (one person has all the knowledge) and a clear indication of which one to talk about.
-Selecting or ordering objects is pretty simple.
+This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO].
+The implicit theoretical claim is that these phenomena hold whenever people interact repeatedly in ways that require reference to some objects that do not have pre-conventionalized names that are adequate to distinguish them in context.
 
-This has been a very useful microcosm for studying referring expressions, expectations about conventions, and what happens with new listeners or speakers.
-The key phenomena observed across these experiments are that pairs develop partner specific referring expressions -- the utterances produced by the speaker to refer to objects get shorter over repetition, the accuracy-speed trade off improves (?
-maybe don't mention this b/c it's not what we look at), and the terms that develop are partner specific -- people converge on terms within partnerships, but differ between partnerships on the terms used.
+However, there are some ways in which this microcosm is not representative of typical language use.
+For one, in everyday communication, establishing reference rarely happens in isolation but is usually a subgoal in a larger conversational goal.
+The overall goal might be cooperative, such as in asking for a piece of kitchen equipment while cooking together, but it can also be adversarial, such as in a negotiation where the assets to be divied up must be referenced.
+In the reference game context, the goal is always cooperative, to match the images or get them in the same order.
 
-Implicitly, these features are thought to be about how referring expressions develop across some wider range of situations where people interact repeatedly in ways that need reference to some objects that do not already have conventional names that are sufficient for reference.
-Basically it needs to be a situation where people don't at the outset have a high degree of agreement on how to refer to the thing.
+The variation in goals can also lead to choices about when and what to refer to and how precise to be, depending on the costs of mistaken reference.
 
-Needs more transition here. 
+Do the results of dyadic reference games extend to more complex communicative interactions where goals are more complex and may not be aligned?
+
+[possible worth acknowledging somewhere that sometimes esp. in legal things and philosophy, fighting over names is the argument, and we're going to ignore that]
+
 ```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Player interface.
@@
 
-@mankewitz2021 tried to bridge the gap between datasets of negotiation (like let's make a deal?) where there is negotiation over easily nameable objects, and the reference game literature where there is reference (but no negotiation) to hard to name objects.
-They created a 3 person game where players had to select what flowers to grow from a set of 6 shown on the screen.
-The values of the flowers were partially hidden, and flowers were all the same color (but different shapes/species).
-Each player could select a flower, but it was only worth the shown value if one player selected it; if multiple did, they got a much lower reward.
-Flower images repeated across the game.
-@mankewitz intended this set-up so that players would need to refer the flowers in order to negotiate who would take what, but in a context where there wasn't a flower to talk about and there wasn't a designated speaker.
+[needs transition, but not sure how]
+
+@mankewitz2021 tried to bridge the gap between datasets of negotiation over easily nameable objects, and the reference game literature where there is reference (but no negotiation) over hard to name objects.
+They created a 3 person game where players each select what flowers to grow from a set of 6 flowers with variable and partially hidden value.
+Players got the value of the flower only if they were the only one to select it, which incentivized coordination and negotiation between players.
+However, all the flowers shown were the same color, so they were not easily nameable, and the flowers repeated, leading to opportunities for ad-hoc conventionalization.
+This game structure encouraged players to describe flowers while never explicitly saying that a player should refer to any specific flower.
 They also had two conditions, one where players within a group won points together (and thus had fully aligned incentives) and another where they won points as individuals (and didn't).
-The paper found ....
-a slight decrease in language over the course of hte game, possibly consistent with reduced referring expressions, but also consistent with developing a negotiation pattern.
+@mankewitz2021 found a slight decrease in language over the course of hte game, possibly consistent with reduced referring expressions, but also consistent with developing a negotiation pattern.
 
-This builds on the ref game literature in a few dimensions.
-The key difference is the more free-form format where reference is implicitly motivated by the task, rather than being an explicit part of the task, and the equal positions of the players (each has the same amount of knowledge) rather than an asymmetric teacher/student.
-Other differences include that the images are natural photos, which has been used in ref games (CITE that management paper), but is not the typical and the presence of more than two participants (which is done, but not modal, citations!).
+Here we extracted the referring expressions from the chat logs of @mankewitz to look for reduction to partner-specific terms.
+We explored whether the phenomena found in classic dyadic iterated reference games extended to this more naturalistic domain, and whether they differed across the individual and shared incentive conditions.
+The key questions we addressed are:
 
-Here we extract the referring expressions from the chat logs of @mankewitz in order to look at reduction as well as semantic trends.
-We see whether the phenomena found in classic dyadic iterated reference games extend to this more naturalistic domain, and whether they differ across the individual and shared incentive conditions.
+1. Do the referring expressions get shorter over time?
 
-TODO list of key points.
+2. Do referring expressions for the same image converge within a group?
+
+3. Do referring expressions for different images diverge within a group?
+
+4. Do referring expressions for the same image diverge between groups?
+
+[TODO fix list of key points to match results.]
 
@@
 We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions. 
-
-## SBERT embeddings
-
-We embedded each of the extracted referring expressions using SBERT [@reimers2019], and used cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
-
-For the pairwise comparisons of utterances during the game, we took the similarities for utterances that were within 2 trials of each other (ex.
-an utterance from trial 10 was compared with utterances from trials 8,9,10,11,12).
-Presentation of flowers was randomized, so different groups saw different flowers combinations on the same number trial.
-The width of the comparison window was a compromise between having enough pairs of utterances while still being able to treat them as coming from one time point in the game.
-
-# Results
-
-We assess the list of key points given at the beginning.
-
-## Reduction of referring expressions
 ```{r wordcount, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
@@
 
+## SBERT embeddings
+
+We embedded each of the extracted referring expressions using SBERT [@reimers2019], and used cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
+
+For the pairwise comparisons of utterances during the game, we took the similarities for utterances that were within 2 trials of each other (ex.
+an utterance from trial 10 was compared with utterances from trials 8,9,10,11,12).
+Presentation of flowers was randomized, so different groups saw different flowers combinations on the same number trial.
+The width of the comparison window was a compromise between having enough pairs of utterances while still being able to treat them as coming from one time point in the game.
+
+# Results
+
+We address each of the questions of interest listed at the end of the Introduction.
+
+## Reduction of referring expressions
+
 ```{r}
@@
 
-Consistent with the results from dyadic iterated reference games, the amount of referring language decreases over the course of the game.
-As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game in both conditions.
-This is due to the individual referring expressions getting shorter (Figure \ref{wordcount}B), while the total number of referring expressions remains that same (Figure \ref{wordcount}C).
+The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
+As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game in both Shared and Individual utility conditions.
+The reduction in referring langauge was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per round remained constant (Figure \ref{wordcount}C).
 
@@
 
-Another coarse metric for how referring expressions are being produced is to look at how many different flowers are talked about each round.
-There are 6 visible on the screen, and participants are incentivized to each pick a different one.
+Another coarse metric for how referring expressions are produced is how many different flowers are referred to each round.
+Each round of the game, there are 6 flowers visible on the screen, and participants want to each pick a different one.
 In general, each trial contained references to 2 or 3 distinct flowers (Individual Utilities: mean of `r  mean_flowers |> pluck(2,1)`, Shared Utilities: `r mean_flowers |> pluck(2,2)`).
@@
 
-This pattern is consistent with each person saying what flower they plan on choosing; some games talked less than this, and others talked more as players queried the worths of various flowers or confirmed the plan of who would pick what.
+The average pattern is consistent with each person saying what flower they plan on choosing.
+Some games talked less than this and risked collisions.
+Other games  talked more as players queried the worths of various flowers or confirmed the plan of who would pick what.
+
+? should there be a visual of these distributions ?
 
 ## Convergence of referring expressions
+
+TODO resummarize what cosine similarity is
+
+[formatting comment -- changes in cos sim tend to be small, so estimates and CredInts look kinda silly some of the time -- could leave or increase sigfigs which is also silly]
+
 ```{r, include=F,}
@@
 
+Reference games have clear structures of repetition, as each image is usually the target once before any are again, and so it makes sense to treat each block as a time point.
+Because of the different structure of the game, trials are more continuous with each other.
+We smooth the random sampling and sparseness of flowers occuring the same trial by treating descriptions from within two trials as being at a comparable time in the game (so descriptions from round 10 are compared with descriptions from rounds 8, 9, 10, 11, and 12).
+
 ### Within games
 
-TODO we compare games within 2 rounds of each other (ex. trial 10 compared to 8, 9, 10, 11, 12)
+One of the key claims is that within a partnership, the referring expressions should converge for the same image, but get more different for different images.
+Convergence would be increasing similarity (higher cosine similarity in later rounds), divergence would be decreasing similarity in later rounds).
 
-One of the key claims is that within a partnership, the referring expressions should converge for the same image, but get more different for different images.
-Convergence would be increasing similiarity (higher cosine similarity in later rounds), divergence would be decreasing similarity in later rounds).
-
-For the same flower within the same game, similarity increased over rounds (trial: `r stats_text(wi_g_wi_f,6)`).
+If the dyadic pattern held, we expected similarity to increase over time for the same image with a group.
+In a model of similarity of utterances within flower and within group, we found that similarity increased over time (trial: `r stats_text(wi_g_wi_f,6)`, see Figure \ref{during} lower middle and lower right panels).
 Utterances were more similar if they were produced by the same participant (`r stats_text(wi_g_wi_f,8)`), although this did not interact with trial number (`r stats_text(wi_g_wi_f,7)`).
@@
 
-Different flowers within the same game have diverging descriptions that get less similar over time (trial: `r stats_text(wi_g_bet_f,6)`).
-Terms are overall slightly less similar in Shared Utility groups (`r stats_text(wi_g_bet_f,2)`), but this does not interact with trial (`r stats_text(wi_g_bet_f,3)`).
-There is not an effect of being said by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) or condition (`r stats_text(wi_g_bet_f,5)`).
+In contrast, we expected descriptions of different flowers to diverge within the same group.
+In a model of similarity between descriptions of different flowers within a group, similarity decreased  over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} upper middle and upper right panels).
+Terms were slightly less similar in Shared Utility groups (`r stats_text(wi_g_bet_f,2)`), but this did not interact with trial (`r stats_text(wi_g_bet_f,3)`).
+There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) or condition (`r stats_text(wi_g_bet_f,5)`).
+
+The main results of convergence of descriptions for the same image and divergence between descriptions of different images matches the reference game phenomena.
 
@@
 
-Utterances describing the same flower in different games diverged slightly (trial: `r stats_text(bet_g_wi_f,4)`).
+In reference games, conventions are partner-specific, so over time, descriptions become more particular to the group.
+If that held, we expected, utterances describing the same flower in different groups to have lower similarities later in the game.
+
+In a model of utterance similarities for the same flower in different games, similiarities decreased slightly over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} lower left panel).
 There was a condition difference with shared utility conditions having more similar descriptions across games (`r stats_text(bet_g_wi_f,2)`); this did not interact with trial (`r stats_text(bet_g_wi_f,3)`).
@@
 TODO this should be hedged *a lot* but we can also cite the study about large groupy things and the crab-like images
+
 ```{r }
@@
 
-```{r toend, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.", fig.height=4, fig.width=6, out.width="50%"}
+```{r}
+in_end <- M_mat %>% 
+  filter(seen %in% c("own")) |> 
+  group_by(color,condition) %>% 
+  mutate(combinedId=str_c(referent, "_",seen,"_", gameId, "_",playerId)) %>% 
+  make_across_df(F_mat, 'cosine') %>% 
+  separate(dim1, c("ref1", "seen1","game1","player1"), convert=T) |> 
+  separate(dim2, c("ref2",  "seen2","game2", "player2"), convert=T) |> 
+ # rename(referent1=dim1,referent2=dim2) %>% 
+  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  ungroup()
+```
+
+```{r toend, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.
+\\label{toend}", fig.height=4, fig.width=6}
 to_end|> 
@@
 to_e_bet_f <- read_rds(here(model_path,"to_end_diffflower_model.rds"))
+
+wi_e <- read_rds(here(model_path, "within_end.rds"))
 ```
+
 ## End expressions
 
+```{r withinend, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="90%", fig.width=4, fig.height=2, fig.cap = "Cosine similarities between flower descriptions provided by participants after the game.
+Descriptions were more similar to each other if they came from the same group and were of the same flower.
+\\label{withinend}"}
+in_end|> 
+  filter(player2!=player1) |> 
+  mutate(distance=case_when(
+    game1==game2 ~ "Same group",
+    T ~ "Different group"),
+    same_flower=ifelse(ref1==ref2, "Same flower", "Different flower"))|> 
+ggplot( aes(x=str_c(same_flower,"\n", distance),y=sim,color=condition))+
+  labs(y="Cosine Similarity", x=NULL)+
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3))+theme(legend.position=c(0.2,.75), legend.title=element_blank(), legend.background=element_blank())+
+        scale_color_brewer(palette="Dark2")
+```
+
 ### game to end
-As a complementary way of looking at how referring expressions change over time as conventions are formed, we can take the post-game flower descriptions (each player was asked to provide what they would call each flower to their teammates) as the conventions.
+As a complementary way of looking at how referring expressions change over time as conventions are formed, we can take the post-game flower descriptions (each player was asked to provide what they would call each flower to their teammates) as the conventions, and look at how the in-game utterances converge to these conventions.
 
-Descriptions become more similar to the conventions over time  (trial: `r stats_text(to_e_wi_f,6)`).
-Utterances were more similar to the convention given by the same person (versu from a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
+Descriptions become more similar to the conventions over time  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} lower panels).
+Utterances were more similar to the convention given by the same person (versus from a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
 Games in the shared utilities condition had utterances that were more similar to their conventions (`r stats_text(to_e_wi_f,2)`), but this did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
 
-There was not a significant divergence from the conventions for other flowers over time (trial: `r stats_text(to_e_bet_f,6)`).
+There was not a significant divergence from the conventions for other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
 Terms are overall slightly less similar in Shared Utility groups (`r stats_text(to_e_bet_f,2)`), but this does not interact with trial (`r stats_text(to_e_bet_f,3)`).
@@
 ### within end
-want to look for more similar with a group than between groups
+One last way of looking at convention formation is to look at how similar the end descriptions were to those of other players, either groupmates or people in different groups (but playing with the same color palette) for the same or different flowers.
+We expected that within a group, descriptions for the same flower would be much more similar, while descriptions of different flowers would be more different, compared to cross-group similarities.
 
-do we want to look at the "other" (flower not in game) at all? No?
+As shown in Figure \ref{withinend}, compared to different flowers described by different games as a baseline, descriptions for the same game are more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower are more similar (even across games, `r stats_text(wi_e,6)`), and there is a large interaction effect, where descriptions of the same flower are very similar among groupmates (`r stats_text(wi_e,7)`).
+Compared with the individual utility condition, shared utility games had more similar descriptions of the same flower between games (`r stats_text(wi_e,3)`).
+Other differences between the two conditions were small.
+
+[probably need to talk about this result]
+[this whole section might be more interpretable if I centered Shared/indiv utilities in the model?]
 
 # Discussion
+TODO!
 
@@
 One difference we see is that groups don’t diverge (from each other). This may be dependent on stimulus properities (are there universal features of some of the images?) and group dynamics
+
+mention the set of departures this made from classic reference game
 

### 2023-01-16 code/annotation_analysis.Rmd
work on paper and redo model summaries

@@
   intervals <- gather_draws(model, `[a-zA-z_(): ]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
-    mutate(lower=round(.lower, digits=2),
-           upper=round(.upper, digits=2),
+    mutate(lower=round(.lower, digits=3),
+           upper=round(.upper, digits=3),
            `Credible Interval`=str_c("[",lower,", ", upper,"]"),
            Term=.variable,
-           Estimate=round(.value, digits=2)) %>% 
+           Estimate=round(.value, digits=3)) %>% 
     select(Term, Estimate, `Credible Interval`)
@@
 
-m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
-show_summary_rstan(m1) |> write_rds(here(model_path,"word_model.rds"))
+m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|flower)+(1|player), data=annotated) |> write_rds(here(model_path,"m1.rds"))
+
 ```
 
-```{r, eval=F, include=F}
-m1 <- read_rds(here(model_path,"m1.rds"))
-m1
+```{r, eval=F}
+ read_rds(here(model_path,"m1.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"word_model.rds"))
+```
 
+```{r}
+foo <- read_rds(here(model_path,"m1.rds"))
+gather_draws(foo, `[a-zA-z_(): ]+`, regex=TRUE) |> select(.variable) |> unique()
 ```
@@
                                       (1|game1)+(1|flower), data=within_game_flower) |> write_rds(here(model_path,"m2.rds"))
+```
 
-show_summary_rstan(m_samegame_sameflower) |> write_rds(here(model_path,"samegame_sameflower_model.rds"))
+```{r, eval=F}
+read_rds(here(model_path,"m2.rds")) |> 
+show_summary_rstan() |> write_rds(here(model_path,"samegame_sameflower_model.rds"))
 
@@
 m_samegame_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=within_game_between_flower) |> write_rds(here(model_path,"m3.rds"))
+```
 
-show_summary_rstan(m_samegame_diffflower) |> write_rds(here(model_path,"samegame_diffflower_model.rds"))
+```{r, eval=F}
+read_rds(here(model_path,"m3.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"samegame_diffflower_model.rds"))
 
@@
 
-```{r}
+```{r, eval=F}
 between_game_within_flower <- flower_all |> 
@@
 m_diffgame_sameflower <-  stan_glm(sim ~ condition*earlier, data=between_game_within_flower) |> write_rds(here(model_path,"m4.rds"))
+```
 
-show_summary_rstan(m_diffgame_sameflower) |> write_rds(here(model_path,"diffgame_sameflower_model.rds"))
+```{r, eval=F}
+read_rds(here(model_path,"m4.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"diffgame_sameflower_model.rds"))
 
@@
 m_to_end_sameflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_same_flower) |> write_rds(here(model_path,"m5.rds"))
+```
 
-show_summary_rstan(m_to_end_sameflower) |> write_rds(here(model_path,"to_end_sameflower_model.rds"))
+```{r, eval=F}
+read_rds(here(model_path,"m5.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"to_end_sameflower_model.rds"))
 
@@
 
-m_to_end_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_diff_flower) |> write_rds(here(model_path,"m5.rds"))
+m_to_end_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_diff_flower) |> write_rds(here(model_path,"m6.rds"))
+```
 
-show_summary_rstan(m_to_end_diffflower) |> write_rds(here(model_path,"to_end_diffflower_model.rds"))
+```{r, eval=F}
+read_rds(here(model_path,"m6.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"to_end_diffflower_model.rds"))
 
@@
 
-m_within_end <-  stan_glm(sim ~ condition*same_flower*same_game, data=within_end) |> write_rds(here(model_path,"m6.rds"))
+m_within_end <-  stan_glm(sim ~ condition*same_flower*same_game, data=within_end) |> write_rds(here(model_path,"m7.rds"))
 
-show_summary_rstan(m_within_end) |> write_rds(here(model_path,"within_end.rds"))
+```
+
+```{r, eval=F}
+
+read_rds(here(model_path,"m7.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"within_end.rds"))
 

### 2023-01-16 write-ups/cogsci23_paper/manuscript.Rmd
work on paper and redo model summaries

@@
 ---
-title: "TODO"
+title: "Communicative reduction in referring expressions in a multi-player negotation game"
 bibliography: [ aaflowers.bib]
@@
 abstract: >
-    TODO
+    Reduction of referring expressions to short conventionalized terms is found robustly in dyadic repeated reference games when the target  images are initially difficult to name.
+Establishing reference is a crucial part of many conversations with different goals and more complex motivations; thus, it is useful to look at how the findings from reference games, where participants share the explicit goal of establishing joint reference, carry over into situations where participants have different overarching goals.
+In the current work, we analyse a dataset where reference was embedded in strategic 3- player negotiation and coordination games.
+In these more complex games, we found that the patterns of reduction and convergence to conventions held across two different incentive conditions, with some suggestive differences between the conditions.
 
 keywords: >
-    Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
+    Convention formation; reference games; reduction of referring expressions; multi-party communication;
 
@@
 
+# d.games |> filter(conditionName!="Old Cooperative") |> group_by(conditionName) |> tally()
 d.rounds <- read_csv(here(read_data_path,"rounds.csv"))
@@
 
-# introduction
-
 Successful communication is grounded in a shared understanding of what utterances mean in the context they are produced in.
@@
 Over repeated reference to the same images, pairs develop shorthand names for the images, leading to shorter descriptions in later repetitions.
-These nicknames conventionalize with a pair, as they tend to stick to the same description, but are ideosyncratic and differ between pairs.
+These nicknames conventionalize with a pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
 
-This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO].
+This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO CITES].
 The implicit theoretical claim is that these phenomena hold whenever people interact repeatedly in ways that require reference to some objects that do not have pre-conventionalized names that are adequate to distinguish them in context.
@@
 
-The variation in goals can also lead to choices about when and what to refer to and how precise to be, depending on the costs of mistaken reference.
-
-Do the results of dyadic reference games extend to more complex communicative interactions where goals are more complex and may not be aligned?
-
-[possible worth acknowledging somewhere that sometimes esp. in legal things and philosophy, fighting over names is the argument, and we're going to ignore that]
+TODO paragraph on on the issue of shared vs.
+privileged knowledge, which has been studied extensively in communication psycholinguistics (e.g., heller, keysar, barr) but IIRC not in the context of creating novel referring expressions in the tangram task.
+so we are blending in this other important element of naturalism, albeit only around the values and not the actual stuff?
 
@@
 
-[needs transition, but not sure how]
-
 @mankewitz2021 tried to bridge the gap between datasets of negotiation over easily nameable objects, and the reference game literature where there is reference (but no negotiation) over hard to name objects.
@@
 They also had two conditions, one where players within a group won points together (and thus had fully aligned incentives) and another where they won points as individuals (and didn't).
-@mankewitz2021 found a slight decrease in language over the course of hte game, possibly consistent with reduced referring expressions, but also consistent with developing a negotiation pattern.
+@mankewitz2021 found a slight decrease in language over the course of the game, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
 
-Here we extracted the referring expressions from the chat logs of @mankewitz to look for reduction to partner-specific terms.
+Historically, studies of convention formation have had to rely on proxies such as reduction and fluency, qualitatively assess similarity of utterances, or manually measure lexical overlap in small datasets to get at the ideas of convergence and divergence.
+More recent work has taken advantage of natural language processing tools for mapping words and sentences to semantic vector spaces in order to quantify semantic similarities between sentences [@hawkins2020, @boyce2022].
+We follow this recent work in using vector embeddings to measure semantic distances between referring expressions in the @mankewitz2021 dataset.
+
+## Current work
+Here we ask whether participants in a complex, multi-player strategic game with differing incentives and shared knoweldge were able to develop conventionalized reference expressions.
 We explored whether the phenomena found in classic dyadic iterated reference games extended to this more naturalistic domain, and whether they differed across the individual and shared incentive conditions.
-The key questions we addressed are:
+The key questions we addressed were:
 
@@
 
-2. Do referring expressions for the same image converge within a group?
+2. How do referring expressions change within and between groups over the course of the games? Do descriptions converge within groups and diverge between groups?
 
-3. Do referring expressions for different images diverge within a group?
-
-4. Do referring expressions for the same image diverge between groups?
-
-[TODO fix list of key points to match results.]
+3. How do the different incentive conditions influence the referring expressions?
 
@@
 
-## Data Acquisition
+## Procedure
 As described in @mankewitz2021, participants played a real-time coordination game in groups of three, implemented using Empirica [@almaatouqEmpiricaVirtualLab2020].
@@
 
-## Exclusions 
-Following @mankewitz2021 we only included games where players finished all 24 trials in the game.
-We only included games where participants had access tt the chat box (@mankewitz2021 also had a no-chat baseline condition).
-This left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the individual utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the shared utilities Condition.
+## Participants
+@mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants for 25 3-person games in each of the individual and shared utilities conditions where they had access to a chat box (there were also no-chat-box baseline conditions, which we did not analyse).
+Some games did not complete the full 24 rounds because participants dropped out early.
+We excluded incomplete games from analyses, which left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the individual utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the shared utilities condition (`r 3*(d.games.final |> nrow())` participants included in the analyses).
 
@@
         scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        scale_color_brewer(palette="Dark2")+
+        ggthemes::scale_color_solarized()+
         guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
@@
         scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        scale_color_brewer(palette="Dark2")
+        ggthemes::scale_color_solarized()
 
@@
 ## SBERT embeddings
+Whereas in the past studies of convention formation have had to rely on manual assessments of similarity or lexical overlap, natural language processing tools now make it possible to quantitatively measure the similarity between phrases.
+Recent work by @Hawkins2020 and @boyce2022 have used these tools to assess the dynamics of convergence to conventions.
 
-We embedded each of the extracted referring expressions using SBERT [@reimers2019], and used cosine distance between embedding pairs as a measure of similarity between the corresponding referring expressions.
+We embedded each of the extracted referring expressions using SBERT [@reimers2019] which maps each utterance to a vector in a high-dimensional semantic vector space.
+We use cosine distance between pairs of vector embeddings as a measure of similarity between the corresponding referring expressions.
+When two vectors are near each other, the angle between them is small, and the cosine of this angle is large.
+Higher cosine values correspond to vectors close together, and thus, reference expressions with more similar meanings.
 
-For the pairwise comparisons of utterances during the game, we took the similarities for utterances that were within 2 trials of each other (ex.
-an utterance from trial 10 was compared with utterances from trials 8,9,10,11,12).
-Presentation of flowers was randomized, so different groups saw different flowers combinations on the same number trial.
-The width of the comparison window was a compromise between having enough pairs of utterances while still being able to treat them as coming from one time point in the game.
+When doing pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same round and the two prior rounds (thus the similarity at round 10 comes from 8-10, 9-10 and 10-10 comparisons).
+This smoothing was not necessary when comparing with the names produced after the game.
+
+## Analytic methods
+TODO this is the section where we talk about modelling details
 
@@
 
-We address each of the questions of interest listed at the end of the Introduction.
+We address each of the questions of interest listed at the end of the Introduction in order.
 
@@
 
-word_count <- read_rds(here(model_path,"word_model.rds"))
+word_count <- read_rds(here(model_path,"word_model.rds")) |> filter(str_detect(Term,"\\[", negate=T)) #because there are problems with a couple mixed effects getting through
 ```
 
-The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
-As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game in both Shared and Individual utility conditions.
-The reduction in referring langauge was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per round remained constant (Figure \ref{wordcount}C).
-
-TODO include model!
-
-Another coarse metric for how referring expressions are produced is how many different flowers are referred to each round.
-Each round of the game, there are 6 flowers visible on the screen, and participants want to each pick a different one.
+Each round of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
 In general, each trial contained references to 2 or 3 distinct flowers (Individual Utilities: mean of `r  mean_flowers |> pluck(2,1)`, Shared Utilities: `r mean_flowers |> pluck(2,2)`).
@@
 
-The average pattern is consistent with each person saying what flower they plan on choosing.
-Some games talked less than this and risked collisions.
-Other games  talked more as players queried the worths of various flowers or confirmed the plan of who would pick what.
-
-? should there be a visual of these distributions ?
+The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
+As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game (trial: `r stats_text(word_count,3)`.
+The reduction in referring language was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per round remained constant (Figure \ref{wordcount}C).
 
 ## Convergence of referring expressions
-
-TODO resummarize what cosine similarity is
-
-[formatting comment -- changes in cos sim tend to be small, so estimates and CredInts look kinda silly some of the time -- could leave or increase sigfigs which is also silly]
 
@@
     player1==player2 ~ "Same Person",
-    game1==game2 ~ "Same Game",
+    game1==game2 ~ "Same Game - Different Person",
     T ~ "Different Games"),
     same_flower=ifelse(ref1==ref2, "Same Flower", "Different Flowers")
-  ) |> 
+  ) |> mutate(same_flower=factor(same_flower,levels=c("Same Flower", "Different Flowers")),
+              distance=factor(distance, levels=c("Same Person", "Same Game - Different Person", "Different Games"))) |> 
 ggplot( aes(x=later,y=sim,color=condition))+     
@@
         scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        scale_color_brewer(palette="Dark2")+
+        ggthemes::scale_color_solarized()+
   labs(x="Trial Number", y="Cosine Similarity")+
@@
 
-Reference games have clear structures of repetition, as each image is usually the target once before any are again, and so it makes sense to treat each block as a time point.
-Because of the different structure of the game, trials are more continuous with each other.
-We smooth the random sampling and sparseness of flowers occuring the same trial by treating descriptions from within two trials as being at a comparable time in the game (so descriptions from round 10 are compared with descriptions from rounds 8, 9, 10, 11, and 12).
+The classic reference game phenomenon of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
+We assess all of these in two ways: 1) comparing referring expressions within the game to other referring expressions from the same timepoint to get a sense of the dynamic changes, and 2) comparing referring expressions within the game to the names players gave the flowers after the game, to track how referring expressions evolve towards these final conventions.
+
+TODO table with examples of flower descriptions over time and cos similarities !!!
 
@@
 
-One of the key claims is that within a partnership, the referring expressions should converge for the same image, but get more different for different images.
-Convergence would be increasing similarity (higher cosine similarity in later rounds), divergence would be decreasing similarity in later rounds).
+Within games, the patterns from dyadic reference games predict that similarity for descriptions of the same image will increase over time while similarity for descriptions of different images will decrease over time.
 
-If the dyadic pattern held, we expected similarity to increase over time for the same image with a group.
-In a model of similarity of utterances within flower and within group, we found that similarity increased over time (trial: `r stats_text(wi_g_wi_f,6)`, see Figure \ref{during} lower middle and lower right panels).
+We found that for descriptions of the same flower, cosine similarity increased over time (trial: `r stats_text(wi_g_wi_f,6)`, see Figure \ref{during} upper left and middle panels).
 Utterances were more similar if they were produced by the same participant (`r stats_text(wi_g_wi_f,8)`), although this did not interact with trial number (`r stats_text(wi_g_wi_f,7)`).
-Games in the shared utilities condition may have had higher similarities overall (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or same participant versus different participant (`r stats_text(wi_g_wi_f,5)`).
 
-In contrast, we expected descriptions of different flowers to diverge within the same group.
-In a model of similarity between descriptions of different flowers within a group, similarity decreased  over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} upper middle and upper right panels).
-Terms were slightly less similar in Shared Utility groups (`r stats_text(wi_g_bet_f,2)`), but this did not interact with trial (`r stats_text(wi_g_bet_f,3)`).
-There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) or condition (`r stats_text(wi_g_bet_f,5)`).
+For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
+There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) .
 
-The main results of convergence of descriptions for the same image and divergence between descriptions of different images matches the reference game phenomena.
+These results mirror are consistent with the reference game phenomenon.
 
@@
 
-In reference games, conventions are partner-specific, so over time, descriptions become more particular to the group.
-If that held, we expected, utterances describing the same flower in different groups to have lower similarities later in the game.
+The theory of partner-specificity predictions that descriptions to the same flower in different groups will diverge over time as each group forms their own group-specific convention.
 
-In a model of utterance similarities for the same flower in different games, similiarities decreased slightly over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} lower left panel).
-There was a condition difference with shared utility conditions having more similar descriptions across games (`r stats_text(bet_g_wi_f,2)`); this did not interact with trial (`r stats_text(bet_g_wi_f,3)`).
-
-TODO: what is our possibly explanation for this?!
-
-This may be a fluke, as there are not a huge number of games in either condition, but if it's robust, it could be that the shared utility framing leads to less idiosyncratic descriptions and instead towards more universal priors.
-TODO this should be hedged *a lot* but we can also cite the study about large groupy things and the crab-like images
+For descriptions of the same flower across games, we found that cosine similarities decreased slightly over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} upper right panel).
+This is weakly consistent with the predictions.
 
@@
     T ~ "Same Game - Different Person"),
-        same_flower=ifelse(ref1==ref2, "Same flower", "Different flower")
-) |> 
+        same_flower=ifelse(ref1==ref2, "Same Flower", "Different Flowers")
+) |> mutate(same_flower=factor(same_flower,levels=c("Same Flower", "Different Flowers")),
+              distance=factor(distance, levels=c("Same Person", "Same Game - Different Person" ))) |> 
 ggplot( aes(x=earlier,y=sim,color=condition))+     
@@
         scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        scale_color_brewer(palette="Dark2")+
+        ggthemes::scale_color_solarized()+
   guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
@@
 
-## End expressions
+### Comparison with post-game names
 
@@
   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3))+theme(legend.position=c(0.2,.75), legend.title=element_blank(), legend.background=element_blank())+
-        scale_color_brewer(palette="Dark2")
+        ggthemes::scale_color_solarized()
 ```
 
-### game to end
-As a complementary way of looking at how referring expressions change over time as conventions are formed, we can take the post-game flower descriptions (each player was asked to provide what they would call each flower to their teammates) as the conventions, and look at how the in-game utterances converge to these conventions.
+After the game, players were asked to provide the description they would use to identify each flower to their teammates.
+We treated these descriptions as the conventionalized names and looked at how the in-game descriptions developed into these conventions.
 
-Descriptions become more similar to the conventions over time  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} lower panels).
-Utterances were more similar to the convention given by the same person (versus from a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
-Games in the shared utilities condition had utterances that were more similar to their conventions (`r stats_text(to_e_wi_f,2)`), but this did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
+Referring descriptions later in a game were more similar to the convention than descriptions earlier in the game  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} upper panels).
+Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
+This is consistent with the expected pattern of convention formation.
 
-There was not a significant divergence from the conventions for other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
-Terms are overall slightly less similar in Shared Utility groups (`r stats_text(to_e_bet_f,2)`), but this does not interact with trial (`r stats_text(to_e_bet_f,3)`).
-There is not an effect of being said by the same person (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
-Similarity to a different convention was slightly higher if was said by the same person in a Shared Utility group (`r stats_text(to_e_bet_f,5)`).
+We also examined how referring expressions related the conventionalized names for other flowers.
+Utterances did not significantly diverge from the conventionalized names of other flower over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
+There was not an effect of being said by the same person who gave the convention (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
 
-### within end
-One last way of looking at convention formation is to look at how similar the end descriptions were to those of other players, either groupmates or people in different groups (but playing with the same color palette) for the same or different flowers.
-We expected that within a group, descriptions for the same flower would be much more similar, while descriptions of different flowers would be more different, compared to cross-group similarities.
+## Individual and Shared Utility conditions
 
-As shown in Figure \ref{withinend}, compared to different flowers described by different games as a baseline, descriptions for the same game are more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower are more similar (even across games, `r stats_text(wi_e,6)`), and there is a large interaction effect, where descriptions of the same flower are very similar among groupmates (`r stats_text(wi_e,7)`).
-Compared with the individual utility condition, shared utility games had more similar descriptions of the same flower between games (`r stats_text(wi_e,3)`).
-Other differences between the two conditions were small.
+The last question of interest is whether there were differences between the individual and shared utility conditions.
+Here we report the extent to which the shared utility condition differed from the individual utility condition.
 
-[probably need to talk about this result]
-[this whole section might be more interpretable if I centered Shared/indiv utilities in the model?]
+There was not a significant difference in the amount of referring language produced in the two conditions (`r stats_text(word_count,2)`).
+
+### Within games
+Games in the shared utilities condition had slightly higher cosine similarities between descriptions of the same flower within games (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or whether the utterances were produced by the same participant (`r stats_text(wi_g_wi_f,5)`).
+As a complement to this, games in the shared utilities condition had slightly lower similarities between descriptions of different flowers within games, but this did not interact with trial (`r stats_text(wi_g_bet_f,3)`) or whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
+Taken together, groups in the shared utility had more internal alignment on the mapping between flowers and names.
+
+Comparing to the end conventions, games in the shared utility condition produced flower descriptions that were more similar to the convention for that flower (`r stats_text(to_e_wi_f,2)`), but this did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
+In the shared utility condition, descriptions were slightly less similar to the conventions for other flowers (`r stats_text(to_e_bet_f,2)`), but this did not interact with trial (`r stats_text(to_e_bet_f,3)`).
+There was an interaction between shared utility condition and the referring expression coming from the same person who gave the convention for the different flower (`r stats_text(to_e_bet_f,5)`).
+
+Taken together, groups in the shared utility had more internal alignment on the mapping between flowers and names.
+
+### Between games
+Across different games, games in the shared utility condition tended to have more similar descriptions for the same flower (`r stats_text(bet_g_wi_f,2)`); this did not interact with trial (`r stats_text(bet_g_wi_f,3)`).
+
+Lastly, we can look at how the conventionalized names compared, both between members of the same group and across players in different groups that saw the same color palette of flowers, and whether this differs by condition.
+
+We treat different flowers described by different games as the baseline condition.
+For individual utility games, descriptions from groupmates are more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower are more similar (even across games, `r stats_text(wi_e,6)`), and there is a large interaction effect, where descriptions of the same flower are very similar among groupmates (`r stats_text(wi_e,7)`).
+
+For shared utility games compared to individual utility games, descriptions are roughly equivalent at baseline (`r stats_text(wi_e,2)`) and for descriptions from groupmates (`r stats_text(wi_e, 5)`).
+Descriptions are much more similar for the same flower (even across games, `r stats_text(wi_e,3)`), but this is balanced out by a negative interaction effect between same flower and same game (`r stats_text(wi_e,4)`).
+
+These findings are suggestive that in the shared utility condition, games differentiated less from each other, and their conventions may have been influenced more by shared priors for how to describe the flowers.
 
 # Discussion
-TODO!
 
-Many of the key reduction findings from tangrams generalize to this situation.
-Specifically, we see utterance reduction over time and w/i group convergence for each image and divergence between images.
-This situation is different in that we have different stimuli (more natural) and the set up is collaborative and more free-form in what is talked about.
-These patterns hold for both the individual and group payoff structures.
+This analysis of the referring expressions in @mankewitz2021 examined three key questions: the reduction of referring expressions, the convergence of referring expressions to conventions, and the differences between the shared and individual utility conditions.
 
-One difference we see is that groups don’t diverge (from each other). This may be dependent on stimulus properities (are there universal features of some of the images?) and group dynamics
+First, we saw reduction, as referring expressions tended to decrease in length over the course of the game.
+Second, we found the expected trifecta of convergence and divergence results.
+Over the course of games, references to the same flower within a game converged both to other descriptions used at that timepoint and to the post-game descriptions.
+This convergence was specific to flower and games, as references to different flowers within a game and  references to the same flower across games both diverged.
 
-mention the set of departures this made from classic reference game
+These results align with the expected pattern of partner-specific convention formation established in dyadic reference games.
+These patterns of reduction and convergence confirm some generalizability of the reference game patterns to the freer form and more naturalistic domain of a negotiation game.
+While still an artificial situation, this game differs from classic reference games in that the set-up was symmetric, with all players having equivalent amounts of knowledge and authority, and that the need for reference was embedding within a more complex goal structure.
+This game also was aytpical for reference games in that it had more than 2 players [although see @yoonAudienceDesignMultiparty2019;@boyce2022] and used photos as target images [although see @weber2003].
 
-Conclusion: The key reference game findings have some generalizability. Settings like this one may be useful for encouraging discussion of a set of images and setting up partial knowledge situations.
+Our last consideration was how these reduction and convergence effects might be moderated by the shared and individual utility conditions.
+We found slight condition differences where being in the shared utility condition led to stronger alignment between players throughout.
+We hypothesize that they might be due to a perceived greater importance of communicating with each other or paying greater attention to what groupmates are seeing.
+In addition, groups in the shared utility condition diverged less from each other.
+This reduced divergence suggests that their descriptions and conventions were less ideosyncratic and perhaps closer to shared underlying priors about how to describe the images.
+We hypothesize that the explicitly more cooperative goal might have led to participants caring more about being understood by their partners.
+This is interesting in light of @guilbeault2021 finding that larger groups tend to all converge to similar conventions whereas smaller groups can support ideosyncratic conventions.
+
+However, we are cautious not to over-interpret the condition differences.
+There were a limited number of groups in each condition, so these effects may have been driven or exacerbated by heterogeneity in which participants were in each group.
+The differences are suggestive, but these ideas will need to be tested further with larger samples and more different goal structures.
+
+Our interpretation of results is limited by the dataset, which is small, especially compared to recent datasets of convention formation in reference games [@hawkins2020, @boyce2022].
+The dataset also has large item-level variation, as the images differ in how nameable the flowers are or whether their are salient features.
+
+In conclusion, this analysis showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
+Additionally, we saw hints that different incentives may lightly alter moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.
 

### 2023-01-17 code/annotation_analysis.Rmd
fussing with tables

@@
   badlist <- c("accept_stat__","divergent__","energy__","n_leapfrog__","sigma", "stepsize__","treedepth__")
-  intervals <- gather_draws(model, `[a-zA-z_(): ]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
+  intervals <- gather_draws(model, `[a-zA-Z_(): ]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
     mutate(lower=round(.lower, digits=3),

### 2023-01-17 write-ups/cogsci23_paper/manuscript.Rmd
fussing with tables

@@
 library(lsa)
+library(kableExtra)
 #rstan_options(auto_write = TRUE)
@@
 
-When doing pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same round and the two prior rounds (thus the similarity at round 10 comes from 8-10, 9-10 and 10-10 comparisons).
+When doing pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same round and the two later rounds (thus the similarity at round 10 comes from 10-10, 10-11 and 10-12 comparisons).
 This smoothing was not necessary when comparing with the names produced after the game.
@@
 ## Analytic methods
-TODO this is the section where we talk about modelling details
+All models were run using rstan [@rstan]. We treatment-coded condition, with individual utilities as the baseline condition, and shared utilities as the marked condition.
 
@@
 ## Reduction of referring expressions
+
+```{r}
+
+include <- tribble(
+  ~gameId, ~name, ~flower, ~trialNum, ~normalized_span,
+  "cK5bFKgBKPiq4oCki", "Laju", "11yellow", 1, "not sure what kind of flower it is but the droopy-ish one",
+  "cK5bFKgBKPiq4oCki", "Repi", "11yellow", 3, "droopy iris flower",
+  "cK5bFKgBKPiq4oCki", "Minu", "11yellow", 20, "droopy",
+  "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 1, "the red middle with spike",
+    "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 2, "the red center",
+    "DTw5mrEvPvfyuEiGw", "Laju", "2yellow", 19, "red middle",
+  "KiRHskdD2iLy8Fh8C", "Repi", "2yellow", 5, "the one with the dark red centre",
+    "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 12, "the one with black background",
+      "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 23, "black background",
+    "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 3, "the big cluster of flowers with the orange in the middle",
+      "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 22, "cluster",
+      "DTw5mrEvPvfyuEiGw", "Repi", "4yellow", 23, "bundle",
+        "KiRHskdD2iLy8Fh8C", "Minu", "4yellow", 23, "multi flowers",
+)
+
+examples <- read_rds(here(model_path,"sbert_examples.rds")) |> left_join(include, by=c("game1"="gameId","ref1"="flower","player1"="name","trialNum1"="trialNum")) |> rename(text1=normalized_span) |> left_join(include, by=c("game2"="gameId","ref2"="flower","player2"="name","trialNum2"="trialNum")) |> rename(text2=normalized_span)
+
+```
+
+```{r examples}
+text_examples <- include |> 
+  mutate(Game=as.factor(gameId) |> as.numeric() |> as.character(),
+         Player=LETTERS[as.factor(name) |> as.numeric()],
+         GamePlayer=str_c(Game,Player),
+         Trial=trialNum+1,
+         Flower=as.factor(flower) |> as.numeric() |> as.character()) |> arrange(Flower,Game,Trial) |> 
+  select(Flower,Game=GamePlayer, Trial,  Expression=normalized_span )
+
+knitr::kable(text_examples, caption = "Examples of descriptions of different flowers from different games, illustrating reduction and convergence phenomenon.
+Games are numbered, letters refer to players within games.
+Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.",  format="latex", booktabs=TRUE, linesep="") %>%
+kable_styling(full_width = F) %>%
+  row_spec(c(3,9), hline_after=T) |> 
+column_spec(4, width = "12em") %>%
+  kable_styling(latex_options="H")
+
+```
+
+```{r sbert-example}
+sbert_include <- tribble(
+  ~text1, ~text2,
+  #"the red center", "the one with the dark red centre",
+  "the red center", "red middle",
+  "droopy iris flower", "droopy",
+  "cluster", "bundle",
+  "droopy iris flower","the red center", 
+  "droopy iris flower","multi flowers", 
+  "red middle", "black background",
+  "droopy", "bundle",
+) |> left_join(examples) |> arrange(desc(sim)) |> mutate(sim=round(sim,2)) |> select(`Expression 1`=text1, `Expression 2`=text2, `Sim`=sim)
+
+knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions.",  format="latex", booktabs=TRUE, linesep="") %>%
+kable_styling(full_width = F) %>%
+column_spec(c(1,2), width = "9em") %>% 
+  kable_styling(latex_options="H")
+
+```
 
@@
               distance=factor(distance, levels=c("Same Person", "Same Game - Different Person", "Different Games"))) |> 
-ggplot( aes(x=later,y=sim,color=condition))+     
+ggplot( aes(x=earlier,y=sim,color=condition))+     
   stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
@@
 
-```{r }
+```{r, cache=T}
 
@@
 
-```{r}
+```{r, cache=T}
 in_end <- M_mat %>% 

### 2023-01-17 write-ups/cogsci23_paper/manuscript.Rmd
fussing with table placement

@@
 
-1. Do the referring expressions get shorter over time?
+1. Do the referring expressions reduce over time?
 
-2. How do referring expressions change within and between groups over the course of the games? Do descriptions converge within groups and diverge between groups?
+2. How do referring expressions change within and between groups over the course of the games?
 
@@
 If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they each got a lower reward instead (Figure \ref{interface}B).
-This incentivized participants to communicate about the flowers in order to coordinate on selecting differenct flowers.
+Players needed to communicate about the flowers in order to coordinate their selections.
 The rewards earned over the course of game translated in a monetary bonus for the participants at the end of the game.
 
-In *individual utility* games, each player earned points for the flowers they selected; in the *shared utility* games, the points earned were averaged together, and all players in a game got the same reward.
+In *individual utility* games, each player earned points for the flowers they selected; in the *shared utility* games, the points were averaged together, and all players in a game got the same reward.
 This made for slightly different incentives; in an individual game, players wanted to maximize the rewards of flowers they selected, and only cared about avoiding collisions with other players' selections; in a shared game, players wanted their teammates to select different high reward flowers, and were indifferent on who selected the highest one.
@@
 After the game, players were asked how they would describe each of the images they had played with to their teammates.
-They were also asked how they would describe each of 4 images from a different color set than the one they had used.
 
 ## Participants
-@mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants for 25 3-person games in each of the individual and shared utilities conditions where they had access to a chat box (there were also no-chat-box baseline conditions, which we did not analyse).
-Some games did not complete the full 24 rounds because participants dropped out early.
+@mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants for 25 3-person games in each of the individual and shared utilities conditions.
+Some games did not complete the full 24 trials because participants dropped out early.
 We excluded incomplete games from analyses, which left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the individual utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the shared utilities condition (`r 3*(d.games.final |> nrow())` participants included in the analyses).
 
-## Textual annotations
-
-```{r, include=F}
-vb <- read_csv(here(annotation_path, "vb_agreement.csv")) %>% rename(vb_span=span,vb_referent=referent) %>% 
-  mutate(vb_span=ifelse(str_starts(vb_span, "the "), str_sub(vb_span, 5, -1), vb_span)) %>% 
-  mutate(row=row_number())
-bl <- read_csv(here(annotation_path, "bl_agreement.csv")) %>% rename(bl_span=span,bl_referent=referent) %>% 
-  mutate(bl_span=str_sub(bl_span, 2,-2)) %>% #remove quote marks
-  mutate(row=row_number())
-
-both <- vb %>% full_join(bl) %>% 
-  mutate(same_span=ifelse(vb_span==bl_span, 1,0),
-         same_ref=ifelse(vb_referent==bl_referent, 1,0)) %>% 
-  select(text,vb_span, bl_span, vb_referent,bl_referent, same_span, same_ref) |> 
-  filter(!(is.na(vb_span)&is.na(bl_span)))
-
-raw_rows <- vb |> select(text) |> nrow()
-vb |> select(gameId) |> unique() |> nrow()
-
-agree_span <- both %>% group_by(same_span) |> tally() |> filter(same_span==1) |> pluck(2)
-
-count <- both %>% filter(!is.na(bl_referent))
-
-agree_target <- count |> group_by(same_ref) %>% tally() |> filter(same_ref==1) |> pluck(2)
-
-```
-
-We annotated the chat transcripts to extract all referring expressions and identify which flower image each expression referred to.
-We also spellchecked and corrected the referring expressions.
-Annotations were done primarily by the first author, with some done by a research assistant.
-
-Two games consisting of `r raw_rows` utterances were annotated by both annotators.
-`r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
-The second annotator coded the target of the referring expression for `r nrow(count)` of the utterances.
-Of these, the two annotators agreed on the target in `r agree_target` (`r round(agree_target/nrow(count)*100)`%) of cases.
-We take this high level of interannotator agreement as an indication that the reference spans and targets were identified in a reliable way.
-
-We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions. 
 ```{r wordcount, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
@@
 
+## Textual annotations
+
+```{r, include=F}
+vb <- read_csv(here(annotation_path, "vb_agreement.csv")) %>% rename(vb_span=span,vb_referent=referent) %>% 
+  mutate(vb_span=ifelse(str_starts(vb_span, "the "), str_sub(vb_span, 5, -1), vb_span)) %>% 
+  mutate(row=row_number())
+bl <- read_csv(here(annotation_path, "bl_agreement.csv")) %>% rename(bl_span=span,bl_referent=referent) %>% 
+  mutate(bl_span=str_sub(bl_span, 2,-2)) %>% #remove quote marks
+  mutate(row=row_number())
+
+both <- vb %>% full_join(bl) %>% 
+  mutate(same_span=ifelse(vb_span==bl_span, 1,0),
+         same_ref=ifelse(vb_referent==bl_referent, 1,0)) %>% 
+  select(text,vb_span, bl_span, vb_referent,bl_referent, same_span, same_ref) |> 
+  filter(!(is.na(vb_span)&is.na(bl_span)))
+
+raw_rows <- vb |> select(text) |> nrow()
+vb |> select(gameId) |> unique() |> nrow()
+
+agree_span <- both %>% group_by(same_span) |> tally() |> filter(same_span==1) |> pluck(2)
+
+count <- both %>% filter(!is.na(bl_referent))
+
+agree_target <- count |> group_by(same_ref) %>% tally() |> filter(same_ref==1) |> pluck(2)
+
+```
+
+We annotated the chat transcripts to extract all referring expressions and identify which flower image each expression referred to.
+We corrected the referring expressions for spelling errors.
+Annotations were done primarily by the first author, with some done by a research assistant.
+
+Two games consisting of `r raw_rows` utterances were annotated by both annotators.
+`r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
+The second annotator coded the target of the referring expression for `r nrow(count)` of the utterances.
+Of these, the two annotators agreed on the target in `r agree_target` (`r round(agree_target/nrow(count)*100)`%) of cases.
+We take this high level of interannotator agreement as an indication that the reference spans and targets were identified reliably.
+
+We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions.
+
 ## SBERT embeddings
-Whereas in the past studies of convention formation have had to rely on manual assessments of similarity or lexical overlap, natural language processing tools now make it possible to quantitatively measure the similarity between phrases.
-Recent work by @Hawkins2020 and @boyce2022 have used these tools to assess the dynamics of convergence to conventions.
+Following recent work by @hawkins2020 and @boyce2022, we used tools for natural langauge processing to quantify similarity between phrases in order to assess the dynamics of convergence to conventions.
 
-We embedded each of the extracted referring expressions using SBERT [@reimers2019] which maps each utterance to a vector in a high-dimensional semantic vector space.
+We embedded each of the extracted referring expressions using SBERT [@reimers2019] which maps each utterance to a vector in a high-dimensional semantic space.
 We use cosine distance between pairs of vector embeddings as a measure of similarity between the corresponding referring expressions.
-When two vectors are near each other, the angle between them is small, and the cosine of this angle is large.
 Higher cosine values correspond to vectors close together, and thus, reference expressions with more similar meanings.
+Some examples of pairs of referring expressions and their similarities are shown in Table \ref{costable}.
 
-When doing pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same round and the two later rounds (thus the similarity at round 10 comes from 10-10, 10-11 and 10-12 comparisons).
+For pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same trial and the two later trials (thus the similarity at trial 10 comes from 10-10, 10-11 and 10-12 comparisons).
 This smoothing was not necessary when comparing with the names produced after the game.
@@
 ## Reduction of referring expressions
+
+```{r}
+possible <- d.rounds.final |> select(-name) |> unique()
+
+per_game <-  annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+
+mean_flowers <- per_game |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
+  select(condition, summ)
+
+per_person <- annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
+  mutate(n=ifelse(is.na(n),0,n))
+mean_person <- per_person |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
+  select(condition, summ)
+```
+
+```{r,eval=F, include=F}
+
+m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
+show_summary_rstan(m1) |> write_rds(here(model_path,"word_model.rds"))
+```
+
+```{r}
+
+word_count <- read_rds(here(model_path,"word_model.rds")) |> filter(str_detect(Term,"\\[", negate=T)) #because there are problems with a couple mixed effects getting through
+```
+
+Each trial of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
+In general, each trial contained references to 2 or 3 distinct flowers (individual utilities: mean of `r  mean_flowers |> pluck(2,1)`, shared utilities: `r mean_flowers |> pluck(2,2)`).
+Most players referred to one flower each trial (individual utilities: `r  mean_person |> pluck(2,1)`, shared utilities: `r mean_person |> pluck(2,2)`).
+
+The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
+As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game (trial: `r stats_text(word_count,3)`).
+The reduction in referring language was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per trial remained constant (Figure \ref{wordcount}C).
+A couple examples of reduction of referring expressions are shown in Table \ref{extable}, flowers 1 and 2.
 
@@
 Games are numbered, letters refer to players within games.
-Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.",  format="latex", booktabs=TRUE, linesep="") %>%
+Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.
+\\label{extable}",  format="latex", booktabs=TRUE, linesep="") %>%
 kable_styling(full_width = F) %>%
@@
 
-knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions.",  format="latex", booktabs=TRUE, linesep="") %>%
+knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions. \\label{costable}",  format="latex", booktabs=TRUE, linesep="") %>%
 kable_styling(full_width = F) %>%
@@
 ```
-
-```{r}
-possible <- d.rounds.final |> select(-name) |> unique()
-
-per_game <-  annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
-  mutate(n=ifelse(is.na(n),0,n))
-
-mean_flowers <- per_game |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
-  select(condition, summ)
-
-per_person <- annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
-  mutate(n=ifelse(is.na(n),0,n))
-mean_person <- per_person |> group_by(condition) |> summarize(mean=mean(n) |> round(2), sd=sd(n) |> round(2), summ=str_c(mean," (sd: ",sd,")")) |> 
-  select(condition, summ)
-```
-
-```{r,eval=F, include=F}
-
-m1 <- stan_lmer(numword ~ trialNum + condition + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))
-show_summary_rstan(m1) |> write_rds(here(model_path,"word_model.rds"))
-```
-
-```{r}
-
-word_count <- read_rds(here(model_path,"word_model.rds")) |> filter(str_detect(Term,"\\[", negate=T)) #because there are problems with a couple mixed effects getting through
-```
-
-Each round of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
-In general, each trial contained references to 2 or 3 distinct flowers (Individual Utilities: mean of `r  mean_flowers |> pluck(2,1)`, Shared Utilities: `r mean_flowers |> pluck(2,2)`).
-Most players referred to one flower each round (Individual Utilities: `r  mean_person |> pluck(2,1)`, Shared Utilities: `r mean_person |> pluck(2,2)`).
-
-The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
-As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game (trial: `r stats_text(word_count,3)`.
-The reduction in referring language was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per round remained constant (Figure \ref{wordcount}C).
-
-## Convergence of referring expressions
 
@@
 ```{r during, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Cosine similarities between SBERT embeddings of utterances produced within 2 trials of each other.
-Utterances were paired with other utterances from different games; from the same game, but a different utterer speaker; or from the same speaker (vertical panels).
-Utterances were either in reference to different flowers or both in reference to the same flower (horizontal panels).
-\\label{during}", fig.height=4, fig.width=10, out.width="100%"}
+Utterances were paired with other utterances from the same speaker; from the same game, but a different speaker; or from a different game (vertical panels).
+Utterances were either in reference to the same flower or different flowers (horizontal panels).
+\\label{during}", fig.height=4, fig.width=10, out.width="80%"}
 flower_all|> 
@@
 
+## Convergence of referring expressions
+
 The classic reference game phenomenon of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
 We assess all of these in two ways: 1) comparing referring expressions within the game to other referring expressions from the same timepoint to get a sense of the dynamic changes, and 2) comparing referring expressions within the game to the names players gave the flowers after the game, to track how referring expressions evolve towards these final conventions.
-
-TODO table with examples of flower descriptions over time and cos similarities !!!
 
@@
 Within games, the patterns from dyadic reference games predict that similarity for descriptions of the same image will increase over time while similarity for descriptions of different images will decrease over time.
+Table \ref{extable} offers examples of how some descriptions changed over time, and Table \ref{costable} gives similarity measures between selected pairs of these descriptions.
 
@@
 The theory of partner-specificity predictions that descriptions to the same flower in different groups will diverge over time as each group forms their own group-specific convention.
+Table \ref{extable} shows how groups descriptions of the same flower may diverge (flower 2), but also how groups might choose related but different descriptions (flower 3).
 
@@
 
+After the game, players were asked to provide the description they would use to identify each flower to their teammates.
+We treated these descriptions as the conventionalized names and looked at how the in-game descriptions developed into these conventions.
+
+Referring descriptions later in a game were more similar to the convention than descriptions earlier in the game  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} upper panels).
+Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
+This is consistent with the expected pattern of convention formation.
+
+We also examined how referring expressions related the conventionalized names for other flowers.
+Utterances did not significantly diverge from the conventionalized names of other flower over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
+There was not an effect of being said by the same person who gave the convention (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
+
 ```{r withinend, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="90%", fig.width=4, fig.height=2, fig.cap = "Cosine similarities between flower descriptions provided by participants after the game.
@@
 ```
-
-After the game, players were asked to provide the description they would use to identify each flower to their teammates.
-We treated these descriptions as the conventionalized names and looked at how the in-game descriptions developed into these conventions.
-
-Referring descriptions later in a game were more similar to the convention than descriptions earlier in the game  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} upper panels).
-Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
-This is consistent with the expected pattern of convention formation.
-
-We also examined how referring expressions related the conventionalized names for other flowers.
-Utterances did not significantly diverge from the conventionalized names of other flower over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
-There was not an effect of being said by the same person who gave the convention (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
-
-## Individual and Shared Utility conditions
+## Individual and shared utility conditions
 

### 2023-01-18 write-ups/cogsci23_paper/manuscript.Rmd
apply Mike's advice

@@
 Successful communication is grounded in a shared understanding of what utterances mean in the context they are produced in.
+Whether the communicative goals are cooperative or adversarial, interlocuters need to establish joint reference to be effective.
 In many cases conventional word meanings are enough, but there are also contexts where objects to be referred to are not easy to distinguish.
 In these situations, establishing reference is more difficult, but no less important.
+<!--For one, in everyday communication, establishing reference rarely happens in isolation but is usually a subgoal in a larger conversational goal.
+The overall goal might be cooperative, such as in asking for a piece of kitchen equipment while cooking together, but it can also be adversarial, such as in a negotiation where the assets to be divied up must be referenced.
+In the reference game context, the goal is always cooperative, to match the images or get them in the same order.
+-->
 
@@
 These nicknames conventionalize with a pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
+This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO CITES].
 
-This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO CITES].
-The implicit theoretical claim is that these phenomena hold whenever people interact repeatedly in ways that require reference to some objects that do not have pre-conventionalized names that are adequate to distinguish them in context.
+The implicit theoretical claim is that reduction to partner-specific conventions occurs whenever people interact repeatedly in ways that require reference to some objects that do not have adequate pre-conventionalized names.
+Do these phenomena generalize?
 
-However, there are some ways in which this microcosm is not representative of typical language use.
-For one, in everyday communication, establishing reference rarely happens in isolation but is usually a subgoal in a larger conversational goal.
-The overall goal might be cooperative, such as in asking for a piece of kitchen equipment while cooking together, but it can also be adversarial, such as in a negotiation where the assets to be divied up must be referenced.
-In the reference game context, the goal is always cooperative, to match the images or get them in the same order.
-
-TODO paragraph on on the issue of shared vs.
-privileged knowledge, which has been studied extensively in communication psycholinguistics (e.g., heller, keysar, barr) but IIRC not in the context of creating novel referring expressions in the tangram task.
-so we are blending in this other important element of naturalism, albeit only around the values and not the actual stuff?
+Our current work begins to address this by looking for these phenomena in a slightly more naturalistic setting.
+One way in which reference games have not been representative of typical language use is in there isolated goal of reference.
+In everyday communication, establishing reference rarely happens in isolation, but is often an instrumental subgoal in a larger conversation.
+The overall task may be cooperative, such as asking for kitchen equipment while cooking together, or adversarial, or such identifying items while negotiating a division of assets.
+In the reference game context, the goal is explicitly cooperative: match corresponding images.
 
@@
 
-@mankewitz2021 tried to bridge the gap between datasets of negotiation over easily nameable objects, and the reference game literature where there is reference (but no negotiation) over hard to name objects.
+@mankewitz2021 started to bridge the gap between reference games with hard-to-name objects and negotation datsets wtih easily nameable objects.
 They created a 3 person game where players each select what flowers to grow from a set of 6 flowers with variable and partially hidden value.
@@
 @mankewitz2021 found a slight decrease in language over the course of the game, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
+Here, we separate out the referring expressions from other language produced in order to check for patterns of convention-formation.
 
-Historically, studies of convention formation have had to rely on proxies such as reduction and fluency, qualitatively assess similarity of utterances, or manually measure lexical overlap in small datasets to get at the ideas of convergence and divergence.
-More recent work has taken advantage of natural language processing tools for mapping words and sentences to semantic vector spaces in order to quantify semantic similarities between sentences [@hawkins2020, @boyce2022].
-We follow this recent work in using vector embeddings to measure semantic distances between referring expressions in the @mankewitz2021 dataset.
-
-## Current work
-Here we ask whether participants in a complex, multi-player strategic game with differing incentives and shared knoweldge were able to develop conventionalized reference expressions.
+Our current work addresses whether participants in a complex, multi-player strategic game with differing incentives and shared knowledge were able to develop conventionalized reference expressions.
 We explored whether the phenomena found in classic dyadic iterated reference games extended to this more naturalistic domain, and whether they differed across the individual and shared incentive conditions.
@@
 
-1. Do the referring expressions reduce over time?
+1. Did the referring expressions reduce over time?
 
-2. How do referring expressions change within and between groups over the course of the games?
+2. How did referring expressions change within and between groups over the course of the games?
 
-3. How do the different incentive conditions influence the referring expressions?
+3. How did the different incentive conditions influence the referring expressions?
+
+Historically, studies of convention formation have had to rely on proxies such as reduction, subjective judgements of similarity, and and manual measures of lexical overlap.
+In contrast, we use natural language processing tools for mapping sentences to semantic vector spaces to quantify semantic similarities between sentences, following recent work by @hawkins2020 and @boyce2022.
 
@@
 
-### Comparison with post-game names
+### Comparison with post-game convention reports
 
@@
 We also examined how referring expressions related the conventionalized names for other flowers.
-Utterances did not significantly diverge from the conventionalized names of other flower over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
+Utterances diverge from the conventionalized names of other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
 There was not an effect of being said by the same person who gave the convention (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
@@
 First, we saw reduction, as referring expressions tended to decrease in length over the course of the game.
-Second, we found the expected trifecta of convergence and divergence results.
-Over the course of games, references to the same flower within a game converged both to other descriptions used at that timepoint and to the post-game descriptions.
+Second, we found the expected set of convergence and divergence patterns.
+References to the same flower within a game converged both to other descriptions used at that timepoint and to the post-game descriptions.
 This convergence was specific to flower and games, as references to different flowers within a game and references to the same flower across games both diverged.
 
-These results align with the expected pattern of partner-specific convention formation established in dyadic reference games.
 These patterns of reduction and convergence confirm some generalizability of the reference game patterns to the freer form and more naturalistic domain of a negotiation game.
-While still an artificial situation, this game differs from classic reference games in that the set-up was symmetric, with all players having equivalent amounts of knowledge and authority, and that the need for reference was embedding within a more complex goal structure.
-This game also was aytpical for reference games in that it had more than 2 players [although see @yoonAudienceDesignMultiparty2019;@boyce2022] and used photos as target images [although see @weber2003].
+While still an artificial situation, this game differed from classic reference games in that the set-up was symmetric, with all players having equivalent amounts of knowledge and authority, and that the need for reference was embedding within a more complex goal structure.
+This game also was aytpical for reference games in that it had more than 2 players [although see @boyce2022; @yoonAudienceDesignMultiparty2019] and used photos as target images [although see @weber2003].
 
-Our last consideration was how these reduction and convergence effects might be moderated by the shared and individual utility conditions.
-We found slight condition differences where being in the shared utility condition led to stronger alignment between players throughout.
-We hypothesize that they might be due to a perceived greater importance of communicating with each other or paying greater attention to what groupmates are seeing.
+Our third consideration was how the different utility conditions moderated the reduction and convergence effects.
+We found slight condition differences where being in the shared utility condition meant more similar descriptions for the same flowers and more different descriptions for different.
 In addition, groups in the shared utility condition diverged less from each other.
-This reduced divergence suggests that their descriptions and conventions were less ideosyncratic and perhaps closer to shared underlying priors about how to describe the images.
-We hypothesize that the explicitly more cooperative goal might have led to participants caring more about being understood by their partners.
-This is interesting in light of @guilbeault2021 finding that larger groups tend to all converge to similar conventions whereas smaller groups can support ideosyncratic conventions.
+This pattern of condition differences is consistent with less idiosyncratic descriptions in the shared utility condition, possibly due to a greater concern with being understood by groupmates.
+The greater alignment between players (and between games) might be due to a perceived greater importance of communicating with each other and understanding what others say, possibly triggered by the cooperative goal.
+The similarity across games is particularly intriguing given @guilbeault2021, which found  that larger groups tend to all converge to similar conventions whereas smaller groups can support idiosyncratic conventions.
 
@@
 
+This dataset is just the start of testing how well findings from reference games generalize to more complex conversational situations.
+One dimension that we were unable to explore here due to the design of the game is the idea of shared versus privileged knowledge (CITATIONS).
+In the flowers game, like in reference games, all players saw the same set of images (and knew they did), which makes making descriptions to flowers easier.
+In real-world situations, interlocuters may have different sets of objects in their context and may not know what is in their interlocuters context.
+An important future direction is studying how repeated reference phenomena interact with different communication forms such as goals and contexts.
+
 In conclusion, this analysis showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
 Additionally, we saw hints that different incentives may lightly alter moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.
+We hope future work will continue to develop our understanding of these phenomena in more natural settings.
 

### 2023-01-26 write-ups/cogsci23_paper/manuscript.Rmd
minor edits

@@
     Reduction of referring expressions to short conventionalized terms is found robustly in dyadic repeated reference games when the target images are initially difficult to name.
-Establishing reference is a crucial part of many conversations with different goals and more complex motivations; thus, it is useful to look at how the findings from reference games, where participants share the explicit goal of establishing joint reference, carry over into situations where participants have different overarching goals.
+Establishing reference is crucial to many conversations with different goals and more complex motivations; thus, it is useful to look at how the findings from reference games, where participants share the explicit goal of establishing joint reference, carry over into situations where participants have different overarching goals.
 In the current work, we analyse a dataset where reference was embedded in strategic 3- player negotiation and coordination games.
@@
 
-The formation of these ad-hoc referring expressions has been studied extensively in the context of dyadic reference games [@clarkReferringCollaborativeProcess1986, ADD MORE CITES].
+The formation of these ad-hoc referring expressions has been studied extensively in the context of dyadic reference games [@clarkReferringCollaborativeProcess1986; @krauss1964changes].
 In these, two participants see a set of images (often abstract shapes) and one person is tasked with identifying an image so the other person can pick it out.
@@
 These nicknames conventionalize with a pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
-This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, etc [TODO CITES].
+This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, and when they change [@brennan1996conceptual; @metzing2003conceptual; @leung2020parents; @weber2003; @yoonAudienceDesignMultiparty2019].
 
-The implicit theoretical claim is that reduction to partner-specific conventions occurs whenever people interact repeatedly in ways that require reference to some objects that do not have adequate pre-conventionalized names.
-Do these phenomena generalize?
+The implicit theoretical claim is that reduction to partner-specific conventions occurs whenever people interact repeatedly in ways that require reference to some objects without adequate conventional names.
+Do these phenomena generalize beyond artificial reference game environments?
 
 Our current work begins to address this by looking for these phenomena in a slightly more naturalistic setting.
-One way in which reference games have not been representative of typical language use is in there isolated goal of reference.
-In everyday communication, establishing reference rarely happens in isolation, but is often an instrumental subgoal in a larger conversation.
-The overall task may be cooperative, such as asking for kitchen equipment while cooking together, or adversarial, or such identifying items while negotiating a division of assets.
+The conversational goals of reference games are not representative of typical language use.
+In everyday communication, establishing reference is often an instrumental subgoal in a larger conversation.
+The overall task may be cooperative, such as asking for kitchen equipment while cooking together, or adversarial, such as identifying items while negotiating a division of assets.
 In the reference game context, the goal is explicitly cooperative: match corresponding images.
@@
 
-@mankewitz2021 started to bridge the gap between reference games with hard-to-name objects and negotation datsets wtih easily nameable objects.
-They created a 3 person game where players each select what flowers to grow from a set of 6 flowers with variable and partially hidden value.
-Players got the value of the flower only if they were the only one to select it, which incentivized coordination and negotiation between players.
+@mankewitz2021 bridged the gap between reference games with hard-to-name objects and negotation datsets with easily nameable objects.
+They created a 3-person game where players each selected what flowers to grow from a set of 6 flowers with variable and partially hidden value.
+Players got the value of the flower only if they were the only one to select it, which incentivized coordination and negotiation.
 However, all the flowers shown were the same color, so they were not easily nameable, and the flowers repeated, leading to opportunities for ad-hoc conventionalization.
-This game structure encouraged players to describe flowers while never explicitly saying that a player should refer to any specific flower.
-They also had two conditions, one where players within a group won points together (and thus had fully aligned incentives) and another where they won points as individuals (and didn't).
-@mankewitz2021 found a slight decrease in language over the course of the game, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
-Here, we separate out the referring expressions from other language produced in order to check for patterns of convention-formation.
+This game structure encouraged players to describe flowers while never explicitly telling a player to refer to any specific flower.
+The experiment had two conditions, one where players within a group won points together (and thus had fully aligned incentives) and another where they won points as individuals (and didn't).
+@mankewitz2021 found a slight decrease in language over the course of the games, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
+Here, we separated out the referring expressions in order to check for patterns of convention-formation.
 
@@
 On each trial, each group saw a set of 6 flower images (Figure \ref{interface}A) .
-Each participant saw the values for four of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant.
+Each participant saw the values for 4 of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant.
 Players could coordinate and discuss using a chat box before each selected a flower.
@@
 Players needed to communicate about the flowers in order to coordinate their selections.
-The rewards earned over the course of game translated in a monetary bonus for the participants at the end of the game.
+The rewards translated in a monetary bonus for the participants at the end of the game.
 
@@
 Each game consisted of 24 trials.
-The use of different flowers of the same color created situations where players did not have established names for the flowers in context and needed to develop shared referring expressions to clearly communicate with their partners.
+Different flowers of the same color created situations where players did not have established names for the flowers in context and needed to develop shared referring expressions to clearly communicate with their partners.
 
-After the game, players were asked how they would describe each of the images they had played with to their teammates.
+After the game, players were asked how they would describe each of the images to their teammates.
 
@@
 
-```{r wordcount, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Amount of words produced in referring expressions across trial in games in both conditions.
+```{r wordcount, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Words produced in referring expressions across trials in both conditions.
 A: Total words of referring language per game per trial.
@@
 `r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
-The second annotator coded the target of the referring expression for `r nrow(count)` of the utterances.
+The second annotator only coded the target of the referring expression for `r nrow(count)` of the utterances.
 Of these, the two annotators agreed on the target in `r agree_target` (`r round(agree_target/nrow(count)*100)`%) of cases.
-We take this high level of interannotator agreement as an indication that the reference spans and targets were identified reliably.
+We take this level of interannotator agreement as an indication that the reference spans and targets were identified reliably.
 
@@
 For pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same trial and the two later trials (thus the similarity at trial 10 comes from 10-10, 10-11 and 10-12 comparisons).
-This smoothing was not necessary when comparing with the names produced after the game.
 
 ## Analytic methods
-All models were run using rstan [@rstan]. We treatment-coded condition, with individual utilities as the baseline condition, and shared utilities as the marked condition.
+All models were run using rstan [@rstan]. We coded condition as 0 for individual utilities (baseline condition) and 1 for shared utilities.
 
@@
 
-Each trial of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
+In each trial of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
 In general, each trial contained references to 2 or 3 distinct flowers (individual utilities: mean of `r  mean_flowers |> pluck(2,1)`, shared utilities: `r mean_flowers |> pluck(2,2)`).
@@
 
-The amount of referring language decreases over the course of the game, consistent with the dyadic reference game pattern.
-As shown in Figure \ref{wordcount}A, the number of words of referring language decreases across the game (trial: `r stats_text(word_count,3)`).
-The reduction in referring language was driven by shorter referring expressions later in the game (Figure \ref{wordcount}B), while the total number of referring expressions per trial remained constant (Figure \ref{wordcount}C).
-A couple examples of reduction of referring expressions are shown in Table \ref{extable}, flowers 1 and 2.
+The amount of referring language decreased over the course of the game, consistent with the dyadic reference game pattern.
+As shown in Figure \ref{wordcount}A, the number of words of referring language decreased across the game (trial: `r stats_text(word_count,3)`).
+The reduction in referring language was driven by shortening referring expressions (Figure \ref{wordcount}B), while the total number of referring expressions per trial remained constant (Figure \ref{wordcount}C).
+See Table \ref{extable}, flowers 1 and 2, for examples of reduction.
 
-```{r}
+## Convergence of referring expressions
 
-include <- tribble(
-  ~gameId, ~name, ~flower, ~trialNum, ~normalized_span,
-  "cK5bFKgBKPiq4oCki", "Laju", "11yellow", 1, "not sure what kind of flower it is but the droopy-ish one",
-  "cK5bFKgBKPiq4oCki", "Repi", "11yellow", 3, "droopy iris flower",
-  "cK5bFKgBKPiq4oCki", "Minu", "11yellow", 20, "droopy",
-  "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 1, "the red middle with spike",
-    "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 2, "the red center",
-    "DTw5mrEvPvfyuEiGw", "Laju", "2yellow", 19, "red middle",
-  "KiRHskdD2iLy8Fh8C", "Repi", "2yellow", 5, "the one with the dark red centre",
-    "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 12, "the one with black background",
-      "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 23, "black background",
-    "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 3, "the big cluster of flowers with the orange in the middle",
-      "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 22, "cluster",
-      "DTw5mrEvPvfyuEiGw", "Repi", "4yellow", 23, "bundle",
-        "KiRHskdD2iLy8Fh8C", "Minu", "4yellow", 23, "multi flowers",
-)
-
-examples <- read_rds(here(model_path,"sbert_examples.rds")) |> left_join(include, by=c("game1"="gameId","ref1"="flower","player1"="name","trialNum1"="trialNum")) |> rename(text1=normalized_span) |> left_join(include, by=c("game2"="gameId","ref2"="flower","player2"="name","trialNum2"="trialNum")) |> rename(text2=normalized_span)
-
-```
-
-```{r examples}
-text_examples <- include |> 
-  mutate(Game=as.factor(gameId) |> as.numeric() |> as.character(),
-         Player=LETTERS[as.factor(name) |> as.numeric()],
-         GamePlayer=str_c(Game,Player),
-         Trial=trialNum+1,
-         Flower=as.factor(flower) |> as.numeric() |> as.character()) |> arrange(Flower,Game,Trial) |> 
-  select(Flower,Game=GamePlayer, Trial,  Expression=normalized_span )
-
-knitr::kable(text_examples, caption = "Examples of descriptions of different flowers from different games, illustrating reduction and convergence phenomenon.
-Games are numbered, letters refer to players within games.
-Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.
-\\label{extable}",  format="latex", booktabs=TRUE, linesep="") %>%
-kable_styling(full_width = F) %>%
-  row_spec(c(3,9), hline_after=T) |> 
-column_spec(4, width = "12em") %>%
-  kable_styling(latex_options="H")
-
-```
-
-```{r sbert-example}
-sbert_include <- tribble(
-  ~text1, ~text2,
-  #"the red center", "the one with the dark red centre",
-  "the red center", "red middle",
-  "droopy iris flower", "droopy",
-  "cluster", "bundle",
-  "droopy iris flower","the red center", 
-  "droopy iris flower","multi flowers", 
-  "red middle", "black background",
-  "droopy", "bundle",
-) |> left_join(examples) |> arrange(desc(sim)) |> mutate(sim=round(sim,2)) |> select(`Expression 1`=text1, `Expression 2`=text2, `Sim`=sim)
-
-knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions. \\label{costable}",  format="latex", booktabs=TRUE, linesep="") %>%
-kable_styling(full_width = F) %>%
-column_spec(c(1,2), width = "9em") %>% 
-  kable_styling(latex_options="H")
-
-```
+The classic reference game phenomenon of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
+We assess all of these in two ways: 1) comparing referring expressions within the game to other referring expressions from the same time to get at the dynamic changes, and 2) comparing referring expressions within the game to the names players gave the flowers after the game, to track how expressions evolve towards these final conventions.
 
@@
 
+```{r}
+wi_g_wi_f <- read_rds(here(model_path,"samegame_sameflower_model.rds"))
+
+wi_g_bet_f <- read_rds(here(model_path,"samegame_diffflower_model.rds"))
+
+bet_g_wi_f <- read_rds(here(model_path,"diffgame_sameflower_model.rds"))
+```
+
+### Within games
+
+Within games, the patterns from dyadic reference games predict that similarity for descriptions of the same image will increase over time while similarity for descriptions of different images will decrease over time.
+Table \ref{extable} offers examples of how some descriptions changed over time, and Table \ref{costable} gives similarity measures between selected pairs of these descriptions.
+
+We found that for descriptions of the same flower, cosine similarity increased over time (trial: `r stats_text(wi_g_wi_f,6)`, see Figure \ref{during} upper left and middle panels).
+Utterances were more similar if they were produced by the same participant (`r stats_text(wi_g_wi_f,8)`), although this did not interact with trial number (`r stats_text(wi_g_wi_f,7)`).
+
+For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
+There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) .
+
+These results are consistent with the reference game phenomenon.
+
+```{r}
+
+include <- tribble(
+  ~gameId, ~name, ~flower, ~trialNum, ~normalized_span,
+  "cK5bFKgBKPiq4oCki", "Laju", "11yellow", 1, "not sure what kind of flower it is but the droopy-ish one",
+  "cK5bFKgBKPiq4oCki", "Repi", "11yellow", 3, "droopy iris flower",
+  "cK5bFKgBKPiq4oCki", "Minu", "11yellow", 20, "droopy",
+  "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 1, "the red middle with spike",
+    "DTw5mrEvPvfyuEiGw", "Minu", "2yellow", 2, "the red center",
+    "DTw5mrEvPvfyuEiGw", "Laju", "2yellow", 19, "red middle",
+  "KiRHskdD2iLy8Fh8C", "Repi", "2yellow", 5, "the one with the dark red centre",
+    "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 12, "the one with black background",
+      "KiRHskdD2iLy8Fh8C", "Laju", "2yellow", 23, "black background",
+    "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 3, "the big cluster of flowers with the orange in the middle",
+      "cK5bFKgBKPiq4oCki", "Laju", "4yellow", 22, "cluster",
+      "DTw5mrEvPvfyuEiGw", "Repi", "4yellow", 23, "bundle",
+        "KiRHskdD2iLy8Fh8C", "Minu", "4yellow", 23, "multi flowers",
+)
+
+examples <- read_rds(here(model_path,"sbert_examples.rds")) |> left_join(include, by=c("game1"="gameId","ref1"="flower","player1"="name","trialNum1"="trialNum")) |> rename(text1=normalized_span) |> left_join(include, by=c("game2"="gameId","ref2"="flower","player2"="name","trialNum2"="trialNum")) |> rename(text2=normalized_span)
+
+```
+
+```{r examples}
+text_examples <- include |> 
+  mutate(Game=as.factor(gameId) |> as.numeric() |> as.character(),
+         Player=LETTERS[as.factor(name) |> as.numeric()],
+         GamePlayer=str_c(Game,Player),
+         Trial=trialNum+1,
+         Flower=as.factor(flower) |> as.numeric() |> as.character()) |> arrange(Flower,Game,Trial) |> 
+  select(Flower,Game=GamePlayer, Trial,  Expression=normalized_span )
+
+knitr::kable(text_examples, caption = "Examples of descriptions of different flowers from different games, illustrating reduction and convergence phenomenon.
+Games are numbered, letters refer to players within games.
+Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.
+\\label{extable}",  format="latex", booktabs=TRUE, linesep="") %>%
+kable_styling(full_width = F) %>%
+  row_spec(c(3,9), hline_after=T) |> 
+column_spec(4, width = "12em") %>%
+  kable_styling(latex_options="H")
+
+```
+
+```{r sbert-example}
+sbert_include <- tribble(
+  ~text1, ~text2,
+  #"the red center", "the one with the dark red centre",
+  "the red center", "red middle",
+  "droopy iris flower", "droopy",
+  "cluster", "bundle",
+  "droopy iris flower","the red center", 
+  "droopy iris flower","multi flowers", 
+  "red middle", "black background",
+  "droopy", "bundle",
+) |> left_join(examples) |> arrange(desc(sim)) |> mutate(sim=round(sim,2)) |> select(`Expression 1`=text1, `Expression 2`=text2, `Sim`=sim)
+
+knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions. \\label{costable}",  format="latex", booktabs=TRUE, linesep="") %>%
+kable_styling(full_width = F) %>%
+column_spec(c(1,2), width = "9em") %>% 
+  kable_styling(latex_options="H")
+
+```
+
 ```{r during, fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Cosine similarities between SBERT embeddings of utterances produced within 2 trials of each other.
@@
 ```
-
-```{r}
-wi_g_wi_f <- read_rds(here(model_path,"samegame_sameflower_model.rds"))
-
-wi_g_bet_f <- read_rds(here(model_path,"samegame_diffflower_model.rds"))
-
-bet_g_wi_f <- read_rds(here(model_path,"diffgame_sameflower_model.rds"))
-```
-
-## Convergence of referring expressions
-
-The classic reference game phenomenon of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
-We assess all of these in two ways: 1) comparing referring expressions within the game to other referring expressions from the same timepoint to get a sense of the dynamic changes, and 2) comparing referring expressions within the game to the names players gave the flowers after the game, to track how referring expressions evolve towards these final conventions.
-
-### Within games
-
-Within games, the patterns from dyadic reference games predict that similarity for descriptions of the same image will increase over time while similarity for descriptions of different images will decrease over time.
-Table \ref{extable} offers examples of how some descriptions changed over time, and Table \ref{costable} gives similarity measures between selected pairs of these descriptions.
-
-We found that for descriptions of the same flower, cosine similarity increased over time (trial: `r stats_text(wi_g_wi_f,6)`, see Figure \ref{during} upper left and middle panels).
-Utterances were more similar if they were produced by the same participant (`r stats_text(wi_g_wi_f,8)`), although this did not interact with trial number (`r stats_text(wi_g_wi_f,7)`).
-
-For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
-There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) .
-
-These results mirror are consistent with the reference game phenomenon.
-
 ### Between games
 
-The theory of partner-specificity predictions that descriptions to the same flower in different groups will diverge over time as each group forms their own group-specific convention.
+The theory of partner-specificity predicts that descriptions to the same flower in different groups will diverge over time as each group forms their own distinct convention.
 Table \ref{extable} shows how groups descriptions of the same flower may diverge (flower 2), but also how groups might choose related but different descriptions (flower 3).
@@
 
-```{r toend, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.
+```{r toend, fig.env="figure", fig.pos = "h", fig.align = "center", out.width="100%", fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.
 \\label{toend}", fig.height=4, fig.width=6}
@@
 
-After the game, players were asked to provide the description they would use to identify each flower to their teammates.
+After the game, players provided the description they would use to identify each flower to their teammates.
 We treated these descriptions as the conventionalized names and looked at how the in-game descriptions developed into these conventions.
@@
 
-We also examined how referring expressions related the conventionalized names for other flowers.
+We also examined how referring expressions related to the conventionalized names for other flowers.
 Utterances diverge from the conventionalized names of other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
-There was not an effect of being said by the same person who gave the convention (`r stats_text(to_e_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(to_e_bet_f,7)`).
+There was no effect of the same person saying the description and convention (`r stats_text(to_e_bet_f,8)`) and no interaction between same person and trial number (`r stats_text(to_e_bet_f,7)`).
 
-```{r withinend, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="90%", fig.width=4, fig.height=2, fig.cap = "Cosine similarities between flower descriptions provided by participants after the game.
-Descriptions were more similar to each other if they came from the same group and were of the same flower.
+```{r withinend, fig.env="figure", fig.pos = "h", fig.align = "center", out.width="90%", fig.width=4, fig.height=2, fig.cap = "Cosine similarities between flower descriptions provided by participants after the game.
 \\label{withinend}"}
@@
 Here we report the extent to which the shared utility condition differed from the individual utility condition.
-
-There was not a significant difference in the amount of referring language produced in the two conditions (`r stats_text(word_count,2)`).
+Overall, there was not a significant difference in the amount of referring language produced in the two conditions (`r stats_text(word_count,2)`).
 
@@
 As a complement to this, games in the shared utilities condition had slightly lower similarities between descriptions of different flowers within games, but this did not interact with trial (`r stats_text(wi_g_bet_f,3)`) or whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
-Taken together, groups in the shared utility had more internal alignment on the mapping between flowers and names.
 
@@
 There was an interaction between shared utility condition and the referring expression coming from the same person who gave the convention for the different flower (`r stats_text(to_e_bet_f,5)`).
-
 Taken together, groups in the shared utility had more internal alignment on the mapping between flowers and names.
@@
 
-These patterns of reduction and convergence confirm some generalizability of the reference game patterns to the freer form and more naturalistic domain of a negotiation game.
-While still an artificial situation, this game differed from classic reference games in that the set-up was symmetric, with all players having equivalent amounts of knowledge and authority, and that the need for reference was embedding within a more complex goal structure.
-This game also was aytpical for reference games in that it had more than 2 players [although see @boyce2022; @yoonAudienceDesignMultiparty2019] and used photos as target images [although see @weber2003].
+These patterns of reduction and convergence confirm some generalizability of the reference game patterns to the freer-form and more naturalistic domain of a negotiation game.
+While still artificial, this game differed from classic reference games in that the set-up was symmetric, with all players having equivalent amounts of knowledge and authority, and that the need for reference was embedded within a more complex goal structure.
+This game also was atypical for reference games in that it had more than 2 players [although see @boyce2022; @yoonAudienceDesignMultiparty2019] and used photos as target images [although see @weber2003].
 
 Our third consideration was how the different utility conditions moderated the reduction and convergence effects.
-We found slight condition differences where being in the shared utility condition meant more similar descriptions for the same flowers and more different descriptions for different.
+We found slight condition differences where being in the shared utility condition meant more similar descriptions for the same flowers and more different descriptions for different flowers.
 In addition, groups in the shared utility condition diverged less from each other.
-This pattern of condition differences is consistent with less idiosyncratic descriptions in the shared utility condition, possibly due to a greater concern with being understood by groupmates.
-The greater alignment between players (and between games) might be due to a perceived greater importance of communicating with each other and understanding what others say, possibly triggered by the cooperative goal.
+This pattern of condition differences is consistent with less idiosyncratic descriptions in the shared utility condition.
+
+The greater alignment between players (and between games) in the shared utility condition might be due to a perceived greater importance of communicating with each other and understanding what others say, possibly triggered by the cooperative goal.
 The similarity across games is particularly intriguing given @guilbeault2021, which found that larger groups tend to all converge to similar conventions whereas smaller groups can support idiosyncratic conventions.
@@
 There were a limited number of groups in each condition, so these effects may have been driven or exacerbated by heterogeneity in which participants were in each group.
-The differences are suggestive, but these ideas will need to be tested further with larger samples and more different goal structures.
+These ideas will need to be tested further with larger samples and more different goal structures.
 
 Our interpretation of results is limited by the dataset, which is small, especially compared to recent datasets of convention formation in reference games [@hawkins2020, @boyce2022].
-The dataset also has large item-level variation, as the images differ in how nameable the flowers are or whether their are salient features.
+The dataset also has large item-level variation, as the images differ in how nameable the flowers are or whether there are salient features.
 
 This dataset is just the start of testing how well findings from reference games generalize to more complex conversational situations.
-One dimension that we were unable to explore here due to the design of the game is the idea of shared versus privileged knowledge (CITATIONS).
+One dimension that we were unable to explore here due to the design of the game is the idea of shared versus privileged knowledge [@keysar2000taking; @heller2008role].
 In the flowers game, like in reference games, all players saw the same set of images (and knew they did), which makes making descriptions to flowers easier.
@@
 In conclusion, this analysis showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
-Additionally, we saw hints that different incentives may lightly alter moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.
+Additionally, we saw hints that different incentives may lightly moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.
 We hope future work will continue to develop our understanding of these phenomena in more natural settings.
@@
 
-# Acknowledgements
-
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
-
 # References
@@
 \noindent
+\small

### 2023-01-26 write-ups/cogsci23_paper/manuscript.Rmd
add model code chunks

@@
 author-information: > 
-    \author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA
-    \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\ Department of Educational Psychology, 1025 W. Johnson Street \\ Madison, WI 53706 USA}
+    \author{{\large \bf Veronica Boyce (vboyce@stanford.edu)} \\ Department of Psychology, Stanford University
+    \AND {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, Stanford University}
 
@@
 #library(rstan)
-library(tidybayes)
+#library(tidybayes)
 library(viridis)
 library(cowplot)
-library(lsa)
+#library(lsa)
 library(kableExtra)
@@
 
+```{r, eval=F}
+within_game_flower <- flower_all |> 
+  filter(game1==game2) |> 
+  filter(ref1==ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(same_player=ifelse(player1==player2,1,0),
+         flower=str_c(color,ref1)) |> 
+  select(sim, condition, earlier, same_player, game1, flower )
+
+m_samegame_sameflower <-  stan_lmer(sim ~ condition*earlier*same_player + 
+                                      (1|game1)+(1|flower), data=within_game_flower) |> write_rds(here(model_path,"m2.rds"))
+
+read_rds(here(model_path,"m2.rds")) |> 
+show_summary_rstan() |> write_rds(here(model_path,"samegame_sameflower_model.rds"))
+
+within_game_between_flower <- flower_all |> 
+  filter(game1==game2) |> 
+  filter(ref1!=ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(same_player=ifelse(player1==player2,1,0)) |> 
+  select(sim, condition, earlier, same_player, game1)
+
+m_samegame_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=within_game_between_flower) |> write_rds(here(model_path,"m3.rds"))
+
+read_rds(here(model_path,"m3.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"samegame_diffflower_model.rds"))
+
+within_game_between_flower <- flower_all |> 
+  filter(game1==game2) |> 
+  filter(ref1!=ref2) |> 
+  filter(later-earlier<3) |> 
+  mutate(same_player=ifelse(player1==player2,1,0)) |> 
+  select(sim, condition, earlier, same_player, game1)
+
+m_samegame_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=within_game_between_flower) |> write_rds(here(model_path,"m3.rds"))
+
+read_rds(here(model_path,"m3.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"samegame_diffflower_model.rds"))
+
+```
+
 ```{r}
@@
 
+# note sbert_examples are generated by running sbert process just on include 
 examples <- read_rds(here(model_path,"sbert_examples.rds")) |> left_join(include, by=c("game1"="gameId","ref1"="flower","player1"="name","trialNum1"="trialNum")) |> rename(text1=normalized_span) |> left_join(include, by=c("game2"="gameId","ref2"="flower","player2"="name","trialNum2"="trialNum")) |> rename(text2=normalized_span)
@@
 
+```{r, eval=F}
+to_end_same_flower <- to_end |> 
+  filter(ref1==ref2) |> 
+  mutate(same_player=ifelse(player1==player2,1,0),
+         flower=str_c(color,ref1)) |> 
+  select(sim, condition, earlier, same_player, gameId, flower)
+
+m_to_end_sameflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_same_flower) |> write_rds(here(model_path,"m5.rds"))
+
+read_rds(here(model_path,"m5.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"to_end_sameflower_model.rds"))
+
+to_end_diff_flower <- to_end |> 
+  filter(ref1!=ref2) |> 
+  mutate(same_player=ifelse(player1==player2,1,0)) |> 
+  select(sim, condition, earlier, same_player, gameId)
+
+m_to_end_diffflower <-  stan_glm(sim ~ condition*earlier*same_player, data=to_end_diff_flower) |> write_rds(here(model_path,"m6.rds"))
+
+read_rds(here(model_path,"m6.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"to_end_diffflower_model.rds"))
+
+within_end <- in_end |> 
+  filter(seen1=="own" & seen2=="own") |> 
+  filter(player1!=player2) |> 
+  mutate(same_flower=ifelse(ref1==ref2, 1, 0),
+         same_game=ifelse(game1==game2, 1,0)) |> 
+  select(sim, condition, same_flower, same_game)
+
+m_within_end <-  stan_glm(sim ~ condition*same_flower*same_game, data=within_end) |> write_rds(here(model_path,"m7.rds"))
+
+read_rds(here(model_path,"m7.rds")) |> show_summary_rstan() |> write_rds(here(model_path,"within_end.rds"))
+
+```
+
 ```{r}
@@
 
-```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
-#img <- png::readPNG("figs/walrus.png")
-#grid::grid.raster(img)
-```
-
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
-#img <- png::readPNG("figs/lab_logo_stanford.png")
-#grid::grid.raster(img)
-```
-
 # References

### 2023-01-27 write-ups/cogsci23_paper/manuscript.Rmd
edits from Mike; fixed a number of incorrect statements in results

@@
 ---
-title: "Communicative reduction in referring expressions in a multi-player negotation game"
+title: "Communicative reduction in referring expressions within a multi-player negotation game"
 bibliography: [ aaflowers.bib]
@@
 abstract: >
-    Reduction of referring expressions to short conventionalized terms is found robustly in dyadic repeated reference games when the target images are initially difficult to name.
-Establishing reference is crucial to many conversations with different goals and more complex motivations; thus, it is useful to look at how the findings from reference games, where participants share the explicit goal of establishing joint reference, carry over into situations where participants have different overarching goals.
-In the current work, we analyse a dataset where reference was embedded in strategic 3- player negotiation and coordination games.
-In these more complex games, we found that the patterns of reduction and convergence to conventions held across two different incentive conditions, with some suggestive differences between the conditions.
+    The ability to form novel conventions is a key signature of efficient linguistic conmmunication.
+Reduction of referring expressions, one measure of convention formation, is found robustly in dyadic repeated reference games when the target images are initially difficult to name.
+In reference games, participants share the explicit goal of establishing joint reference.
+However, establishing reference is a key subgoal of many conversations where interlocuters have more complex goals.
+In the current work, we analyze a dataset where reference was embedded in strategic 3- player negotiation and coordination games.
+In these more complex games, we found that the patterns of reduction and convergence to conventions still held across two different incentive conditions, with some suggestive differences between the conditions.
 
 keywords: >
-    Convention formation; reference games; reduction of referring expressions; multi-party communication;
+    Convention formation; reference games; reduction of referring expressions; multi-party communication
 
@@
 
-Successful communication is grounded in a shared understanding of what utterances mean in the context they are produced in.
+Successful communication is grounded in a shared understanding of what utterances mean in the context of their production.
 Whether the communicative goals are cooperative or adversarial, interlocuters need to establish joint reference to be effective.
-In many cases conventional word meanings are enough, but there are also contexts where objects to be referred to are not easy to distinguish.
+In many cases, conventional word meanings are enough, but there are also contexts where objects to be referred to are not easy to distinguish.
 In these situations, establishing reference is more difficult, but no less important.
-<!--For one, in everyday communication, establishing reference rarely happens in isolation but is usually a subgoal in a larger conversational goal.
-The overall goal might be cooperative, such as in asking for a piece of kitchen equipment while cooking together, but it can also be adversarial, such as in a negotiation where the assets to be divied up must be referenced.
-In the reference game context, the goal is always cooperative, to match the images or get them in the same order.
--->
 
-The formation of these ad-hoc referring expressions has been studied extensively in the context of dyadic reference games [@clarkReferringCollaborativeProcess1986; @krauss1964changes].
+The formation of ad-hoc referring expressions has been studied extensively in the context of dyadic reference games [@clarkReferringCollaborativeProcess1986; @krauss1964changes].
 In these, two participants see a set of images (often abstract shapes) and one person is tasked with identifying an image so the other person can pick it out.
@@
 These nicknames conventionalize with a pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
-This has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, and when they change [@brennan1996conceptual; @metzing2003conceptual; @leung2020parents; @weber2003; @yoonAudienceDesignMultiparty2019].
+This task has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, and when they change [@brennan1996conceptual; @metzing2003conceptual; @leung2020parents; @weber2003; @yoonAudienceDesignMultiparty2019].
 
-The implicit theoretical claim is that reduction to partner-specific conventions occurs whenever people interact repeatedly in ways that require reference to some objects without adequate conventional names.
-Do these phenomena generalize beyond artificial reference game environments?
+The implicit theoretical claim is that reduction to partner-specific conventions occurs *whenever* people interact repeatedly in ways that require reference to some objects without adequate conventional names.
+But do these phenomena generalize beyond artificial reference game environments?
 
-Our current work begins to address this by looking for these phenomena in a slightly more naturalistic setting.
+Our current work begins to address this question  by looking for reduction phenomena in a slightly more naturalistic setting.
 The conversational goals of reference games are not representative of typical language use.
@@
 The overall task may be cooperative, such as asking for kitchen equipment while cooking together, or adversarial, such as identifying items while negotiating a division of assets.
-In the reference game context, the goal is explicitly cooperative: match corresponding images.
+In contrast, in the reference game context, the goal is explicitly cooperative: match corresponding images.
 
@@
 Panel B shows the feedback stage, after all players have selected.
+Players see player icons on the value bars indicating who chose what.
 When multiple players select the same flower, they recieve a lower value rather than what is shown.
@@
 
-@mankewitz2021 bridged the gap between reference games with hard-to-name objects and negotation datsets with easily nameable objects.
+We build on work by @mankewitz2021 who created a task with the aim of bridging the gap between reference games with hard-to-name objects and negotiation datsets with easily nameable objects.
 They created a 3-person game where players each selected what flowers to grow from a set of 6 flowers with variable and partially hidden value.
@@
 @mankewitz2021 found a slight decrease in language over the course of the games, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
-Here, we separated out the referring expressions in order to check for patterns of convention-formation.
+Here, we address this gap by separating out the referring expressions in order to check for patterns of convention-formation.
 
-Our current work addresses whether participants in a complex, multi-player strategic game with differing incentives and shared knowledge were able to develop conventionalized reference expressions.
-We explored whether the phenomena found in classic dyadic iterated reference games extended to this more naturalistic domain, and whether they differed across the individual and shared incentive conditions.
+Our work addresses whether participants in a complex, multi-player strategic game with differing incentives and shared knowledge are able to develop conventionalized reference expressions.
+We explore whether the phenomena found in classic dyadic iterated reference games extends to this more naturalistic domain, and whether the results differ across the individual and shared incentive conditions.
 The key questions we addressed were:
@@
 We take this level of interannotator agreement as an indication that the reference spans and targets were identified reliably.
-
 We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions.
@@
 For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
-There was not an effect of utterances being produced by the same person (`r stats_text(wi_g_bet_f,8)`), or interactions between who said it and trial number (`r stats_text(wi_g_bet_f,7)`) .
+Utterances being produced by the same person did not have an effect (`r stats_text(wi_g_bet_f,8)`) or an interaction with trial number (`r stats_text(wi_g_bet_f,7)`) .
 
-These results are consistent with the reference game phenomenon.
+These results replicate and generalize the pattern found in the prior literature [@hawkins2020].
 
@@
 Games are numbered, letters refer to players within games.
-Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.
-\\label{extable}",  format="latex", booktabs=TRUE, linesep="") %>%
+Images of flowers can be found in Figure 1: flower 1 upper left, flower 2 lower center, flower 3 lower left.\\label{extable}",  format="latex", booktabs=TRUE, linesep="") %>%
 kable_styling(full_width = F) %>%
@@
 
-For descriptions of the same flower across games, we found that cosine similarities decreased slightly over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} upper right panel).
-This is weakly consistent with the predictions.
+For descriptions of the same flower across games, we found that cosine similarities decreased over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} upper right panel).
+This decrease is consistent with the predictions.
 
@@
 Referring descriptions later in a game were more similar to the convention than descriptions earlier in the game  (trial: `r stats_text(to_e_wi_f,6)`, see Figure \ref{toend} upper panels).
-Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), although this did not interact with trial number (`r stats_text(to_e_wi_f,7)`).
+Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), and this increased over time (`r stats_text(to_e_wi_f,7)`).
 This is consistent with the expected pattern of convention formation.
@@
 ### Within games
-Games in the shared utilities condition had slightly higher cosine similarities between descriptions of the same flower within games (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or whether the utterances were produced by the same participant (`r stats_text(wi_g_wi_f,5)`).
-As a complement to this, games in the shared utilities condition had slightly lower similarities between descriptions of different flowers within games, but this did not interact with trial (`r stats_text(wi_g_bet_f,3)`) or whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
+Games in the shared utilities condition had non-significantly more similar descriptions of the same flower within games (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or whether the utterances were produced by the same participant (`r stats_text(wi_g_wi_f,5)`).
+Games in the shared utilities condition did have lower similarities between descriptions of different flowers within games (`r stats_text(wi_g_bet_f,2)`), but this was moderated by trial (`r stats_text(wi_g_bet_f,3)`) and whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
 
-Comparing to the end conventions, games in the shared utility condition produced flower descriptions that were more similar to the convention for that flower (`r stats_text(to_e_wi_f,2)`), but this did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
-In the shared utility condition, descriptions were slightly less similar to the conventions for other flowers (`r stats_text(to_e_bet_f,2)`), but this did not interact with trial (`r stats_text(to_e_bet_f,3)`).
-There was an interaction between shared utility condition and the referring expression coming from the same person who gave the convention for the different flower (`r stats_text(to_e_bet_f,5)`).
-Taken together, groups in the shared utility had more internal alignment on the mapping between flowers and names.
+Comparing to the end conventions, games in the shared utility condition produced flower descriptions that were more similar to the convention for that flower (`r stats_text(to_e_wi_f,2)`), but condition did not interact with trial number (`r stats_text(to_e_wi_f,3)`) and did not significantly interact with same participant versus different participant (`r stats_text(to_e_wi_f,5)`).
+In the shared utility condition, descriptions were slightly more similar to the conventions for other flowers (`r stats_text(to_e_bet_f,2)`), but condition did not interact with trial (`r stats_text(to_e_bet_f,3)`).
+This similarity was even higher if the description and convention came from the same person (`r stats_text(to_e_bet_f,5)`).
+As a whole, groups in the shared utility may have had more internal alignment on the mapping between flowers and names, but this pattern was inconsistent.
 
 ### Between games
-Across different games, games in the shared utility condition tended to have more similar descriptions for the same flower (`r stats_text(bet_g_wi_f,2)`); this did not interact with trial (`r stats_text(bet_g_wi_f,3)`).
+Across different games, games in the shared utility condition tended to have more similar descriptions for the same flower (`r stats_text(bet_g_wi_f,2)`), and this increased over time (`r stats_text(bet_g_wi_f,3)`).
 
-Lastly, we can look at how the conventionalized names compared, both between members of the same group and across players in different groups that saw the same color palette of flowers, and whether this differs by condition.
+Lastly, we can look at how the conventionalized names compared, both between members of the same group and across players in different groups that saw the same color palette of flowers, and whether these comparisons differ by condition (Figure \ref{withinend}).
 
-We treat different flowers described by different games as the baseline condition.
-For individual utility games, descriptions from groupmates are more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower are more similar (even across games, `r stats_text(wi_e,6)`), and there is a large interaction effect, where descriptions of the same flower are very similar among groupmates (`r stats_text(wi_e,7)`).
+We treated different flowers described by different games as the baseline condition.
+For individual utility games, descriptions from groupmates were more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower were more similar (even across games, `r stats_text(wi_e,6)`), and there was a large interaction effect, where descriptions of the same flower were very similar among groupmates (`r stats_text(wi_e,7)`).
 
-For shared utility games compared to individual utility games, descriptions are roughly equivalent at baseline (`r stats_text(wi_e,2)`) and for descriptions from groupmates (`r stats_text(wi_e, 5)`).
-Descriptions are much more similar for the same flower (even across games, `r stats_text(wi_e,3)`), but this is balanced out by a negative interaction effect between same flower and same game (`r stats_text(wi_e,4)`).
+For shared utility games compared to individual utility games, descriptions were roughly equivalent at baseline (`r stats_text(wi_e,2)`) and for descriptions from groupmates (`r stats_text(wi_e, 5)`).
+Descriptions were much more similar for the same flower (even across games, `r stats_text(wi_e,3)`), but this was balanced out by a negative interaction effect between same flower and same game (`r stats_text(wi_e,4)`).
 
-These findings are suggestive that in the shared utility condition, games differentiated less from each other, and their conventions may have been influenced more by shared priors for how to describe the flowers.
+These findings are suggestive that in the shared utility condition, games differentiated less from each other, and their conventions may have been influenced more by shared priors for how to describe the flowers than the conventions in individual utility games.
 
 # Discussion
-
+Our main question was whether the reduction phenomena found in dyadic reference games would occur within a negotiation game.
 This analysis of the referring expressions in @mankewitz2021 examined three key questions: the reduction of referring expressions, the convergence of referring expressions to conventions, and the differences between the shared and individual utility conditions.
@@
 Our third consideration was how the different utility conditions moderated the reduction and convergence effects.
-We found slight condition differences where being in the shared utility condition meant more similar descriptions for the same flowers and more different descriptions for different flowers.
+We found some slight and hard to interpret condition differences within games.
 In addition, groups in the shared utility condition diverged less from each other.
@@
 
-Our interpretation of results is limited by the dataset, which is small, especially compared to recent datasets of convention formation in reference games [@hawkins2020, @boyce2022].
-The dataset also has large item-level variation, as the images differ in how nameable the flowers are or whether there are salient features.
+Our interpretation of results is limited by the dataset, which is small, especially compared to recent datasets of convention formation in reference games [@hawkins2020; @boyce2022].
+The dataset also has large item-level variation, as the images differ in how nameable the flowers are or how salient background features are.
 
@@
 In the flowers game, like in reference games, all players saw the same set of images (and knew they did), which makes making descriptions to flowers easier.
-In real-world situations, interlocuters may have different sets of objects in their context and may not know what is in their interlocuters context.
-An important future direction is studying how repeated reference phenomena interact with different communication forms such as goals and contexts.
+In real-world situations, interlocuters may have different sets of objects in their context and may not know what is in their interlocuter's context.
+An important future direction is studying how repeated reference phenomena interact with different communcation goals and contexts.
 
-In conclusion, this analysis showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
+In conclusion, our analysis here showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
 Additionally, we saw hints that different incentives may lightly moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.

### 2023-01-30 write-ups/cogsci23_paper/manuscript.Rmd
fix typos

@@
 abstract: >
-    The ability to form novel conventions is a key signature of efficient linguistic conmmunication.
+    The ability to form novel conventions is a key signature of efficient linguistic communication.
 Reduction of referring expressions, one measure of convention formation, is found robustly in dyadic repeated reference games when the target images are initially difficult to name.
@@
 In the current work, we analyze a dataset where reference was embedded in strategic 3- player negotiation and coordination games.
-In these more complex games, we found that the patterns of reduction and convergence to conventions still held across two different incentive conditions, with some suggestive differences between the conditions.
+In these more complex games, we found that the patterns of reduction and convergence to conventions still held across two different incentive conditions, with some modest differences between the conditions.
 
@@
 Successful communication is grounded in a shared understanding of what utterances mean in the context of their production.
-Whether the communicative goals are cooperative or adversarial, interlocuters need to establish joint reference to be effective.
+Whether the communicative goals are cooperative or adversarial, interlocutors need to establish joint reference to be effective.
 In many cases, conventional word meanings are enough, but there are also contexts where objects to be referred to are not easy to distinguish.
@@
 Over repeated reference to the same images, pairs develop shorthand names for the images, leading to shorter descriptions in later repetitions.
-These nicknames conventionalize with a pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
+These nicknames conventionalize within each pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
 This task has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, and when they change [@brennan1996conceptual; @metzing2003conceptual; @leung2020parents; @weber2003; @yoonAudienceDesignMultiparty2019].
@@
 
-We build on work by @mankewitz2021 who created a task with the aim of bridging the gap between reference games with hard-to-name objects and negotiation datsets with easily nameable objects.
+We build on work by @mankewitz2021 who created a task with the aim of bridging the gap between reference games with hard-to-name objects and negotiation datasets with easily nameable objects.
 They created a 3-person game where players each selected what flowers to grow from a set of 6 flowers with variable and partially hidden value.
@@
 Our work addresses whether participants in a complex, multi-player strategic game with differing incentives and shared knowledge are able to develop conventionalized reference expressions.
-We explore whether the phenomena found in classic dyadic iterated reference games extends to this more naturalistic domain, and whether the results differ across the individual and shared incentive conditions.
+We explore whether the phenomena found in classic dyadic iterated reference games extend to this more naturalistic domain, and whether the results differ across the individual and shared incentive conditions.
 The key questions we addressed were:
@@
 
-Historically, studies of convention formation have had to rely on proxies such as reduction, subjective judgements of similarity, and and manual measures of lexical overlap.
+Historically, studies of convention formation have had to rely on proxies such as reduction, subjective judgments of similarity, and manual measures of lexical overlap.
 In contrast, we use natural language processing tools for mapping sentences to semantic vector spaces to quantify semantic similarities between sentences, following recent work by @hawkins2020 and @boyce2022.
@@
 Each participant saw the values for 4 of the flowers (represented as a colored bar), such that each flower's value was hidden from one participant.
-Players could coordinate and discuss using a chat box before each selected a flower.
+Players could coordinate and discuss using a chat box before each player selected a flower.
 If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they each got a lower reward instead (Figure \ref{interface}B).
 Players needed to communicate about the flowers in order to coordinate their selections.
-The rewards translated in a monetary bonus for the participants at the end of the game.
+The rewards translated into a monetary bonus for the participants at the end of the game.
 
@@
 ## Participants
-@mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants for 25 3-person games in each of the individual and shared utilities conditions.
+@mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants total for 25 3-person games in each of the individual and shared utilities conditions.
 Some games did not complete the full 24 trials because participants dropped out early.
@@
 Two games consisting of `r raw_rows` utterances were annotated by both annotators.
-`r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
+`r nrow(both)` utterances were identified as containing reference expressions by at least one annotator; of these, the annotators agreed on the exact reference expression for `r agree_span` (`r round(agree_span/nrow(both)*100)`%) of the cases.
 The second annotator only coded the target of the referring expression for `r nrow(count)` of the utterances.
 Of these, the two annotators agreed on the target in `r agree_target` (`r round(agree_target/nrow(count)*100)`%) of cases.
-We take this level of interannotator agreement as an indication that the reference spans and targets were identified reliably.
+We take this level of inter-annotator agreement as an indication that the reference spans and targets were identified reliably.
 We extracted a total of `r annotated |> filter(numword>0) |>  tally()` referring expressions.
@@
 ## SBERT embeddings
-Following recent work by @hawkins2020 and @boyce2022, we used tools for natural langauge processing to quantify similarity between phrases in order to assess the dynamics of convergence to conventions.
+Following recent work by @hawkins2020 and @boyce2022, we used tools for natural language processing to quantify similarity between phrases in order to assess the dynamics of convergence to conventions.
 
 We embedded each of the extracted referring expressions using SBERT [@reimers2019] which maps each utterance to a vector in a high-dimensional semantic space.
-We use cosine distance between pairs of vector embeddings as a measure of similarity between the corresponding referring expressions.
+We used cosine distance between pairs of vector embeddings as a measure of similarity between the corresponding referring expressions.
 Higher cosine values correspond to vectors close together, and thus, reference expressions with more similar meanings.
@@
 The theory of partner-specificity predicts that descriptions to the same flower in different groups will diverge over time as each group forms their own distinct convention.
-Table \ref{extable} shows how groups descriptions of the same flower may diverge (flower 2), but also how groups might choose related but different descriptions (flower 3).
+Table \ref{extable} shows how descriptions of the same flower may diverge (flower 2), but also how groups might choose related but different descriptions (flower 3).
 
@@
 Games in the shared utilities condition had non-significantly more similar descriptions of the same flower within games (`r stats_text(wi_g_wi_f,2)`), but this did not interact with trial number (`r stats_text(wi_g_wi_f,3)`) or whether the utterances were produced by the same participant (`r stats_text(wi_g_wi_f,5)`).
-Games in the shared utilities condition did have lower similarities between descriptions of different flowers within games (`r stats_text(wi_g_bet_f,2)`), but this was moderated by trial (`r stats_text(wi_g_bet_f,3)`) and whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
+Games in the shared utilities condition did have lower similarities between descriptions of different flowers within games (`r stats_text(wi_g_bet_f,2)`), but this was moderated by trial (`r stats_text(wi_g_bet_f,3)`) and by whether the utterances were produced by the same player (`r stats_text(wi_g_bet_f,5)`).
 
@@
 This similarity was even higher if the description and convention came from the same person (`r stats_text(to_e_bet_f,5)`).
-As a whole, groups in the shared utility may have had more internal alignment on the mapping between flowers and names, but this pattern was inconsistent.
+As a whole, groups in the shared utility condition may have had more internal alignment on the mapping between flowers and names, but this pattern was inconsistent.
 
@@
 
-Lastly, we can look at how the conventionalized names compared, both between members of the same group and across players in different groups that saw the same color palette of flowers, and whether these comparisons differ by condition (Figure \ref{withinend}).
+### Comparison of end conventions
+Lastly, we looked at how the conventionalized names compared, both between members of the same group and across players in different groups that saw the same color palette of flowers, and whether these comparisons differed by condition (Figure \ref{withinend}).
 
 We treated different flowers described by different games as the baseline condition.
-For individual utility games, descriptions from groupmates were more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower were more similar (even across games, `r stats_text(wi_e,6)`), and there was a large interaction effect, where descriptions of the same flower were very similar among groupmates (`r stats_text(wi_e,7)`).
+For individual games, descriptions from groupmates were more similar (even for different flowers, `r stats_text(wi_e,8)`), descriptions of the same flower were more similar (even across games, `r stats_text(wi_e,6)`), and there was a large interaction effect, where descriptions of the same flower were very similar among groupmates (`r stats_text(wi_e,7)`).
 
@@
 
-These findings are suggestive that in the shared utility condition, games differentiated less from each other, and their conventions may have been influenced more by shared priors for how to describe the flowers than the conventions in individual utility games.
+These findings are suggestive that in the shared utility condition, games differentiated less from each other, and their conventions may have been influence more by shared priors for how to describe the flowers more than the conventions in individual utility games were.
 
@@
 Second, we found the expected set of convergence and divergence patterns.
-References to the same flower within a game converged both to other descriptions used at that timepoint and to the post-game descriptions.
+References to the same flower within a game converged both to other descriptions used at that time point and to the post-game descriptions.
 This convergence was specific to flower and games, as references to different flowers within a game and references to the same flower across games both diverged.
@@
 Our third consideration was how the different utility conditions moderated the reduction and convergence effects.
-We found some slight and hard to interpret condition differences within games.
+We found some modest condition differences within games.
 In addition, groups in the shared utility condition diverged less from each other.
@@
 One dimension that we were unable to explore here due to the design of the game is the idea of shared versus privileged knowledge [@keysar2000taking; @heller2008role].
-In the flowers game, like in reference games, all players saw the same set of images (and knew they did), which makes making descriptions to flowers easier.
-In real-world situations, interlocuters may have different sets of objects in their context and may not know what is in their interlocuter's context.
-An important future direction is studying how repeated reference phenomena interact with different communcation goals and contexts.
+In the flowers game, like in reference games, all players saw the same set of images (and knew they did).
+In real-world situations, interlocutors may have different sets of objects in their context and may not know what is in their communication partner's context.
+An important future direction is studying how repeated reference phenomena interact with different communication goals and contexts.
 
@@
 \noindent
-\small

### 2023-02-01 write-ups/cogsci23_paper/manuscript.Rmd
final!

@@
 ---
-title: "Communicative reduction in referring expressions within a multi-player negotation game"
+title: "Communicative reduction in referring expressions within a multi-player negotiation game"
 bibliography: [ aaflowers.bib]
@@
 Over repeated reference to the same images, pairs develop shorthand names for the images, leading to shorter descriptions in later repetitions.
-These nicknames conventionalize within each pair, as they tend to stick to the same description, but are idiosyncratic and differ between pairs.
+These nicknames conventionalize within each pair, as pairs tend to stick to the same description, but nicknames are idiosyncratic and differ between pairs.
 This task has been a well-studied microcosm for understanding reference which has proven useful for testing theories about how referring expressions originate, how expressions are designed, and when they change [@brennan1996conceptual; @metzing2003conceptual; @leung2020parents; @weber2003; @yoonAudienceDesignMultiparty2019].
@@
 The implicit theoretical claim is that reduction to partner-specific conventions occurs *whenever* people interact repeatedly in ways that require reference to some objects without adequate conventional names.
+Indeed, reduction does occur regardless of modality, having been found in both oral and written communication [@brennan1996conceptual; @hawkins2020].
 But do these phenomena generalize beyond artificial reference game environments?
@@
 
-We build on work by @mankewitz2021 who created a task with the aim of bridging the gap between reference games with hard-to-name objects and negotiation datasets with easily nameable objects.
+We build on work by @mankewitz2021, who created a task with the aim of bridging the gap between reference games with hard-to-name objects and negotiation datasets with easily nameable objects.
 They created a 3-person game where players each selected what flowers to grow from a set of 6 flowers with variable and partially hidden value.
@@
 Our work addresses whether participants in a complex, multi-player strategic game with differing incentives and shared knowledge are able to develop conventionalized reference expressions.
-We explore whether the phenomena found in classic dyadic iterated reference games extend to this more naturalistic domain, and whether the results differ across the individual and shared incentive conditions.
+We explore whether the phenomena found in classic dyadic iterated reference games can also be found in this more naturalistic domain, and whether the results differ across the individual and shared incentive conditions.
 The key questions we addressed were:
 
-1. Did the referring expressions reduce over time?
+1. Did referring expressions reduce over time?
 
@@
 
-3. How did the different incentive conditions influence the referring expressions?
+3. How did different incentive conditions influence referring expression formation?
 
-Historically, studies of convention formation have had to rely on proxies such as reduction, subjective judgments of similarity, and manual measures of lexical overlap.
+Historically, studies of convention formation have had to rely on proxies for convention formation such as reduction, subjective judgments of similarity, and manual measures of lexical overlap.
 In contrast, we use natural language processing tools for mapping sentences to semantic vector spaces to quantify semantic similarities between sentences, following recent work by @hawkins2020 and @boyce2022.
@@
 
-For pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same trial and the two later trials (thus the similarity at trial 10 comes from 10-10, 10-11 and 10-12 comparisons).
+For pairwise comparisons between referring expressions produced during the game, we smoothed out the random sampling and sparseness of which flowers were presented on the same trial by comparing descriptions with those from the same trial and the two later trials (thus the similarity at trial 10 comes from comparisons of trial 10 to trials 10, 11, and 12).
 
@@
 
-We address each of the questions of interest listed at the end of the Introduction in order.
+We address each of the questions of interest in order: 1) did referring expressions reduce over time, 2) how referring expressions changed during games, and 3) how the incentive manipulation influenced referring expressions.
 
@@
 In each trial of the game, there were 6 flowers visible on the screen, and participants wanted to each pick a different one.
-In general, each trial contained references to 2 or 3 distinct flowers (individual utilities: mean of `r  mean_flowers |> pluck(2,1)`, shared utilities: `r mean_flowers |> pluck(2,2)`).
-Most players referred to one flower each trial (individual utilities: `r  mean_person |> pluck(2,1)`, shared utilities: `r mean_person |> pluck(2,2)`).
+In general, each trial contained references to 2 or 3 distinct flowers: individual utilities: mean of `r  mean_flowers |> pluck(2,1)`, shared utilities: `r mean_flowers |> pluck(2,2)`.
+Most players referred to one flower each trial: individual utilities: `r  mean_person |> pluck(2,1)`, shared utilities: `r mean_person |> pluck(2,2)`.
 
@@
 
-For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
-Utterances being produced by the same person did not have an effect (`r stats_text(wi_g_bet_f,8)`) or an interaction with trial number (`r stats_text(wi_g_bet_f,7)`) .
-
-These results replicate and generalize the pattern found in the prior literature [@hawkins2020].
-
 ```{r}
@@
 ```
+
+For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
+Utterances being produced by the same person did not have an effect (`r stats_text(wi_g_bet_f,8)`) or an interaction with trial number (`r stats_text(wi_g_bet_f,7)`).
+
+These results replicate and generalize the pattern found in prior literature: convergence in expressions for the same referent  games and divergence between referents [@hawkins2020].
+
 ### Between games
@@
 For descriptions of the same flower across games, we found that cosine similarities decreased over time  (trial: `r stats_text(bet_g_wi_f,4)`, see Figure \ref{during} upper right panel).
-This decrease is consistent with the predictions.
+This between-game decrease is also consistent with prior reports in more restricted reference games.
 
@@
 Utterances were more similar to the convention given by the same person (versus the convention given by a groupmate) (`r stats_text(to_e_wi_f,8)`), and this increased over time (`r stats_text(to_e_wi_f,7)`).
-This is consistent with the expected pattern of convention formation.
+This finding is consistent with the expected pattern of convention formation.
 
@@
 ```{r withinend, fig.env="figure", fig.pos = "h", fig.align = "center", out.width="90%", fig.width=4, fig.height=2, fig.cap = "Cosine similarities between flower descriptions provided by participants after the game.
+Error bars are 95\\% CIs, but some are hidden by the point size.
 \\label{withinend}"}
@@
 In addition, groups in the shared utility condition diverged less from each other.
-This pattern of condition differences is consistent with less idiosyncratic descriptions in the shared utility condition.
+This pattern of condition differences is consistent with less idiosyncratic descriptions in the shared utility condition, and the modest magnitude of the effect is consistent with the relatively subtle nature of this manipulation, which did not have a large effect on choice behavior [@mankewitz2021].
 

### 2023-04-12 write-ups/cogsci23_paper/manuscript.Rmd
cogsci final

@@
 output: cogsci2016::cogsci_paper
-#final-submission: \cogscifinalcopy
+final-submission: \cogscifinalcopy
 ---
@@
                       fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
+                      echo=F, warning=F, cache=T, 
                       message=F, sanitize = T)
@@
 @mankewitz2021 found a slight decrease in language over the course of the games, but they did not conduct a detailed analysis to assess whether the decreases in words were associated with shorter and more conventionalized referring expressions.
+However, this pattern of decreased language production is actually totally consistent with increased task familiarity, fatigue, and/or less communication about overall strategy, with no change in the actual referring language.
 Here, we address this gap by separating out the referring expressions in order to check for patterns of convention-formation.
@@
 
-This paper presents a reanalysis of data from @mankewitz2021. We describe both the original data collection and the additional data processing done for reanalysis.
+This paper presents a reanalysis of data from @mankewitz2021.
+We describe both the original data collection and the additional data processing done for reanalysis.
+The materials, data, and code are all available at https://github.com/vboyce/AA-flowers.
 
@@
 Players could coordinate and discuss using a chat box before each player selected a flower.
+There were no restrictions on what participants could say in the chat.
+
 If one player selected a flower, it was worth the shown reward; if multiple players collided and selected the same flower, they each got a lower reward instead (Figure \ref{interface}B).
@@
 @mankewitz2021 recruited `r 3 *(d.games |> filter(conditionName!="Old Cooperative") |> nrow())` participants total for 25 3-person games in each of the individual and shared utilities conditions.
-Some games did not complete the full 24 trials because participants dropped out early.
+Games took roughly 20 minutes with wide variability (5-40 minutes).
+If a participant disconnected, the game ended, so some games did not complete the full 24 trials as participants dropped out early.
 We excluded incomplete games from analyses, which left us with `r d.games.final |> filter(condition=="Individual Utilities") |> nrow()` games in the individual utilities condition and `r d.games.final |> filter(condition=="Shared Utilities") |> nrow()` in the shared utilities condition (`r 3*(d.games.final |> nrow())` participants included in the analyses).
@@
 
-        top <- plot_grid(all_words, words, utts,nrow=1, labels="AUTO")
+        top <- plot_grid(all_words, words, utts,nrow=1, labels="AUTO", hjust=c(-5, -4, -4), vjust=1.8)
         plot_grid(top, legend, nrow=2, rel_heights=c(1,.15))
@@
 
+Participants employed a range of reference strategies, including referring to the flowers by a common name (ex.
+rose, iris), referring to the number or groupings of flowers, referring to properties of the flowers (ex.
+spiky), referring to background elements (ex.
+white house), and analogies to other objects (ex.
+flame).
+Descriptions sometimes combined multiple of these approaches.
+
 The amount of referring language decreased over the course of the game, consistent with the dyadic reference game pattern.
@@
 
-The classic reference game phenomenon of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
+The classic reference game phenomena of convergence to partner-specific referring expressions can be quantified using similarity metrics as the combination of three patterns occurring over time: 1) within a group, descriptions of the same flower converge, 2) within a group, descriptions of different flowers diverge, and 3) between groups, descriptions of the same flower diverge.
 We assess all of these in two ways: 1) comparing referring expressions within the game to other referring expressions from the same time to get at the dynamic changes, and 2) comparing referring expressions within the game to the names players gave the flowers after the game, to track how expressions evolve towards these final conventions.
@@
 
+For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
+Utterances being produced by the same person did not have an effect (`r stats_text(wi_g_bet_f,8)`) or an interaction with trial number (`r stats_text(wi_g_bet_f,7)`).
+
 ```{r}
@@
 
-knitr::kable(text_examples, caption = "Examples of descriptions of different flowers from different games, illustrating reduction and convergence phenomenon.
+knitr::kable(text_examples, caption = "Examples of descriptions of different flowers from different games, illustrating reduction and convergence phenomena.
 Games are numbered, letters refer to players within games.
@@
 column_spec(4, width = "12em") %>%
-  kable_styling(latex_options="H")
-
-```
-
-```{r sbert-example}
-sbert_include <- tribble(
-  ~text1, ~text2,
-  #"the red center", "the one with the dark red centre",
-  "the red center", "red middle",
-  "droopy iris flower", "droopy",
-  "cluster", "bundle",
-  "droopy iris flower","the red center", 
-  "droopy iris flower","multi flowers", 
-  "red middle", "black background",
-  "droopy", "bundle",
-) |> left_join(examples) |> arrange(desc(sim)) |> mutate(sim=round(sim,2)) |> select(`Expression 1`=text1, `Expression 2`=text2, `Sim`=sim)
-
-knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions. \\label{costable}",  format="latex", booktabs=TRUE, linesep="") %>%
-kable_styling(full_width = F) %>%
-column_spec(c(1,2), width = "9em") %>% 
   kable_styling(latex_options="H")
@@
 ```
-
-For descriptions of different flowers, cosine similarity decreased over time (trial: `r stats_text(wi_g_bet_f,6)`, see Figure \ref{during} lower left and middle panels).
-Utterances being produced by the same person did not have an effect (`r stats_text(wi_g_bet_f,8)`) or an interaction with trial number (`r stats_text(wi_g_bet_f,7)`).
 
@@
   ungroup()
-```
-
-```{r toend, fig.env="figure", fig.pos = "h", fig.align = "center", out.width="100%", fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.
-\\label{toend}", fig.height=4, fig.width=6}
-to_end|> 
-  mutate(distance=case_when(
-    player1==player2 ~ "Same Person",
-    T ~ "Same Game - Different Person"),
-        same_flower=ifelse(ref1==ref2, "Same Flower", "Different Flowers")
-) |> mutate(same_flower=factor(same_flower,levels=c("Same Flower", "Different Flowers")),
-              distance=factor(distance, levels=c("Same Person", "Same Game - Different Person" ))) |> 
-ggplot( aes(x=earlier,y=sim,color=condition))+     
-  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
-  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
-        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
-        ggthemes::scale_color_solarized()+
-  guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
-
 ```
@@
 
-We also examined how referring expressions related to the conventionalized names for other flowers.
-Utterances diverge from the conventionalized names of other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
-There was no effect of the same person saying the description and convention (`r stats_text(to_e_bet_f,8)`) and no interaction between same person and trial number (`r stats_text(to_e_bet_f,7)`).
+```{r sbert-example}
+sbert_include <- tribble(
+  ~text1, ~text2,
+  #"the red center", "the one with the dark red centre",
+  "the red center", "red middle",
+  "droopy iris flower", "droopy",
+  "cluster", "bundle",
+  "droopy iris flower","the red center", 
+  "droopy iris flower","multi flowers", 
+  "red middle", "black background",
+  "droopy", "bundle",
+) |> left_join(examples) |> arrange(desc(sim)) |> mutate(sim=round(sim,2)) |> select(`Expression 1`=text1, `Expression 2`=text2, `Sim`=sim)
+
+knitr::kable(sbert_include, caption = "Examples of cosine similarities between pairs of descriptions. \\label{costable}",  format="latex", booktabs=TRUE, linesep="") %>%
+kable_styling(full_width = F) %>%
+column_spec(c(1,2), width = "9em") %>% 
+  kable_styling(latex_options="H")
+
+```
 
@@
 ```
+
+```{r toend, fig.env="figure", fig.pos = "h", fig.align = "center", out.width="100%", fig.cap = "Cosine similarities between SBERT embeddings of utterances produced during a game and the post-game descriptions of the flowers.
+\\label{toend}", fig.height=4, fig.width=6}
+to_end|> 
+  mutate(distance=case_when(
+    player1==player2 ~ "Same Person",
+    T ~ "Same Game - Different Person"),
+        same_flower=ifelse(ref1==ref2, "Same Flower", "Different Flowers")
+) |> mutate(same_flower=factor(same_flower,levels=c("Same Flower", "Different Flowers")),
+              distance=factor(distance, levels=c("Same Person", "Same Game - Different Person" ))) |> 
+ggplot( aes(x=earlier,y=sim,color=condition))+     
+  stat_summary(aes(group=str_c(color,condition), color=condition),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.3, geom="point")+
+  geom_smooth(formula=y~poly(x,2))+facet_grid(same_flower~distance)+theme(legend.position="bottom", legend.title=element_blank())+
+        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
+        ggthemes::scale_color_solarized()+
+  guides(color = guide_legend(override.aes = list(linetype = NA, alpha=1, fill=NA, size=5) ) )
+
+```
+
+We also examined how referring expressions related to the conventionalized names for other flowers.
+Utterances diverge from the conventionalized names of other flowers over time (trial: `r stats_text(to_e_bet_f,6)`, see Figure \ref{toend} lower panels).
+There was no effect of the same person saying the description and convention (`r stats_text(to_e_bet_f,8)`) and no interaction between same person and trial number (`r stats_text(to_e_bet_f,7)`).
+
 ## Individual and shared utility conditions
@@
 
-In conclusion, our analysis here showed that the phenomenon of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
+In conclusion, our analysis here showed that the phenomena of reduction and conventionalization occurs even in complex games with different incentives, asymmetric knowledge, and more open-ended negotiation and dialogue than that found in reference games.
 Additionally, we saw hints that different incentives may lightly moderate the formation of conventions, but that reduction and convention formation took place regardless of incentive condition.

### 2023-06-14 write-ups/cogsci23_poster/document.tex
draft poster

\documentclass[10pt, xcolor=table, dvipsnames]{beamer}
\usepackage[scale=3]{beamerposter}
\setlength{\paperwidth}{72in} % A0 width: 46.8in
\setlength{\textwidth}{72in}
\setlength{\paperheight}{45 in} % A0 height: 33.1in
\setlength{\textwidth}{45 in}
\usepackage[overlay,absolute]{textpos}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows, fit}
\usepackage{tcolorbox}
\tcbuselibrary{raster}
\usepackage{enumitem}
\setlist{label=\textbullet} 
\usepackage{multicol}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\TPGrid[20mm,20mm]{1}{1}
\tcbset{%
	noparskip,
	colback=white, %background color of the box
	colframe=black, %color of frame and title background
	coltext=black, %color of body text
	coltitle=black, %color of title text 
	colbacktitle=Violet!30,
	boxrule=4pt,
	fonttitle=\bfseries,
}
\usepackage{xcolor}
\usepackage{soul}
\makeatletter
\let\HL\hl
\renewcommand\hl{%
	\let\set@color\beamerorig@set@color
	\let\reset@color\beamerorig@reset@color
	\HL}
\makeatother
\usepackage{pifont}
\newcommand{\cmark}{{\color{ForestGreen}\ding{51}}}
\newcommand{\xmark}{{\color{red}\ding{55}}}%
\newcommand{\qmark}{{\color{blue}\textbf{?}}}%
\begin{document}
	\begin{textblock}{1}(0.0,0)
		\begin{tcolorbox}[colback=Violet!30]{}

			\begin{centering} 
				\bigskip

				{\LARGE\textbf{Communicative reduction in referring expressions \\ within a
						multi-player negotiation game}}\\
				{\Large Veronica Boyce, Michael C. Frank} \\
				{Stanford University, send correspondence to vboyce@mit.edu}

			\end{centering}
		\end{tcolorbox}
	\end{textblock}

	\begin{textblock}{.32}(0,0.16)
		\begin{tcolorbox}[title= {\centering Goals }]
			\begin{small}\centering
				\bigskip
			The ability to form novel conventions is a key signature of efficient
			linguistic communication.
 In dyadic reference games, we commonly observe 
 \begin{itemize}
 	\item \textbf{reduction} as utterances shorter over time,
 	 \item \textbf{convergence} within groups to a shared nick name,
 	  \item \textbf{divergence} between groups as each focuses on different features.
   \end{itemize}  CITATIONS

Do these phenomena occur in strategic games with more complex goals?

			\end{small}
		\end{tcolorbox}
	\end{textblock}

	\begin{textblock}{.32}[0,1](0,1 )
		\begin{tcolorbox}[title={\centering Experimental design }]

			\includegraphics[clip,width=\textwidth]{interface-1.pdf}

			\begin{small}
				 During selection (A) each participant sees 6 flowers, 4 with value bars. Players can use a chat box to communicate before making selections. Then they see feedback (B) indicating who chose what.  When multiple players select the same flower, they recieve a lower value rather than what is shown.

				In \emph{individual utility} games (18 games), each player earned points for the
				flowers they selected; in the \emph{shared utility} games (21 games), the points
				were averaged together, and all players in a game got the same reward.
								Each group plays 24 trials with images drawn from a set of 12 flowers. This data was first presented in  Mankewitz et al. (2021).

								We measure distance between two utterances by the cosine of the angle between their embedings using SBERT (CITATION).

			\end{small}

		\end{tcolorbox}
	\end{textblock}

	\begin{textblock}{.41}[0,0](.33,.16)
	\begin{tcolorbox}[title={\centering Referring expressions reduce over time}]  
		\includegraphics[width=.95\textwidth]{wordcount-1.pdf}

	\begin{small}	Over time individual utterances shorten, but the same number of referring expressions are produced.     \end{small}                                              
	\end{tcolorbox}
\end{textblock}

		\begin{textblock}{.41}[0,0](.33,.43)

		\begin{tcolorbox}[title={\centering Expressions converge within groups}]  
			\includegraphics[width=.95\textwidth]{during-1.pdf}                            			                                                  
			\begin{small}

				Descriptions to the same flower in nearby trials increase in similarity with a group.

				Similarity decreases slightly for the same flower across games.

				Flower names become more distinctive as descriptions for two different flowers become less similar.

			\end{small}
		\end{tcolorbox}
	\end{textblock}

	\begin{textblock}{.41}[0,1](.33,1)
	\begin{tcolorbox}[title={ \centering 					Incentive conditions differed a bit}]  
		\begin{minipage}{.45\textwidth}
			\includegraphics[width=\textwidth]{withinend-1.pdf}    
		\end{minipage}
	\hfill
	\noindent
	\begin{minipage}{.45\textwidth}
				\begin{small}

				Individual and shared utilities conditions differed slightly in divergence rate during games (above). In post-game descriptions (left), shared utility games show greater cross-game similarity for the same flowers than individual utility.

				\end{small}
	\end{minipage}

	\end{tcolorbox}
\end{textblock}

		\begin{textblock}{.25}[1,0](1,.16)
	\begin{tcolorbox}[title={\centering Examples}]

		\centering
		\begin{footnotesize}
			Descriptions of different flowers from different games, illustrating reduction and convergence phenomenon. Images from experimental design: flower 1 upper left, flower 2 lower center, flower 3 lower left.
			\vspace{-1em}

			\centering
			\begin{tabular}[t]{llr>{\raggedright\arraybackslash}p{16em}}
				\toprule
				Flower & Game & Trial & Expression\\
				\midrule
				1 & 1A & 2 & not sure what kind of flower it is but the droopy-ish one\\
				1 & 1C & 4 & droopy iris flower\\
				1 & 1B & 21 & droopy\\
				\midrule
				2 & 2B & 2 & the red middle with spike\\
				2 & 2B & 3 & the red center\\
				2 & 2A & 20 & red middle\\
				2 & 3C & 6 & the one with the dark red centre\\
				2 & 3A & 13 & the one with black background\\
				2 & 3A & 24 & black background\\
				\midrule
				3 & 1A & 4 & the big cluster of flowers with the orange in the middle\\
				3 & 1A & 23 & cluster\\
				3 & 2C & 24 & bundle\\
				3 & 3B & 24 & multi flowers\\
				\bottomrule
			\end{tabular}

			\vspace{2em}
			Cosine similarities between pairs of descriptions.

			\vspace{-1em}
			\begin{tabular}[t]{>{\raggedright\arraybackslash}p{9em}>{\raggedright\arraybackslash}p{9em}r}
				\toprule
				Expression 1 & Expression 2 & Sim\\
				\midrule
				the red center & red middle & 0.78\\
				droopy iris flower & multi flowers & 0.56\\
				droopy iris flower & droopy & 0.56\\
				cluster & bundle & 0.25\\
				red middle & black background & 0.25\\
				droopy iris flower & the red center & 0.09\\
				droopy & bundle & 0.03\\
				\bottomrule
			\end{tabular}
		\end{footnotesize}

	\end{tcolorbox}
\end{textblock}
	\begin{textblock}{.25}[1,1](1,1)
		\begin{tcolorbox}[title= {\centering Conclusion}]
			\begin{small} Reduction and convergence patterns generalize to freer-form and
			more naturalistic domain of a negotiation game
			\end{small}
			\begin{scriptsize}

				\textbf{References: }\noindent
				TODO
			\end{scriptsize}
\end{tcolorbox}
\end{textblock}

\end{document}

### 2023-06-15 write-ups/cogsci23_poster/document.tex
work on poster

@@
 				{\Large Veronica Boyce, Michael C. Frank} \\
-				{Stanford University, send correspondence to vboyce@mit.edu}
+				{Stanford University, contact: vboyce@stanford.edu}
 
@@
 
-	\begin{textblock}{.32}(0,0.16)
+	\begin{textblock}{.32}(0,0.15)
 		\begin{tcolorbox}[title= {\centering Goals }]
+			\vspace{.5em}
 			\begin{small}\centering
 				\bigskip
-			The ability to form novel conventions is a key signature of efficient
-			linguistic communication.
- In dyadic reference games, we commonly observe 
+			\textbf{The ability to form novel conventions is a key signature of efficient
+			linguistic communication.}
+
+			\vspace{.5em}
+ In dyadic reference games (Clark \& Wilkes-Gibbs 1986, Hawkins et al. 2020), speaker-listener pairs show the following patterns:
  \begin{itemize}
  	\item \textbf{reduction} as utterances shorter over time,
- 	 \item \textbf{convergence} within groups to a shared nick name,
- 	  \item \textbf{divergence} between groups as each focuses on different features.
-   \end{itemize}  CITATIONS
+ 	 \item \textbf{convergence} within groups to a shared nickname,
+ 	  \item \textbf{divergence} between groups as different nicknames develop.
+   \end{itemize}
 
-Do these phenomena occur in strategic games with more complex goals?
+\vspace{.5em}
+\textbf{Do these patterns occur for reference expressions in strategic games with more complex goals?}
+\vspace{.5em}
 
@@
 		\begin{tcolorbox}[title={\centering Experimental design }]
+\vspace{.5em}
 
@@
 
+			\vspace{.5em}
 			\begin{small}
-				 During selection (A) each participant sees 6 flowers, 4 with value bars.
-Players can use a chat box to communicate before making selections.
-Then they see feedback (B) indicating who chose what.
-When multiple players select the same flower, they recieve a lower value rather than what is shown.
+
+				Participants were assigned to 3-player games. 
+				 During selection (A) each participant saw 6 flowers, 4 with value bars.
+Players could communicate in a chat box before each selecting a flower.
+Then they saw feedback (B) indicating who chose what.
+When multiple players selected the same flower, they received a lower value rather than the indicated value.
+
+\vspace{.5em}
 
@@
 				were averaged together, and all players in a game got the same reward.
-								Each group plays 24 trials with images drawn from a set of 12 flowers. This data was first presented in  Mankewitz et al. (2021).
-
-								We measure distance between two utterances by the cosine of the angle between their embedings using SBERT (CITATION).
+								Each group played 24 trials with images drawn from a set of 12 flowers. This data was first presented in  Mankewitz et al. (2021).
 
@@
 
+		\vspace{.5em}
 		\end{tcolorbox}
@@
 
-	\begin{textblock}{.41}[0,0](.33,.16)
-	\begin{tcolorbox}[title={\centering Referring expressions reduce over time}]  
-		\includegraphics[width=.95\textwidth]{wordcount-1.pdf}
+	\begin{textblock}{.41}[0,0](.33,.15)
+	\begin{tcolorbox}[title={\centering Referring expressions reduced over time}]  
+		\vspace{.5em}
+		\centering
+		\includegraphics[width=.9\textwidth]{wordcount-1.pdf}
 
-	\begin{small}	Over time individual utterances shorten, but the same number of referring expressions are produced.     \end{small}                                              
+	\begin{small}	Descriptions got shorter over time, but the number of descriptions stayed constant.   \end{small}                                              
 	\end{tcolorbox}
@@
 
-		\begin{textblock}{.41}[0,0](.33,.43)
+		\begin{textblock}{.41}[0,0](.33,.4)
 
-		\begin{tcolorbox}[title={\centering Expressions converge within groups}]  
-			\includegraphics[width=.95\textwidth]{during-1.pdf}                            			                                                  
+		\begin{tcolorbox}[title={\centering Referring expressions converged within groups}]  
+			\vspace{.5em}
+			\centering
+			\includegraphics[width=.9\textwidth]{during-1.pdf}                            			                                                  
 			\begin{small}
 
-				Descriptions to the same flower in nearby trials increase in similarity with a group.
+					Similarity between two descriptions was measured by the cosine between their S-BERT embeddings (Reimers \& Gurevych, 2019).
+				\begin{itemize}
 
-				Similarity decreases slightly for the same flower across games.
+				\item Similarity increased for the same flower within games (top left, center).
 
-				Flower names become more distinctive as descriptions for two different flowers become less similar.
+				\item Similarity decreased slightly for the same flower across games (top right).
 
+				\item Descriptions for different flowers became more distinct (bottom).
+				\end{itemize}
 			\end{small}
+		\vspace{.5em}
 		\end{tcolorbox}
@@
 	\begin{textblock}{.41}[0,1](.33,1)
-	\begin{tcolorbox}[title={ \centering 					Incentive conditions differed a bit}]  
-		\begin{minipage}{.45\textwidth}
+	\begin{tcolorbox}[title={ \centering 					Shared utilities games diverged less}]  
+		\begin{minipage}{.4\textwidth}
+			\vspace{.5em}
 			\includegraphics[width=\textwidth]{withinend-1.pdf}    
@@
 	\noindent
-	\begin{minipage}{.45\textwidth}
+	\begin{minipage}{.55\textwidth}
 				\begin{small}
+					\vspace{.5em}
+			Shared utility games diverged less from each other in how to refer to each flower, both during game (above, top right) and in post-game descriptions (left), possibly due to the shared utilities encouraging coordination and thus comprehension.
 
-				Individual and shared utilities conditions differed slightly in divergence rate during games (above).
-In post-game descriptions (left), shared utility games show greater cross-game similarity for the same flowers than individual utility.
+				\vspace{.2em}
+				\end{small}
 
-				\end{small}
 	\end{minipage}
@@
 
-		\begin{textblock}{.25}[1,0](1,.16)
+		\begin{textblock}{.25}[1,0](1,.15)
 	\begin{tcolorbox}[title={\centering Examples}]
+		\vspace{.2em}
 
@@
 		\begin{footnotesize}
-			Descriptions of different flowers from different games, illustrating reduction and convergence phenomenon.
-Images from experimental design: flower 1 upper left, flower 2 lower center, flower 3 lower left.
+			Descriptions of different flowers from different games. Images shown in  experimental design: flower 1 upper left, flower 2 lower center, flower 3 lower left.
 			\vspace{-1em}
@@
 
-			\vspace{2em}
+			\vspace{1em}
 			Cosine similarities between pairs of descriptions.
 
-			\vspace{-1em}
+			\vspace{-1.5em}
 			\begin{tabular}[t]{>{\raggedright\arraybackslash}p{9em}>{\raggedright\arraybackslash}p{9em}r}
@@
 		\end{footnotesize}
-
+		\vspace{.5em}
 	\end{tcolorbox}
@@
 		\begin{tcolorbox}[title= {\centering Conclusion}]
-			\begin{small} Reduction and convergence patterns generalize to freer-form and
-			more naturalistic domain of a negotiation game
+			\vspace{.5em}
+
+			\begin{small}
+
+				 Reduction and 
+				and conventionalization occur even in strategic games with more open-ended negotiation and more complex goals than reference games.
+
 			\end{small}
-			\begin{scriptsize}
 
-				\textbf{References: }\noindent
-				TODO
-			\end{scriptsize}
+		\vspace{.5em}
+		\begin{tiny}
+			References: Clark, H., \& Wilkes-Gibbs, D. (1986). Referring as a collaborative
+			process. $\bullet$
+			Reimers, N., \& Gurevych, I. (2019). Sentence-{BERT}:
+				{Sentence Embeddings} using {Siamese BERT-Networks}. $\bullet$ 
+Hawkins, R., Frank, M., \& Goodman, N. (2020). Characterizing
+			the dynamics of learning in repeated reference games. $\bullet$
+		Mankewitz, J., et al. (2021). Multi-party referential communication in
+	complex strategic games.
+
+			\end{tiny}
+
 \end{tcolorbox}
